### 2024-02-05 README.Rmd
start a bit of analysis

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# archyjobads

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/archyjobads/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): Quarto source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `paper.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: data](/analysis/data): Data used in the analysis.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder:
    supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2024-02-05 analysis/paper/paper.qmd
start a bit of analysis

---
title: "Title Goes Here"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

# Introduction

Here is a citation [@Marwick2017]

# Background

# Methods

# Results

<!-- Here's some example analysis code: -->

```{r}
#| label: get-data

library(tidyverse)
library(ggbeeswarm)
# This CSV file was downloaded from our data sheet here
# https://docs.google.com/spreadsheets/d/1Jwe3UqJyedrV-QWlwR_44__t4xBVrCfxGyhXdi3E0sg/edit?resourcekey#gid=1686084773
# note that you may need to download it again to get the latest updates!

jobdata <- read_csv(here::here('analysis/data/raw_data/Tenure Track Job Advertisements in Archaeology (Responses) - Form Responses 1.csv')) %>% 
  # simplify the column names 
  janitor::clean_names()

total_number_of_ads_in_our_sample <- nrow(jobdata)
```

We have `r total_number_of_ads_in_our_sample` job advertisements in our sample

```{r}
#| label: fig-how-many-jobs-per-year
#| fig-cap: "Number of tenure-track archaeology faculty job ads posted each year"

# we can get the year from the URL to the Academic Job Ads Wiki

year_ad_posted <- 
jobdata %>% 
  pull(url_to_data_source_e_g_paste_in_url_to_the_jobs_wiki_page) %>% 
  str_extract(.,  "[[0-9]]{4}-[[0-9]]{4}")

jobdata <- 
  jobdata %>% 
  mutate(year_ad_posted = year_ad_posted) %>% 
  drop_na()

ggplot(jobdata) +
  aes(year_ad_posted) +
  geom_bar() +
  scale_x_discrete(name = "Academic Year") +
  theme_minimal()
```

@fig-how-many-jobs-per-year shows how many jobs per year in our sample

```{r}
#| label: ratio-tt-to-non-tt

# ratio of tenure-track to untenured positions

base_url <- "http://academicjobs.wikia.com/wiki/Archaeology_Jobs_"

# starts at 2010-2011

years <- map_chr(2010:2019, ~str_glue('{.x}-{.x +1}'))
# though it seems to start at 2007-8: https://academicjobs.fandom.com/wiki/Archaeology_07-08

urls_for_each_year <- str_glue('{base_url}{years}')

library(rvest)

# all years
urls_for_each_year_headers <- 
map(urls_for_each_year,
    ~.x %>% 
      read_html() %>% 
      html_nodes('.mw-headline') %>% 
      html_text())

# what are the different sections? I reviewed every year to see what 
# headings were used to indicate TT and non-TT jobs
tt_sections <- c("TENURE-TRACK POSITIONS",
                "TENURE-TRACK OR TENURED / FULL-TIME POSITIONS",
               "Tenure-Track or Tenured / Full-time Position",
               "ASSISTANT PROFESSOR OR OPEN RANK",
               "TENURE TRACK ASSISTANT PROFESSOR OR OPEN RANK",
               "TENURED ASSOCIATE OR FULL PROFESSOR",
               "ASSOCIATE OR FULL PROFESSOR")

non_tt_sections <- c("NON-TENURE-TRACK POSITIONS", 
                     "VISITING POSITIONS / Limited-Term Appointments / Postdocs", 
                     "Visiting Positions / Limited-Term Appointments / Postdocs",
                     "VISITING POSITIONS / LIMITED TERM APPOINTMENTS / POSTDOCS",   
                     "VISITING POSITIONS / LIMITED-TERM APPOINTMENTS / POSTDOCS / PART-TIME POSITIONS",
                     "VISITING POSITIONS")

end_non_tt_sections <- c("DISCUSSION, RUMORS AND SPECULATION",   
                         "DISCUSSION, RUMORS, SPECULATION",
                         "General Discussion, Rumors, and Speculation" )

# what to do about this?
# "COMPLETED SEARCHES"

# this is a function that extracts text from each section with a heading
# that matches those above, and counts how many positions are in each section
library(stringr)
get_counts_tt_non_tt <- function(x){

tt_start      <- which(str_detect(x, 
                                  paste(tt_sections, 
                                        collapse = "|")))[1] 
non_tt_start   <- which(str_detect(x, 
                                   paste(non_tt_sections, 
                                         collapse = "|")))[1] 
non_tt_end    <- which(str_detect(x, 
                                  paste(end_non_tt_sections, 
                                        collapse = "|")))[1] 
n_tt_jobs     <- length( x[(tt_start + 1) : (non_tt_start - 1) ] )
n_non_tt_jobs <- length( x[(non_tt_start + 1) : (non_tt_end  - 1) ] )

return(list(n_tt_jobs = n_tt_jobs,
            n_non_tt_jobs = n_non_tt_jobs))
}

# calculate ratio
ratios_of_tt_to_non_tt_jobs <- 
map_df(urls_for_each_year_headers,
    get_counts_tt_non_tt) %>% 
  mutate(ratio = n_tt_jobs / n_non_tt_jobs) %>% 
  mutate(year = str_replace(str_sub( urls_for_each_year, -9),
                            "-", "-\n"))

# draw plot
  ggplot(ratios_of_tt_to_non_tt_jobs) +
  aes(year, 
      group = 1,
      ratio) +
  geom_line() +
  geom_hline(yintercept = 1,
             colour = "red") +
  labs(y = "Ratio of tenure-track to non-tenure track\njobs in Archaeology",
       x = "") +
  theme_bw(base_size = 16)

```

```{r}
#| label: documents-requested

integer_breaks <- function(n = 5, ...) {
  fxn <- function(x) {
    breaks <- floor(pretty(x, n, ...))
    names(breaks) <- attr(breaks, "labels")
    breaks
  }
  return(fxn)
}

jobdata %>% 
  select(starts_with("documents_requested")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name, value) %>% 
  tally() %>% 
  mutate(value = case_when(
    value == "Not requested in the job ad" ~ 0,
    value == "One" ~ 1,
    value == "Two (e.g. two syllabi)" ~ 2,
    value == "Three" ~ 3,
    .default = 0
  )) %>% 
  ggplot() +
  aes(value,
      n) +
  geom_col() +
  facet_wrap( ~ name) 
```

```{r}
#| label: fig-requirements-over-time

jobdata_requirements <- 
jobdata %>% 
  select(year_ad_posted,
         starts_with("documents_requested")) %>% 
  pivot_longer(-year_ad_posted) %>% 
  mutate(value = case_when(
    value == "Not requested in the job ad" ~ 0,
    value == "One" ~ 1,
    value == "Two (e.g. two syllabi)" ~ 2,
    value == "Three" ~ 3,
    .default = 0
  ))  %>% 
  # trim names a bit
  mutate(name = str_remove(name, "documents_requested_"))

jobdata_requirements_means <- 
jobdata_requirements %>% # average number requested per year
  group_by(year_ad_posted, 
           name) %>% 
  summarise(mean_n = mean(value))

ggplot(jobdata_requirements_means) +
  aes(year_ad_posted, 
      mean_n,
      group = name) +
  geom_line() +
  geom_jitter(data = jobdata_requirements,
             aes(year_ad_posted, 
                 value),
             alpha = 0.1,
             height = 0.2,
             width =  0.1) +
  facet_wrap(~name,
             scales = "free_y") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

```{r}
#| label: fig-requirements-by-position

# do the requirements differ for associate positions 
jobdata %>% 
  mutate(position_title = case_when(
    str_detect(title_of_position_tenure_track_jobs_only, "Associate") ~ "Associate",
    str_detect(title_of_position_tenure_track_jobs_only, "Assistant") ~ "Assistant")) %>% 
  select(position_title,
         starts_with("documents_requested")) %>% 
  pivot_longer(-position_title) %>% 
  mutate(value = case_when(
    value == "Not requested in the job ad" ~ 0,
    value == "One" ~ 1,
    value == "Two (e.g. two syllabi)" ~ 2,
    value == "Three" ~ 3,
    .default = 0
  )) %>% 
  filter(!is.na(position_title)) %>% 
  ggplot() +
  aes(position_title,
      value
      ) +
    geom_jitter(height = 0.05) +
  facet_wrap( ~ name)
```

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2024-02-09 analysis/paper/paper.qmd
minor edits and comments

@@
   str_extract(.,  "[[0-9]]{4}-[[0-9]]{4}")
+ # fix for 2021-22 and 2023
 
@@
   mutate(year_ad_posted = year_ad_posted) %>% 
-  drop_na()
+  drop_na(year_ad_posted)
 
@@
 # starts at 2010-2011
+# fix for 2021-22
 
@@
 #| label: documents-requested
-
-integer_breaks <- function(n = 5, ...) {
-  fxn <- function(x) {
-    breaks <- floor(pretty(x, n, ...))
-    names(breaks) <- attr(breaks, "labels")
-    breaks
-  }
-  return(fxn)
-}
 

### 2024-02-16 analysis/paper/sketches.R
minor updates and add sketches

@@ -1,0 +1,40 @@
+
+set_of_bigrams <- c("material culture",
+                    "resource management",
+                    "environmental archaeology")
+
+jobdata$topical_focus_of_position
+
+jobdata$topic_bigram <-
+str_match(jobdata$topical_focus_of_position,
+          paste0(set_of_bigrams, collapse = "|"))[ , 1]
+
+# check it
+jobdata$topical_focus_of_position[1:5]
+
+
+#----------------------------------------------------
+
+lookup_table <-
+  tribble(~"Category", ~"From the data",
+"Africa", "Africa; sub-Saharan Africa; North Africa",
+"Americas", "Western Hemisphere; New World",
+"Northeastern US", "The Northeast; New York; Pennsylvania; Middle Atlantic; the mid-Atlantic; the Northeast or southeastern Canada; Arctic-North Atlantic region; the Atlantic world; the Chesapeake region; the greater Northeast (Quebec and neighboring regions, including the eastern Subarctic and Arctics, Atlantic provinces, northwestern United States, and lower Great Lakes); North Atlantic; The Northeastern United States; Northeast; the Northeastern U.S.; the Mid-Atlantic; the North Atlantic")
+
+lookup_table_long <-
+lookup_table %>%
+  separate_longer_delim(`From the data`, ";") %>%
+  mutate(`From the data` = str_squish(`From the data`))
+
+jobdata$geographic_focus_of_position
+
+# getting only the first match of geographic entity
+jobdata$geo_first_match <-
+str_match(jobdata$geographic_focus_of_position,
+          paste0(lookup_table_long$`From the data`, collapse = "|"))[ , 1]
+
+
+jobdata <-
+left_join(jobdata,
+          lookup_table_long,
+          by = join_by(geo_first_match == `From the data`))

### 2024-02-16 analysis/paper/paper.qmd
fix for 2021-22

@@
   pull(url_to_data_source_e_g_paste_in_url_to_the_jobs_wiki_page) %>% 
-  str_extract(.,  "[[0-9]]{4}-[[0-9]]{4}")
- # fix for 2021-22 and 2023
+  str_extract(.,  "[[0-9]]{4}-[[0-9]]{4}|2021-22") %>% 
+  str_replace("2021-22",  "2021-2022")
+ # fix for 2021-22 DONE!
+ # fix for and 2023
 

### 2024-02-28 analysis/paper/sketches.R
add sketches

@@ -38,3 +38,74 @@ jobdata <-
 left_join(jobdata,
           lookup_table_long,
           by = join_by(geo_first_match == `From the data`))
+
+#----------------------------------------
+
+jobdata$title_of_position_tenure_track_jobs_only
+
+we_care_about <-
+  c("Assistant", "Asst.",
+    "Associate",
+    "Full",
+    "Curator",
+    "Open rank",
+    "Curator"
+    )
+
+jobdata %>%
+  mutate(job_title_simple = case_when(
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "Assistant|Asst.") ~ "Assistant Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "Associate|Assoc.") ~ "Associate Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "Full") ~ "Full Professor")) %>%
+  select(title_of_position_tenure_track_jobs_only,
+         job_title_simple) %>%
+  mutate(job_title_simple = case_when(
+    str_detect(title_of_position_tenure_track_jobs_only,
+               "Assistant & Associate") ~ "Assistant or Associate Professor",
+    .default = job_title_simple
+    )) %>% View
+
+
+
+
+
+jobdata %>%
+  filter(title_of_position_tenure_track_jobs_only %in% c("Assistant", "Associate"))
+
+
+jobdata %>%
+  filter(str_detect(title_of_position_tenure_track_jobs_only, paste(c(c("Assistant", "Associate")),collapse = '|'))) %>%
+  select(title_of_position_tenure_track_jobs_only)
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2024-03-01 analysis/paper/paper.qmd
fix for 2023 year
new approach to tt/ntt ratio

@@
  # fix for 2021-22 DONE!
- # fix for and 2023
+ # fix for 2023 DONE!
 
@@
 # ratio of tenure-track to untenured positions
+# base URL changes after 2018_2019
 
-base_url <- "http://academicjobs.wikia.com/wiki/Archaeology_Jobs_"
+base_url_to_2019 <- "https://academicjobs.fandom.com/wiki/Archaeology_Jobs_"
+base_url_after_2020 <- "https://academicjobs.fandom.com/wiki/Archaeology_"
 
@@
 # fix for 2021-22
+# base UR
 
-years <- map_chr(2010:2019, ~str_glue('{.x}-{.x +1}'))
+years_to_2019 <- map_chr(2010:2019, ~str_glue('{.x}-{.x +1}'))
+years_after_2020 <- map_chr(2020:2023, ~str_glue('{.x}-{.x +1}'))
 # though it seems to start at 2007-8: https://academicjobs.fandom.com/wiki/Archaeology_07-08
 
-urls_for_each_year <- str_glue('{base_url}{years}')
+# make a set of URLs for each page for each year
+urls_for_each_year <- c(str_glue('{base_url_to_2019}{years_to_2019}'), 
+                        str_glue('{base_url_after_2020}{years_after_2020}')) %>% 
+    str_replace("2021-2022", "2021-22")
 
@@
 
-# what are the different sections? I reviewed every year to see what 
-# headings were used to indicate TT and non-TT jobs
-tt_sections <- c("TENURE-TRACK POSITIONS",
-                "TENURE-TRACK OR TENURED / FULL-TIME POSITIONS",
-               "Tenure-Track or Tenured / Full-time Position",
-               "ASSISTANT PROFESSOR OR OPEN RANK",
-               "TENURE TRACK ASSISTANT PROFESSOR OR OPEN RANK",
-               "TENURED ASSOCIATE OR FULL PROFESSOR",
-               "ASSOCIATE OR FULL PROFESSOR")
+# keep only headings that are actual jobs, they include the terms:
+job_headings <- c("college", "university")
 
-non_tt_sections <- c("NON-TENURE-TRACK POSITIONS", 
-                     "VISITING POSITIONS / Limited-Term Appointments / Postdocs", 
-                     "Visiting Positions / Limited-Term Appointments / Postdocs",
-                     "VISITING POSITIONS / LIMITED TERM APPOINTMENTS / POSTDOCS",   
-                     "VISITING POSITIONS / LIMITED-TERM APPOINTMENTS / POSTDOCS / PART-TIME POSITIONS",
-                     "VISITING POSITIONS")
+total_number_of_jobs_per_year <- 
+  map(urls_for_each_year_headers,
+      ~str_subset(tolower(.x),
+          paste0(job_headings, collapse = "|")))
 
-end_non_tt_sections <- c("DISCUSSION, RUMORS AND SPECULATION",   
-                         "DISCUSSION, RUMORS, SPECULATION",
-                         "General Discussion, Rumors, and Speculation" )
+total_number_of_jobs_per_year_n <- 
+map_int(total_number_of_jobs_per_year, length)
 
-# what to do about this?
-# "COMPLETED SEARCHES"
+total_number_of_jobs_per_year_tbl <- 
+tibble(
+  url_to_data_source_e_g_paste_in_url_to_the_jobs_wiki_page = urls_for_each_year,
+  total_number_of_jobs_per_year = total_number_of_jobs_per_year_n
+)
 
-# this is a function that extracts text from each section with a heading
-# that matches those above, and counts how many positions are in each section
-library(stringr)
-get_counts_tt_non_tt <- function(x){
-
-tt_start      <- which(str_detect(x, 
-                                  paste(tt_sections, 
-                                        collapse = "|")))[1] 
-non_tt_start   <- which(str_detect(x, 
-                                   paste(non_tt_sections, 
-                                         collapse = "|")))[1] 
-non_tt_end    <- which(str_detect(x, 
-                                  paste(end_non_tt_sections, 
-                                        collapse = "|")))[1] 
-n_tt_jobs     <- length( x[(tt_start + 1) : (non_tt_start - 1) ] )
-n_non_tt_jobs <- length( x[(non_tt_start + 1) : (non_tt_end  - 1) ] )
-
-return(list(n_tt_jobs = n_tt_jobs,
-            n_non_tt_jobs = n_non_tt_jobs))
-}
-
-# calculate ratio
-ratios_of_tt_to_non_tt_jobs <- 
-map_df(urls_for_each_year_headers,
-    get_counts_tt_non_tt) %>% 
-  mutate(ratio = n_tt_jobs / n_non_tt_jobs) %>% 
-  mutate(year = str_replace(str_sub( urls_for_each_year, -9),
-                            "-", "-\n"))
+# count of TT jobs per year from our manual data collection,
+# join with our total number of all jobs by scraping
+count_of_tt_jobs_per_year_from_our_form <- 
+jobdata %>% 
+  group_by(url_to_data_source_e_g_paste_in_url_to_the_jobs_wiki_page) %>% 
+  tally() %>% 
+  left_join(total_number_of_jobs_per_year_tbl) %>% 
+  rename(n_tt_jobs  = n,
+         n_total_jobs = total_number_of_jobs_per_year) %>% 
+  mutate(n_non_tt_jobs = n_total_jobs - n_tt_jobs,
+         ratio_tt_2_ntt = n_tt_jobs / n_non_tt_jobs) %>% 
+  mutate(year = str_extract(url_to_data_source_e_g_paste_in_url_to_the_jobs_wiki_page,  "[[0-9]]{4}-[[0-9]]{4}|2021-22")) %>%  
+  mutate(year = ifelse(year =="2021-22",  "2021-2022", year))
 
 # draw plot
-  ggplot(ratios_of_tt_to_non_tt_jobs) +
+  ggplot(count_of_tt_jobs_per_year_from_our_form) +
   aes(year, 
       group = 1,
-      ratio) +
+      ratio_tt_2_ntt) +
   geom_line() +

### 2024-03-01 analysis/paper/sketches.R

@@ -38,3 +38,61 @@ jobdata <-
 left_join(jobdata,
           lookup_table_long,
           by = join_by(geo_first_match == `From the data`))
+
+#---------------------------------------------------
+
+
+# what are the different sections? I reviewed every year to see what
+# headings were used to indicate TT and non-TT jobs
+tt_sections <- c("TENURE-TRACK POSITIONS",
+                 "TENURE-TRACK OR TENURED / FULL-TIME POSITIONS",
+                 "Tenure-Track or Tenured / Full-time Position",
+                 "ASSISTANT PROFESSOR OR OPEN RANK",
+                 "TENURE TRACK ASSISTANT PROFESSOR OR OPEN RANK",
+                 "TENURED ASSOCIATE OR FULL PROFESSOR",
+                 "ASSOCIATE OR FULL PROFESSOR")
+
+non_tt_sections <- c("NON-TENURE-TRACK POSITIONS",
+                     "VISITING POSITIONS / Limited-Term Appointments / Postdocs",
+                     "Visiting Positions / Limited-Term Appointments / Postdocs",
+                     "VISITING POSITIONS / LIMITED TERM APPOINTMENTS / POSTDOCS",
+                     "VISITING POSITIONS / LIMITED-TERM APPOINTMENTS / POSTDOCS / PART-TIME POSITIONS",
+                     "VISITING POSITIONS")
+
+end_non_tt_sections <- c("DISCUSSION, RUMORS AND SPECULATION",
+                         "DISCUSSION, RUMORS, SPECULATION",
+                         "General Discussion, Rumors, and Speculation" )
+
+# what to do about this?
+# "COMPLETED SEARCHES"
+
+# this is a function that extracts text from each section with a heading
+# that matches those above, and counts how many positions are in each section
+library(stringr)
+get_counts_tt_non_tt <- function(x){
+
+  tt_start      <- which(str_detect(x,
+                                    paste(tt_sections,
+                                          collapse = "|")))[1]
+  non_tt_start   <- which(str_detect(x,
+                                     paste(non_tt_sections,
+                                           collapse = "|")))[1]
+  non_tt_end    <- which(str_detect(x,
+                                    paste(end_non_tt_sections,
+                                          collapse = "|")))[1]
[…]

### 2024-03-01 analysis/paper/sketches.R
resolve merge conflict

@@ -39,7 +39,7 @@ left_join(jobdata,
           lookup_table_long,
           by = join_by(geo_first_match == `From the data`))
 
-<<<<<<< HEAD
+
 #---------------------------------------------------
 
 
@@ -95,7 +95,7 @@ ratios_of_tt_to_non_tt_jobs <-
   mutate(ratio = n_tt_jobs / n_non_tt_jobs) %>%
   mutate(year = str_replace(str_sub( urls_for_each_year, -9),
                             "-", "-\n"))
-=======
+
 #----------------------------------------
 
 jobdata$title_of_position_tenure_track_jobs_only

### 2024-03-01 analysis/paper/paper.qmd
by rank by year

@@
 @fig-how-many-jobs-per-year shows how many jobs per year in our sample
+
+```{r}
+#| label: fig-how-many-jobs-per-year-by-rank
+#| fig-cap: "Proportion of tenure-track archaeology faculty jobs per year by rank"
+
+# how many jobs of each rank per year?
+
+jobdata <-
+jobdata %>%
+  # simplify rank descriptions
+  mutate(title_of_position_tenure_track_jobs_only = tolower(title_of_position_tenure_track_jobs_only)) %>%
+  mutate(job_title_simple = case_when(
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "assistant prof|asst. prof|asst prof") ~ "Assistant Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "associate prof|assoc. prof") ~ "Associate Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "full prof") ~ "Full Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "assistant or associate prof|assistant/associate prof") ~ "Assistant or Associate Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "open rank|open-rank|assistant, associate, or full prof|assistant prof, associate prof, or prof") ~ "Open Rank",
+   .default = "other"))
+
+# explore over time
+jobdata %>%
+  group_by(year_ad_posted) %>%
+  count(job_title_simple) %>%
+  mutate(prop = n / sum(n)) %>%
+  ggplot() +
+  aes(year_ad_posted,
+      prop,
+      group = job_title_simple,
+      colour = job_title_simple) +
+  geom_line()
+
+```
 

### 2024-03-01 analysis/paper/sketches.R

@@ -104,56 +104,49 @@ we_care_about <-
   c("Assistant", "Asst.",
     "Associate",
     "Full",
-    "Curator",
-    "Open rank",
-    "Curator"
+    "Open rank"
     )
 
+jobdata <-
 jobdata %>%
+  # simplify rank descriptions
+  mutate(title_of_position_tenure_track_jobs_only = tolower(title_of_position_tenure_track_jobs_only)) %>%
   mutate(job_title_simple = case_when(
    str_detect(title_of_position_tenure_track_jobs_only,
-              "Assistant|Asst.") ~ "Assistant Professor",
+              "assistant prof|asst. prof|asst prof") ~ "Assistant Professor",
    str_detect(title_of_position_tenure_track_jobs_only,
-              "Associate|Assoc.") ~ "Associate Professor",
+              "associate prof|assoc. prof") ~ "Associate Professor",
    str_detect(title_of_position_tenure_track_jobs_only,
-              "Full") ~ "Full Professor")) %>%
-  select(title_of_position_tenure_track_jobs_only,
-         job_title_simple) %>%
-  mutate(job_title_simple = case_when(
-    str_detect(title_of_position_tenure_track_jobs_only,
-               "Assistant & Associate") ~ "Assistant or Associate Professor",
-    .default = job_title_simple
-    )) %>% View
+              "full prof") ~ "Full Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "assistant or associate prof|assistant/associate prof") ~ "Assistant or Associate Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "open rank|open-rank|assistant, associate, or full prof|assistant prof, associate prof, or prof") ~ "Open Rank",
+   .default = "other"))
 
-
-
-
-
+# explore over time
 jobdata %>%
-  filter(title_of_position_tenure_track_jobs_only %in% c("Assistant", "Associate"))
+  group_by(year_ad_posted) %>%
+  count(job_title_simple) %>%
+  mutate(prop = n / sum(n)) %>%
+  ggplot() +
+  aes(year_ad_posted,
+      prop,
+      group = job_title_simple,
+      colour = job_title_simple) +
+  geom_line()
 
+#-----------------------------------------------------------------------------
 
-jobdata %>%
[…]

### 2024-03-01 analysis/paper/paper.qmd
added sketch of the geographic foci analysis

--- 2024-03-01 analysis/paper/paper.qmd
@@
 ```{r}
+#| label: fig-how-many-jobs-per-year-by-rank
+#| fig-cap: "Proportion of tenure-track archaeology faculty jobs per year by rank"
+
+# how many jobs of each rank per year?
+
+jobdata <-
+jobdata %>%
+  # simplify rank descriptions
+  mutate(title_of_position_tenure_track_jobs_only = tolower(title_of_position_tenure_track_jobs_only)) %>%
+  mutate(job_title_simple = case_when(
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "assistant prof|asst. prof|asst prof") ~ "Assistant Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "associate prof|assoc. prof") ~ "Associate Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "full prof") ~ "Full Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "assistant or associate prof|assistant/associate prof") ~ "Assistant or Associate Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "open rank|open-rank|assistant, associate, or full prof|assistant prof, associate prof, or prof") ~ "Open Rank",
+   .default = "Other"))
+
+# explore over time
+jobdata %>%
+  group_by(year_ad_posted) %>%
+  count(job_title_simple) %>%
+  mutate(prop = n / sum(n)) %>%
+  ggplot() +
+  aes(year_ad_posted,
+      prop,
+      group = job_title_simple,
+      colour = job_title_simple) +
+  geom_line() +
+  theme_minimal() +
+  theme(legend.position = c(0.2, 0.8))
+
+```
+
+```{r}
 #| label: ratio-tt-to-non-tt
@@
              colour = "red") +
-  labs(y = "Ratio of tenure-track to non-tenure track\njobs in Archaeology",
+  labs(y = "Ratio of tenure-track to\nnon-tenure track and othe\njobs in Archaeology",
        x = "") +

### 2024-03-01 analysis/paper/sketches.R

@@ -148,8 +148,43 @@ jobdata %>%
 
 # geographic focus by year
 
+library(googlesheets4)
 
+geographic_foci <-
+read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0")
 
+geographic_foci_clean <-
+  map(
+   str_split(geographic_foci$`From the data`, ";"),
+   ~.x %>%
+     str_squish() %>%
+     stri_remove_empty())
+
+jobdata_geo <-
+  jobdata %>%
+  select(geographic_focus_of_position)
+
+jobdata_geo <-
+  # add one column for each geo region in our categories
+cbind(jobdata_geo,
+      setNames( lapply(geographic_foci$Category, function(x) x=NA),
+                geographic_foci$Category) )
+
+for(i in 1:length(geographic_foci$Category)){
+
+  this_location <- geographic_foci$Category[i]
+
+  # create the pattern to search for
+  x <- paste0(geographic_foci_clean[[i]], collapse = "|")
+
+  # do the search through all the job ads for that pattern
+  y <- str_detect(jobdata_geo$geographic_focus_of_position,
+             x)
+
+  # assign back to our data frame in the appropiate location column
+  jobdata_geo[, this_location] <- y
+
+}
 
 
 

### 2024-03-03 analysis/paper/sketches.R
exploring geo over time

@@ -149,6 +149,7 @@ jobdata %>%
 # geographic focus by year
 
 library(googlesheets4)
+library(stringi)
 
 geographic_foci <-
 read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0")
@@ -167,12 +168,12 @@ jobdata_geo <-
 jobdata_geo <-
   # add one column for each geo region in our categories
 cbind(jobdata_geo,
-      setNames( lapply(geographic_foci$Category, function(x) x=NA),
-                geographic_foci$Category) )
+      setNames( lapply(geographic_foci$Category2, function(x) x=NA),
+                geographic_foci$Category2) )
 
-for(i in 1:length(geographic_foci$Category)){
+for(i in 1:length(geographic_foci$Category2)){
 
-  this_location <- geographic_foci$Category[i]
+  this_location <- geographic_foci$Category2[i]
 
   # create the pattern to search for
   x <- paste0(geographic_foci_clean[[i]], collapse = "|")
@@ -185,6 +186,85 @@ for(i in 1:length(geographic_foci$Category)){
   jobdata_geo[, this_location] <- y
 
 }
+
+united_states_regions <-
+c( "Northeastern US",
+   "Midwest US",
+   "Southeast US",
+   "Southwest US",
+   "Western US",
+   "Southern US",
+   "Eastern US"  )
+
+jobdata_geo_year <-
+jobdata %>%
+  bind_cols(jobdata_geo) %>%
+  select(year_ad_posted,
+         geographic_foci$Category2) %>%
+  pivot_longer(-year_ad_posted) %>%
+  drop_na()
+
+# how many times each location mentioned?
+jobdata_geo_year %>%
+  group_by(name) %>%
+  summarise(n = sum(value)) %>%
+  arrange(desc(n)) %>%
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  theme_minimal() +
+  coord_flip()
+
+# explore trends over time. put a point on the max year
+jobdata_geo_year_tally <-
+jobdata_geo_year %>%
+  # exclude those with <20 ads
+  filter(!name %in% c("Canada",
+                      "Americas",
+                      "Arctic",
+                      "Oceania",
+                      "Eastern US",
+                      "Southern US",
+                      "Midwest US",
+                      "Northeastern US"
+                      )) %>%
[…]

### 2024-03-05 analysis/paper/sketches.R
stash

@@ -181,7 +181,7 @@ for(i in 1:length(geographic_foci$Category)){
   y <- str_detect(jobdata_geo$geographic_focus_of_position,
              x)
 
-  # assign back to our data frame in the appropiate location column
+  # assign back to our data frame in the appropriate location column
   jobdata_geo[, this_location] <- y
 
 }

### 2024-03-05 analysis/paper/paper.qmd
finesse requirements per year

--- 2024-03-01 analysis/paper/paper.qmd
@@
 library(tidyverse)
+library(here)
 library(ggbeeswarm)
@@
   scale_x_discrete(name = "Academic Year") +
+  ylab("Number of tenure track job ads") +
   theme_minimal()
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-how-many-jobs-per-year.png"),
+       bg ="white")
 ```
@@
 @fig-how-many-jobs-per-year shows how many jobs per year in our sample
+
+```{r}
+#| label: fig-how-many-jobs-per-year-by-rank
+#| fig-cap: "Proportion of tenure-track archaeology faculty jobs per year by rank"
+
+# how many jobs of each rank per year?
+
+jobdata <-
+jobdata %>%
+  # simplify rank descriptions
+  mutate(title_of_position_tenure_track_jobs_only = tolower(title_of_position_tenure_track_jobs_only)) %>%
+  mutate(job_title_simple = case_when(
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "assistant prof|asst. prof|asst prof") ~ "Assistant Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "associate prof|assoc. prof") ~ "Associate Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "full prof") ~ "Full Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "assistant or associate prof|assistant/associate prof") ~ "Assistant or Associate Professor",
+   str_detect(title_of_position_tenure_track_jobs_only,
+              "open rank|open-rank|assistant, associate, or full prof|assistant prof, associate prof, or prof") ~ "Open Rank",
+   .default = "Other (Curator, Director, etc.)"))
+
+# explore over time
+jobdata %>%
+  group_by(year_ad_posted) %>%
+  count(job_title_simple) %>%
+  mutate(prop = n / sum(n)) %>%
+  ggplot() +
+  aes(year_ad_posted,
+      prop,
+      group = job_title_simple,
+      colour = job_title_simple) +
+  geom_line() +
+  theme_minimal() +
+  xlab("Year") +
+  ylab("Proprtion of all tenure-track ads") +
+  theme(legend.position = c(0.5, 0.5)) +
+  guides(colour = guide_legend(nrow=2,
+                               byrow=TRUE,
+                               "Job title"))
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-prop-by-job-title-per-year.png"),
+       bg ="white")
+
+```
 
@@
 
-years_to_2019 <- map_chr(2010:2019, ~str_glue('{.x}-{.x +1}'))
-years_after_2020 <- map_chr(2020:2023, ~str_glue('{.x}-{.x +1}'))
+years_to_2019 <- map_chr(2013:2019, ~str_glue('{.x}-{.x +1}'))
+years_after_2020 <- map_chr(2020:2022, ~str_glue('{.x}-{.x +1}'))
 # though it seems to start at 2007-8: https://academicjobs.fandom.com/wiki/Archaeology_07-08
@@
   tally() %>% 
-  left_join(total_number_of_jobs_per_year_tbl) %>% 
+  right_join(total_number_of_jobs_per_year_tbl) %>% 
   rename(n_tt_jobs  = n,
@@
              colour = "red") +
-  labs(y = "Ratio of tenure-track to non-tenure track\njobs in Archaeology",
+  annotate("text", 
+           x = 3, 
+           y = 1.1, 
+           label = "1:1 ratio",
+           colour = "red") +
+  labs(y = "Ratio of tenure-track to non-tenure track\nand other jobs in Archaeology",
        x = "") +
-  theme_bw(base_size = 16)
+  theme_bw(base_size = 14)
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-ratio-tt-2-ntt-jobs-per-year.png"),
+       bg ="white")
 
@@
   # trim names a bit
-  mutate(name = str_remove(name, "documents_requested_"))
+  mutate(name = str_remove(name, "documents_requested_")) %>% 
+  mutate(name = str_replace_all(name, "_", " "))
 
@@
 
+integer_breaks <- function(n = 5, ...) {
+  fxn <- function(x) {
+    breaks <- floor(pretty(x, n, ...))
+    names(breaks) <- attr(breaks, "labels")
+    breaks
+  }
+  return(fxn)
+}
+
 ggplot(jobdata_requirements_means) +
@@
       group = name) +
-  geom_line() +
+  geom_line(linewidth = 2) +
   geom_jitter(data = jobdata_requirements,
@@
              scales = "free_y") +
+  xlab("Year") +
+  ylab("Number requested in job ad") +
+  scale_y_continuous(breaks = integer_breaks()) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-requirements-per-year.png"),
+       bg ="white")
 
@@
 # do the requirements differ for associate positions 
+jobdata_requirements_by_rank <- 
 jobdata %>% 
   mutate(position_title = case_when(
-    str_detect(title_of_position_tenure_track_jobs_only, "Associate") ~ "Associate",
-    str_detect(title_of_position_tenure_track_jobs_only, "Assistant") ~ "Assistant")) %>% 
+    str_detect(title_of_position_tenure_track_jobs_only, 
+               "associate") ~ "associate",
+    str_detect(title_of_position_tenure_track_jobs_only, 
+               "assistant") ~ "assistant",
+    str_detect(title_of_position_tenure_track_jobs_only, 
+               "full") ~ "full"))  %>% 
   select(position_title,
@@
   )) %>% 
-  filter(!is.na(position_title)) %>% 
+  filter(!is.na(position_title))
+
+jobdata_requirements_by_rank_means <- 
+  jobdata_requirements_by_rank %>% 
+  group_by(position_title,
+           name) %>% 
+  summarise(mean = mean(value))
+
 ggplot() +
+    geom_jitter(data = jobdata_requirements_by_rank,
                   aes(position_title, 
-      value
-      ) +
-    geom_jitter(height = 0.05) +
-  facet_wrap( ~ name)
+                      value),
+                height = 0.05,
+                alpha = 0.1) +
+    geom_point(data = jobdata_requirements_by_rank_means,
+               aes(position_title,
+                   mean),
+               size = 4,
+               colour = "red") +
+  facet_wrap( ~ name,
+              scales = "free_y")
+```
+
+```{r}
+#| label: fig-geographic-focus-by-year
+
+# geographic focus by year
+
+library(googlesheets4)
+library(stringi)
+
+geographic_foci <-
+read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0")
+
+geographic_foci_clean <-
+  map(
+   str_split(geographic_foci$`From the data`, ";"),
+   ~.x %>%
+     str_squish() %>%
+     stri_remove_empty())
+
+jobdata_geo <-
+  jobdata %>%
+  select(geographic_focus_of_position)
+
+jobdata_geo <-
+  # add one column for each geo region in our categories
+cbind(jobdata_geo,
+      setNames( lapply(geographic_foci$Category2, function(x) x=NA),
+                geographic_foci$Category2) )
+
+for(i in 1:length(geographic_foci$Category2)){
+
+  this_location <- geographic_foci$Category2[i]
+
+  # create the pattern to search for
+  x <- paste0(geographic_foci_clean[[i]], collapse = "|")
+
+  # do the search through all the job ads for that pattern
+  y <- str_detect(jobdata_geo$geographic_focus_of_position,
+             x)
+
+  # assign back to our data frame in the appropriate location column
+  jobdata_geo[, this_location] <- y
+
+}
+
+united_states_regions <-
+c( "Northeastern US",
+   "Midwest US",
+   "Southeast US",
+   "Southwest US",
+   "Western US",
+   "Southern US",
+   "Eastern US"  )
+
+jobdata_geo_year <-
+jobdata %>%
+  bind_cols(jobdata_geo) %>%
+  select(year_ad_posted,
+         geographic_foci$Category2) %>%
+  pivot_longer(-year_ad_posted) %>%
+  drop_na()
+
+# how many times each location mentioned?
+jobdata_geo_year %>%
+  group_by(name) %>%
+  summarise(n = sum(value)) %>%
+  arrange(desc(n)) %>%
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  theme_minimal() +
+  coord_flip()
+
+# explore trends over time. put a point on the max year
+jobdata_geo_year_tally <-
+jobdata_geo_year %>%
+  # exclude those with <20 ads
+  filter(!name %in% c("Canada",
+                      "Americas",
+                      "Arctic",
+                      "Oceania",
+                      "Eastern US",
+                      "Southern US",
+                      "Midwest US",
+                      "Northeastern US"
+                      )) %>%
+  group_by(year_ad_posted,
+           name) %>%
+  summarise(n = sum(value)) %>%
+  mutate(prop = n / sum(n))
+
+jobdata_geo_year_tally_max <-
+  jobdata_geo_year_tally %>%
+  group_by(
+           name ) %>%
+  filter(prop == max(prop))
+
+library(ggrepel)
+
+ggplot() +
+  geom_line(data =  jobdata_geo_year_tally,
+            aes(year_ad_posted,
+                 prop,
+                 group = name,
+                 colour = name) ) +
+   geom_point(data = jobdata_geo_year_tally_max,
+              aes(year_ad_posted,
+                  prop,
+                  group = name,
+                  colour = name),
+              size = 3) +
+   geom_text_repel(data = jobdata_geo_year_tally_max,
+             aes(year_ad_posted,
+                 prop ,
+                 group = name,
+                 colour = name,
+                 label = name),
+             nudge_y = 0.01,
+             bg.color = "white",
+             bg.r = 0.25,
+             size = 3) +
+  xlab("Year") +
+  ylab("Proportion of all ads") +
+  theme_minimal()
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-geo-focus-by-year.png"),
+       bg ="white")
+
+# what about within the US
+jobdata_geo_us_year <-
+  jobdata %>%
+  bind_cols(jobdata_geo) %>%
+  select(year_ad_posted,
+         geographic_foci$Category2) %>%
+  pivot_longer(-year_ad_posted) %>%
+  drop_na()
+
 ```

### 2024-03-05 analysis/paper/sketches.R

@@ -138,7 +138,7 @@ jobdata %>%
 
 #-----------------------------------------------------------------------------
 
-# topic by year
+# topical focus by year
 
 
 
@@ -146,122 +146,6 @@ jobdata %>%
 
 #-----------------------------------------------------------------------------
 
-# geographic focus by year
-
-library(googlesheets4)
-library(stringi)
-
-geographic_foci <-
-read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0")
-
-geographic_foci_clean <-
-  map(
-   str_split(geographic_foci$`From the data`, ";"),
-   ~.x %>%
-     str_squish() %>%
-     stri_remove_empty())
-
-jobdata_geo <-
-  jobdata %>%
-  select(geographic_focus_of_position)
-
-jobdata_geo <-
-  # add one column for each geo region in our categories
-cbind(jobdata_geo,
-      setNames( lapply(geographic_foci$Category2, function(x) x=NA),
-                geographic_foci$Category2) )
-
-for(i in 1:length(geographic_foci$Category2)){
-
-  this_location <- geographic_foci$Category2[i]
-
-  # create the pattern to search for
-  x <- paste0(geographic_foci_clean[[i]], collapse = "|")
-
-  # do the search through all the job ads for that pattern
-  y <- str_detect(jobdata_geo$geographic_focus_of_position,
-             x)
-
-  # assign back to our data frame in the appropriate location column
-  jobdata_geo[, this_location] <- y
-
-}
-
-united_states_regions <-
-c( "Northeastern US",
-   "Midwest US",
-   "Southeast US",
-   "Southwest US",
-   "Western US",
-   "Southern US",
-   "Eastern US"  )
-
-jobdata_geo_year <-
-jobdata %>%
-  bind_cols(jobdata_geo) %>%
-  select(year_ad_posted,
-         geographic_foci$Category2) %>%
-  pivot_longer(-year_ad_posted) %>%
-  drop_na()
-
-# how many times each location mentioned?
-jobdata_geo_year %>%
-  group_by(name) %>%
-  summarise(n = sum(value)) %>%
-  arrange(desc(n)) %>%
-  ggplot() +
-  aes(reorder(name, n),
-      n)+
-  geom_col() +
-  theme_minimal() +
-  coord_flip()
-
[…]

### 2024-03-05 analysis/paper/paper.qmd
topic and geo foci plots

@@
   geom_col() +
+  xlab("") +
   theme_minimal() +
@@
                  group = name,
-                 colour = name) ) +
+                 colour = name),
+            size = 2,
+            position=position_jitter(w=0.05, h=0.0)) +
    geom_point(data = jobdata_geo_year_tally_max,
@@
                   colour = name),
-              size = 3) +
+              size = 5) +
    geom_text_repel(data = jobdata_geo_year_tally_max,
@@
   drop_na()
+
+```
+
+```{r}
+
+# topical focus by year
+
+library(googlesheets4)
+library(stringi)
+
+topical_foci <-
+  read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0",
+             sheet = "topic")
+
+topical_foci_clean <-
+  map(
+    str_split(topical_foci$`From the data`, ";"),
+    ~.x %>%
+      str_squish() %>%
+      stri_remove_empty() %>%
+      str_to_lower)
+
+jobdata_topic <-
+  jobdata %>%
+  select(topical_focus_of_position) %>%
+  mutate(topical_focus_of_position = str_to_lower(topical_focus_of_position))
+
+jobdata_topic <-
+  # add one column for each topic in our categories
+  cbind(jobdata_topic,
+        setNames( lapply(topical_foci$Category, function(x) x=NA),
+                  topical_foci$Category) )
+
+for(i in 1:length(topical_foci$Category)){
+
+  this_topic <- topical_foci$Category[i]
+
+  # create the pattern to search for
+  x <- paste0(topical_foci_clean[[i]], collapse = "|")
+
+  # do the search through all the job ads for that pattern
+  y <- str_detect(jobdata_topic$topical_focus_of_position,
+                  x)
+
+  # assign back to our data frame in the appropriate location column
+  jobdata_topic[, this_topic] <- y
+
+}
+
+jobdata_topic_year <-
+  jobdata %>%
+  bind_cols(jobdata_topic) %>%
+  select(year_ad_posted,
+         topical_foci$Category) %>%
+  pivot_longer(-year_ad_posted) %>%
+  drop_na()
+
+# how many times each location mentioned?
+jobdata_topic_year %>%
+  group_by(name) %>%
+  summarise(n = sum(value)) %>%
+  arrange(desc(n)) %>%
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  xlab("") +
+  theme_minimal() +
+  coord_flip()
+
+# explore trends over time. put a point on the max year
+jobdata_topic_year_tally <-
+  jobdata_topic_year %>%
+  # exclude those with <20 ads
+  filter(!name %in% c("Digital Archaeology",
+                      "Pleistocene archaeology",
+                      "Mesoamerican Archaeology",
+                      "Biological anthropology",
+                      "Archaeological theory",
+                      "Evolutionary anthropology",
+                      "North American archaeology"
+  )) %>%
+  group_by(year_ad_posted,
+           name) %>%
+  summarise(n = sum(value)) %>%
+  mutate(prop = n / sum(n))
+
+jobdata_topic_year_tally_max <-
+  jobdata_topic_year_tally %>%
+  group_by(
+    name ) %>%
+  filter(prop == max(prop))
+
+library(ggrepel)
+
+ggplot() +
+  geom_line(data =  jobdata_topic_year_tally,
+            aes(year_ad_posted,
+                prop,
+                group = name,
+                colour = name),
+            size = 2,
+            position=position_jitter(w=0.05, h=0.0)) +
+  geom_point(data = jobdata_topic_year_tally_max,
+             aes(year_ad_posted,
+                 prop,
+                 group = name,
+                 colour = name),
+             size = 10) +
+  geom_text_repel(data = jobdata_topic_year_tally_max,
+                  aes(year_ad_posted,
+                      prop ,
+                      group = name,
+                      colour = name,
+                      label = name),
+                  nudge_y = 0.01,
+                  bg.color = "white",
+                  bg.r = 0.25,
+                  size = 3) +
+  xlab("Year") +
+  ylab("Proportion of all ads") +
+  theme_minimal()
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-topic-focus-by-year.png"),
+       bg ="white")
 

### 2024-03-05 analysis/paper/sketches.R

@@ -138,10 +138,6 @@ jobdata %>%
 
 #-----------------------------------------------------------------------------
 
-# topical focus by year
-
-
-
 
 
 #-----------------------------------------------------------------------------

### 2024-03-07 analysis/paper/paper.qmd
explore smoothing on topic time trends

@@
 ggplot() +
-  geom_line(data =  jobdata_topic_year_tally,
+  geom_smooth(data =  jobdata_topic_year_tally,
             aes(year_ad_posted,
@@
             size = 2,
-            position=position_jitter(w=0.05, h=0.0)) +
+            position=position_jitter(w=0.05, h=0.0),
+            se = FALSE) +
   geom_point(data = jobdata_topic_year_tally_max,

### 2024-03-07 analysis/paper/sketches.R
stash

@@ -138,6 +138,7 @@ jobdata %>%
 
 #-----------------------------------------------------------------------------
 
+1+1
 
 
 #-----------------------------------------------------------------------------

### 2024-03-08 analysis/paper/paper.qmd
exploring smoothers

--- 2024-03-05 analysis/paper/paper.qmd
@@
             "fig-how-many-jobs-per-year.png"),
+       height = 7,
+       width = 22,
        bg ="white")
@@
 ```{r}
-#| label: documents-requested
-
-jobdata %>% 
-  select(starts_with("documents_requested")) %>% 
-  pivot_longer(everything()) %>% 
-  group_by(name, value) %>% 
-  tally() %>% 
-  mutate(value = case_when(
-    value == "Not requested in the job ad" ~ 0,
-    value == "One" ~ 1,
-    value == "Two (e.g. two syllabi)" ~ 2,
-    value == "Three" ~ 3,
-    .default = 0
-  )) %>% 
-  ggplot() +
-  aes(value,
-      n) +
-  geom_col() +
-  facet_wrap( ~ name) 
-```
-
-```{r}
 #| label: fig-requirements-over-time
@@
       group = name) +
-  geom_line(linewidth = 2) +
+  geom_smooth(linewidth = 2,
+              colour = "black") +
   geom_jitter(data = jobdata_requirements,
@@
   facet_wrap( ~ name,
-              scales = "free_y")
+              scales = "free_y") +
+  theme_minimal()
 ```
@@
 }
+
+# BM TODO: check for job ads that have a location, but we're not getting it
 
@@
 ggplot() +
-  geom_line(data =  jobdata_geo_year_tally,
+  geom_smooth(data =  jobdata_geo_year_tally,
             aes(year_ad_posted,
@@
             size = 2,
-            position=position_jitter(w=0.05, h=0.0)) +
+           se = FALSE #,
+           # position=position_jitter(w=0.05, h=0.0)
+           ) +
    geom_point(data = jobdata_geo_year_tally_max,
@@
             size = 2,
-            position=position_jitter(w=0.05, h=0.0)) +
+           # position=position_jitter(w=0.05, h=0.0),
+            se = FALSE) +
   geom_point(data = jobdata_topic_year_tally_max,

### 2024-03-08 analysis/paper/sketches.R

@@ -139,6 +139,24 @@ jobdata %>%
 #-----------------------------------------------------------------------------
 
 
+jobdata %>%
+  select(starts_with("documents_requested")) %>%
+  pivot_longer(everything()) %>%
+  group_by(name, value) %>%
+  tally() %>%
+  mutate(value = case_when(
+    value == "Not requested in the job ad" ~ 0,
+    value == "One" ~ 1,
+    value == "Two (e.g. two syllabi)" ~ 2,
+    value == "Three" ~ 3,
+    .default = 0
+  )) %>%
+  ggplot() +
+  aes(value,
+      n) +
+  geom_col() +
+  facet_wrap( ~ name)
+
 
 #-----------------------------------------------------------------------------
 

### 2024-04-02 analysis/paper/paper.qmd
update raw data

@@
             "fig-requirements-per-year.png"),
-       bg ="white")
+       bg ="white",
+       h = 5, # experiment with h and w to get the right size and proportion 
+       w = 8,
+       units = "in",
+       dpi = 600) # make the image nice and crisp
 
@@
             size = 2,
+           # n = 10,
+           span = 0.7,
            # position=position_jitter(w=0.05, h=0.0),
             se = FALSE) +
-  geom_point(data = jobdata_topic_year_tally_max,
-             aes(year_ad_posted,
-                 prop,
-                 group = name,
-                 colour = name),
-             size = 10) +
-  geom_text_repel(data = jobdata_topic_year_tally_max,
-                  aes(year_ad_posted,
-                      prop ,
-                      group = name,
-                      colour = name,
-                      label = name),
-                  nudge_y = 0.01,
-                  bg.color = "white",
-                  bg.r = 0.25,
-                  size = 3) +
+  # geom_point(data = jobdata_topic_year_tally_max,
+  #            aes(year_ad_posted,
+  #                prop,
+  #                group = name,
+  #                colour = name),
+  #            size = 10) +
+  # geom_text_repel(data = jobdata_topic_year_tally_max,
+  #                 aes(year_ad_posted,
+  #                     prop ,
+  #                     group = name,
+  #                     colour = name,
+  #                     label = name),
+  #                 nudge_y = 0.01,
+  #                 bg.color = "white",
+  #                 bg.r = 0.25,
+  #                 size = 3) +
+  # explore direct labels
   xlab("Year") +

### 2024-04-02 analysis/paper/paper.qmd
nice panel of basic things over time

@@
 
+fig_how_many_jobs_per_year <- 
 ggplot(jobdata) +
@@
   scale_x_discrete(name = "Academic Year") +
-  ylab("Number of tenure track job ads") +
-  theme_minimal()
+  ylab("Number of\ntenure track job ads") +
+  theme_minimal(base_size = 14)
 
-ggsave(here("analysis",
-            "figures", 
-            "fig-how-many-jobs-per-year.png"),
-       height = 7,
-       width = 22,
-       bg ="white")
 ```
@@
 # explore over time
+fig_prop_by_job_title_per_year <- 
 jobdata %>%
@@
       colour = job_title_simple) +
-  geom_line() +
-  theme_minimal() +
-  xlab("Year") +
-  ylab("Proprtion of all tenure-track ads") +
+  geom_line(linewidth = 2) +
+  theme_minimal(base_size = 14) +
+  xlab("") +
+  ylab("Proportion of\nall tenure-track ads") +
   theme(legend.position = c(0.5, 0.5)) +
@@
                                byrow=TRUE,
-                               "Job title"))
-
-ggsave(here("analysis",
-            "figures", 
-            "fig-prop-by-job-title-per-year.png"),
-       bg ="white")
+                               "Job title")) +
+   guides(x = "none")
 
@@
 # draw plot
+fig_ratio_tt_2_ntt_jobs_per_year <- 
   ggplot(count_of_tt_jobs_per_year_from_our_form) +
@@
       ratio_tt_2_ntt) +
-  geom_line() +
+  geom_line(linewidth = 2) +
   geom_hline(yintercept = 1,
@@
            x = 3, 
-           y = 1.1, 
+           y = 1.3, 
            label = "1:1 ratio",
            colour = "red") +
-  labs(y = "Ratio of tenure-track to non-tenure track\nand other jobs in Archaeology",
+  labs(y = "Ratio of tenure-track to\nnon-tenure track and other",
        x = "") +
-  theme_bw(base_size = 14)
+  theme_minimal(base_size = 14) +
+   guides(x = "none")
+
+```
+
+```{r}
+# save these three plots as one set
+library(cowplot)
+plot_grid(
+  fig_ratio_tt_2_ntt_jobs_per_year,
+  fig_prop_by_job_title_per_year,
+  fig_how_many_jobs_per_year,
+  ncol = 1,
+  align = "hv",
+  axis = "lr"
+)
 
@@
             "figures", 
-            "fig-ratio-tt-2-ntt-jobs-per-year.png"),
-       bg ="white")
+            "fig-panel-per-year.png"),
+       bg ="white",
+       h = 11, # experiment with h and w to get the right size and proportion 
+       w = 20,
+       units = "in",
+       dpi = 900) # make the image nice and crisp
 

### 2024-04-05 analysis/paper/paper.qmd

@@
   facet_wrap(~name,
-             scales = "free_y") +
+             scales = "free_y",
+             nrow = 2) +
   xlab("Year") +
@@
   scale_y_continuous(breaks = integer_breaks()) +
-  theme_minimal() +
+  theme_minimal(base_size = 14) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
@@
        bg ="white",
-       h = 5, # experiment with h and w to get the right size and proportion 
-       w = 8,
+       h = 10, # experiment with h and w to get the right size and proportion 
+       w = 25,
        units = "in",
-       dpi = 600) # make the image nice and crisp
+       dpi = 900) # make the image nice and crisp
 
@@
             size = 2,
-           se = FALSE #,
-           # position=position_jitter(w=0.05, h=0.0)
+           se = FALSE 
            ) +
-   geom_point(data = jobdata_geo_year_tally_max,
-              aes(year_ad_posted,
-                  prop,
-                  group = name,
-                  colour = name),
-              size = 5) +
-   geom_text_repel(data = jobdata_geo_year_tally_max,
-             aes(year_ad_posted,
-                 prop ,
-                 group = name,
-                 colour = name,
-                 label = name),
-             nudge_y = 0.01,
-             bg.color = "white",
-             bg.r = 0.25,
-             size = 3) +
   xlab("Year") +
   ylab("Proportion of all ads") +
-  theme_minimal()
+  guides(colour = guide_legend("Geographic\nfocus",
+                               label.position = "bottom")) +
+    theme_minimal( base_size = 14) +
+  theme(legend.position="bottom")
 
@@
             "fig-geo-focus-by-year.png"),
-       bg ="white")
+       bg ="white",
+       h = 10, # experiment with h and w to get the right size and proportion 
+       w = 12,
+       units = "in",
+       dpi = 900) # make the image nice and crisp)
 
@@
             size = 2,
-           # n = 10,
            span = 0.7,
-           # position=position_jitter(w=0.05, h=0.0),
             se = FALSE) +
-  # geom_point(data = jobdata_topic_year_tally_max,
-  #            aes(year_ad_posted,
-  #                prop,
-  #                group = name,
-  #                colour = name),
-  #            size = 10) +
-  # geom_text_repel(data = jobdata_topic_year_tally_max,
-  #                 aes(year_ad_posted,
-  #                     prop ,
-  #                     group = name,
-  #                     colour = name,
-  #                     label = name),
-  #                 nudge_y = 0.01,
-  #                 bg.color = "white",
-  #                 bg.r = 0.25,
-  #                 size = 3) +
-  # explore direct labels
   xlab("Year") +
   ylab("Proportion of all ads") +
-  theme_minimal()
+  guides(colour = guide_legend("Topic\nfocus",
+                               label.position = "bottom")) +
+    theme_minimal( base_size = 14) +
+  theme(legend.position="bottom")
 
@@
             "fig-topic-focus-by-year.png"),
-       bg ="white")
+       bg ="white",
+       h = 10, # experiment with h and w to get the right size and proportion
+       w = 12,
+       units = "in",
+       dpi = 900) # make the image nice and crisp))
 

### 2024-04-05 universities.Rmd
maps code

```{r}
# read in packages and datasets
install.packages("usmap")
library(usmap) #import the package
library(ggplot2) #use ggplot2 to add layer for visualization
```

```{r}
# chloropleth map visualization

# Carnegie classification data 
universities <- read_excel("cc.xlsx")

# create new dataframe 
univ_and_state <- select(universities, name, state)
head(univ_and_state)

# change column names in jobdata so it is the same as cc data set
names(jobdata)[6] <- "University Name"
names(univ_and_state)[1] <- "University Name"

# remove () from jobdata
jobdata$`University Name` <-  gsub(r"{\s*\([^\)]+\)}","",as.character(jobdata$`University Name`))

# merge datasets according to values in University Name
merged <- merge(jobdata, univ_and_state, "University Name") # only loading 264 universities of 553 obs.

# sum of occurrences of ads per each university 
counts <- data.frame(table(jobdata$`University Name`))
names(counts)[1] <- "University Name"
counts_per_state <- merge(counts, univ_and_state, "University Name")

# sum per state
state_counts <- data.frame(table(merged$state))
names(state_counts)[1] <- "state"

# plot freq in ads / state
usmap::plot_usmap(data = state_counts, values = "Freq") +
  labs(title = "Ad Frequency Per Universities in Each State from 2012-2013 to 2022-2023")

ggsave(here("analysis",
            "figures", 
            "freq-by-state-map.png"),
       bg ="white",
       h = 10, # experiment with h and w to get the right size and proportion
       w = 12,
       units = "in",
       dpi = 900) # make the image nice and crisp))
```

```{r}
# get 50 states
all_states <-
all_states
```

```{r}
# counts per state 2012-2023 & 2022-2023

states <- unique(merged$state)
names(states)[1] <- "state"

# 2012 - 2013 subset
subset_2012_2013 <- merged %>%
  filter(year_ad_posted == "2012-2013")

subset_2022_2023 <- merged %>%
  filter(year_ad_posted == "2022-2023")

state_counts_2012_2013 <- 
  data.frame(table(subset_2012_2013$state))

state_counts_2022_2023 <- data.frame(table(subset_2022_2023$state))

# rename to specify year

names(state_counts_2012_2013)[1] <- "state"
names(state_counts_2012_2013)[2] <- "freq_2012_2013"

names(state_counts_2022_2023)[1] <- 
  "state"
names(state_counts_2022_2023)[2] <-
  "freq_2022_2023"

# merge all states and count 
x <- merge(states, state_counts_2012_2013, "state", all = TRUE)
x[is.na(x)] <- 0

merged_state_counts <- merge(x, state_counts_2022_2023, "state", all = TRUE)
merged_state_counts[is.na(merged_state_counts)] <- 0

# create new column

merged_state_counts$perc_change <- (merged_state_counts$freq_2022_2023 / merged_state_counts$freq_2012_2013)

merged_state_counts$Change <- ifelse(merged_state_counts$freq_2012_2013 < merged_state_counts$freq_2022_2023, 1, -1)

# help with if value stays the same set equql to 0, set as =< ?

#merged_state_counts$Change <- ifelse(merged_state_counts$freq_2012_2013 =< merged_state_counts$freq_2022_2023, 1, -1)

# set states with no change to 0
merged_state_counts$Change[is.na(merged_state_counts$Change)] <- 0

# create map of change 
usmap::plot_usmap(data = merged_state_counts, values = "Change") +
  labs(title = "Increase or Decrease in ad Counts per State from 2012-2013 to 2022-2023")

ggsave(here("analysis",
            "figures", 
            "change-map.png"),
       bg ="white",
       h = 10, # experiment with h and w to get the right size and proportion
       w = 12,
       units = "in",
       dpi = 900) # make the image nice and crisp))

```

### 2024-04-05 analysis/paper/paper.qmd

@@
   ylab("Number of\ntenure track job ads") +
-  theme_minimal(base_size = 14)
+  theme_minimal(base_size = 14) +
+  guides(x = "none")
 
@@
 
-years_to_2019 <- map_chr(2013:2019, ~str_glue('{.x}-{.x +1}'))
+years_to_2019 <- map_chr(2012:2019, ~str_glue('{.x}-{.x +1}'))
 years_after_2020 <- map_chr(2020:2022, ~str_glue('{.x}-{.x +1}'))
@@
        x = "") +
-  theme_minimal(base_size = 14) +
-   guides(x = "none")
+  theme_minimal(base_size = 14) #+
+   #guides(x = "none")
 
@@
 plot_grid(
+  fig_how_many_jobs_per_year,
+  fig_prop_by_job_title_per_year,
   fig_ratio_tt_2_ntt_jobs_per_year,
-  fig_prop_by_job_title_per_year,
-  fig_how_many_jobs_per_year,
+
   ncol = 1,
   align = "hv",
-  axis = "lr"
+  axis = "lr",
+  labels = "AUTO"
 )

### 2024-04-05 analysis/paper/paper.qmd
update requirements fig

@@
 
+# look at only these requirements because the others are flat
+
+intresting_requirements <- 
+c("cover letter",
+  "cv",
+  "names of recommenders",
+  "diversity statement",
+  "research statement",
+  "teaching statement")
+
 jobdata_requirements <- 
@@
   mutate(name = str_remove(name, "documents_requested_")) %>% 
-  mutate(name = str_replace_all(name, "_", " "))
+  mutate(name = str_replace_all(name, "_", " ")) %>% 
+  filter(name %in% intresting_requirements)
 
@@
              scales = "free_y",
-             nrow = 2) +
+             nrow = 3) +
   xlab("Year") +
@@
 
-# how many times each location mentioned?
+# how many times each topic mentioned?
 jobdata_topic_year %>%

### 2024-04-06 analysis/paper/paper.qmd
stash

--- 2024-04-05 analysis/paper/paper.qmd
@@
            colour = "red") +
-  labs(y = "Ratio of tenure-track to\nnon-tenure track and other",
+  labs(y = "Ratio of tenure-track-to-non-tenure track and other",
        x = "") +
@@
        bg ="white",
-       h = 10, # experiment with h and w to get the right size and proportion
-       w = 12,
+       h = 5, # experiment with h and w to get the right size and proportion
+       w = 7,
        units = "in",

### 2024-04-06 analysis/paper/paper.qmd
changed colour palette, resized figures

--- 2024-04-05 analysis/paper/paper.qmd
@@
   geom_bar() +
-  scale_x_discrete(name = "Academic Year") +
+  scale_x_discrete(name = "") +
   ylab("Number of\ntenure track job ads") +
@@
   theme(legend.position = c(0.5, 0.5)) +
+  scale_colour_brewer(palette = "Dark2") +
   guides(colour = guide_legend(nrow=2,
@@
            colour = "red") +
-  labs(y = "Ratio of tenure-track to\nnon-tenure track and other",
+  labs(y = "Ratio of tenure-track\nto non-tenure track and other",
        x = "") +
-  theme_minimal(base_size = 14) #+
+  theme_minimal(base_size = 14) +
+  scale_x_discrete(name = "Academic Year") 
    #guides(x = "none")
@@
   mutate(name = str_replace_all(name, "_", " ")) %>% 
-  filter(name %in% intresting_requirements)
+  filter(name %in% intresting_requirements) %>% 
+  mutate(name = str_wrap(name, 10),
+         year_ad_posted = str_replace(year_ad_posted, "-", "\n"))
 
@@
              scales = "free_y",
-             nrow = 3) +
+             nrow = 1) +
   xlab("Year") +
@@
   theme_minimal(base_size = 14) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
+  theme(axis.text.x = element_text(size = 8),
+        strip.text = element_text( size = 20))
 
@@
        h = 10, # experiment with h and w to get the right size and proportion 
-       w = 25,
+       w = 20,
        units = "in",
@@
                  colour = name),
-            size = 2,
+            size = 3,
            se = FALSE 
@@
   ylab("Proportion of all ads") +
+  scale_colour_brewer(palette = "Dark2") +
   guides(colour = guide_legend("Geographic\nfocus",
@@
                 colour = name),
-            size = 2,
+            size = 3,
            span = 0.7,
@@
   ylab("Proportion of all ads") +
+  scale_colour_brewer(palette = "Dark2") +
   guides(colour = guide_legend("Topic\nfocus",

### 2024-04-06 analysis/paper/paper.qmd
redo the US state map to use most of the data, rather than half of it

@@
 
+```{r}
+# Draw of map to show which states have done the most hiring in our sample
+
+# get the text in parentheses after the university name that gives the
+# state or country abb
+uni_state_country <-
+jobdata %>%
+ select(name_of_hiring_university) %>%
+  mutate(state_country = regmatches(name_of_hiring_university,
+                                    gregexpr( "(?<=\\().+?(?=\\))",
+                                              name_of_hiring_university,
+                                              perl = T))) %>%
+  unnest(state_country)
+
+# tally to get counts:
+uni_state_country_tally <-
+uni_state_country %>%
+  group_by(state_country) %>%
+  tally(sort = TRUE)
+
+# did we get all the job ads?
+sum(uni_state_country_tally$n) # 547 most of them, there are 552 ads in our data
+
+state_to_st <- function(x){
+       c(state.abb, 'DC')[match(x, c(state.name, 'District of Columbia'))]
+}
+
+state_name_and_abb <-
+enframe(state.name, value = 'state_name') %>%
+       mutate(state_abbr = state_to_st(state_name))
+
+# filter to get US states only
+uni_state_country_tally_us <-
+uni_state_country_tally %>%
+  filter(state_country %in% state.abb) %>%
+  select(state_abbr = state_country, n) %>%
+  # make sure we have all states in the dataframe
+  # even those with no jobs
+  right_join(state_name_and_abb) %>%
+  select(state = state_name, n, state_abbr) %>%
+  mutate(state = tolower(state)) %>%
+  mutate(n = ifelse(is.na(n), 0, n))
+
+library(ggplot2)
+library(fiftystater)
+library(tidyverse)
+library(ggrepel)
+
+data("fifty_states")
+
+ggplot(data= uni_state_country_tally_us,
+       aes(map_id = state)) +
+  geom_map(aes(fill = n),
+           color= "black",
+           linewidth = 0.1,
+           map = fifty_states) +
+  expand_limits(x = fifty_states$long,
+                y = fifty_states$lat) +
+  coord_map() +
+  geom_text_repel(data = fifty_states %>%
+              group_by(id) %>%
+              summarise(lat = mean(c(max(lat), min(lat))),
+                        long = mean(c(max(long), min(long)))) %>%
+              mutate(state = id) %>%
+              left_join(uni_state_country_tally_us,
+                        by = "state"),
+            aes(x = long,
+                y = lat,
+                label = n,
+                bg.color = "white",
+                bg.r = 0.1),
+            force = 0,
+            force_pull = 100)+
+  scale_x_continuous(breaks = NULL) +
+  scale_y_continuous(breaks = NULL) +
+  labs(x = "",
+       y = "") +
+  theme(legend.position = "bottom",
+        panel.background = element_blank()) +
+  scale_fill_viridis_c()
+
+ggsave(here("analysis",
+            "figures",
+            "fig-us-state-map.png"),
+       bg ="white",
+       h = 10, # experiment with h and w to get the right size and proportion
+       w = 12,
+       units = "in",
+       dpi = 900) # make the image nice and crisp))
+
+```
+
 # Discussion

### 2024-04-06 analysis/paper/sketches.R

@@ -160,9 +160,95 @@ jobdata %>%
 
 #-----------------------------------------------------------------------------
 
+# Draw of map to show which states have done the most hiring in our sample
+
+# get the text in parentheses after the university name that gives the
+# state or country abb
+uni_state_country <-
+jobdata %>%
+ select(name_of_hiring_university) %>%
+  mutate(state_country = regmatches(name_of_hiring_university,
+                                    gregexpr( "(?<=\\().+?(?=\\))",
+                                              name_of_hiring_university,
+                                              perl = T))) %>%
+  unnest(state_country)
+
+# tally to get counts:
+uni_state_country_tally <-
+uni_state_country %>%
+  group_by(state_country) %>%
+  tally(sort = TRUE)
+
+# did we get all the job ads?
+sum(uni_state_country_tally$n) # 547 most of them, there are 552 ads in our data
+
+state_to_st <- function(x){
+       c(state.abb, 'DC')[match(x, c(state.name, 'District of Columbia'))]
+}
+
+state_name_and_abb <-
+enframe(state.name, value = 'state_name') %>%
+       mutate(state_abbr = state_to_st(state_name))
+
+# filter to get US states only
+uni_state_country_tally_us <-
+uni_state_country_tally %>%
+  filter(state_country %in% state.abb) %>%
+  select(state_abbr = state_country, n) %>%
+  # make sure we have all states in the dataframe
+  # even those with no jobs
+  right_join(state_name_and_abb) %>%
+  select(state = state_name, n, state_abbr) %>%
+  mutate(state = tolower(state)) %>%
+  mutate(n = ifelse(is.na(n), 0, n))
+
+library(ggplot2)
+library(fiftystater)
+library(tidyverse)
+library(ggrepel)
+
+data("fifty_states")
+
+ggplot(data= uni_state_country_tally_us,
+       aes(map_id = state)) +
+  geom_map(aes(fill = n),
+           color= "black",
+           linewidth = 0.1,
+           map = fifty_states) +
+  expand_limits(x = fifty_states$long,
+                y = fifty_states$lat) +
+  coord_map() +
+  geom_text_repel(data = fifty_states %>%
+              group_by(id) %>%
+              summarise(lat = mean(c(max(lat), min(lat))),
+                        long = mean(c(max(long), min(long)))) %>%
+              mutate(state = id) %>%
+              left_join(uni_state_country_tally_us,
+                        by = "state"),
+            aes(x = long,
[…]

### 2024-04-07 analysis/paper/paper.qmd
update data with state abb for a few schools that were missed, redraw map

@@
 
-total_number_of_ads_in_our_sample <- nrow(jobdata)
+total_number_of_ads_in_our_sample <- nrow(jobdata) # 550
 ```
@@
 # state or country abb
-uni_state_country <-
-jobdata %>%
+uni_state_country <- # 550
+jobdata %>%  # 550 rows
  select(name_of_hiring_university) %>%
@@
 # did we get all the job ads?
-sum(uni_state_country_tally$n) # 547 most of them, there are 552 ads in our data
+# sum(uni_state_country_tally$n) # 550 all of them
 
@@
   mutate(n = ifelse(is.na(n), 0, n))
+
+# how many jobs ads now?
+# sum(uni_state_country_tally_us$n) # 433, 78% of the total
 

### 2024-04-17 analysis/paper/paper.qmd
stash

@@
 
+```{r}
+# method focus by year
+
+library(googlesheets4)
+library(stringi)
+
+method_foci <-
+  read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0",
+             sheet = "method")
+
+method_foci_clean <-
+  map(
+    str_split(method_foci$`From the data`, ";"),
+    ~.x %>%
+      str_squish() %>%
+      stri_remove_empty() %>%
+      str_to_lower)
+
+jobdata_method <-
+  jobdata %>%
+  select(methods_focus_of_position) %>%
+  mutate(methods_focus_of_position = str_to_lower(methods_focus_of_position))
+
+jobdata_method <-
+  # add one column for each topic in our categories
+  cbind(jobdata_method,
+        setNames( lapply(method_foci$Category, function(x) x=NA),
+                  method_foci$Category) )
+
+for(i in 1:length(method_foci$Category)){
+
+  this_method <- method_foci$Category[i]
+
+  # create the pattern to search for
+  x <- paste0(method_foci_clean[[i]], collapse = "|")
+
+  # do the search through all the job ads for that pattern
+  y <- str_detect(jobdata_method$methods_focus_of_position,
+                  x)
+
+  # assign back to our data frame in the appropriate location column
+  jobdata_method[, this_method] <- y
+
+}
+
+jobdata_method_year <-
+  jobdata %>%
+  bind_cols(jobdata_method) %>%
+  select(year_ad_posted,
+         method_foci$Category) %>%
+  pivot_longer(-year_ad_posted) %>%
+  drop_na()
+
+# how many times each topic mentioned?
+jobdata_method_year %>%
+  group_by(name) %>%
+  summarise(n = sum(value)) %>%
+  arrange(desc(n)) %>%
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  xlab("") +
+  theme_minimal() +
+  coord_flip()
+
+# explore trends over time. put a point on the max year
+jobdata_method_year_tally <-
+  jobdata_method_year %>%
+  # exclude those with <20 ads
+  filter(!name %in% c("Digital Archaeology",
+                      "Pleistocene archaeology",
+                      "Mesoamerican Archaeology",
+                      "Biological anthropology",
+                      "Archaeological theory",
+                      "Evolutionary anthropology",
+                      "North American archaeology"
+  )) %>%
+  group_by(year_ad_posted,
+           name) %>%
+  summarise(n = sum(value)) %>%
+  mutate(prop = n / sum(n))
+
+jobdata_method_year_tally_max <-
+  jobdata_method_year_tally %>%
+  group_by(
+    name ) %>%
+  filter(prop == max(prop))
+
+library(ggrepel)
+
+ggplot() +
+  geom_smooth(data =  jobdata_method_year_tally,
+            aes(year_ad_posted,
+                prop,
+                group = name,
+                colour = name),
+            size = 3,
+           span = 0.7,
+            se = FALSE) +
+  xlab("Year") +
+  ylab("Proportion of all ads") +
+  scale_colour_brewer(palette = "Dark2") +
+  guides(colour = guide_legend("Method\nfocus",
+                               label.position = "bottom")) +
+    theme_minimal( base_size = 14) +
+  theme(legend.position="bottom")
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-topic-focus-by-year.png"),
+       bg ="white",
+       h = 10, # experiment with h and w to get the right size and proportion
+       w = 12,
+       units = "in",
+       dpi = 900) # make the image nice and crisp))
+```
+
 # Discussion

### 2024-04-17 analysis/supplementary-materials/universities.Rmd

--- 2024-04-05 universities.Rmd
@@
 # read in packages and datasets
-install.packages("usmap")
+# install.packages("usmap")
 library(usmap) #import the package
@@
 # Carnegie classification data 
-universities <- read_excel("cc.xlsx")
+universities <- readxl::read_excel("cc.xlsx")
 
@@
 univ_and_state <- select(universities, name, state)
-head(univ_and_state)
+# head(univ_and_state)
 
@@
 # remove () from jobdata
-jobdata$`University Name` <-  gsub(r"{\s*\([^\)]+\)}","",as.character(jobdata$`University Name`))
+jobdata$`University Name` <- 
+  gsub(r"{\s*\([^\)]+\)}",
+       "",
+       as.character(jobdata$`University Name`))
 
@@
 # plot freq in ads / state
-usmap::plot_usmap(data = state_counts, values = "Freq") +
+usmap::plot_usmap(data = state_counts, 
+                  values = "Freq") +
   labs(title = "Ad Frequency Per Universities in Each State from 2012-2013 to 2022-2023")

### 2024-04-17 analysis/paper/paper.qmd
resolve merge conflict
stash

--- 2024-04-07 analysis/paper/paper.qmd
@@
 geographic_foci <-
-read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0")
+read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0", sheet = "geography")
 
@@
 united_states_regions <-
-c( "Northeastern US",
-   "Midwest US",
-   "Southeast US",
-   "Southwest US",
-   "Western US",
-   "Southern US",
-   "Eastern US"  )
+str_subset(geographic_foci$Category2, "US")
 
@@
  # exclude those with <20 ads
-  filter(!name %in% c("Canada",
-                      "Americas",
-                      "Arctic",
+  filter(!name %in% c("Canada & Arctic",
                       "Oceania",
-                      "Eastern US",
-                      "Southern US",
+                      "Southeast US",
+                      "Southwest US",
+                      "Western US",
                       "Midwest US",
@@
 # what about within the US
-jobdata_geo_us_year <-
-  jobdata %>%
-  bind_cols(jobdata_geo) %>%
-  select(year_ad_posted,
-         geographic_foci$Category2) %>%
-  pivot_longer(-year_ad_posted) %>%
-  drop_na()
+# how many times each location mentioned?
+jobdata_geo_year %>%
+  group_by(name) %>%
+  summarise(n = sum(value)) %>%
+  arrange(desc(n)) %>%
+  filter(name %in% united_states_regions) %>% 
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  ylab("Number of ads") +
+  xlab("") +
+  theme_minimal(base_size = 24) +
+  coord_flip()
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-geo-us-focus-by-year.png"),
+       bg ="white",
+       h = 5, # experiment with h and w to get the right size and proportion 
+       w = 6,
+       units = "in",
+       dpi = 900) # make the image nice and crisp)
 

### 2024-04-17 analysis/paper/paper.qmd
correct the file name for methods figure

@@
             "figures", 
-            "fig-topic-focus-by-year.png"),
+            "fig-method-focus-by-year.png"),
        bg ="white",

### 2024-04-17 analysis/paper/paper.qmd
minor edit to methods

@@
   ylab("Proportion of all ads") +
-  scale_colour_brewer(palette = "Dark2") +
   guides(colour = guide_legend("Method\nfocus",

### 2024-04-18 analysis/paper/paper.qmd
uploaded carneige classification

@@
   # exclude those with <20 ads
-  filter(!name %in% c("Digital Archaeology",
-                      "Pleistocene archaeology",
-                      "Mesoamerican Archaeology",
-                      "Biological anthropology",
-                      "Archaeological theory",
-                      "Evolutionary anthropology",
-                      "North American archaeology"
+  filter(!name %in% c("Material culture analysis"
   )) %>%
@@
 
+```{r}
+# Carneige Classification
+library(googlesheets4)
+library(stringi)
+
+CC <-
+  read_sheet( "https://docs.google.com/spreadsheets/d/1Oe_3X-OR7iuUThb6Tp48prBpbHrWjU4osmHKi1Hbqgc/edit#gid=1922578308",
+             sheet = "Values1")
+Basic2021 <-
+  read_sheet( "https://docs.google.com/spreadsheets/d/1Oe_3X-OR7iuUThb6Tp48prBpbHrWjU4osmHKi1Hbqgc/edit#gid=1922578308",
+             sheet = "Data1")
+
+uni_name <- # 550
+jobdata %>%  # 550 rows
+ select(name_of_hiring_university) %>%
+ mutate(name_of_hiring_university = str_replace(name_of_hiring_university, "\\s*\\(.*?\\)$", ""))
+
+view(uni_name)
+
+```
+
 # Discussion

### 2024-04-19 analysis/paper/paper.qmd
notes on joining the Carnegie university names with our university names

@@
 ```{r}
-# Carneige Classification
+# Carnegie Classification
 library(googlesheets4)
@@
   read_sheet( "https://docs.google.com/spreadsheets/d/1Oe_3X-OR7iuUThb6Tp48prBpbHrWjU4osmHKi1Hbqgc/edit#gid=1922578308",
-             sheet = "Data1")
+             sheet = "Data1") %>% 
+  # remove a bunch of text from uni names to improve our joins
+  mutate(name = str_remove(name, 
+                           " Campus Immersion|-Main Campus"))
 
@@
  jobdata %>%  # 550 rows
- select(name_of_hiring_university) %>%
- mutate(name_of_hiring_university = str_replace(name_of_hiring_university, "\\s*\\(.*?\\)$", ""))
+  # remove parentheses and their contents
+  mutate(name = str_squish(str_replace(name_of_hiring_university,
+                            "\\s*\\(.*?\\)$", 
+                            ""))) %>% 
+  mutate(name = str_replace(name,
+                            "California State University,[[:space:]]",
+                            "California State University-")) %>%
 
-view(uni_name)
+  left_join(Basic2021, 
+            keep = TRUE )
+
+# how many schools did we match with the CC data
+uni_name %>% 
+  filter(!is.na(basic2021)) %>% 
+  nrow() # 245, 290, 300, 313, getting there!  / 500
+
+# which schools don't match with the CC data?
+uni_name %>% 
+  select(name.x, name.y)  %>% 
+  filter(is.na(name.y)) %>% 
+  distinct() %>% View
+
+# for each of these, we need to check
+# - In the US or not? If not in the US, we can ignore, if in the US...
+# - Compare the uni name in the uni_name data frame to the Basic2021 dataframe.
+Sometimes there are simple punctuation differences, but most differences seem to be because the Basic2021 data have a location near the university name, which our raw data currently lack.
+# - For example, our spreadsheet we had "Texas A&M University", but in the Basic2021 dataframe we see three possible matches:
+# Texas A&M University-Texarkana
+# Texas A&M University-San Antonio
+# Texas A&M University-Central Texas
+# so we need to take a look at the details of the job ad, and find the details to confirm which is the correct match for the "Texas A&M University" in our data. 
+# Once we've confirmed the correct match, we update the text in the appropriate row of our raw data, column F of this sheet https://docs.google.com/spreadsheets/d/1Jwe3UqJyedrV-QWlwR_44__t4xBVrCfxGyhXdi3E0sg/edit?resourcekey#gid=1686084773 to match what the university name is in the Basic2021 dataframe.
+So the full university name in our spreadsheet becomes "Texas A&M University-San Antonio (TX)", then we can match the university name when we join, and we keep the state abbreviation for mapping.
+I've done this for several schools, can you please continue it?
+# - if it's not clear which university is the correct match to our data, choose the one with the highest CC rank.
+For example, for three U of Texas campuses, the ranks are 19, 18, 19.
+I could not confirm from the details of the job ad, so I choose Texas A&M University-San Antonio because it has the rank of 18, higher than the other two.
+It's most likely that the biggest and higher ranking campus will hire an archaeology professor, because smaller campuses tend not to have archaeologists.
 

### 2024-04-19 analysis/paper/paper.qmd
more notes on matching the Carnegie data

@@
 It's most likely that the biggest and higher ranking campus will hire an archaeology professor, because smaller campuses tend not to have archaeologists.
+# - then download our raw data as CSV, move it into this repo, and re-run the code to read it in and join with the CC data, and see how many more jobs match with the CC school names.
+On our SAA poster we say we have 433 US jobs, so we should be able to match around that many.
+The rest are international unis that wont be in the CC data
 

### 2024-04-22 analysis/paper/paper.qmd
minor update

--- 2024-04-18 analysis/paper/paper.qmd
@@
 
+# data from https://carnegieclassifications.acenet.edu/resource/2021-update-public-file/
 CC <-
@@
   read_sheet( "https://docs.google.com/spreadsheets/d/1Oe_3X-OR7iuUThb6Tp48prBpbHrWjU4osmHKi1Hbqgc/edit#gid=1922578308",
-             sheet = "Data1")
+             sheet = "Data1") %>% 
+  # remove a bunch of text from uni names to improve our joins
+  mutate(name = str_remove(name, 
+                           " Campus Immersion|-Main Campus"))
 
-uni_name <- # 550
-jobdata %>%  # 550 rows
- select(name_of_hiring_university) %>%
- mutate(name_of_hiring_university = str_replace(name_of_hiring_university, "\\s*\\(.*?\\)$", ""))
+uni_name <- # 547
+ jobdata %>%  # 547 rows
+  # remove parentheses and their contents
+  mutate(name = str_squish(str_replace(name_of_hiring_university,
+                            "\\s*\\(.*?\\)$", 
+                            ""))) %>% 
+  mutate(name = str_replace(name,
+                            "California State University,[[:space:]]",
+                            "California State University-")) %>%
 
-view(uni_name)
+  left_join(Basic2021, 
+            keep = TRUE )
+
+# how many schools did we match with the CC data
+uni_name %>% 
+  filter(!is.na(basic2021)) %>% 
+  nrow() # 245, 290, 300, 313, 404 / 433 for US unis
+
+# which schools don't match with the CC data?
+uni_name %>% 
+  select(name.x, name.y)  %>% 
+  filter(is.na(name.y)) %>% 
+  distinct()
+
+uni_name_tally <- 
+uni_name %>% 
+  left_join(CC,
+            join_by("basic2021" == "Value")) %>% 
+  group_by(Category) %>% 
+  tally(sort = TRUE) %>% 
+  drop_na()
+
+ggplot(uni_name_tally) +
+  aes(reorder(Category, n), n) +
+  geom_col() +
+  coord_flip() +
+  xlab("") +
+  theme_minimal(base_size = 14)
 

### 2024-04-22 analysis/paper/paper.qmd
stash

--- 2024-04-19 analysis/paper/paper.qmd
@@
   read_sheet( "https://docs.google.com/spreadsheets/d/1Oe_3X-OR7iuUThb6Tp48prBpbHrWjU4osmHKi1Hbqgc/edit#gid=1922578308",
-             sheet = "Data1") %>% 
+             sheet = "Data1") 
   # remove a bunch of text from uni names to improve our joins
-  mutate(name = str_remove(name, 
-                           " Campus Immersion|-Main Campus"))
 

### 2024-05-01 README.Rmd
update readme

@@
 
-This repository contains the data and code for our paper:
+This repository contains the data and code for our poster presented at the Society of American Archaeology Annual Meeting in New Orleans 17-21 April 2024:
 
-> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>
-
-Our pre-print is online here:
-
-> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
-
-### How to cite
-
-Please cite this compendium as:
-
-> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>
+> Poole, Anne Marie, Ailin Zhang, Setareh Shafizadeh, Jess Beck, Ben Marwick, 2024.
+Careers in Ruins: Academic Archaeology Job Trends From 2013 – 2023.
+Poster presented at the SAA annual meeting in New Orleans, 17-21 April 2024
 
@@
 
-  - [:file\_folder: paper](/analysis/paper): Quarto source document
+  - [:file\_folder: paper](/analysis/paper): PDF file of the poster we presented at the SAA. Quarto source document
     for manuscript. Includes code to reproduce the figures and tables
-    generated by the analysis. It also has a rendered version,
-    `paper.docx`, suitable for reading (the code is replaced by figures
-    and tables in this file)
+    generated by the analysis. 
   - [:file\_folder: data](/analysis/data): Data used in the analysis.

### 2024-05-03 analysis/paper/paper.qmd
add SAA poster and update readme
update figure sizes for URS poster

@@
        bg ="white",
-       h = 11, # experiment with h and w to get the right size and proportion 
+       h = 15, # experiment with h and w to get the right size and proportion 
        w = 20,
@@
        h = 10, # experiment with h and w to get the right size and proportion 
-       w = 12,
+       w = 20,
        units = "in",
@@
        bg ="white",
-       h = 5, # experiment with h and w to get the right size and proportion 
-       w = 6,
+       h = 10, # experiment with h and w to get the right size and proportion 
+       w = 20,
        units = "in",
@@
        h = 10, # experiment with h and w to get the right size and proportion
-       w = 12,
+       w = 20,
        units = "in",
@@
   filter(!is.na(basic2021)) %>% 
-  nrow() # 245, 290, 300, 313, 404 / 433 for US unis
+  nrow() # 245, 290, 300, 313, 404, 430 / 433 for US unis
 
@@
 
-ggplot(uni_name_tally) +
+uni_name_tally %>% 
+  mutate(Category = str_wrap(Category, width = 30)) %>% 
+ggplot() +
   aes(reorder(Category, n), n) +
@@
   theme_minimal(base_size = 14)
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-carnegie-classification.png"),
+       bg ="white",
+       h = 7, # experiment with h and w to get the right size and proportion
+       w = 12,
+       units = "in",
+       dpi = 900) # make the image nice and crisp))
 

### 2024-05-06 analysis/paper/paper.qmd
base size 28

@@
                                label.position = "bottom")) +
-    theme_minimal( base_size = 14) +
+    theme_minimal( base_size = 28) +
   theme(legend.position="bottom")

### 2024-05-06 README.Rmd

@@
 
-# archyjobads
+# Careers in Ruins: Academic Archaeology Job Trends From 2013 – 2023
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/archyjobads/master?urlpath=rstudio)
+[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/archyjobads/master?urlpath=rstudio) [![DOI:10.17605/OSF.IO/J58K](https://img.shields.io/badge/DOI-10.17605/OSF.IO/J58KH-brightgreen.svg)](http://doi.org/10.17605/OSF.IO/J58KH)
 
@@
 Careers in Ruins: Academic Archaeology Job Trends From 2013 – 2023.
-Poster presented at the SAA annual meeting in New Orleans, 17-21 April 2024
+Poster presented at the SAA annual meeting in New Orleans, 17-21 April 2024.
+doi: 10.17605/OSF.IO/J58KH
+
+![](README-fig-001.png)
 

### 2024-05-06 analysis/paper/paper.qmd
stash
28 for all panel plots

@@
 
+# how many times each topic mentioned?
+jobdata_method_year_tally_max %>%
+
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  xlab("") +
+  theme_minimal() +
+  coord_flip()
+
 library(ggrepel)
@@
                                label.position = "bottom")) +
-    theme_minimal( base_size = 14) +
+    theme_minimal( base_size = 28) +
+    scale_colour_brewer(palette = "Dark2") +
   theme(legend.position="bottom")

### 2024-05-06 analysis/paper/paper.qmd
methods over time

@@
 library(stringi)
+1
 
@@
   # exclude those with <20 ads
-  filter(!name %in% c("Material culture analysis"
+  filter(!name %in% c("Material culture analysis",
+                      "Ceramic analysis"
   )) %>%
@@
   mutate(prop = n / sum(n))
-
-jobdata_method_year_tally_max <-
-  jobdata_method_year_tally %>%
-  group_by(
-    name ) %>%
-  filter(prop == max(prop))
-
-# how many times each topic mentioned?
-jobdata_method_year_tally_max %>%
-
-  ggplot() +
-  aes(reorder(name, n),
-      n)+
-  geom_col() +
-  xlab("") +
-  theme_minimal() +
-  coord_flip()
 
@@
 
+view(jobdata)
 ```

### 2024-05-06 analysis/paper/paper.qmd
increase text size on panel job numbers/prop per year

@@
   ylab("Number of\ntenure track job ads") +
-  theme_minimal(base_size = 14) +
+  theme_minimal(base_size = 28) +
   guides(x = "none")
@@
   geom_line(linewidth = 2) +
-  theme_minimal(base_size = 14) +
+  theme_minimal(base_size = 28) +
   xlab("") +
@@
        x = "") +
-  theme_minimal(base_size = 14) +
+  theme_minimal(base_size = 28) +
   scale_x_discrete(name = "Academic Year") 

### 2024-05-13 analysis/paper/paper.qmd
EEO

@@
 
+```{r}
+# EEO
+library(stringr)
+library(dplyr)
+library(stopwords)
+library(tidyverse)
+library(tm)
+
+stopwords <- stopwords("en")
+
+EEO <- jobdata %>%
+  select(equal_employment_opportunity_statement)
+
+remove_stopwords <- function(text, stopwords) {
+  pattern <- paste0("\\b(", paste(stopwords, collapse = "|"), ")\\b")
+  str_remove_all(text, pattern)
+}
+
+EEO_wordcount <- EEO %>%
+  mutate(word_count = str_count(equal_employment_opportunity_statement, "\\S+")) %>%
+  print()
+
+EEO_clean <- EEO %>%
+  mutate(equal_employment_opportunity_statement = str_to_lower(equal_employment_opportunity_statement), 
+         equal_employment_opportunity_statement = str_remove_all(equal_employment_opportunity_statement, "[[:punct:]]"),
+         equal_employment_opportunity_statement = remove_stopwords(equal_employment_opportunity_statement, stopwords))
+
+EEO_word <- EEO_clean %>%
+   unnest_tokens(word, text) %>%
+  anti_join(get_stopwords()) %>%
+  count(word, sort = TRUE)
+```
+
+```{r}
+ggplot(word_data, aes(x = reorder(word, n), y = n)) +
+  geom_col() +
+  coord_flip() +  # Flips the axes for easier reading of text
+  labs(x = "Words", y = "Frequency", title = "Word Frequency")
+```
+
 # Discussion

### 2024-05-13 analysis/paper/paper.qmd
exploring EEO

@@
 library(tidyverse)
+library(tidytext)
 library(tm)
@@
 EEO <- jobdata %>%
-  select(equal_employment_opportunity_statement)
+  select(equal_employment_opportunity_statement) %>% 
+  drop_na
 
@@
 
+ggplot(EEO_wordcount) +
+  aes(word_count) +
+  geom_histogram()
+
 EEO_clean <- EEO %>%
@@
 
+# this doesn't work, need help to troubleshoot
 EEO_word <- EEO_clean %>%
-   unnest_tokens(word, text) %>%
+   unnest_tokens(word, 
+                 equal_employment_opportunity_statement) %>%
   anti_join(get_stopwords()) %>%

### 2024-05-13 analysis/paper/paper.qmd
investigating words of interest in the EEO

@@
   count(word, sort = TRUE)
+
+# here's the list of words of interest
+community_word <- c("women","color","religion","race","age","veteran","disability")
+
+# prepare a dataframe to store the count of words of interest
+df <- data.frame(matrix(ncol = length(community_word), 
+                        nrow = 0))
+colnames(df) <- community_word
+
+for(i in 1: nrow(EEO_clean)){
+  df[i, ] <- str_detect(EEO_clean[i, ], 
+                   community_word)
+}
+
+# detect if the words of interest is present or not.
+EEO_clean_community_word <- 
+  bind_cols(EEO_clean, df) %>% 
+  rowwise() %>% 
+  mutate(community_word_tally = sum(c_across(all_of(community_word)), 
+                                    na.rm = T))
+
 ```

### 2024-05-13 analysis/paper/paper.qmd
EEO

--- 2024-05-13 analysis/paper/paper.qmd
@@
 library(stringi)
-1
 
@@
 EEO_word <- EEO_clean %>%
-   unnest_tokens(word, text) %>%
+   unnest_tokens(word, 
+                 equal_employment_opportunity_statement) %>%
   anti_join(get_stopwords()) %>%
-  count(word, sort = TRUE)
-```
-
-```{r}
-ggplot(word_data, aes(x = reorder(word, n), y = n)) +
+  count(word, sort = TRUE) %>%
+  drop_na()
+
+community_word <- c("women","color","religion","race","age","veteran","disability")
+
+```
+
+```{r}
+ggplot(EEO_word, aes(x = reorder(word, n), y = n)) +
   geom_col() +
-  coord_flip() +  # Flips the axes for easier reading of text
+  coord_flip() +  
   labs(x = "Words", y = "Frequency", title = "Word Frequency")

### 2024-05-13 analysis/paper/paper.qmd
plotting EEO word freq

--- 2024-05-13 analysis/paper/paper.qmd
@@
 ```{r}
-ggplot(word_data, aes(x = reorder(word, n), y = n)) +
+
+EEO_word %>% 
+  filter(n >= 20) %>% 
+ggplot() +
+  aes(x = reorder(word, n), y = n) +
   geom_col() +

### 2024-05-24 analysis/paper/paper.qmd

--- 2024-05-13 analysis/paper/paper.qmd
@@
   anti_join(get_stopwords()) %>%
+
+  count(word, sort = TRUE) %>%
+  drop_na()
+
+# here's the list of words of interest
+community_word <- c("women","color","religion","race","age","veteran","disability","citizenship","ancestry","national origin","nondiscrimination","sex","gender","marital status","genetic information","breastfeeding","sexual orientation","discriminate","transgender status","gender expression","veteran")
 count(word, sort = TRUE)
-
-# here's the list of words of interest
-community_word <- c("women","color","religion","race","age","veteran","disability")
 
@@
 ```{r}
-ggplot(word_data, aes(x = reorder(word, n), y = n)) +
+ggplot(EEO_clean, aes(x = reorder(word, n), y = n)) +
   geom_col() +
-  coord_flip() +  # Flips the axes for easier reading of text
-  labs(x = "Words", y = "Frequency", title = "Word Frequency")
+  labs(x = "Words", y = "Frequency", title = "Word Frequency") +
+  ylim(10,50)
 ```

### 2024-05-24 analysis/paper/paper.qmd

@@
   anti_join(get_stopwords()) %>%
-
   count(word, sort = TRUE) %>%

### 2024-05-24 analysis/paper/paper.qmd
exploring EEO words

--- 2024-05-13 analysis/paper/paper.qmd
@@
 
-# this doesn't work, need help to troubleshoot
 EEO_word <- EEO_clean %>%
@@
   anti_join(get_stopwords()) %>%
-  count(word, sort = TRUE)
+  count(word, sort = TRUE) %>%
+  drop_na()
 
 # here's the list of words of interest
-community_word <- c("women","color","religion","race","age","veteran","disability")
+community_word <-
+  c(
+    "women",
+    "color",
+    "religion",
+    "race",
+    "age",
+    "veteran",
+    "disability",
+    "citizenship",
+    "ancestry",
+    "national origin",
+    "sex",
+    "gender",
+    "marital status",
+    "genetic information",
+    "breastfeeding",
+    "sexual orientation",
+    "transgender status",
+    "gender expression"
+  )
 
@@
   mutate(community_word_tally = sum(c_across(all_of(community_word)), 
-                                    na.rm = T))
-
-```
-
-```{r}
-
-EEO_word %>% 
-  filter(n >= 20) %>% 
-ggplot() +
-  aes(x = reorder(word, n), y = n) +
+                                    na.rm = T)) %>% 
+  bind_cols(EEO_wordcount) %>% 
+  mutate(eeo_words_total_ratio = community_word_tally / word_count)
+
+ggplot(EEO_clean_community_word) +
+  aes(community_word_tally,
+      word_count,
+      size = eeo_words_total_ratio) +
+  geom_point()
+
+```
+
+```{r}
+ggplot(EEO_word %>% 
+         filter(n > 20), 
+       aes(x = reorder(word, n), 
+           y = n)) +
   geom_col() +
-  coord_flip() +  # Flips the axes for easier reading of text
-  labs(x = "Words", y = "Frequency", title = "Word Frequency") 
+  labs(x = "Words", 
+       y = "Frequency", 
+       title = "Word Frequency") +
+  coord_flip()
 ```

### 2024-05-24 analysis/paper/paper.qmd
explore QS rankings

@@
 
+```{r}
+# explore QS rankings
+
+qs_rankings <- 
+  readxl::read_excel("analysis/data/raw_data/2024 QS WUR by Subject - Public Results (for qs.com)_48.xlsx", sheet = "Archaeology", skip = 12)
+
+jobdata %>% 
+  mutate(Institution = str_remove(name_of_hiring_university,
+                                  " \\(.*\\)")) %>% 
+  mutate(Institution = case_when(
+    # modify the school name to match with the name in the QS list
+    Institution == "Ohio State University-Main Campus" ~ "The Ohio State University",
+    # ... 
+    .default = as.character(Institution)
+  )) %>% 
+  select(Institution) %>% 
+  left_join(qs_rankings) %>% 
+  drop_na(`2024`) %>% 
+  distinct(Institution, .keep_all = TRUE) %>% 
+  View
+
+```
+
 # Discussion

### 2024-05-28 analysis/paper/paper.qmd
update on community word

--- 2024-05-24 analysis/paper/paper.qmd
@@
 EEO <- jobdata %>%
-  select(equal_employment_opportunity_statement) %>% 
+  select(equal_employment_opportunity_statement, year_ad_posted) %>% 
   drop_na
@@
 
-EEO_wordcount <- EEO %>%
-  mutate(word_count = str_count(equal_employment_opportunity_statement, "\\S+")) %>%
-  print()
-
-ggplot(EEO_wordcount) +
-  aes(word_count) +
-  geom_histogram()
-
 EEO_clean <- EEO %>%
@@
 
-# this doesn't work, need help to troubleshoot
-EEO_word <- EEO_clean %>%
-   unnest_tokens(word, 
-                 equal_employment_opportunity_statement) %>%
-  anti_join(get_stopwords()) %>%
-  count(word, sort = TRUE) %>%
-  drop_na()
+#word count for EEO statement
+EEO_wordcount <- EEO %>%
+  mutate(word_count = str_count(equal_employment_opportunity_statement, "\\S+")) %>%
+  print()
+
+jobdata_EEO_year_tally <-
+  EEO_wordcount %>%
+  group_by(year_ad_posted,
+           word_count) %>%
+  summarise(n = sum(value)) %>%
+  mutate(prop = n / sum(n))
+
+ggplot(EEO_wordcount) +
+  aes(word_count, year_ad_posted) +
+  geom_histogram()
 
 # here's the list of words of interest
-community_word <- c("women","color","religion","race","age","veteran","disability","citizenship","ancestry","national origin","nondiscrimination","sex","gender","marital status","genetic information","breastfeeding","sexual orientation","discriminate","transgender status","gender expression","veteran")
-count(word, sort = TRUE)
+community_word <- c("women","color","religion","race","age","veteran","disability","citizenship","national origin","sex","gender","marital status","genetic information","breastfeeding","sexual orientation","transgender status","gender expression","gender identity","ancestry","income assignment for child support","arrest and court record","domestic or sexual violence victim status","national guard absence","visible minorities","aboriginal persons","indigenous","LGBTQ2S+","pregnancy","childbirth","faith","genetic or family medical history","medical condition","nationality","ethnicity","parental status","accent","ethnic","creed","political affiliation","immigration status")
 
@@
 
-```
-
-```{r}
-ggplot(EEO_clean, aes(x = reorder(word, n), y = n)) +
-  geom_col() +
-  labs(x = "Words", y = "Frequency", title = "Word Frequency") +
-  ylim(10,50)
+ggplot(EEO_clean_community_word) +
+  aes(community_word_tally) +
+  geom_histogram()
 ```

### 2024-05-31 analysis/paper/paper.qmd
QS

@@
 
+fig_ave_length_EEO_per_year <- 
+ggplot(jobdata_EEO_year_tally) +
+  aes(year_ad_posted) +
+  geom_bar() +
+  scale_x_discrete(name = "") +
+  ylab("Average length of EEO statement by year") +
+  theme_minimal(base_size = 28) +
+  guides(x = "none")
+
 ggplot(EEO_wordcount) +
@@
 community_word <- c("women","color","religion","race","age","veteran","disability","citizenship","national origin","sex","gender","marital status","genetic information","breastfeeding","sexual orientation","transgender status","gender expression","gender identity","ancestry","income assignment for child support","arrest and court record","domestic or sexual violence victim status","national guard absence","visible minorities","aboriginal persons","indigenous","LGBTQ2S+","pregnancy","childbirth","faith","genetic or family medical history","medical condition","nationality","ethnicity","parental status","accent","ethnic","creed","political affiliation","immigration status")
+
+EEO_word <- EEO_clean %>%
+   unnest_tokens(word, 
+                 equal_employment_opportunity_statement) %>%
+  anti_join(get_stopwords()) %>%
+  count(word, sort = TRUE) %>%
+  drop_na()
 
@@
   mutate(community_word_tally = sum(c_across(all_of(community_word)), 
-                                    na.rm = T))
+                                    na.rm = T)) %>% 
+  bind_cols(EEO_wordcount) %>% 
+  mutate(eeo_words_total_ratio = community_word_tally / word_count)
+
+ggplot(EEO_clean_community_word) +
+  aes(community_word_tally,
+      word_count,
+      size = eeo_words_total_ratio) +
+  geom_point()
 
@@
 
+```{r}
+ggplot(EEO_word %>% 
+         filter(n > 20), 
+       aes(x = reorder(word, n), 
+           y = n)) +
+  geom_col() +
+  labs(x = "Words", 
+       y = "Frequency", 
+       title = "Word Frequency") +
+  coord_flip()
+```
+
+```{r}
+# explore QS rankings
+
+qs_rankings <- 
+  readxl::read_excel("analysis/data/raw_data/2024 QS WUR by Subject - Public Results (for qs.com)_48.xlsx", sheet = "Archaeology", skip = 12)
+
+jobdata %>% 
+  mutate(Institution = str_remove(name_of_hiring_university,
+                                  " \\(.*\\)")) %>% 
+  mutate(Institution = case_when(
+    # modify the school name to match with the name in the QS list
+    Institution == "Ohio State University-Main Campus" ~ "The Ohio State University",
+  "Cambridge" ~ "University of Cambridge",
+  "Cambridge University" ~ "University of Cambridge",
+  "Keble College" ~ "University of Oxford",
+  "University of California-Berkeley" ~ "University of California, Berkeley (UCB)",
+"University of California-Los Angeles" ~ "University of California, Los Angeles (UCLA)",
+"University of Tuebingen" ~ "Eberhard Karls Universität Tübingen",
+"Universität Tübingen" ~ "Eberhard Karls Universität Tübingen",
+"Arizona State University Campus Immersion" ~ "Arizona State University",
+"Kiel University" ~ "Christian-Albrechts-University zu Kiel",
+"University of Arizona" ~ "The University of Arizona",
+"The Pennsylvania State University" ~ "Pennsylvania State University",
+"University of Bologna" ~ "Alma Mater Studiorum - University of Bologna",
+"University of Sydney" ~ "The University of Sydney",
+"University of Copenhagen, Saxo Institute" ~ "University of Copenhagen",
+"The University of Texas at Austin" ~ "University of Texas at Austin",
+
+    .default = as.character(Institution)
+  )) %>% 
+  select(Institution) %>% 
+  left_join(qs_rankings) %>% 
+  drop_na(`2024`) %>% 
+  distinct(Institution, .keep_all = TRUE) %>% 
+  View()
+
+```
+
 # Discussion

### 2024-05-31 analysis/paper/paper.qmd
EEO

@@
   EEO_wordcount %>%
-  group_by(year_ad_posted,
-           word_count) %>%
-  summarise(n = sum(value)) %>%
-  mutate(prop = n / sum(n))
+  group_by(year_ad_posted) %>%
+  summarise(n = mean(word_count))
 
@@
 ggplot(jobdata_EEO_year_tally) +
-  aes(year_ad_posted) +
-  geom_bar() +
-  scale_x_discrete(name = "") +
+  aes(year_ad_posted,n) +
+  geom_col() +
   ylab("Average length of EEO statement by year") +
@@
   guides(x = "none")
+
+ggplot(EEO_wordcount) +
+  aes(year_ad_posted, word_count) +
+  geom_boxplot() +
+  geom_quasirandom()
 

### 2024-09-17 analysis/paper/paper.qmd
diversity over time to compare with EEO
write methods section

--- 2024-05-28 analysis/paper/paper.qmd
@@
 ---
-title: "Title Goes Here"
+title: "Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology"
 author:
-  - Jane Doe:
+  - Ben Marwick:
       correspondence: "yes"
-      email: janedoe@fosg.org
-      orcid: 0000-0003-1689-0557
+      email: bmarwick@uw.edu
+      orcid: 0000-0001-7879-4531
       institute:
-        - fosg
-        - fop
-  - John Q. Doe:
-      institute: fosg
-      orcid: 0000-0003-1689-0558
-  - Peder Ås:
-      institute: fosg
-      orcid: 0000-0003-1689-0559
-  - Juan Pérez:
-      orcid: 0000-0003-1689-0551
+        - uw
+  - Anne Marie Poole:
+      institute: uw
+      orcid: xxx
+  - Ailin Zhang:
+      institute: uw
+      orcid: xxx
+  - Setareh Shafizadeh:
+      orcid: 0000-0003-2231-4310
+      institute: uw
+  - Jess Beck:
+      orcid: 0000-0002-7387-2307
       institute:
-        - name: Acme Corporation
-  - Max Mustermann:
-      orcid: 0000-0003-1689-0552
+        - ucd
 institute:
-  - fosg:
-      name: Formatting Open Science Group
-      address: 23 Science Street, Eureka, Mississippi, USA
-  - fop: Federation of Planets
+  - uw: Department of Anthropology, University of Washington, Seattle, USA
+  - ucd: School of Archaeology, University College Dublin, Ireland
+
 title-block-published: "Last updated"  
@@
   fig-path: "../figures/"
-  fig-dpi: 600
+  fig-dpi: 900
 filters:
@@
 abstract: |
-  Text of abstract
+  Academic careers are frequently sought by archaeology graduate students.
+Job listing websites often serve as the first place for these students when seeking academic positions.
+We examined tenure-track job advertisements over the past decade to gain insights into the academic job market for archaeologists.
+Using data from the community-edited Academic Jobs Wiki for Archaeology, we examine changes in the academic job market over time.
+We studied the text of 449 job ads posted from 2013-2023.
+Our analysis focuses on shifts in archaeological topics and methods requested in job ads.
+We investigate whether the burden on applicants has changed over time: do institutions request more information and documents from applicants at the initial stages of application, compared to a decade ago?
+We also examine whether there is an increasing trend in job advertisements highlighting diversity and inclusivity, thereby encouraging a broader range of applicants.
+Additionally, we assess the influence of socio-political factors on the changing focus of research topics in the field.
+This research aims to assist current and future archaeology students and graduates in better understanding the job market and the requirements of employers, thereby aiding them in effectively preparing for their applications for positions in archaeology.
 keywords: |
@@
 
+<!-- Instructors to authors for AAQ
+
+https://documents.saa.org/container/docs/default-source/doc-publications/style-guide/saa-style-guide_english_updated_2021_final08023c15928949dabd02faafb269fb1c.pdf?sfvrsn=c1f41c1b_11
+
+"AAQ recommends 10,000 words for ARTICLES... Total word count includes
+the title page, abstracts (English and second language), keywords (English and second language), text, acknowledgments, funding statements, statements of data availability and competing interests, lists of online supplemental material, figure and table captions, text of all tables (for Excel tables, copy and paste into a Word document to check the word count), notes, and the references cited section"
+
+Our notes document: https://docs.google.com/document/d/12FNfuAPIdZPBlKTEZK_HEWM8GwR_ZHyS/edit Ailin: https://docs.google.com/document/d/13OzXGFdZFraEb60ZXGwMUcRxCRKZxAQG/edit & https://docs.google.com/document/d/1NUF6dD8EnStlM6OameVvI9v6ehtf9IpWNt5SCYZgHPQ/edit
+
+Our SAA poster: Poole et al. https://osf.io/p5uzg & https://docs.google.com/presentation/d/1bljiJiE_JmywxRCaUGh4NkkcaMnQNh2uIBUcmxlp-yw/edit#slide=id.g26d90de87be_2_0
+
+Our URS23 poster: Zhang et al. https://docs.google.com/presentation/d/1u9OGBJNRilNBXAsysiY3z4Q5_qOQa4fxxzBfVlZnQYk/edit#slide=id.g2d1d7d9d51d_0_0
+
+-->
+
 # Introduction
 
-Here is a citation [@Marwick2017]
-
-# Background
-
 # Methods
 
+Our primary data source is the Archaeology Academic Jobs Wiki.
+Originating in 2007, this is a set of freely accessible web pages that anyone can edit (anonymously or with a free user account) hosted by Fandom, a for-profit company.
+The Archaeology pages are part of the Academic Jobs Wiki, which coordinates similar collaboratively-edited resources for around 40 academic disciplines.
+The coordinators and contributors are nearly all anonymous or pseudonymous.
+Typically contributors copy and paste the text of job ads from other sources, such as the _Chronicle of Higher Education_, _Higher Ed Jobs_, and university websites, into the wiki, collecting ads originally posted in numerous different locations.
+Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position, such as a tally of how many people have applied, the dates of events such as requests for more materials, interviews, offer made, rejection notices, etc.
+Contributors also edit the page to ask and answer questions about the positions and the application process.
+These comments make the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market.
+Because of its reputation for aggregating ads from diverse sources and rapidly-updated information that is not available elsewhere, the Academic Jobs Wiki has a large community of users that keep it updated and accurate has become an authoritative data source for studies of hiring trends in academia [e.g.
+@musial2018five] and a widely recommended resource [e.g.
+@lightfoot2021preparing].
+
+For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023 we read the text and recorded into a Google form the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methods foci on the position.
+The topical focus is what we understood as the primary intellectual focus of the position.
+The geographic focus is the region of the world that the ideal candidate has scholarly expertise on.
+The methods focus is the data-generating sub-field of archaeology that is mentioned in the ad.
+We recorded the type and number of documents requested in each ad (e.g.
+cover letter, CV, statements on research, teaching, diversity, syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad.
+
+After completing primary data collection, we studied the topical, geographic, and methods foci of each ad and collaboratively and manually reduced the variation in the raw data into 10-15 categories to simplify analysis and visualisation.
+Our topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
+Our methods categories were: Archaeobotany, Archaeometry, Bioarchaeology, Ceramic analysis, Computational and Digital archaeology, Geoarchaeology, Landscape analysis, Lithic analysis, Material culture analysis, and Zooarchaeology.
+Ads could have multiple or none of these three foci, and some of the foci overlap.
+Some topics include geographic regions because this is how they are typically understood by archaeologists.
+For example Mesoamerican archaeology is understood to refer to a specific time period and geographic region.
+While these overlaps can make the data challenging to interpret, in our view it reflects the complex realities of  how search committees express their needs in searching for new faculty, and is insightful in how it reveals intersections between different foci.
+
 # Results
 
-<!-- Here's some example analysis code: -->
+- basic trends
+- where are the hiring universities?
+- what Carnegie classifications of the hiring universities?
+
+## Geographic trends over time
+
+- what regions of the world are mentioned in the ads?
+
+## Topic trends over time
+
+## Method trends over time
+
+## Instructions to applicants over time
 
@@
 EEO <- jobdata %>%
-  select(equal_employment_opportunity_statement, year_ad_posted) %>% 
+  select(equal_employment_opportunity_statement) %>% 
   drop_na
@@
 
+EEO_wordcount <- EEO %>%
+  mutate(word_count = str_count(equal_employment_opportunity_statement, "\\S+")) %>%
+  print()
+
+ggplot(EEO_wordcount) +
+  aes(word_count) +
+  geom_histogram()
+
 EEO_clean <- EEO %>%
@@
 
-#word count for EEO statement
-EEO_wordcount <- EEO %>%
-  mutate(word_count = str_count(equal_employment_opportunity_statement, "\\S+")) %>%
-  print()
-
-jobdata_EEO_year_tally <-
-  EEO_wordcount %>%
-  group_by(year_ad_posted,
-           word_count) %>%
-  summarise(n = sum(value)) %>%
-  mutate(prop = n / sum(n))
-
-ggplot(EEO_wordcount) +
-  aes(word_count, year_ad_posted) +
-  geom_histogram()
+# this doesn't work, need help to troubleshoot
+EEO_word <- EEO_clean %>%
+   unnest_tokens(word, 
+                 equal_employment_opportunity_statement) %>%
+  anti_join(get_stopwords()) %>%
+  count(word, sort = TRUE)
 
 # here's the list of words of interest
-community_word <- c("women","color","religion","race","age","veteran","disability","citizenship","national origin","sex","gender","marital status","genetic information","breastfeeding","sexual orientation","transgender status","gender expression","gender identity","ancestry","income assignment for child support","arrest and court record","domestic or sexual violence victim status","national guard absence","visible minorities","aboriginal persons","indigenous","LGBTQ2S+","pregnancy","childbirth","faith","genetic or family medical history","medical condition","nationality","ethnicity","parental status","accent","ethnic","creed","political affiliation","immigration status")
+community_word <- c("women","color","religion","race","age","veteran","disability")
 
@@
 
-ggplot(EEO_clean_community_word) +
-  aes(community_word_tally) +
-  geom_histogram()
+```
+
+```{r}
+
+EEO_word %>% 
+  filter(n >= 20) %>% 
+ggplot() +
+  aes(x = reorder(word, n), y = n) +
+  geom_col() +
+  coord_flip() +  # Flips the axes for easier reading of text
+  labs(x = "Words", y = "Frequency", title = "Word Frequency") 
 ```

### 2024-09-17 analysis/paper/paper.qmd

@@
 
-# this doesn't work, need help to troubleshoot
+#word count for EEO statement
+EEO_wordcount <- EEO %>%
+  mutate(word_count = str_count(equal_employment_opportunity_statement, "\\S+")) %>%
+  print()
+
+jobdata_EEO_year_tally <-
+  EEO_wordcount %>%
+  group_by(year_ad_posted) %>%
+  summarise(n = mean(word_count))
+
+fig_ave_length_EEO_per_year <- 
+ggplot(jobdata_EEO_year_tally) +
+  aes(year_ad_posted,n) +
+  geom_col() +
+  ylab("Average length of EEO statement by year") +
+  theme_minimal(base_size = 28) +
+  guides(x = "none")
+
+ggplot(EEO_wordcount) +
+  aes(year_ad_posted, word_count) +
+  geom_boxplot() +
+  geom_quasirandom()
+
+ggplot(jobdata_requirements_means %>% 
+         filter(name == "diversity\nstatement") ) +
+  aes(year_ad_posted, 
+      mean_n) +
+  geom_col()
+
+# here's the list of words of interest
+community_word <- c("women","color","religion","race","age","veteran","disability","citizenship","national origin","sex","gender","marital status","genetic information","breastfeeding","sexual orientation","transgender status","gender expression","gender identity","ancestry","income assignment for child support","arrest and court record","domestic or sexual violence victim status","national guard absence","visible minorities","aboriginal persons","indigenous","LGBTQ2S+","pregnancy","childbirth","faith","genetic or family medical history","medical condition","nationality","ethnicity","parental status","accent","ethnic","creed","political affiliation","immigration status")
+
 EEO_word <- EEO_clean %>%
@@
   anti_join(get_stopwords()) %>%
-  count(word, sort = TRUE)
-
-# here's the list of words of interest
-community_word <- c("women","color","religion","race","age","veteran","disability")
+  count(word, sort = TRUE) %>%
+  drop_na()
 
@@
   mutate(community_word_tally = sum(c_across(all_of(community_word)), 
-                                    na.rm = T))
-
-```
-
-```{r}
-
-EEO_word %>% 
-  filter(n >= 20) %>% 
-ggplot() +
-  aes(x = reorder(word, n), y = n) +
+                                    na.rm = T)) %>% 
+  bind_cols(EEO_wordcount) %>% 
+  mutate(eeo_words_total_ratio = community_word_tally / word_count)
+
+ggplot(EEO_clean_community_word) +
+  aes(community_word_tally,
+      word_count,
+      size = eeo_words_total_ratio) +
+  geom_point()
+
+ggplot(EEO_clean_community_word) +
+  aes(community_word_tally) +
+  geom_histogram()
+```
+
+```{r}
+ggplot(EEO_word %>% 
+         filter(n > 20), 
+       aes(x = reorder(word, n), 
+           y = n)) +
   geom_col() +
-  coord_flip() +  # Flips the axes for easier reading of text
-  labs(x = "Words", y = "Frequency", title = "Word Frequency") 
+  labs(x = "Words", 
+       y = "Frequency", 
+       title = "Word Frequency") +
+  coord_flip()
+```
+
+```{r}
+# explore QS rankings
+
+qs_rankings <- 
+  readxl::read_excel("analysis/data/raw_data/2024 QS WUR by Subject - Public Results (for qs.com)_48.xlsx", sheet = "Archaeology", skip = 12)
+
+jobdata %>% 
+  mutate(Institution = str_remove(name_of_hiring_university,
+                                  " \\(.*\\)")) %>% 
+  mutate(Institution = case_when(
+    # modify the school name to match with the name in the QS list
+    Institution == "Ohio State University-Main Campus" ~ "The Ohio State University",
+  "Cambridge" ~ "University of Cambridge",
+  "Cambridge University" ~ "University of Cambridge",
+  "Keble College" ~ "University of Oxford",
+  "University of California-Berkeley" ~ "University of California, Berkeley (UCB)",
+"University of California-Los Angeles" ~ "University of California, Los Angeles (UCLA)",
+"University of Tuebingen" ~ "Eberhard Karls Universität Tübingen",
+"Universität Tübingen" ~ "Eberhard Karls Universität Tübingen",
+"Arizona State University Campus Immersion" ~ "Arizona State University",
+"Kiel University" ~ "Christian-Albrechts-University zu Kiel",
+"University of Arizona" ~ "The University of Arizona",
+"The Pennsylvania State University" ~ "Pennsylvania State University",
+"University of Bologna" ~ "Alma Mater Studiorum - University of Bologna",
+"University of Sydney" ~ "The University of Sydney",
+"University of Copenhagen, Saxo Institute" ~ "University of Copenhagen",
+"The University of Texas at Austin" ~ "University of Texas at Austin",
+
+    .default = as.character(Institution)
+  )) %>% 
+  select(Institution) %>% 
+  left_join(qs_rankings) %>% 
+  drop_na(`2024`) %>% 
+  distinct(Institution, .keep_all = TRUE) %>% 
+  View()
+
 ```

### 2024-09-18 analysis/paper/paper.qmd
article qmd for the journal article

--- 2024-05-31 analysis/paper/paper.qmd
@@
 ---
-title: "Title Goes Here"
+title: "Careers in Ruins: Academic Archaeology Job Trends From 2013 - 2023"
 author:
-  - Jane Doe:
+  - Ben Marwick:
       correspondence: "yes"
-      email: janedoe@fosg.org
-      orcid: 0000-0003-1689-0557
+      email: bmarwick@uw.edu
+      orcid: 0000-0001-7879-4531
       institute:
-        - fosg
-        - fop
-  - John Q. Doe:
-      institute: fosg
-      orcid: 0000-0003-1689-0558
-  - Peder Ås:
-      institute: fosg
-      orcid: 0000-0003-1689-0559
-  - Juan Pérez:
-      orcid: 0000-0003-1689-0551
+        - uw
+  - Anne Marie Poole:
+      institute: uw
+      orcid: xxx
+  - Ailin Zhang:
+      institute: uw
+      orcid: xxx
+  - Setareh Shafizadeh:
+      orcid: 0000-0003-2231-4310
+      institute: uw
+  - Jess Beck:
+      orcid: 0000-0002-7387-2307
       institute:
-        - name: Acme Corporation
-  - Max Mustermann:
-      orcid: 0000-0003-1689-0552
+        - ucd
 institute:
-  - fosg:
-      name: Formatting Open Science Group
-      address: 23 Science Street, Eureka, Mississippi, USA
-  - fop: Federation of Planets
+  - uw: Department of Anthropology, University of Washington, Seattle, USA
+  - ucd: School of Archaeology, University College Dublin, Ireland
+
 title-block-published: "Last updated"  
@@
 abstract: |
-  Text of abstract
+
 keywords: |
@@
 
-# Introduction
-
-Here is a citation [@Marwick2017]
-
-# Background
-
-# Methods
-
-# Results
-
-<!-- Here's some example analysis code: -->
+This document contains the code used to generate the figures for our SAA poster.
 
@@
 
-ggplot(EEO_wordcount) +
-  aes(word_count, year_ad_posted) +
-  geom_histogram()
+ggplot(jobdata_requirements_means %>% 
+         filter(name == "diversity\nstatement") ) +
+  aes(year_ad_posted, 
+      mean_n) +
+  geom_col()
 

### 2024-10-07 analysis/paper/article.qmd
article qmd for the journal article
starting the results

--- 2024-09-17 analysis/paper/paper.qmd
@@
 execute:
-  echo: true
+  echo: false
   warning: false
@@
 
+The aim of this paper is to explore the demand-side of the academic job market for archaeologists in the United States.
+We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.
+
+# Background
+
 # Methods
@@
 
-- basic trends
-- where are the hiring universities?
-- what Carnegie classifications of the hiring universities?
-
-## Geographic trends over time
-
-- what regions of the world are mentioned in the ads?
-
-## Topic trends over time
-
-## Method trends over time
-
-## Instructions to applicants over time
-
 ```{r}
@@
 total_number_of_ads_in_our_sample <- nrow(jobdata) # 550
-```
-
-We have `r total_number_of_ads_in_our_sample` job advertisements in our sample
+
+# for text on our figures
+base_size <- 6
+```
 
@@
 #| label: fig-how-many-jobs-per-year
-#| fig-cap: "Number of tenure-track archaeology faculty job ads posted each year"
 
@@
 
-fig_how_many_jobs_per_year <- 
-ggplot(jobdata) +
-  aes(year_ad_posted) +
-  geom_bar() +
-  scale_x_discrete(name = "") +
-  ylab("Number of\ntenure track job ads") +
-  theme_minimal(base_size = 28) +
-  guides(x = "none")
-
-```
-
-@fig-how-many-jobs-per-year shows how many jobs per year in our sample
+```
 
@@
 #| label: fig-how-many-jobs-per-year-by-rank
-#| fig-cap: "Proportion of tenure-track archaeology faculty jobs per year by rank"
 
@@
   aes(year_ad_posted,
-      prop,
+      n,
       group = job_title_simple,
-      colour = job_title_simple) +
-  geom_line(linewidth = 2) +
-  theme_minimal(base_size = 28) +
+      fill = job_title_simple) +
+ # geom_line(linewidth = 2) +
+  geom_col() +
+ theme_minimal(base_size = base_size) +
   xlab("") +
-  ylab("Proportion of\nall tenure-track ads") +
-  theme(legend.position = c(0.5, 0.5)) +
-  scale_colour_brewer(palette = "Dark2") +
-  guides(colour = guide_legend(nrow=2,
+  ylab("Number of\ntenure track job ads") +
+  scale_fill_brewer(palette = "Dark2") +
+  guides(fill = guide_legend(nrow=2,
                                byrow=TRUE,
-                               "Job title")) +
+                               "Job title",
+                                keywidth=0.1,
+                 keyheight=0.1,
+                 default.unit="inch")) +
+    theme(legend.position = c(0.5, 0.85)) +
   guides(x = "none")
@@
 #| label: ratio-tt-to-non-tt
-
-# ratio of tenure-track to untenured positions
-# base URL changes after 2018_2019
 
@@
     str_replace("2021-2022", "2021-22")
+```
+
+```{r}
+#| eval: false
 
 library(rvest)
+
+# scrape the web pages to get the 
+# ratio of tenure-track to untenured positions
+# base URL changes after 2018_2019
+# We did this once and then saved the output so we don't scrape every
+# time the quarto document is rendered. So this code block is
+# eval: false to avoid repeated and unnesecary scraping
 
@@
       html_text())
+
+# save this list so we don't have to scrape again
+library(rlist)
+list.save(urls_for_each_year_headers, 
+          here('analysis/data/raw_data/urls_for_each_year_headers.yaml'))
+```
+
+```{r}
+#| label: ratio-tt-to-non-tt-jobs
+
+library(rlist)
+urls_for_each_year_headers <-  list.load( here('analysis/data/raw_data/urls_for_each_year_headers.yaml'))
 
@@
   mutate(year = str_extract(url_to_data_source_e_g_paste_in_url_to_the_jobs_wiki_page,  "[[0-9]]{4}-[[0-9]]{4}|2021-22")) %>%  
-  mutate(year = ifelse(year =="2021-22",  "2021-2022", year))
+  mutate(year = ifelse(year =="2021-22",  "2021-2022", year))  %>% 
+  # add line break for prettier plotting
+  mutate(year_ad_posted_break = str_replace(year, "-", "-\n"))
 
@@
   ggplot(count_of_tt_jobs_per_year_from_our_form) +
-  aes(year, 
+  aes(year_ad_posted_break, 
       group = 1,
       ratio_tt_2_ntt) +
-  geom_line(linewidth = 2) +
+  #geom_line(linewidth = 2) +
+  geom_col() +
   geom_hline(yintercept = 1,
@@
   annotate("text", 
-           x = 3, 
-           y = 1.3, 
+           x = 6, 
+           y = 1.4, 
            label = "1:1 ratio",
-           colour = "red") +
+           colour = "red",
+           size = 3) +
   labs(y = "Ratio of tenure-track\nto non-tenure track and other",
        x = "") +
-  theme_minimal(base_size = 28) +
+  theme_minimal(base_size = base_size) +
   scale_x_discrete(name = "Academic Year") 
@@
 ```{r}
+#| label: combine-three-basic-plots
+
 # save these three plots as one set
@@
 plot_grid(
-  fig_how_many_jobs_per_year,
   fig_prop_by_job_title_per_year,
@@
   axis = "lr",
-  labels = "AUTO"
+  labels = "AUTO",
+  label_size = 8
 )
@@
 
+```
+
+```{r}
+#| label: show-basic-plots
+#| fig-cap: "A: total number of job ads posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank.
+B: Ratio of tenure-track to non-tenure-track positions over time."
+
+# show figure from PNG file
+knitr::include_graphics("../figures/fig-panel-per-year.png")
+```
+
+We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023.
+Figure @show-basic-plots shows the count of ads for each year.
+Assistant Professor jobs are consistently the most common title and rank, and open rank or full professor the least frequent.
+The ratio of tenure-track to non-tenure track positions is generally well above one, with
+
+- basic trends
+- where are the hiring universities?
+- what Carnegie classifications of the hiring universities?
+
+## Geographic trends over time
+
+- what regions of the world are mentioned in the ads?
+
+## Topic trends over time
+
+## Method trends over time
+
+## Instructions to applicants over time
+
+```{r}
+# stop rendering at this point because the code below is not working
+knitr::knit_exit()
 ```

### 2024-10-10 analysis/paper/article.qmd
basic characteristics of institutions

@@
 # for text on our figures
-base_size <- 6
+base_size <- 12
 ```
@@
 ```{r}
-#| label: combine-three-basic-plots
+#| label: combine-basic-plots
 
@@
 library(cowplot)
+p <- 
 plot_grid(
@@
 
-ggsave(here("analysis",
+ggsave(filename = here("analysis",
             "figures", 
@@
        bg ="white",
+      plot = p,
        h = 15, # experiment with h and w to get the right size and proportion 
@@
 ```{r}
-#| label: show-basic-plots
+#| label: fig-show-basic-plots
 #| fig-cap: "A: total number of job ads posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank.
@@
 # show figure from PNG file
-knitr::include_graphics("../figures/fig-panel-per-year.png")
+knitr::include_graphics(here("analysis",
+            "figures", 
+            "fig-panel-per-year.png"))
 ```
@@
 We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023.
-Figure @show-basic-plots shows the count of ads for each year.
-Assistant Professor jobs are consistently the most common title and rank, and open rank or full professor the least frequent.
-The ratio of tenure-track to non-tenure track positions is generally well above one, with
-
-- basic trends
-- where are the hiring universities?
-- what Carnegie classifications of the hiring universities?
-
-## Geographic trends over time
-
-- what regions of the world are mentioned in the ads?
-
-## Topic trends over time
-
-## Method trends over time
-
-## Instructions to applicants over time
-
-```{r}
-# stop rendering at this point because the code below is not working
-knitr::knit_exit()
-```
-
-```{r}
-#| label: fig-requirements-over-time
-
-# look at only these requirements because the others are flat
-
-intresting_requirements <- 
-c("cover letter",
-  "cv",
-  "names of recommenders",
-  "diversity statement",
-  "research statement",
-  "teaching statement")
-
-jobdata_requirements <- 
-jobdata %>% 
-  select(year_ad_posted,
-         starts_with("documents_requested")) %>% 
-  pivot_longer(-year_ad_posted) %>% 
-  mutate(value = case_when(
-    value == "Not requested in the job ad" ~ 0,
-    value == "One" ~ 1,
-    value == "Two (e.g. two syllabi)" ~ 2,
-    value == "Three" ~ 3,
-    .default = 0
-  ))  %>% 
-  # trim names a bit
-  mutate(name = str_remove(name, "documents_requested_")) %>% 
-  mutate(name = str_replace_all(name, "_", " ")) %>% 
-  filter(name %in% intresting_requirements) %>% 
-  mutate(name = str_wrap(name, 10),
-         year_ad_posted = str_replace(year_ad_posted, "-", "\n"))
-
-jobdata_requirements_means <- 
-jobdata_requirements %>% # average number requested per year
-  group_by(year_ad_posted, 
-           name) %>% 
-  summarise(mean_n = mean(value))
-
-integer_breaks <- function(n = 5, ...) {
-  fxn <- function(x) {
-    breaks <- floor(pretty(x, n, ...))
-    names(breaks) <- attr(breaks, "labels")
-    breaks
-  }
-  return(fxn)
-}
-
-ggplot(jobdata_requirements_means) +
-  aes(year_ad_posted, 
-      mean_n,
-      group = name) +
-  geom_smooth(linewidth = 2,
-              colour = "black") +
-  geom_jitter(data = jobdata_requirements,
-             aes(year_ad_posted, 
-                 value),
-             alpha = 0.1,
-             height = 0.2,
-             width =  0.1) +
-  facet_wrap(~name,
-             scales = "free_y",
-             nrow = 1) +
-  xlab("Year") +
-  ylab("Number requested in job ad") +
-  scale_y_continuous(breaks = integer_breaks()) +
-  theme_minimal(base_size = 14) +
-  theme(axis.text.x = element_text(size = 8),
-        strip.text = element_text( size = 20))
-
-ggsave(here("analysis",
-            "figures", 
-            "fig-requirements-per-year.png"),
-       bg ="white",
-       h = 10, # experiment with h and w to get the right size and proportion 
-       w = 20,
-       units = "in",
-       dpi = 900) # make the image nice and crisp
-
-```
-
-```{r}
-#| label: fig-requirements-by-position
-
-# do the requirements differ for associate positions 
-jobdata_requirements_by_rank <- 
-jobdata %>% 
-  mutate(position_title = case_when(
-    str_detect(title_of_position_tenure_track_jobs_only, 
-               "associate") ~ "associate",
-    str_detect(title_of_position_tenure_track_jobs_only, 
-               "assistant") ~ "assistant",
-    str_detect(title_of_position_tenure_track_jobs_only, 
-               "full") ~ "full"))  %>% 
-  select(position_title,
-         starts_with("documents_requested")) %>% 
-  pivot_longer(-position_title) %>% 
-  mutate(value = case_when(
-    value == "Not requested in the job ad" ~ 0,
-    value == "One" ~ 1,
-    value == "Two (e.g. two syllabi)" ~ 2,
-    value == "Three" ~ 3,
-    .default = 0
-  )) %>% 
-  filter(!is.na(position_title))
-
-jobdata_requirements_by_rank_means <- 
-  jobdata_requirements_by_rank %>% 
-  group_by(position_title,
-           name) %>% 
-  summarise(mean = mean(value))
-
-ggplot() +
-    geom_jitter(data = jobdata_requirements_by_rank,
-                  aes(position_title, 
-                      value),
-                height = 0.05,
-                alpha = 0.1) +
-    geom_point(data = jobdata_requirements_by_rank_means,
-               aes(position_title,
-                   mean),
-               size = 4,
-               colour = "red") +
-  facet_wrap( ~ name,
-              scales = "free_y") +
-  theme_minimal()
-```
-
-```{r}
-#| label: fig-geographic-focus-by-year
-
-# geographic focus by year
-
-library(googlesheets4)
-library(stringi)
-
-geographic_foci <-
-read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0", sheet = "geography")
-
-geographic_foci_clean <-
-  map(
-   str_split(geographic_foci$`From the data`, ";"),
-   ~.x %>%
-     str_squish() %>%
-     stri_remove_empty())
-
-jobdata_geo <-
-  jobdata %>%
-  select(geographic_focus_of_position)
-
-jobdata_geo <-
-  # add one column for each geo region in our categories
-cbind(jobdata_geo,
-      setNames( lapply(geographic_foci$Category2, function(x) x=NA),
-                geographic_foci$Category2) )
-
-for(i in 1:length(geographic_foci$Category2)){
-
-  this_location <- geographic_foci$Category2[i]
-
-  # create the pattern to search for
-  x <- paste0(geographic_foci_clean[[i]], collapse = "|")
-
-  # do the search through all the job ads for that pattern
-  y <- str_detect(jobdata_geo$geographic_focus_of_position,
-             x)
-
-  # assign back to our data frame in the appropriate location column
-  jobdata_geo[, this_location] <- y
-
-}
-
-# BM TODO: check for job ads that have a location, but we're not getting it
-
-united_states_regions <-
-str_subset(geographic_foci$Category2, "US")
-
-jobdata_geo_year <-
-jobdata %>%
-  bind_cols(jobdata_geo) %>%
-  select(year_ad_posted,
-         geographic_foci$Category2) %>%
-  pivot_longer(-year_ad_posted) %>%
-  drop_na()
-
-# how many times each location mentioned?
-jobdata_geo_year %>%
-  group_by(name) %>%
-  summarise(n = sum(value)) %>%
-  arrange(desc(n)) %>%
-  ggplot() +
-  aes(reorder(name, n),
-      n)+
-  geom_col() +
-  xlab("") +
-  theme_minimal() +
-  coord_flip()
-
-# explore trends over time. put a point on the max year
-jobdata_geo_year_tally <-
-jobdata_geo_year %>%
- # exclude those with <20 ads
-  filter(!name %in% c("Canada & Arctic",
-                      "Oceania",
-                      "Southeast US",
-                      "Southwest US",
-                      "Western US",
-                      "Midwest US",
-                      "Northeastern US"
-                      )) %>%
-  group_by(year_ad_posted,
-           name) %>%
-  summarise(n = sum(value)) %>%
-  mutate(prop = n / sum(n))
-
-jobdata_geo_year_tally_max <-
-  jobdata_geo_year_tally %>%
-  group_by(
-           name ) %>%
-  filter(prop == max(prop))
-
-library(ggrepel)
-
-ggplot() +
-  geom_smooth(data =  jobdata_geo_year_tally,
-            aes(year_ad_posted,
-                 prop,
-                 group = name,
-                 colour = name),
-            size = 3,
-           se = FALSE 
-           ) +
-  xlab("Year") +
-  ylab("Proportion of all ads") +
-  scale_colour_brewer(palette = "Dark2") +
-  guides(colour = guide_legend("Geographic\nfocus",
-                               label.position = "bottom")) +
-    theme_minimal( base_size = 28) +
-  theme(legend.position="bottom")
-
-ggsave(here("analysis",
-            "figures", 
-            "fig-geo-focus-by-year.png"),
-       bg ="white",
-       h = 10, # experiment with h and w to get the right size and proportion 
-       w = 20,
-       units = "in",
-       dpi = 900) # make the image nice and crisp)
-
-# what about within the US
-# how many times each location mentioned?
-jobdata_geo_year %>%
-  group_by(name) %>%
-  summarise(n = sum(value)) %>%
-  arrange(desc(n)) %>%
-  filter(name %in% united_states_regions) %>% 
-  ggplot() +
-  aes(reorder(name, n),
-      n)+
-  geom_col() +
-  ylab("Number of ads") +
-  xlab("") +
-  theme_minimal(base_size = 24) +
-  coord_flip()
-
-ggsave(here("analysis",
-            "figures", 
-            "fig-geo-us-focus-by-year.png"),
-       bg ="white",
-       h = 10, # experiment with h and w to get the right size and proportion 
-       w = 20,
-       units = "in",
-       dpi = 900) # make the image nice and crisp)
-
-```
-
-```{r}
-
-# topical focus by year
-
-library(googlesheets4)
-library(stringi)
-
-topical_foci <-
-  read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0",
-             sheet = "topic")
-
-topical_foci_clean <-
-  map(
-    str_split(topical_foci$`From the data`, ";"),
-    ~.x %>%
-      str_squish() %>%
-      stri_remove_empty() %>%
-      str_to_lower)
-
-jobdata_topic <-
-  jobdata %>%
-  select(topical_focus_of_position) %>%
-  mutate(topical_focus_of_position = str_to_lower(topical_focus_of_position))
-
-jobdata_topic <-
-  # add one column for each topic in our categories
-  cbind(jobdata_topic,
-        setNames( lapply(topical_foci$Category, function(x) x=NA),
-                  topical_foci$Category) )
-
-for(i in 1:length(topical_foci$Category)){
-
-  this_topic <- topical_foci$Category[i]
-
-  # create the pattern to search for
-  x <- paste0(topical_foci_clean[[i]], collapse = "|")
-
-  # do the search through all the job ads for that pattern
-  y <- str_detect(jobdata_topic$topical_focus_of_position,
-                  x)
-
-  # assign back to our data frame in the appropriate location column
-  jobdata_topic[, this_topic] <- y
-
-}
-
-jobdata_topic_year <-
-  jobdata %>%
-  bind_cols(jobdata_topic) %>%
-  select(year_ad_posted,
-         topical_foci$Category) %>%
-  pivot_longer(-year_ad_posted) %>%
-  drop_na()
-
-# how many times each topic mentioned?
-jobdata_topic_year %>%
-  group_by(name) %>%
-  summarise(n = sum(value)) %>%
-  arrange(desc(n)) %>%
-  ggplot() +
-  aes(reorder(name, n),
-      n)+
-  geom_col() +
-  xlab("") +
-  theme_minimal() +
-  coord_flip()
-
-# explore trends over time. put a point on the max year
-jobdata_topic_year_tally <-
-  jobdata_topic_year %>%
-  # exclude those with <20 ads
-  filter(!name %in% c("Digital Archaeology",
-                      "Pleistocene archaeology",
-                      "Mesoamerican Archaeology",
-                      "Biological anthropology",
-                      "Archaeological theory",
-                      "Evolutionary anthropology",
-                      "North American archaeology"
-  )) %>%
-  group_by(year_ad_posted,
-           name) %>%
-  summarise(n = sum(value)) %>%
-  mutate(prop = n / sum(n))
-
-jobdata_topic_year_tally_max <-
-  jobdata_topic_year_tally %>%
-  group_by(
-    name ) %>%
-  filter(prop == max(prop))
-
-library(ggrepel)
-
-ggplot() +
-  geom_smooth(data =  jobdata_topic_year_tally,
-            aes(year_ad_posted,
-                prop,
-                group = name,
-                colour = name),
-            size = 3,
-           span = 0.7,
-            se = FALSE) +
-  xlab("Year") +
-  ylab("Proportion of all ads") +
-  scale_colour_brewer(palette = "Dark2") +
-  guides(colour = guide_legend("Topic\nfocus",
-                               label.position = "bottom")) +
-    theme_minimal( base_size = 28) +
-  theme(legend.position="bottom")
-
-ggsave(here("analysis",
-            "figures", 
-            "fig-topic-focus-by-year.png"),
-       bg ="white",
-       h = 10, # experiment with h and w to get the right size and proportion
-       w = 20,
-       units = "in",
-       dpi = 900) # make the image nice and crisp))
-
-```
+@fig-show-basic-plots shows the count of ads for each year.
+Assistant Professor jobs are consistently the most common title and rank, and open rank or full professor are the least frequent.
+The ratio of tenure-track to non-tenure track positions is generally well above one.
+Only 2013-2014 had more non-tenure track positions than tenure track, which was followed by an upward trend peaking at 2018-2019 and the declining again into the present.
+
+## Characteristics of the hiring institutions
 
@@
 
+p_map <- 
 ggplot(data= uni_state_country_tally_us,
@@
             force = 0,
-            force_pull = 100)+
+            force_pull = 100,
+            size = 3)+
   scale_x_continuous(breaks = NULL) +
@@
 
-ggsave(here("analysis",
+ggsave(plot = p_map, 
+       filename = here("analysis",
             "figures",
@@
        w = 12,
+       units = "in",
+       dpi = 900) # make the image nice and crisp))
+```
+
+```{r}
+# Carnegie Classification
+library(stringi)
+library(readxl)
+
+# data from https://carnegieclassifications.acenet.edu/resource/2021-update-public-file/
+CC <-
+  read_excel(here::here('analysis/data/raw_data/CCIHE2021-PublicData.xlsx'),
+              sheet = "Values1")
+Basic2021 <-
+  read_excel(here::here('analysis/data/raw_data/CCIHE2021-PublicData.xlsx'),
+             sheet = "Data1") 
+  # remove a bunch of text from uni names to improve our joins
+
+uni_name <- # 547
+ jobdata %>%  # 547 rows
+  # remove parentheses and their contents
+  mutate(name = str_squish(str_replace(name_of_hiring_university,
+                            "\\s*\\(.*?\\)$", 
+                            ""))) %>% 
+  mutate(name = str_replace(name,
+                            "California State University,[[:space:]]",
+                            "California State University-")) %>%
+
+  left_join(Basic2021, 
+            keep = TRUE )
+
+# how many schools did we match with the CC data
+uni_name %>% 
+  filter(!is.na(basic2021)) %>% 
+  nrow() # 245, 290, 300, 313, 404, 430 / 433 for US unis
+
+# which schools don't match with the CC data?
+no_match <- 
+uni_name %>% 
+  select(name.x, name.y)  %>% 
+  filter(is.na(name.y)) %>% 
+  distinct()
+
+uni_name_tally <- 
+uni_name %>% 
+  left_join(CC,
+            join_by("basic2021" == "Value")) %>% 
+  group_by(Category) %>% 
+  tally(sort = TRUE) %>% 
+  drop_na()
+
+p_c <- 
+uni_name_tally %>% 
+  mutate(Category = str_wrap(Category, width = 30)) %>% 
+ggplot() +
+  aes(reorder(Category, n), n) +
+  geom_col() +
+  coord_flip() +
+  xlab("") +
+  theme_minimal(base_size = 14)
+
+library(cowplot)
+p_map_c <- 
+ggdraw(p_c) +
+  draw_plot(p_map, .4, .1, .65, .65) +
+  draw_plot_label(
+    c("A", "B"),
+    c(0, 0.45),
+    c(1, 0.65),
+    size = 12
+  )
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-map-and-carnegie-classification.png"),
+       plot = p_map_c,
+       bg ="white",
+       h = 7, # experiment with h and w to get the right size and proportion
+       w = 12,
+       units = "in",
+       dpi = 900) # make the image nice and crisp))
+
+```
+
+```{r}
+#| label: fig-show-map-of-hiring-institution-and-cc
+#| fig-cap: "A: Frequency of hiring institution by Carnegie classification.
+B: Inset shows map of the United States showing the count of tenure-track job ads posted by all insititutions in each state during 2013-2023"
+
+# show figure from PNG file
+knitr::include_graphics(here("analysis",
+            "figures",
+            "fig-map-and-carnegie-classification.png"))
+```
+
+Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of institutions according to their Carnegie Classification, which is a framework for classifying US colleges and universities according to the types of degrees awarded, levels of activity such as research, and topical foci.
+Doctoral universities with high and very high research activity are by far the most active with hiring archaeology faculty.
+Associate's colleges, also known as community colleges, rarely post job ads for archaeology faculty.
+
+Panel B of @fig-show-map-of-hiring-institution-and-cc shows the geographic distribution of the hiring institutions.
+California posted almost twice as many job ads as the next most active states.
+After California, the states that posted the most ads during 2013-2023 include New York, Texas, and Pennsylvania, and Florida.
+These top five states also correspond to the top five most populous US states, indicating that rates of hiring is approximately proportional to population density.
+Similarly, the lowest counts of job ads were observed in states with the lowest populations: North Dakota, South Dakota, Alaska, and Nebraska.
+No institutions in Montana posted a job ad during this period.
+
+## Geographic trends over time in job ads
+
+```{r}
+#| label: fig-geographic-focus-by-year
+
+# geographic focus by year
+
+library(stringi)
+
+geographic_foci <-
+  read_csv(here("analysis/data/raw_data/Tenure Track Job Advertisements in Archaeology - geography.csv"))
+
+geographic_foci_clean <-
+  map(
+   str_split(geographic_foci$`From the data`, ";"),
+   ~.x %>%
+     str_squish() %>%
+     stri_remove_empty())
+
+jobdata_geo <-
+  jobdata %>%
+  select(geographic_focus_of_position)
+
+jobdata_geo <-
+  # add one column for each geo region in our categories
+cbind(jobdata_geo,
+      setNames( lapply(geographic_foci$Category2, function(x) x=NA),
+                geographic_foci$Category2) )
+
+for(i in 1:length(geographic_foci$Category2)){
+
+  this_location <- geographic_foci$Category2[i]
+
+  # create the pattern to search for
+  x <- paste0(geographic_foci_clean[[i]], collapse = "|")
+
+  # do the search through all the job ads for that pattern
+  y <- str_detect(jobdata_geo$geographic_focus_of_position,
+             x)
+
+  # assign back to our data frame in the appropriate location column
+  jobdata_geo[, this_location] <- y
+
+}
+
+# BM TODO: check for job ads that have a location, but we're not getting it
+
+united_states_regions <-
+str_subset(geographic_foci$Category2, "US")
+
+jobdata_geo_year <-
+jobdata %>%
+  bind_cols(jobdata_geo) %>%
+  select(year_ad_posted,
+         geographic_foci$Category2) %>%
+  pivot_longer(-year_ad_posted) %>%
+  drop_na()
+
+# how many times each location mentioned?
+jobdata_geo_year %>%
+  group_by(name) %>%
+  summarise(n = sum(value)) %>%
+  arrange(desc(n)) %>%
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  xlab("") +
+  theme_minimal() +
+  coord_flip()
+
+# explore trends over time. put a point on the max year
+jobdata_geo_year_tally <-
+jobdata_geo_year %>%
+ # exclude those with <20 ads
+  filter(!name %in% c("Canada & Arctic",
+                      "Oceania",
+                      "Southeast US",
+                      "Southwest US",
+                      "Western US",
+                      "Midwest US",
+                      "Northeastern US"
+                      )) %>%
+  group_by(year_ad_posted,
+           name) %>%
+  summarise(n = sum(value)) %>%
+  mutate(prop = n / sum(n))
+
+jobdata_geo_year_tally_max <-
+  jobdata_geo_year_tally %>%
+  group_by(
+           name ) %>%
+  filter(prop == max(prop))
+
+library(ggrepel)
+
+ggplot() +
+  geom_smooth(data =  jobdata_geo_year_tally,
+            aes(year_ad_posted,
+                 prop,
+                 group = name,
+                 colour = name),
+            size = 3,
+           se = FALSE 
+           ) +
+  xlab("Year") +
+  ylab("Proportion of all ads") +
+  scale_colour_brewer(palette = "Dark2") +
+  guides(colour = guide_legend("Geographic\nfocus",
+                               label.position = "bottom")) +
+    theme_minimal( base_size = base_size) +
+  theme(legend.position="bottom")
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-geo-focus-by-year.png"),
+       bg ="white",
+       h = 10, # experiment with h and w to get the right size and proportion 
+       w = 20,
+       units = "in",
+       dpi = 900) # make the image nice and crisp)
+
+# what about within the US
+# how many times each location mentioned?
+jobdata_geo_year %>%
+  group_by(name) %>%
+  summarise(n = sum(value)) %>%
+  arrange(desc(n)) %>%
+  filter(name %in% united_states_regions) %>% 
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  ylab("Number of ads") +
+  xlab("") +
+  theme_minimal(base_size = 24) +
+  coord_flip()
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-geo-us-focus-by-year.png"),
+       bg ="white",
+       h = 10, # experiment with h and w to get the right size and proportion 
+       w = 20,
+       units = "in",
+       dpi = 900) # make the image nice and crisp)
+
+```
+
+- what regions of the world are mentioned in the ads?
+
+- basic trends
+- where are the hiring universities?
+- what Carnegie classifications of the hiring universities?
+
+## Topic trends over time
+
+## Method trends over time
+
+## Instructions to applicants over time
+
+```{r}
+# stop rendering at this point because the code below is not working
+knitr::knit_exit()
+```
+
+```{r}
+#| label: fig-requirements-over-time
+
+# look at only these requirements because the others are flat
+
+intresting_requirements <- 
+c("cover letter",
+  "cv",
+  "names of recommenders",
+  "diversity statement",
+  "research statement",
+  "teaching statement")
+
+jobdata_requirements <- 
+jobdata %>% 
+  select(year_ad_posted,
+         starts_with("documents_requested")) %>% 
+  pivot_longer(-year_ad_posted) %>% 
+  mutate(value = case_when(
+    value == "Not requested in the job ad" ~ 0,
+    value == "One" ~ 1,
+    value == "Two (e.g. two syllabi)" ~ 2,
+    value == "Three" ~ 3,
+    .default = 0
+  ))  %>% 
+  # trim names a bit
+  mutate(name = str_remove(name, "documents_requested_")) %>% 
+  mutate(name = str_replace_all(name, "_", " ")) %>% 
+  filter(name %in% intresting_requirements) %>% 
+  mutate(name = str_wrap(name, 10),
+         year_ad_posted = str_replace(year_ad_posted, "-", "\n"))
+
+jobdata_requirements_means <- 
+jobdata_requirements %>% # average number requested per year
+  group_by(year_ad_posted, 
+           name) %>% 
+  summarise(mean_n = mean(value))
+
+integer_breaks <- function(n = 5, ...) {
+  fxn <- function(x) {
+    breaks <- floor(pretty(x, n, ...))
+    names(breaks) <- attr(breaks, "labels")
+    breaks
+  }
+  return(fxn)
+}
+
+ggplot(jobdata_requirements_means) +
+  aes(year_ad_posted, 
+      mean_n,
+      group = name) +
+  geom_smooth(linewidth = 2,
+              colour = "black") +
+  geom_jitter(data = jobdata_requirements,
+             aes(year_ad_posted, 
+                 value),
+             alpha = 0.1,
+             height = 0.2,
+             width =  0.1) +
+  facet_wrap(~name,
+             scales = "free_y",
+             nrow = 1) +
+  xlab("Year") +
+  ylab("Number requested in job ad") +
+  scale_y_continuous(breaks = integer_breaks()) +
+  theme_minimal(base_size = 14) +
+  theme(axis.text.x = element_text(size = 8),
+        strip.text = element_text( size = 20))
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-requirements-per-year.png"),
+       bg ="white",
+       h = 10, # experiment with h and w to get the right size and proportion 
+       w = 20,
+       units = "in",
+       dpi = 900) # make the image nice and crisp
+
+```
+
+```{r}
+#| label: fig-requirements-by-position
+
+# do the requirements differ for associate positions 
+jobdata_requirements_by_rank <- 
+jobdata %>% 
+  mutate(position_title = case_when(
+    str_detect(title_of_position_tenure_track_jobs_only, 
+               "associate") ~ "associate",
+    str_detect(title_of_position_tenure_track_jobs_only, 
+               "assistant") ~ "assistant",
+    str_detect(title_of_position_tenure_track_jobs_only, 
+               "full") ~ "full"))  %>% 
+  select(position_title,
+         starts_with("documents_requested")) %>% 
+  pivot_longer(-position_title) %>% 
+  mutate(value = case_when(
+    value == "Not requested in the job ad" ~ 0,
+    value == "One" ~ 1,
+    value == "Two (e.g. two syllabi)" ~ 2,
+    value == "Three" ~ 3,
+    .default = 0
+  )) %>% 
+  filter(!is.na(position_title))
+
+jobdata_requirements_by_rank_means <- 
+  jobdata_requirements_by_rank %>% 
+  group_by(position_title,
+           name) %>% 
+  summarise(mean = mean(value))
+
+ggplot() +
+    geom_jitter(data = jobdata_requirements_by_rank,
+                  aes(position_title, 
+                      value),
+                height = 0.05,
+                alpha = 0.1) +
+    geom_point(data = jobdata_requirements_by_rank_means,
+               aes(position_title,
+                   mean),
+               size = 4,
+               colour = "red") +
+  facet_wrap( ~ name,
+              scales = "free_y") +
+  theme_minimal()
+```
+
+```{r}
+
+# topical focus by year
+
+library(googlesheets4)
+library(stringi)
+
+topical_foci <-
+  read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0",
+             sheet = "topic")
+
+topical_foci_clean <-
+  map(
+    str_split(topical_foci$`From the data`, ";"),
+    ~.x %>%
+      str_squish() %>%
+      stri_remove_empty() %>%
+      str_to_lower)
+
+jobdata_topic <-
+  jobdata %>%
+  select(topical_focus_of_position) %>%
+  mutate(topical_focus_of_position = str_to_lower(topical_focus_of_position))
+
+jobdata_topic <-
+  # add one column for each topic in our categories
+  cbind(jobdata_topic,
+        setNames( lapply(topical_foci$Category, function(x) x=NA),
+                  topical_foci$Category) )
+
+for(i in 1:length(topical_foci$Category)){
+
+  this_topic <- topical_foci$Category[i]
+
+  # create the pattern to search for
+  x <- paste0(topical_foci_clean[[i]], collapse = "|")
+
+  # do the search through all the job ads for that pattern
+  y <- str_detect(jobdata_topic$topical_focus_of_position,
+                  x)
+
+  # assign back to our data frame in the appropriate location column
+  jobdata_topic[, this_topic] <- y
+
+}
+
+jobdata_topic_year <-
+  jobdata %>%
+  bind_cols(jobdata_topic) %>%
+  select(year_ad_posted,
+         topical_foci$Category) %>%
+  pivot_longer(-year_ad_posted) %>%
+  drop_na()
+
+# how many times each topic mentioned?
+jobdata_topic_year %>%
+  group_by(name) %>%
+  summarise(n = sum(value)) %>%
+  arrange(desc(n)) %>%
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  xlab("") +
+  theme_minimal() +
+  coord_flip()
+
+# explore trends over time. put a point on the max year
+jobdata_topic_year_tally <-
+  jobdata_topic_year %>%
+  # exclude those with <20 ads
+  filter(!name %in% c("Digital Archaeology",
+                      "Pleistocene archaeology",
+                      "Mesoamerican Archaeology",
+                      "Biological anthropology",
+                      "Archaeological theory",
+                      "Evolutionary anthropology",
+                      "North American archaeology"
+  )) %>%
+  group_by(year_ad_posted,
+           name) %>%
+  summarise(n = sum(value)) %>%
+  mutate(prop = n / sum(n))
+
+jobdata_topic_year_tally_max <-
+  jobdata_topic_year_tally %>%
+  group_by(
+    name ) %>%
+  filter(prop == max(prop))
+
+library(ggrepel)
+
+ggplot() +
+  geom_smooth(data =  jobdata_topic_year_tally,
+            aes(year_ad_posted,
+                prop,
+                group = name,
+                colour = name),
+            size = 3,
+           span = 0.7,
+            se = FALSE) +
+  xlab("Year") +
+  ylab("Proportion of all ads") +
+  scale_colour_brewer(palette = "Dark2") +
+  guides(colour = guide_legend("Topic\nfocus",
+                               label.position = "bottom")) +
+    theme_minimal( base_size = 28) +
+  theme(legend.position="bottom")
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-topic-focus-by-year.png"),
+       bg ="white",
+       h = 10, # experiment with h and w to get the right size and proportion
+       w = 20,
        units = "in",
@@
        dpi = 900) # make the image nice and crisp))
-```
-
-```{r}
-# Carnegie Classification
-library(googlesheets4)
-library(stringi)
-
-# data from https://carnegieclassifications.acenet.edu/resource/2021-update-public-file/
-CC <-
-  read_sheet( "https://docs.google.com/spreadsheets/d/1Oe_3X-OR7iuUThb6Tp48prBpbHrWjU4osmHKi1Hbqgc/edit#gid=1922578308",
-             sheet = "Values1")
-Basic2021 <-
-  read_sheet( "https://docs.google.com/spreadsheets/d/1Oe_3X-OR7iuUThb6Tp48prBpbHrWjU4osmHKi1Hbqgc/edit#gid=1922578308",
-             sheet = "Data1") 
-  # remove a bunch of text from uni names to improve our joins
-
-uni_name <- # 547
- jobdata %>%  # 547 rows
-  # remove parentheses and their contents
-  mutate(name = str_squish(str_replace(name_of_hiring_university,
-                            "\\s*\\(.*?\\)$", 
-                            ""))) %>% 
-  mutate(name = str_replace(name,
-                            "California State University,[[:space:]]",
-                            "California State University-")) %>%
-
-  left_join(Basic2021, 
-            keep = TRUE )
-
-# how many schools did we match with the CC data
-uni_name %>% 
-  filter(!is.na(basic2021)) %>% 
-  nrow() # 245, 290, 300, 313, 404, 430 / 433 for US unis
-
-# which schools don't match with the CC data?
-uni_name %>% 
-  select(name.x, name.y)  %>% 
-  filter(is.na(name.y)) %>% 
-  distinct()
-
-uni_name_tally <- 
-uni_name %>% 
-  left_join(CC,
-            join_by("basic2021" == "Value")) %>% 
-  group_by(Category) %>% 
-  tally(sort = TRUE) %>% 
-  drop_na()
-
-uni_name_tally %>% 
-  mutate(Category = str_wrap(Category, width = 30)) %>% 
-ggplot() +
-  aes(reorder(Category, n), n) +
-  geom_col() +
-  coord_flip() +
-  xlab("") +
-  theme_minimal(base_size = 14)
-
-ggsave(here("analysis",
-            "figures", 
-            "fig-carnegie-classification.png"),
-       bg ="white",
-       h = 7, # experiment with h and w to get the right size and proportion
-       w = 12,
-       units = "in",
-       dpi = 900) # make the image nice and crisp))
-
-view(jobdata)
 ```

### 2024-10-11 analysis/paper/article.qmd
finish geo trends

@@
            label = "1:1 ratio",
-           colour = "red",
+           colour = "white",
            size = 3) +
@@
 plot_grid(
-  fig_prop_by_job_title_per_year,
-  fig_ratio_tt_2_ntt_jobs_per_year,
+  fig_prop_by_job_title_per_year + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
+  fig_ratio_tt_2_ntt_jobs_per_year + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
 
@@
       plot = p,
-       h = 15, # experiment with h and w to get the right size and proportion 
-       w = 20,
+       h = 5, # experiment with h and w to get the right size and proportion 
+       w = 7,
        units = "in",
@@
                             "California State University-")) %>% 
-
   left_join(Basic2021, 
@@
 # how many schools did we match with the CC data
+hm_schools_match <- 
 uni_name %>% 
@@
 # how many times each location mentioned?
+p_locations <- 
 jobdata_geo_year %>%
@@
   summarise(n = sum(value)) %>%
-  mutate(prop = n / sum(n))
+  mutate(prop = n / sum(n)) %>% 
+  # add line break for prettier plotting
+  mutate(year_ad_posted_break = str_replace(year_ad_posted, "-", "-\n"))
 
@@
   jobdata_geo_year_tally %>%
-  group_by(
-           name ) %>%
+  group_by(name ) %>%
   filter(prop == max(prop))
@@
 
+p_geo_year <- 
 ggplot() +
+  geom_jitter(data =  jobdata_geo_year_tally,
+              alpha = 0.3, size = 4,
+            aes(year_ad_posted_break,
+                 prop,
+                 group = name,
+                 colour = name)) +
   geom_smooth(data =  jobdata_geo_year_tally,
-            aes(year_ad_posted,
+            aes(year_ad_posted_break,
                  prop,
@@
                  colour = name),
-            size = 3,
+            linewidth = 3,
            se = FALSE 
            ) +
-  xlab("Year") +
+  xlab("") +
   ylab("Proportion of all ads") +
@@
 
+p_geo <- 
+plot_grid(p_locations, 
+          p_geo_year,
+          ncol = 1,
+          align = 'v', 
+          axis = 'l',
+          rel_heights =  c(0.5, 1.5),
+          labels = "AUTO")
+
 ggsave(here("analysis",
@@
        h = 10, # experiment with h and w to get the right size and proportion 
-       w = 20,
+       w = 8,
        units = "in",
@@
 # how many times each location mentioned?
+p <- 
 jobdata_geo_year %>%
@@
   coord_flip()
-
-ggsave(here("analysis",
+```
+
+```{r}
+#| label: fig-show-geo-trends
+#| fig-cap: "A: Frequency of locations mentioned in the text of the job ads.
+B: Popularity of locations in job ads over time.
+Individual data points are show, overlain by a locally weighted regression line for each location to indicate temporal trends."
+
+# show figure from PNG file
+knitr::include_graphics(here("analysis",
             "figures",
-            "fig-geo-us-focus-by-year.png"),
-       bg ="white",
-       h = 10, # experiment with h and w to get the right size and proportion 
-       w = 20,
-       units = "in",
-       dpi = 900) # make the image nice and crisp)
-
-```
-
-- what regions of the world are mentioned in the ads?
-
-- basic trends
-- where are the hiring universities?
-- what Carnegie classifications of the hiring universities?
+            "fig-geo-focus-by-year.png"))
+```
+
+We recorded all geographic regions mentioned in the text where the successful applicant should have expertise and be research active.
+Our analysis focuses on those locations mentioned in 20 or more ads Overall, American locations dominate.
+Panel A of @fig-show-geo-trends shows that a single region of the US, the Southwest, occurs in more job ads than every other part of the world except for the Mediterranean.
+The Southwest includes Arizona and New Mexico, with portions of California, Colorado, Nevada, Oklahoma, Texas, and Utah.
+It is archaeologically significant as the home of the Ancestral Pueblo, Hohokam, and Mogollon peoples who practiced irrigation agriculture and lived in relatively large settlements, compared to other regions of the US.
+The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, also in relatively large numbers.
+The Mediterranean is prominent because it is the location that is often mentioned in job ads focused on classical archaeology (i.e.
+archaeology of Bronze and Iron Age Italy and Greece).
+
+Demand for jobs focusing on the Americas is generally high over time, with a peak in 2019-2020, and a decrease since then.
+Demand for jobs focusing on Africa was very low until 2019-2020, peaking in 2020-2021.
+The proportion of ads with a geographic focus on the Mediterranean has varied substantially, peaking at 2016 and at its lowest in 2019, showing an inverse pattern of the Americas.
+Asia and India, the Near East and Europe are consistently rare as a geographical focus in job ads.
+Asia and India, Africa and the Americas appear correlated with each other, while the Near East and Mediterranean are correlated in an opposite trend.
+
+## Method trends over time
 
 ## Topic trends over time
-
-## Method trends over time
 

### 2024-10-12 analysis/paper/article.qmd
add co-occurrance plot and stat

@@
 
-After completing primary data collection, we studied the topical, geographic, and methods foci of each ad and collaboratively and manually reduced the variation in the raw data into 10-15 categories to simplify analysis and visualisation.
+After completing primary data collection, we studied the topical, geographic, and methods foci of each ad and collaboratively and manually reduced the variation in the raw data into 10-15 categories appearing in 20 (for topics and geography) or 10 (for methods) or more job ads to simplify analysis and visualisation.
+Full details of the category reduction are in our Supplementary Materials.
 Our topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
@@
 For example Mesoamerican archaeology is understood to refer to a specific time period and geographic region.
+Similarly, digital archaeology we recorded as both a method (when a job ad has a clearly distinct topical focus, such as historic archaeology) and a topic (when there is no other topics mentioned in the job ad).
 While these overlaps can make the data challenging to interpret, in our view it reflects the complex realities of  how search committees express their needs in searching for new faculty, and is insightful in how it reveals intersections between different foci.
+
+The entire R code (R Core Team, 2021) and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx to enable re-use of materials and improve reproducibility and transparency (Marwick, 2017).
+All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
+The code is released under the MIT license, the data as CC-0, and the figures as CC-BY, to enable maximum re-use.
 
@@
           ncol = 1,
-          align = 'v', 
-          axis = 'l',
           rel_heights =  c(0.5, 1.5),
@@
 
+```{r}
+# method focus by year
+
+library(stringi)
+
+method_foci <-
+  read_csv(here("analysis/data/raw_data/Tenure Track Job Advertisements in Archaeology - method.csv"))
+
+method_foci_clean <-
+  map(
+    str_split(method_foci$`From the data`, ";"),
+    ~.x %>%
+      str_squish() %>%
+      stri_remove_empty() %>%
+      str_to_lower)
+
+jobdata_method <-
+  jobdata %>%
+  select(methods_focus_of_position) %>%
+  mutate(methods_focus_of_position = str_to_lower(methods_focus_of_position))
+
+jobdata_method <-
+  # add one column for each topic in our categories
+  cbind(jobdata_method,
+        setNames( lapply(method_foci$Category, function(x) x=NA),
+                  method_foci$Category) )
+
+for(i in 1:length(method_foci$Category)){
+
+  this_method <- method_foci$Category[i]
+
+  # create the pattern to search for
+  x <- paste0(method_foci_clean[[i]], collapse = "|")
+
+  # do the search through all the job ads for that pattern
+  y <- str_detect(jobdata_method$methods_focus_of_position,
+                  x)
+
+  # assign back to our data frame in the appropriate location column
+  jobdata_method[, this_method] <- y
+
+}
+
+jobdata_method_year <-
+  jobdata %>%
+  bind_cols(jobdata_method) %>%
+  select(year_ad_posted,
+         method_foci$Category) %>%
+  pivot_longer(-year_ad_posted) %>%
+  drop_na()  %>% 
+  mutate(name = case_when(
+    name == "Computational and Digital archaeology" ~ "Computational and\ndigital archaeology",
+    name == "Material culture analysis" ~ "Material culture\nanalysis",
+    .default = name))
+
+# how many times each topic mentioned?
+p_method <- 
+jobdata_method_year %>%
+  group_by(name) %>%
+  summarise(n = sum(value)) %>%
+  arrange(desc(n)) %>%
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  xlab("") +
+    theme_minimal(base_size = base_size) +
+  theme(axis.text.y  = element_text(size = 7)) +
+  coord_flip()
+
+# explore trends over time. put a point on the max year
+jobdata_method_year_tally <-
+  jobdata_method_year %>%
+  # exclude those with <20 ads
+  filter(!name %in% c("Material culture\nanalysis",
+                      "Ceramic analysis"
+  )) %>%
+  group_by(year_ad_posted,
+           name) %>%
+  summarise(n = sum(value)) %>%
+  mutate(prop = n / sum(n)) %>% 
+  # add line break for prettier plotting
+  mutate(year_ad_posted_break = str_replace(year_ad_posted, "-", "-\n"))
+
+library(ggrepel)
+
+p_method_trend <- 
+ggplot() +
+    geom_jitter(data =  jobdata_method_year_tally,
+              alpha = 0.3, size = 4,
+            aes(year_ad_posted_break,
+                 prop,
+                 group = name,
+                 colour = name)) +
+  geom_smooth(data =  jobdata_method_year_tally,
+            aes(year_ad_posted_break,
+                prop,
+                group = name,
+                colour = name),
+            size = 3,
+            se = FALSE) +
+  xlab("Year") +
+  ylab("Proportion of all ads") +
+  guides(colour = guide_legend("Method\nfocus",
+                               label.position = "bottom")) +
+    theme_minimal( base_size = base_size) +
+  scale_colour_brewer(palette = "Dark2") +
+  theme(legend.position="bottom")
+
+p_meth <- 
+plot_grid(p_method, 
+          p_method_trend,
+          ncol = 1,
+          rel_heights =  c(0.5, 1.5),
+          labels = "AUTO")
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-method-focus-by-year.png"),
+       plot = p_meth, 
+       bg ="white",
+       h = 10, # experiment with h and w to get the right size and proportion
+       w = 8,
+       units = "in",
+       dpi = 900) # make the image nice and crisp))
+```
+
+```{r}
+#| label: fig-show-metho-trends
+#| fig-cap: "A: Frequency of methods mentioned in the text of the job ads.
+B: Popularity of methods in job ads over time.
+Individual data points are show, overlain by a locally weighted regression line for each location to indicate temporal trends."
+
+# show figure from PNG file
+knitr::include_graphics(here("analysis",
+            "figures",
+            "fig-method-focus-by-year.png"))
+```
+
+Landscape archaeology, encompassing GIS and remote sensing, has remained prominent compared to other methods [@fig-show-metho-trends].
+Methods focused on a specific element of the archaeological record, such as Lithic analysis, Zooarchaeology and Ceramics are among the least frequently mentioned in job ads.
+Instead we see the more popular methods here are ones that are relevant to multiple elements of the archaeological record (e.g.
+Archaeobotany encompasses macroscopic and microscopic plant remains; Bioarchaeology may include skeletal analysis, isotopes, protiens, etc.).
+
+Landscape archaeology, although dominant has fluctuated over the years and has been on a downtrend since 2018-2019.
+Computational and digital archaeology is the second most represented method, showing an overall increasing trend, particularly since 2020-2021.
+Archaeobotany shows a strong cyclical trend, with a rise and fall and rise again over our study period.
+Archaeometry and Geoarchaeology have maintained a relatively low but steady presence in job ads, peaking in 2017-2018 and 2018-2019 and declining thereafter.
+Lithic analysis and Zooarchaeology are also mentioned relatively infrequently in job ads and show an inverse correlation with each other after 2018-2019.
+
 ## Topic trends over time
 
-## Instructions to applicants over time
+```{r}
+
+# topical focus by year
+library(stringi)
+
+topical_foci <-
+  read_csv(here("analysis/data/raw_data/Tenure Track Job Advertisements in Archaeology - topic.csv"))
+
+topical_foci_clean <-
+  map(
+    str_split(topical_foci$`From the data`, ";"),
+    ~.x %>%
+      str_squish() %>%
+      stri_remove_empty() %>%
+      str_to_lower)
+
+jobdata_topic <-
+  jobdata %>%
+  select(topical_focus_of_position) %>%
+  mutate(topical_focus_of_position = str_to_lower(topical_focus_of_position))
+
+jobdata_topic <-
+  # add one column for each topic in our categories
+  cbind(jobdata_topic,
+        setNames( lapply(topical_foci$Category, function(x) x=NA),
+                  topical_foci$Category) )
+
+for(i in 1:length(topical_foci$Category)){
+
+  this_topic <- topical_foci$Category[i]
+
+  # create the pattern to search for
+  x <- paste0(topical_foci_clean[[i]], collapse = "|")
+
+  # do the search through all the job ads for that pattern
+  y <- str_detect(jobdata_topic$topical_focus_of_position,
+                  x)
+
+  # assign back to our data frame in the appropriate location column
+  jobdata_topic[, this_topic] <- y
+
+}
+
+jobdata_topic_year <-
+  jobdata %>%
+  bind_cols(jobdata_topic) %>%
+  select(year_ad_posted,
+         topical_foci$Category) %>%
+  pivot_longer(-year_ad_posted) %>%
+  drop_na() %>% 
+  # add line break for prettier plotting
+  mutate(year_ad_posted_break = str_replace(year_ad_posted, "-", "-\n"))
+
+# how many times each topic mentioned?
+p_topic <- 
+jobdata_topic_year %>%
+  group_by(name) %>%
+  summarise(n = sum(value)) %>%
+  arrange(desc(n)) %>%
+  ggplot() +
+  aes(reorder(name, n),
+      n)+
+  geom_col() +
+  xlab("") +
+  theme_minimal() +
+  coord_flip()
+
+# explore trends over time. put a point on the max year
+jobdata_topic_year_tally <-
+  jobdata_topic_year %>%
+  # exclude those with <20 ads
+  filter(!name %in% c("Digital Archaeology",
+                      "Digital archaeology",
+                      "Pleistocene archaeology",
+                      "Mesoamerican Archaeology",
+                      "Biological anthropology",
+                      "Archaeological theory",
+                      "Evolutionary anthropology",
+                      "North American archaeology"
+  )) %>%
+  group_by(year_ad_posted,
+           name) %>%
+  summarise(n = sum(value)) %>%
+  mutate(prop = n / sum(n))  %>% 
+  # add line break for prettier plotting
+  mutate(year_ad_posted_break = str_replace(year_ad_posted, "-", "-\n"))
+
+jobdata_topic_year_tally_max <-
+  jobdata_topic_year_tally %>%
+  group_by(
+    name ) %>%
+  filter(prop == max(prop))
+
+library(ggrepel)
+
+p_topic_trend <- 
+ggplot() +
+      geom_jitter(data =  jobdata_topic_year_tally,
+              alpha = 0.3, size = 4,
+            aes(year_ad_posted_break,
+                 prop,
+                 group = name,
+                 colour = name)) +
+  geom_smooth(data =  jobdata_topic_year_tally,
+            aes(year_ad_posted_break,
+                prop,
+                group = name,
+                colour = name),
+            size = 3,
+            se = FALSE) +
+  xlab("Year") +
+  ylab("Proportion of all ads") +
+  scale_colour_brewer(palette = "Dark2") +
+  guides(colour = guide_legend("Topic\nfocus",
+                               label.position = "bottom")) +
+    theme_minimal( base_size = base_size) +
+  theme(legend.position="bottom")
+
+p_topi <- 
+plot_grid(p_topic, 
+          p_topic_trend,
+          ncol = 1,
+          rel_heights =  c(0.5, 1.5),
+          labels = "AUTO")
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-topic-focus-by-year.png"),
+       plot = p_topi, 
+       bg ="white",
+       h = 10, # experiment with h and w to get the right size and proportion
+       w = 8,
+       units = "in",
+       dpi = 900) # make the image nice and crisp))
+```
+
+```{r}
+#| label: fig-show-topi-trends
+#| fig-cap: "A: Frequency of methods mentioned in the text of the job ads.
+B: Popularity of methods in job ads over time.
+Individual data points are show, overlain by a locally weighted regression line for each location to indicate temporal trends."
+
+# show figure from PNG file
+knitr::include_graphics(here("analysis",
+            "figures",
+            "fig-topic-focus-by-year.png"))
+```
+
+The most frequently mentioned topic in the job ads in Environmental archaeology [@fig-show-topi-trends], which is our category for phrases found in the text of ads such as human-environmental dynamics, interaction between humans and their environments, environmental change, climate change, historical ecology, ecological knowledge,human ecology, and ecological systems.
+Public archaeology is the second most frequent topic overall, this includes phrases such as cultural resource management, cultural heritage, heritage studies, museum studies, human rights, community engaged, historic preservation; social justice, community-based,  repatriation, and community-engaged archaeology.
+The least frequent topics are Pleistocene archaeology, e.g.
+human origins; hunter-gatherer archaeology, and digital archaeology.
+
+```{r}
+#| label: richness-in-ads
+
+# geo co-occurance
+jobdata_geom_cooc <- 
+  jobdata_geo %>% 
+  select(-geographic_focus_of_position) %>% 
+  drop_na()  %>% 
+  mutate_all(as.numeric)
+
+# topic co-occurance
+jobdata_topic_cooc <- 
+jobdata_topic %>% 
+  select(-topical_focus_of_position) %>% 
+  drop_na()  %>% 
+  mutate_all(as.numeric)
+
+# method co-occurance
+jobdata_method_cooc <- 
+  jobdata_method %>% 
+  select(-methods_focus_of_position) %>% 
+  drop_na()  %>% 
+  mutate_all(as.numeric)
+
+kw_test_richness_input <- 
+map(
+  list(jobdata_topic_cooc,
+       jobdata_method_cooc,
+       jobdata_geom_cooc),
+  ~rowSums(.x)) %>% 
+  imap(~ data.frame(group = .y, value = .x)) %>%
+  bind_rows()
+
+kw_test_richness <- 
+  kruskal.test(kw_test_richness_input, 
+               value ~ group)
+
+# from https://stackoverflow.com/a/61758180/1036500 for pretty printing
+expSup <- function(w, digits=0) {
+  sprintf(paste0("%.", digits, "f*x*10^%d^"), w/10^floor(log10(abs(w))), floor(log10(abs(w))))
+}
+```
+
+```{r}
+# plot topic co-occurance
+
+# Calculate co-occurrence matrix
+cooccurrence_matrix <- t(jobdata_topic_cooc) %*% as.matrix(jobdata_topic_cooc)
+
+# Convert to a tidy format for easy inspection and manipulation
+cooccurrence_tidy <- as.data.frame(as.table(as.matrix(cooccurrence_matrix)))
+
+# Rename the columns for better readability
+colnames(cooccurrence_tidy) <- c("Variable1", "Variable2", "Cooccurrence")
+
+# View the top co-occurring variable pairs
+cooccurrence_tidy_df <- 
+cooccurrence_tidy %>%
+  filter(Variable1 != Variable2) %>%  # Exclude diagonal (self-co-occurrence)
+  filter(as.numeric(Variable1) < as.numeric(Variable2)) %>% 
+  arrange(desc(Cooccurrence)) %>% 
+  drop_na()
+
+library(viridis)
+
+# Plot heatmap of co-occurrence matrix
+p_coo <- 
+ggplot(cooccurrence_tidy_df, 
+       aes(Variable1, Variable2, fill = Cooccurrence)) +
+  geom_tile() +
+  scale_fill_viridis(option = "viridis") +
+  theme_minimal(base_size  = base_size) +
+  guides(fill = guide_colorbar(title.position = "top")) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        axis.text.y = element_text(angle = 45, hjust = 1),
+        legend.position = c(.7,.2),
+        legend.direction = "horizontal") +
+  coord_equal() +
+  labs(x = "", y = "")
+
+ggsave(here("analysis",
+            "figures", 
+            "fig-topic-cooc-heatmap.png"),
+       plot = p_coo, 
+       bg ="white",
+       h = 8, # experiment with h and w to get the right size and proportion
+       w = 8,
+       units = "in",
+       dpi = 900) # make the image nice and crisp))
+
+```
+
+```{r}
+#| label: fig-show-cooc
+#| fig-cap: "Heatmap of topic co-occurrance in job ads."
+
+# show figure from PNG file
+knitr::include_graphics(here("analysis",
+            "figures",
+            "fig-topic-cooc-heatmap.png"))
+```
+
+Job ads frequently mention multiple topics in a single call for applications, more often than they they multiple methods or geographic locations.
+A Kruskal-Wallis test indicated significantly higher richness in topics compared to richness of geographic locations or methods in job ads (χ2 (df = `r unname(kw_test_richness$df)`, N = `r nrow(kw_test_richness_input)`)  = `r round(unname(kw_test_richness$statistic), 2)`, p = `r expSup(unname(kw_test_richness$p.value))`).
+@fig-show-cooc shows topic co-occurances in our sample.
 
@@
 ```
+
+## Instructions to applicants over time
 

### 2024-10-14 analysis/paper/article.qmd
writing about topic trends over time

@@
 #| fig-cap: "A: Frequency of methods mentioned in the text of the job ads.
-B: Popularity of methods in job ads over time.
-Individual data points are show, overlain by a locally weighted regression line for each location to indicate temporal trends."
+B: Popularity of methods in job ads over time for topics that appear in 20 or more ads.
+Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
 
@@
 human origins; hunter-gatherer archaeology, and digital archaeology.
+
+In the years 2019-2020 and 2020-2021 there are striking changes in the popularity of topics in job ads.
+Indigenous and historical archaeology became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
+Conversely, archaeological science, which was popular during 2012-2017, was rarely mentioned in job ads during 2019-2021.
+Ancient Europe and the Mediterranean, which was consistently one of the least frequently mentioned topics, virtually disappeared from job ads during 2019-2021.
+Mentions  of complex societies in ads consistently decrease over the study period, while mentions of public archaeology consistently increase.
+The topic of Environmental Archaeology remains consistently high over time.
 
@@
 
-Job ads frequently mention multiple topics in a single call for applications, more often than they they multiple methods or geographic locations.
+Job ads are frequently rich in topics, that is, they mention multiple topics in a single call for applications, more often than they mention multiple methods or geographic locations.
 A Kruskal-Wallis test indicated significantly higher richness in topics compared to richness of geographic locations or methods in job ads (χ2 (df = `r unname(kw_test_richness$df)`, N = `r nrow(kw_test_richness_input)`)  = `r round(unname(kw_test_richness$statistic), 2)`, p = `r expSup(unname(kw_test_richness$p.value))`).
-@fig-show-cooc shows topic co-occurances in our sample.
+@fig-show-cooc shows topic co-occurrences in our sample.
+Indigenous and historical archaeology often occurs in job ads with Public archaeology and North American archaeology.
+Complex societies and environmental archaeology were frequently found in the same ads.
+Biological archaeology, archaeological science, and evolutionary archaeology are another cluster of topics that frequently occur together.
+Other topics are relatively isolated, for example, Pleistocene archaeology and digital archaeology rarely occur with other topics.
 

### 2024-10-16 analysis/paper/article.qmd
working on requirements

@@
 
-```{r}
-# stop rendering at this point because the code below is not working
-knitr::knit_exit()
-```
-
 ## Instructions to applicants over time
@@
 ```{r}
-#| label: fig-requirements-over-time
+#| label: requirements-over-time
 
@@
   filter(name %in% intresting_requirements) %>% 
-  mutate(name = str_wrap(name, 10),
-         year_ad_posted = str_replace(year_ad_posted, "-", "\n"))
+  mutate(name = str_wrap(name, 30),
+         year_ad_posted = str_replace(year_ad_posted, "-", "\n")) %>% 
+  # add line break for prettier plotting
+  mutate(year_ad_posted_break = str_replace(year_ad_posted, "-", "-\n"))
 
@@
            name) %>% 
-  summarise(mean_n = mean(value))
+  summarise(mean_n = mean(value)) %>% 
+  # add line break for prettier plotting
+  mutate(year_ad_posted_break = str_replace(year_ad_posted, "-", "-\n"))
 
@@
 
+p_req <- 
 ggplot(jobdata_requirements_means) +
-  aes(year_ad_posted, 
+  aes(year_ad_posted_break, 
       mean_n,
@@
   geom_jitter(data = jobdata_requirements,
-             aes(year_ad_posted, 
+             aes(year_ad_posted_break, 
                  value),
@@
              scales = "free_y",
-             nrow = 1) +
-  xlab("Year") +
+             nrow = 2) +
+  xlab("") +
   ylab("Number requested in job ad") +
@@
   theme(axis.text.x = element_text(size = 8),
-        strip.text = element_text( size = 20))
-
-ggsave(here("analysis",
+        strip.text = element_text( size = 16))
+
+ggsave(filename = here("analysis",
             "figures", 
             "fig-requirements-per-year.png"),
+       plot = p_req,
        bg ="white",
-       h = 10, # experiment with h and w to get the right size and proportion 
-       w = 20,
+       h = 6, # experiment with h and w to get the right size and proportion 
+       w = 12,
        units = "in",
@@
 ```{r}
-#| label: fig-requirements-by-position
+#| label: fig-requirements-over-time
+#| fig-cap: "Changing requirements in job ads over time.
+Each point represents an individual job ad.
+The points have been jittered to avoid overlap.
+The black line is a locally weighted regression line to identify temporal trends.
+The grey region is the 95% confidence interval for the regression line."
+
+# show figure from PNG file
+knitr::include_graphics(here("analysis",
+            "figures",
+            "fig-requirements-per-year.png"))
+```
+
+Over our ten year study period there have been substantial changes in the instructions to applicants in terms of the type and number of documents that are requested by the search committees [@fig-requirements-over-time].
+A cover letter and CV are less requested in more recent years, especially after 2018.
+The requirement for a diversity statement is rare until 2019-2020, peaks around 2020-2021, then decreases towards the present.
+Requests for names of recommenders (either zero or three names, rarely only two names) reaches a maximum during 2019-2020 then decreases towards the present.
+The requirement for a research statement and teaching statement increases after 2015-2016, and becomes more frequent in job ads in more recent years.
+Requests for course descriptions, syllabus samples, teaching evaluations, transcripts and writing samples are consistently low over time (not shown here).
+
+```{r}
+#| label: requirements-by-position
 
@@
   )) %>% 
-  filter(!is.na(position_title))
+  filter(!is.na(position_title))   %>% 
+  mutate(position_title = str_to_title(position_title)) %>% 
+    # trim names a bit
+  mutate(name = str_remove(name, "documents_requested_")) %>% 
+  mutate(name = str_replace_all(name, "_", " "))
 
@@
 
+p_req_rank <- 
 ggplot() +
@@
               scales = "free_y") +
+   scale_y_continuous(breaks = integer_breaks()) +
+  xlab("") +
+  ylab("Number requested in job ad") +
   theme_minimal()
-```
-
-```{r}
-
-# topical focus by year
-
-library(googlesheets4)
-library(stringi)
-
-topical_foci <-
-  read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0",
-             sheet = "topic")
-
-topical_foci_clean <-
-  map(
-    str_split(topical_foci$`From the data`, ";"),
-    ~.x %>%
-      str_squish() %>%
-      stri_remove_empty() %>%
-      str_to_lower)
-
-jobdata_topic <-
-  jobdata %>%
-  select(topical_focus_of_position) %>%
-  mutate(topical_focus_of_position = str_to_lower(topical_focus_of_position))
-
-jobdata_topic <-
-  # add one column for each topic in our categories
-  cbind(jobdata_topic,
-        setNames( lapply(topical_foci$Category, function(x) x=NA),
-                  topical_foci$Category) )
-
-for(i in 1:length(topical_foci$Category)){
-
-  this_topic <- topical_foci$Category[i]
-
-  # create the pattern to search for
-  x <- paste0(topical_foci_clean[[i]], collapse = "|")
-
-  # do the search through all the job ads for that pattern
-  y <- str_detect(jobdata_topic$topical_focus_of_position,
-                  x)
-
-  # assign back to our data frame in the appropriate location column
-  jobdata_topic[, this_topic] <- y
-
-}
-
-jobdata_topic_year <-
-  jobdata %>%
-  bind_cols(jobdata_topic) %>%
-  select(year_ad_posted,
-         topical_foci$Category) %>%
-  pivot_longer(-year_ad_posted) %>%
-  drop_na()
-
-# how many times each topic mentioned?
-jobdata_topic_year %>%
-  group_by(name) %>%
-  summarise(n = sum(value)) %>%
-  arrange(desc(n)) %>%
-  ggplot() +
-  aes(reorder(name, n),
-      n)+
-  geom_col() +
-  xlab("") +
-  theme_minimal() +
-  coord_flip()
-
-# explore trends over time. put a point on the max year
-jobdata_topic_year_tally <-
-  jobdata_topic_year %>%
-  # exclude those with <20 ads
-  filter(!name %in% c("Digital Archaeology",
-                      "Pleistocene archaeology",
-                      "Mesoamerican Archaeology",
-                      "Biological anthropology",
-                      "Archaeological theory",
-                      "Evolutionary anthropology",
-                      "North American archaeology"
-  )) %>%
-  group_by(year_ad_posted,
-           name) %>%
-  summarise(n = sum(value)) %>%
-  mutate(prop = n / sum(n))
-
-jobdata_topic_year_tally_max <-
-  jobdata_topic_year_tally %>%
-  group_by(
-    name ) %>%
-  filter(prop == max(prop))
-
-library(ggrepel)
-
-ggplot() +
-  geom_smooth(data =  jobdata_topic_year_tally,
-            aes(year_ad_posted,
-                prop,
-                group = name,
-                colour = name),
-            size = 3,
-           span = 0.7,
-            se = FALSE) +
-  xlab("Year") +
-  ylab("Proportion of all ads") +
-  scale_colour_brewer(palette = "Dark2") +
-  guides(colour = guide_legend("Topic\nfocus",
-                               label.position = "bottom")) +
-    theme_minimal( base_size = 28) +
-  theme(legend.position="bottom")
-
-ggsave(here("analysis",
-            "figures", 
-            "fig-topic-focus-by-year.png"),
+
+ggsave(filename = here("analysis",
+            "figures", 
+            "fig-requirements-by-rank.png"),
+       plot = p_req_rank,
        bg ="white",
-       h = 10, # experiment with h and w to get the right size and proportion
-       w = 20,
+       h = 6, # experiment with h and w to get the right size and proportion 
+       w = 8,
        units = "in",
-       dpi = 900) # make the image nice and crisp))
-
-```
-
-```{r}
-# method focus by year
-
-library(googlesheets4)
-library(stringi)
-
-method_foci <-
-  read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0",
-             sheet = "method")
-
-method_foci_clean <-
-  map(
-    str_split(method_foci$`From the data`, ";"),
-    ~.x %>%
-      str_squish() %>%
-      stri_remove_empty() %>%
-      str_to_lower)
-
-jobdata_method <-
-  jobdata %>%
-  select(methods_focus_of_position) %>%
-  mutate(methods_focus_of_position = str_to_lower(methods_focus_of_position))
-
-jobdata_method <-
-  # add one column for each topic in our categories
-  cbind(jobdata_method,
-        setNames( lapply(method_foci$Category, function(x) x=NA),
-                  method_foci$Category) )
-
-for(i in 1:length(method_foci$Category)){
-
-  this_method <- method_foci$Category[i]
-
-  # create the pattern to search for
-  x <- paste0(method_foci_clean[[i]], collapse = "|")
-
-  # do the search through all the job ads for that pattern
-  y <- str_detect(jobdata_method$methods_focus_of_position,
-                  x)
-
-  # assign back to our data frame in the appropriate location column
-  jobdata_method[, this_method] <- y
-
-}
-
-jobdata_method_year <-
-  jobdata %>%
-  bind_cols(jobdata_method) %>%
-  select(year_ad_posted,
-         method_foci$Category) %>%
-  pivot_longer(-year_ad_posted) %>%
-  drop_na()
-
-# how many times each topic mentioned?
-jobdata_method_year %>%
-  group_by(name) %>%
-  summarise(n = sum(value)) %>%
-  arrange(desc(n)) %>%
-  ggplot() +
-  aes(reorder(name, n),
-      n)+
-  geom_col() +
-  xlab("") +
-  theme_minimal() +
-  coord_flip()
-
-# explore trends over time. put a point on the max year
-jobdata_method_year_tally <-
-  jobdata_method_year %>%
-  # exclude those with <20 ads
-  filter(!name %in% c("Material culture analysis",
-                      "Ceramic analysis"
-  )) %>%
-  group_by(year_ad_posted,
-           name) %>%
-  summarise(n = sum(value)) %>%
-  mutate(prop = n / sum(n))
-
-library(ggrepel)
-
-ggplot() +
-  geom_smooth(data =  jobdata_method_year_tally,
-            aes(year_ad_posted,
-                prop,
-                group = name,
-                colour = name),
-            size = 3,
-           span = 0.7,
-            se = FALSE) +
-  xlab("Year") +
-  ylab("Proportion of all ads") +
-  guides(colour = guide_legend("Method\nfocus",
-                               label.position = "bottom")) +
-    theme_minimal( base_size = 28) +
-  scale_colour_brewer(palette = "Dark2") +
-  theme(legend.position="bottom")
-
-ggsave(here("analysis",
-            "figures", 
-            "fig-method-focus-by-year.png"),
-       bg ="white",
-       h = 10, # experiment with h and w to get the right size and proportion
-       w = 20,
-       units = "in",
-       dpi = 900) # make the image nice and crisp))
-```
-
-```{r}
-# EEO
-library(stringr)
-library(dplyr)
-library(stopwords)
-library(tidyverse)
-library(tidytext)
-library(tm)
-
-stopwords <- stopwords("en")
-
-EEO <- jobdata %>%
-  select(equal_employment_opportunity_statement, year_ad_posted) %>% 
-  drop_na
-
-remove_stopwords <- function(text, stopwords) {
-  pattern <- paste0("\\b(", paste(stopwords, collapse = "|"), ")\\b")
-  str_remove_all(text, pattern)
-}
-
-EEO_clean <- EEO %>%
-  mutate(equal_employment_opportunity_statement = str_to_lower(equal_employment_opportunity_statement), 
-         equal_employment_opportunity_statement = str_remove_all(equal_employment_opportunity_statement, "[[:punct:]]"),
-         equal_employment_opportunity_statement = remove_stopwords(equal_employment_opportunity_statement, stopwords))
-
-#word count for EEO statement
-EEO_wordcount <- EEO %>%
-  mutate(word_count = str_count(equal_employment_opportunity_statement, "\\S+")) %>%
-  print()
-
-jobdata_EEO_year_tally <-
-  EEO_wordcount %>%
-  group_by(year_ad_posted) %>%
-  summarise(n = mean(word_count))
-
-fig_ave_length_EEO_per_year <- 
-ggplot(jobdata_EEO_year_tally) +
-  aes(year_ad_posted,n) +
-  geom_col() +
-  ylab("Average length of EEO statement by year") +
-  theme_minimal(base_size = 28) +
-  guides(x = "none")
-
-ggplot(EEO_wordcount) +
-  aes(year_ad_posted, word_count) +
-  geom_boxplot() +
-  geom_quasirandom()
-
-ggplot(jobdata_requirements_means %>% 
-         filter(name == "diversity\nstatement") ) +
-  aes(year_ad_posted, 
-      mean_n) +
-  geom_col()
-
-# here's the list of words of interest
-community_word <- c("women","color","religion","race","age","veteran","disability","citizenship","national origin","sex","gender","marital status","genetic information","breastfeeding","sexual orientation","transgender status","gender expression","gender identity","ancestry","income assignment for child support","arrest and court record","domestic or sexual violence victim status","national guard absence","visible minorities","aboriginal persons","indigenous","LGBTQ2S+","pregnancy","childbirth","faith","genetic or family medical history","medical condition","nationality","ethnicity","parental status","accent","ethnic","creed","political affiliation","immigration status")
-
-EEO_word <- EEO_clean %>%
-   unnest_tokens(word, 
-                 equal_employment_opportunity_statement) %>%
-  anti_join(get_stopwords()) %>%
-  count(word, sort = TRUE) %>%
-  drop_na()
-
-# prepare a dataframe to store the count of words of interest
-df <- data.frame(matrix(ncol = length(community_word), 
-                        nrow = 0))
-colnames(df) <- community_word
-
-for(i in 1: nrow(EEO_clean)){
-  df[i, ] <- str_detect(EEO_clean[i, ], 
-                   community_word)
-}
-
-# detect if the words of interest is present or not.
-EEO_clean_community_word <- 
-  bind_cols(EEO_clean, df) %>% 
-  rowwise() %>% 
-  mutate(community_word_tally = sum(c_across(all_of(community_word)), 
-                                    na.rm = T)) %>% 
-  bind_cols(EEO_wordcount) %>% 
-  mutate(eeo_words_total_ratio = community_word_tally / word_count)
-
-ggplot(EEO_clean_community_word) +
-  aes(community_word_tally,
-      word_count,
-      size = eeo_words_total_ratio) +
-  geom_point()
-
-ggplot(EEO_clean_community_word) +
-  aes(community_word_tally) +
-  geom_histogram()
-```
-
-```{r}
-ggplot(EEO_word %>% 
-         filter(n > 20), 
-       aes(x = reorder(word, n), 
-           y = n)) +
-  geom_col() +
-  labs(x = "Words", 
-       y = "Frequency", 
-       title = "Word Frequency") +
-  coord_flip()
-```
-
-```{r}
-# explore QS rankings
-
-qs_rankings <- 
-  readxl::read_excel("analysis/data/raw_data/2024 QS WUR by Subject - Public Results (for qs.com)_48.xlsx", sheet = "Archaeology", skip = 12)
-
-jobdata %>% 
-  mutate(Institution = str_remove(name_of_hiring_university,
-                                  " \\(.*\\)")) %>% 
-  mutate(Institution = case_when(
-    # modify the school name to match with the name in the QS list
-    Institution == "Ohio State University-Main Campus" ~ "The Ohio State University",
-  "Cambridge" ~ "University of Cambridge",
-  "Cambridge University" ~ "University of Cambridge",
-  "Keble College" ~ "University of Oxford",
-  "University of California-Berkeley" ~ "University of California, Berkeley (UCB)",
-"University of California-Los Angeles" ~ "University of California, Los Angeles (UCLA)",
-"University of Tuebingen" ~ "Eberhard Karls Universität Tübingen",
-"Universität Tübingen" ~ "Eberhard Karls Universität Tübingen",
-"Arizona State University Campus Immersion" ~ "Arizona State University",
-"Kiel University" ~ "Christian-Albrechts-University zu Kiel",
-"University of Arizona" ~ "The University of Arizona",
-"The Pennsylvania State University" ~ "Pennsylvania State University",
-"University of Bologna" ~ "Alma Mater Studiorum - University of Bologna",
-"University of Sydney" ~ "The University of Sydney",
-"University of Copenhagen, Saxo Institute" ~ "University of Copenhagen",
-"The University of Texas at Austin" ~ "University of Texas at Austin",
-
-    .default = as.character(Institution)
-  )) %>% 
-  select(Institution) %>% 
-  left_join(qs_rankings) %>% 
-  drop_na(`2024`) %>% 
-  distinct(Institution, .keep_all = TRUE) %>% 
-  View()
-
-```
+       dpi = 900) # make the image nice and crisp
+```
+
+```{r}
+#| label: fig-requirements-by-position
+#| fig-cap: "Variation in application requirements by academic position.
+Each point represents an individual job ad.
+The points have been jittered to avoid overlap.
+The red point is the mean value for each group."
+
+# show figure from PNG file
+knitr::include_graphics(here("analysis",
+            "figures",
+            "fig-requirements-by-rank.png"))
+```
+
+@fig-requirements-by-position
 

### 2024-10-22 analysis/paper/article.qmd
writing the discussion

@@
 
+<!--
+
+cite these
+
+- https://www.anthropology-news.org/articles/hello-to-tristes-tropes/
+
+-->
+
 # Methods
@@
 These top five states also correspond to the top five most populous US states, indicating that rates of hiring is approximately proportional to population density.
+These top five states for job ads are also the top five states with highest number of degrees awarded in Anthropology (U.S.
+Department of Education nd).
 Similarly, the lowest counts of job ads were observed in states with the lowest populations: North Dakota, South Dakota, Alaska, and Nebraska.
@@
 In the years 2019-2020 and 2020-2021 there are striking changes in the popularity of topics in job ads.
-Indigenous and historical archaeology became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
+Indigenous and historical archaeology (which includes archaeology of enslaved people) became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
 Conversely, archaeological science, which was popular during 2012-2017, was rarely mentioned in job ads during 2019-2021.
@@
 
-@fig-requirements-by-position
+Instructions to applicants vary by the rank of the position advertised [@fig-requirements-by-position].
+Assistant professor positions tend to have the most complex instructions with many different types of documents required.
+Conversely, full professor positions have minimal requirements.
+The documents that are most frequently requested across the three ranks are cover letter and CV.
 
 # Discussion
+
+Our results reveal some of these shifts, providing insights into the dynamics of topical, methods, and geographic foci over time.
+While these dynamics are deeply felt by people actively seeking faculty jobs, we believe this is the first time they have been quantified at such a large scale.
+Trends in job ads reflect broader shifts in intellectual and practical priorities about archaeology, undergraduate education, and the process of hiring professors.
+The demand for archaeology faculty, indicated by the total number of tenure-track jobs, maybe explained by a variety of processes.
+The downward trend in tenure-track positions during 2013-2019 related to decline in undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a].
+The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities adapted to online instruction and assessment in an effort to minimize the spread of the virus.
+In addition to a dip in job ads, our results show that 2019-2021 was a major inflection point in the popularity of topics and regions in job ads.
+Indigenous and historical archaeology and the Americas strongly increase in popularity at this time, while archaeological science, complex societies, the Mediterranean and Near East and show declines.
+Similarly, the number of job ads with a geographic focus on the Americas and Africa peaks during this time.
+These shifts in the topics and geographic foci of job ads were likely been influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices.
+COVID-19 negatively impacted Black, American Indian and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US.
+The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic with the murder of George Floyd by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
+Mass protests objecting to Floyd's murder generated widespread concern about racial inequities and stimulated a broad interest in addressing systemic racial injustice.
+Our data suggest that US universities participated in this movement by adjusting their hiring plans to prioritize recruiting archaeologists working on topics relevant to Black and Indigenous communities.
+Many universities may have hoped to hire Black and Indigenous archaeologists as part of their effort to tackle systemic racism.
+However as it is illegal in the US for universities use race or ethnicity as a primary factor in hiring (due to the 1964 Civil Rights Act), universities tailored the content of the job ads to focus on topics where Black and Indigenous researchers are most numerous, ensuring that these researchers would be well-represented in the pool of applicants.
+
+This striking change in topics and geographic foci of 2019-2021 occurred against a backdrop of several longer-term trends.
+We found a long-term decline in topics such as the archaeology of complex societies and archaeological science, Mesoamerica and South America as a geographic focus, and methods relating to landscape archaeology.
+These trends are harder to explain as we cannot link their start to a discrete moment like COVID-19.
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism.
+This might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical methods for compositional analysis.
+A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience in the past may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
+Other relevant factors may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation, and an increasing emphasis on more job-oriented degrees that prepare students for working in cultural heritage management rather than advanced research.
+
+Our data on the requirements for applicants shows that application complexity has gradually increased over time.
+This trend is especially pronounced for Assistant Professor positions which make more demands on applicants than Associate and Full Professor positions.
+We saw a growing demand for research and teaching statements.
+Demand for diversity statements shows a unique trajectory, peaking in 2020-2021 and declining into the present.
+This may relate to the intersecting concerns emerging during the COVID-19 pandemic, which have declined in job ads as the pandemic moves into the past.
+The decline in recent years for names of recommenders in the initial application may be a response to criticisms of the burden on the applicant of preparing numerous complex job applications [e.g.
+@dennisWorstAnthroJob2022].
+In recognition of this burden, not only on applicants but also on colleagues writing letters of recommendation over and over to support applicants, many hiring committees now follow the recommendations of @dennisWorstAnthroJob2022 in only requesting names and letters of recommendation at later stages of the hiring process, if at all.
+
+A key limitation of our research is that it does not include an analysis of the academic profiles of those who were eventually hired from the job ads.
+Our results reveal collective aspirations for the future of the field, but do not show how these worked out in the topical and geographic foci of the people who were eventually hired.
 

### 2024-10-31 analysis/paper/article.qmd
text in background section

@@
 
+The academic job market is a source of uncertainty, both for job seekers, as they look for ads that are well-matched to their niche of skills and experience that have taken years to cultivate and refine, and for hiring committees as they sift through sometimes hundreds of applicants in the hope of finding a candidate that meets all of their needs.
+For many academics there is a perception that the last two to three decades have seen the job market become increasingly difficult for job seekers.
+In many fields it is common to hear that there are fewer jobs available relative to the number of PhD graduates, higher numbers of short-term temporary positions relative to secure tenure-track positions, and increasingly complex application requirements with specific statements (e.g.
+teaching, research, and diversity) that need to be tailed for each application, making the process of applying for jobs a full-time job of its own.
+
+In anthropology in the US the number of doctoral anthropology graduates has increased by about 70% over the past 30 years, but the number of new faculty positions has not increased proportionally [@speakman2018choosing].
+Among biological anthropologists, @Passalacqua_2018 found a ratio of 0.81 PhDs to job academic advertisements per year, concluding that academic positions in biological anthropology are barely at sustainable levels.
+This echos findings from other fields, for example in biomedical sciences where there is one tenure-track position in the US for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
+PhD graduates who do not go directly into a tenure-track faculty position after graduation often go into a series of short-term appointments as part-time or limited contract instructors, also known as adjunct or contingent faculty.
+Rising numbers of contingent faculty has long been a concern in US higher education [@Trevithick2010].
+Data from 2021–22 Faculty indicate that more than 60% of faculty positions in US universities were held by non-tenure track full time or part time contingent faculty members [@colby2022annual].
+Contingent faculty positions are precarious because they provide low or no health and retirement benefits, little or nothing in the way of professional development, and lower salaries relative to tenure-track positions.
+PhD graduates who have a sequence of short-term academic positions are often disadvantaged financially, due to low compensation combined with expensive frequent relocation, socially due to isolation from family and community, and professionally due to being ineligible for many decision-making roles at the universities they work at [@platzerAcademicPrecarityAmerican2018].
+
+Research into
+
+abysmal
+
 <!--
@@
 Because of its reputation for aggregating ads from diverse sources and rapidly-updated information that is not available elsewhere, the Academic Jobs Wiki has a large community of users that keep it updated and accurate has become an authoritative data source for studies of hiring trends in academia [e.g.
-@musial2018five] and a widely recommended resource [e.g.
+@Passalacqua_2018; @musial2018five] and a widely recommended resource [e.g.
 @lightfoot2021preparing].
@@
 The downward trend in tenure-track positions during 2013-2019 related to decline in undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a].
-The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities adapted to online instruction and assessment in an effort to minimize the spread of the virus.
+The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus.
+
 In addition to a dip in job ads, our results show that 2019-2021 was a major inflection point in the popularity of topics and regions in job ads.
@@
 A key limitation of our research is that it does not include an analysis of the academic profiles of those who were eventually hired from the job ads.
-Our results reveal collective aspirations for the future of the field, but do not show how these worked out in the topical and geographic foci of the people who were eventually hired.
+Our results reveal collective aspirations for the future of the field, but only from those who are securely employed in it as tenured professors serving on hiring committees at US universities.
+Our results do not show how these aspirations worked out in the topical and geographic foci of the people who were eventually hired for the positions studied here.
+Future work should consider interviewing faculty hired during our study period to match up people to the ads they responded to.
+If the successful applicants can be identified, then the match between the details of the job ad and the applicant's research can be analysed.
+Then we can assess how effective the hiring process is for driving change in the discipline by setting the topics of courses that will be taught in undergraduate and graduate curricula, and research that will be supported by universities.
 

### 2024-11-01 analysis/paper/article.qmd
writing background

@@
 
-Research into
-
-abysmal
+Studies of who holds tenure-track faculty positions in archaeology at US universities, shows that success in job applications is strongly determined by where applicants get their PhD from.
+Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing].
+Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
+This network analysis further found that programs with large endowments and with faculty who hold prestigious awards and are widely cited by other scholars produce the majority of tenured and tenure-track faculty in anthropology.
+This pattern of inequality is not unique to archaeology [@wapmanQuantifyingHierarchyDynamics2022a], but these studies draw attention to the paradox of a discipline that has many members committed to fighting social inequalities, and yet continues to reproduce its own systematic inequalities and hierarchies through hiring practices that favour an elite minority of graduates with prestigious institutional affiliations.
+
+One way that some hiring committees are tackling these biases towards prestigious institutions is to provided detailed instructions to applicants on how to prepare their application materials so that the hiring committee can evaluate and rank the specific accomplishments of candidate across a variety of categories, rather than ranking based on prestige of the candidates' graduate schools.
+This has resulted in job ads that are highly prescriptive in the types of documents that applicants should submit.
+For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements about their teaching and research record and plans.
+In a comparison of job ads from 1999–2000 and 2019–2020 published in _Anthropology News_ @HelloTristesTropes2021 noticed a trend toward increasingly asking for more materials from applicants.
+They found that twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy.
+In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported two job ads with extreme requirements, requesting nine and thirteen documents from applicants.
+A notable requirement that has emerged in recent years is a diversity statement, where the applicant describes their knowledge of, prior contributions to, and future goals for advancing diversity, equity, inclusion.
+This requirement has been much-debated due to an experiment during 2016-2022 at several University of California campuses that used diversity statements as the first cut for selecting candidates for tenure-track faculty positions [@soucek2021diversity].
+Only candidates that scored highly on their diversity statements would have the rest of their application evaluated.
+This experiment generated intense and widespread public debate about merits and risks of requiring and using diversity statements, drawing attention to the challenges of hiring faculty from underrepresented minorities, and resulting in some prestigious universities no longer requiring diversity statements for job applicants [@guidenAreDiversityStatements].
 
@@
 cite these
-
-- https://www.anthropology-news.org/articles/hello-to-tristes-tropes/
 

### 2024-11-02 analysis/paper/article.qmd
writing discussion and conclusion

@@
 
-The academic job market is a source of uncertainty, both for job seekers, as they look for ads that are well-matched to their niche of skills and experience that have taken years to cultivate and refine, and for hiring committees as they sift through sometimes hundreds of applicants in the hope of finding a candidate that meets all of their needs.
+The academic job market is a source of uncertainty, especially for job seekers, as they look for ads that are well-matched to their niche of skills and experience that have taken years to cultivate and refine.
+Hiring committees also face uncertainty as they strive to equitably sift through sometimes hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realise their visions of their program and the discipline.
 For many academics there is a perception that the last two to three decades have seen the job market become increasingly difficult for job seekers.
@@
 Rising numbers of contingent faculty has long been a concern in US higher education [@Trevithick2010].
-Data from 2021–22 Faculty indicate that more than 60% of faculty positions in US universities were held by non-tenure track full time or part time contingent faculty members [@colby2022annual].
+Data from the American Association of University Professors' 2021–22 faculty survey indicate that more than 60% of faculty positions in US universities were held by non-tenure track full time or part time contingent faculty members [@colby2022annual].
 Contingent faculty positions are precarious because they provide low or no health and retirement benefits, little or nothing in the way of professional development, and lower salaries relative to tenure-track positions.
@@
 
-One way that some hiring committees are tackling these biases towards prestigious institutions is to provided detailed instructions to applicants on how to prepare their application materials so that the hiring committee can evaluate and rank the specific accomplishments of candidate across a variety of categories, rather than ranking based on prestige of the candidates' graduate schools.
-This has resulted in job ads that are highly prescriptive in the types of documents that applicants should submit.
+One way that some hiring committees are tackling these biases towards prestigious institutions is by providing detailed instructions to applicants on how to prepare their application materials so that the hiring committee can focus on evaluating and ranking the specific accomplishments of candidates across common categories, rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program.
+This has resulted in job ads that are often highly prescriptive in the types of documents that applicants should submit.
 For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements about their teaching and research record and plans.
@@
 They found that twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy.
-In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported two job ads with extreme requirements, requesting nine and thirteen documents from applicants.
+In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported two job ads with particularly extreme requirements, requesting nine and thirteen documents from applicants.
+This high burden on applicants disproportionately favors people with more time and financial resources to prepare the required materials.
+
 A notable requirement that has emerged in recent years is a diversity statement, where the applicant describes their knowledge of, prior contributions to, and future goals for advancing diversity, equity, inclusion.
-This requirement has been much-debated due to an experiment during 2016-2022 at several University of California campuses that used diversity statements as the first cut for selecting candidates for tenure-track faculty positions [@soucek2021diversity].
+This requirement has been much-debated largely due to an experiment in 2016-2022 at several University of California campuses that used diversity statements as the first cut for selecting candidates for tenure-track faculty positions [@soucek2021diversity].
 Only candidates that scored highly on their diversity statements would have the rest of their application evaluated.
-This experiment generated intense and widespread public debate about merits and risks of requiring and using diversity statements, drawing attention to the challenges of hiring faculty from underrepresented minorities, and resulting in some prestigious universities no longer requiring diversity statements for job applicants [@guidenAreDiversityStatements].
-
-<!--
-
-cite these
-
--->
+This experiment generated intense and widespread public debate about merits and risks of requiring and using diversity statements.
+These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted a variety of approaches to evaluating diversity, including some universities dropping requirements for diversity statements in job applications [@guidenAreDiversityStatements].
 
@@
 
-Landscape archaeology, although dominant has fluctuated over the years and has been on a downtrend since 2018-2019.
+Landscape archaeology, although dominant, has fluctuated over the years and has been on a downtrend since 2018-2019.
 Computational and digital archaeology is the second most represented method, showing an overall increasing trend, particularly since 2020-2021.
@@
 The demand for archaeology faculty, indicated by the total number of tenure-track jobs, maybe explained by a variety of processes.
-The downward trend in tenure-track positions during 2013-2019 related to decline in undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a].
+Overall we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-14.
+This is at odds with prior work finding that contingent positions make up more than half of faculty positions.
+This discrepancy may be due to advertising for short-term positions typically having much more limited circulation than advertising for tenure-track jobs, and thus less likely to be added to the Academic Jobs Wiki.
+The downward trend in tenure-track positions during 2013-2019 maybe related to decline in undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a].
 The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus.
@@
 Many universities may have hoped to hire Black and Indigenous archaeologists as part of their effort to tackle systemic racism.
-However as it is illegal in the US for universities use race or ethnicity as a primary factor in hiring (due to the 1964 Civil Rights Act), universities tailored the content of the job ads to focus on topics where Black and Indigenous researchers are most numerous, ensuring that these researchers would be well-represented in the pool of applicants.
+However as it is illegal in the US for universities use race or ethnicity as a primary factor in hiring (due to the 1964 Civil Rights Act), universities appear to have tailored the content of the job ads to focus on topics where they expect Black and Indigenous researchers to be most numerous, ensuring that these researchers would be well-represented in the pool of applicants.
 
@@
 This might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical methods for compositional analysis.
-A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience in the past may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
 Other relevant factors may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation, and an increasing emphasis on more job-oriented degrees that prepare students for working in cultural heritage management rather than advanced research.
+A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience which may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
 
 Our data on the requirements for applicants shows that application complexity has gradually increased over time.
-This trend is especially pronounced for Assistant Professor positions which make more demands on applicants than Associate and Full Professor positions.
-We saw a growing demand for research and teaching statements.
+This trend is especially pronounced for Assistant Professor positions, which make more demands on applicants than Associate and Full Professor positions.
+We saw a growing demand for research and teaching statements, consistent with previous work.
 Demand for diversity statements shows a unique trajectory, peaking in 2020-2021 and declining into the present.
-This may relate to the intersecting concerns emerging during the COVID-19 pandemic, which have declined in job ads as the pandemic moves into the past.
+This may relate to the intersecting concerns emerging during the COVID-19 pandemic, which may have declined in job ads as the pandemic moves into the past, in addition to the debates surrounding the experiment with diversity statements in hiring at some UC campuses.
 The decline in recent years for names of recommenders in the initial application may be a response to criticisms of the burden on the applicant of preparing numerous complex job applications [e.g.
@@
 
+Our results show that job ads are a highly nuanced record of disciplinary history, showing shifting trends in topics, methods, and geographic foci, with higher chronological resolution than many other sources, such as undergraduate textbooks [@lyman2010american].
+The short periods of time spanning the rise and fall of some of the foci we observed has implications for prospective graduate students.
+An area that is growing in popularity as a student begins their PhD may have peaked and be in decline before they graduate.
+
+Jobs ads also demonstrate how archaeologists draw on current events and politics to prioritize topics, regions and methods in their hiring plans.
+This is consistent with prior findings, such @wilk1985ancient's analysis of publications that found trends in explanations of Mayan collapse during the 1960s-70 that correspond with major contemporary American political and social issues.
+This responsiveness to current events reveals that many archaeologists accept that our narratives of the past are entangled in the social and political milieu in which we work.
+Furthermore, it shows an active effort by archaeologists to take control of interpreting the past.
+
 # Acknowledgements

### 2024-11-05 analysis/paper/article.qmd
discussion and conclusion

@@
 Similarly, the number of job ads with a geographic focus on the Americas and Africa peaks during this time.
-These shifts in the topics and geographic foci of job ads were likely been influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices.
+These shifts in the topics and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices.
 COVID-19 negatively impacted Black, American Indian and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US.
-The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic with the murder of George Floyd by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
+The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic with the murder of George Floyd in Minneapolis by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
 Mass protests objecting to Floyd's murder generated widespread concern about racial inequities and stimulated a broad interest in addressing systemic racial injustice.
@@
 Many universities may have hoped to hire Black and Indigenous archaeologists as part of their effort to tackle systemic racism.
-However as it is illegal in the US for universities use race or ethnicity as a primary factor in hiring (due to the 1964 Civil Rights Act), universities appear to have tailored the content of the job ads to focus on topics where they expect Black and Indigenous researchers to be most numerous, ensuring that these researchers would be well-represented in the pool of applicants.
+However as it is illegal in the US for universities use race or ethnicity as a primary factor in hiring (due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity), universities appear to have tailored the content of the job ads to focus on topics where they expect Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
 
@@
 We found a long-term decline in topics such as the archaeology of complex societies and archaeological science, Mesoamerica and South America as a geographic focus, and methods relating to landscape archaeology.
-These trends are harder to explain as we cannot link their start to a discrete moment like COVID-19.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism.
-This might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical methods for compositional analysis.
+These trends are harder to explain as we cannot link their start to a moment like COVID-19.
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism.
+This might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for compositional analysis.
 Other relevant factors may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation, and an increasing emphasis on more job-oriented degrees that prepare students for working in cultural heritage management rather than advanced research.
-A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience which may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
-
-Our data on the requirements for applicants shows that application complexity has gradually increased over time.
+A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience, which may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
+
+Our data on the requirements for applicants support prior findings that application complexity has gradually increased over time.
 This trend is especially pronounced for Assistant Professor positions, which make more demands on applicants than Associate and Full Professor positions.
@@
 Demand for diversity statements shows a unique trajectory, peaking in 2020-2021 and declining into the present.
-This may relate to the intersecting concerns emerging during the COVID-19 pandemic, which may have declined in job ads as the pandemic moves into the past, in addition to the debates surrounding the experiment with diversity statements in hiring at some UC campuses.
+This may relate to the intersecting concerns emerging during the COVID-19 pandemic, which may have declined in job ads as the most urgent time of the pandemic moves into the past.
+Another factor here may be the debates surrounding the experiment with diversity statements in hiring at some UC campuses during 2016-2022.
+Universities that previously required a diversity statement dropped that requirement after 2022 [@guidenAreDiversityStatements].
+In our data we observed one univeristy .
 The decline in recent years for names of recommenders in the initial application may be a response to criticisms of the burden on the applicant of preparing numerous complex job applications [e.g.

### 2024-11-05 analysis/paper/article.qmd
minor edit to conclusion

@@
 Instead we see the more popular methods here are ones that are relevant to multiple elements of the archaeological record (e.g.
-Archaeobotany encompasses macroscopic and microscopic plant remains; Bioarchaeology may include skeletal analysis, isotopes, protiens, etc.).
+Archaeobotany encompasses macroscopic and microscopic plant remains; Bioarchaeology may include skeletal analysis, isotopes, proteins, etc.).
 
@@
 #| label: fig-show-topi-trends
-#| fig-cap: "A: Frequency of methods mentioned in the text of the job ads.
-B: Popularity of methods in job ads over time for topics that appear in 20 or more ads.
+#| fig-cap: "A: Frequency of topics mentioned in the text of the job ads.
+B: Popularity of topics in job ads over time for topics that appear in 20 or more ads.
 Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
@@
 The short periods of time spanning the rise and fall of some of the foci we observed has implications for prospective graduate students.
-An area that is growing in popularity as a student begins their PhD may have peaked and be in decline before they graduate.
+An topic that is growing in popularity as a student begins their PhD may have peaked and be in decline before they graduate.
+Methods seem to have a much lower frequency of change in popularity, relative to topical and geographic foci.
+One possible implication of this is that a graduate student who has invested in developing technical expertise in a method during their studies, in addition to a topic and region, might be less exposed to the whims of the job market than a student without a distinct area of technical expertise.
 

### 2024-11-06 analysis/paper/article.qmd
minor edits

--- 2024-11-05 analysis/paper/article.qmd
@@
 This discrepancy may be due to advertising for short-term positions typically having much more limited circulation than advertising for tenure-track jobs, and thus less likely to be added to the Academic Jobs Wiki.
+Many of these short-term positions are not advertised nationally, but only through local email lists and are filled by people close to the hiring department, such as recently graduated students.
 The downward trend in tenure-track positions during 2013-2019 maybe related to decline in undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a].
@@
 Then we can assess how effective the hiring process is for driving change in the discipline by setting the topics of courses that will be taught in undergraduate and graduate curricula, and research that will be supported by universities.
+We also recognize the possibility that the Academic Jobs Wiki might not capture all available positions or might over-represent certain types of institutions.
 

### 2024-11-06 analysis/paper/article.qmd
minor edits

@@
 Using data from the community-edited Academic Jobs Wiki for Archaeology, we examine changes in the academic job market over time.
-We studied the text of 449 job ads posted from 2013-2023.
-Our analysis focuses on shifts in archaeological topics and methods requested in job ads.
+We studied the text of 431 job ads posted from 2013-2023.
+Our analysis focuses on shifts in archaeological topics, methods and geographic regions mentioned in job ads.
 We investigate whether the burden on applicants has changed over time: do institutions request more information and documents from applicants at the initial stages of application, compared to a decade ago?
-We also examine whether there is an increasing trend in job advertisements highlighting diversity and inclusivity, thereby encouraging a broader range of applicants.
-Additionally, we assess the influence of socio-political factors on the changing focus of research topics in the field.
+We assess the influence of socio-political factors on the changing focus of research topics in the field.
 This research aims to assist current and future archaeology students and graduates in better understanding the job market and the requirements of employers, thereby aiding them in effectively preparing for their applications for positions in archaeology.
@@
 
-# Introduction
-
-The aim of this paper is to explore the demand-side of the academic job market for archaeologists in the United States.
-We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.
-
-# Background
-
-The academic job market is a source of uncertainty, especially for job seekers, as they look for ads that are well-matched to their niche of skills and experience that have taken years to cultivate and refine.
-Hiring committees also face uncertainty as they strive to equitably sift through sometimes hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realise their visions of their program and the discipline.
-For many academics there is a perception that the last two to three decades have seen the job market become increasingly difficult for job seekers.
-In many fields it is common to hear that there are fewer jobs available relative to the number of PhD graduates, higher numbers of short-term temporary positions relative to secure tenure-track positions, and increasingly complex application requirements with specific statements (e.g.
-teaching, research, and diversity) that need to be tailed for each application, making the process of applying for jobs a full-time job of its own.
-
-In anthropology in the US the number of doctoral anthropology graduates has increased by about 70% over the past 30 years, but the number of new faculty positions has not increased proportionally [@speakman2018choosing].
-Among biological anthropologists, @Passalacqua_2018 found a ratio of 0.81 PhDs to job academic advertisements per year, concluding that academic positions in biological anthropology are barely at sustainable levels.
-This echos findings from other fields, for example in biomedical sciences where there is one tenure-track position in the US for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
-PhD graduates who do not go directly into a tenure-track faculty position after graduation often go into a series of short-term appointments as part-time or limited contract instructors, also known as adjunct or contingent faculty.
-Rising numbers of contingent faculty has long been a concern in US higher education [@Trevithick2010].
-Data from the American Association of University Professors' 2021–22 faculty survey indicate that more than 60% of faculty positions in US universities were held by non-tenure track full time or part time contingent faculty members [@colby2022annual].
-Contingent faculty positions are precarious because they provide low or no health and retirement benefits, little or nothing in the way of professional development, and lower salaries relative to tenure-track positions.
-PhD graduates who have a sequence of short-term academic positions are often disadvantaged financially, due to low compensation combined with expensive frequent relocation, socially due to isolation from family and community, and professionally due to being ineligible for many decision-making roles at the universities they work at [@platzerAcademicPrecarityAmerican2018].
-
-Studies of who holds tenure-track faculty positions in archaeology at US universities, shows that success in job applications is strongly determined by where applicants get their PhD from.
-Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing].
-Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
-This network analysis further found that programs with large endowments and with faculty who hold prestigious awards and are widely cited by other scholars produce the majority of tenured and tenure-track faculty in anthropology.
-This pattern of inequality is not unique to archaeology [@wapmanQuantifyingHierarchyDynamics2022a], but these studies draw attention to the paradox of a discipline that has many members committed to fighting social inequalities, and yet continues to reproduce its own systematic inequalities and hierarchies through hiring practices that favour an elite minority of graduates with prestigious institutional affiliations.
-
-One way that some hiring committees are tackling these biases towards prestigious institutions is by providing detailed instructions to applicants on how to prepare their application materials so that the hiring committee can focus on evaluating and ranking the specific accomplishments of candidates across common categories, rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program.
-This has resulted in job ads that are often highly prescriptive in the types of documents that applicants should submit.
-For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements about their teaching and research record and plans.
-In a comparison of job ads from 1999–2000 and 2019–2020 published in _Anthropology News_ @HelloTristesTropes2021 noticed a trend toward increasingly asking for more materials from applicants.
-They found that twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy.
-In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported two job ads with particularly extreme requirements, requesting nine and thirteen documents from applicants.
-This high burden on applicants disproportionately favors people with more time and financial resources to prepare the required materials.
-
-A notable requirement that has emerged in recent years is a diversity statement, where the applicant describes their knowledge of, prior contributions to, and future goals for advancing diversity, equity, inclusion.
-This requirement has been much-debated largely due to an experiment in 2016-2022 at several University of California campuses that used diversity statements as the first cut for selecting candidates for tenure-track faculty positions [@soucek2021diversity].
-Only candidates that scored highly on their diversity statements would have the rest of their application evaluated.
-This experiment generated intense and widespread public debate about merits and risks of requiring and using diversity statements.
-These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted a variety of approaches to evaluating diversity, including some universities dropping requirements for diversity statements in job applications [@guidenAreDiversityStatements].
-
-# Methods
-
-Our primary data source is the Archaeology Academic Jobs Wiki.
-Originating in 2007, this is a set of freely accessible web pages that anyone can edit (anonymously or with a free user account) hosted by Fandom, a for-profit company.
-The Archaeology pages are part of the Academic Jobs Wiki, which coordinates similar collaboratively-edited resources for around 40 academic disciplines.
-The coordinators and contributors are nearly all anonymous or pseudonymous.
-Typically contributors copy and paste the text of job ads from other sources, such as the _Chronicle of Higher Education_, _Higher Ed Jobs_, and university websites, into the wiki, collecting ads originally posted in numerous different locations.
-Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position, such as a tally of how many people have applied, the dates of events such as requests for more materials, interviews, offer made, rejection notices, etc.
-Contributors also edit the page to ask and answer questions about the positions and the application process.
-These comments make the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market.
-Because of its reputation for aggregating ads from diverse sources and rapidly-updated information that is not available elsewhere, the Academic Jobs Wiki has a large community of users that keep it updated and accurate has become an authoritative data source for studies of hiring trends in academia [e.g.
-@Passalacqua_2018; @musial2018five] and a widely recommended resource [e.g.
-@lightfoot2021preparing].
-
-For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023 we read the text and recorded into a Google form the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methods foci on the position.
-The topical focus is what we understood as the primary intellectual focus of the position.
-The geographic focus is the region of the world that the ideal candidate has scholarly expertise on.
-The methods focus is the data-generating sub-field of archaeology that is mentioned in the ad.
-We recorded the type and number of documents requested in each ad (e.g.
-cover letter, CV, statements on research, teaching, diversity, syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad.
-
-After completing primary data collection, we studied the topical, geographic, and methods foci of each ad and collaboratively and manually reduced the variation in the raw data into 10-15 categories appearing in 20 (for topics and geography) or 10 (for methods) or more job ads to simplify analysis and visualisation.
-Full details of the category reduction are in our Supplementary Materials.
-Our topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
-Our methods categories were: Archaeobotany, Archaeometry, Bioarchaeology, Ceramic analysis, Computational and Digital archaeology, Geoarchaeology, Landscape analysis, Lithic analysis, Material culture analysis, and Zooarchaeology.
-Ads could have multiple or none of these three foci, and some of the foci overlap.
-Some topics include geographic regions because this is how they are typically understood by archaeologists.
-For example Mesoamerican archaeology is understood to refer to a specific time period and geographic region.
-Similarly, digital archaeology we recorded as both a method (when a job ad has a clearly distinct topical focus, such as historic archaeology) and a topic (when there is no other topics mentioned in the job ad).
-While these overlaps can make the data challenging to interpret, in our view it reflects the complex realities of  how search committees express their needs in searching for new faculty, and is insightful in how it reveals intersections between different foci.
-
-The entire R code (R Core Team, 2021) and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx to enable re-use of materials and improve reproducibility and transparency (Marwick, 2017).
-All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
-The code is released under the MIT license, the data as CC-0, and the figures as CC-BY, to enable maximum re-use.
-
-# Results
-
 ```{r}
@@
 base_size <- 12
+
+n_jobs_mentioning_gender <- 
+jobdata %>% 
+  select(topical_focus_of_position) %>% 
+      filter(str_detect(topical_focus_of_position, 
+                        "gender")) %>% 
+  nrow()
 ```
@@
 
-year_ad_posted <- 
+year_ad_posted_vec <- 
 jobdata %>% 
@@
   jobdata %>% 
-  mutate(year_ad_posted = year_ad_posted) %>% 
+  mutate(year_ad_posted = year_ad_posted_vec) %>% 
   drop_na(year_ad_posted)
@@
 
-We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023.
+# Introduction
+
+The aim of this paper is to explore the demand-side of the academic job market for archaeologists in the United States.
+We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.
+
+# Background
+
+The academic job market is a source of uncertainty, especially for job seekers, as they look for ads that are well-matched to their niche of skills and experience that have taken years to cultivate and refine.
+Hiring committees also face uncertainty as they strive to equitably sift through sometimes hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realise their visions of their program and the discipline.
+For many academics there is a perception that the last two to three decades have seen the job market become increasingly difficult for job seekers.
+In many fields it is common to hear that there are fewer jobs available relative to the number of PhD graduates, higher numbers of short-term temporary positions relative to secure tenure-track positions, and increasingly complex application requirements with specific statements (e.g.
+teaching, research, and diversity) that need to be tailed for each application, making the process of applying for jobs a full-time job of its own.
+
+In anthropology in the US the number of doctoral anthropology graduates has increased by about 70% over the past 30 years, but the number of new faculty positions has not increased proportionally [@speakman2018choosing].
+Among biological anthropologists, @Passalacqua_2018 found a ratio of 0.81 PhDs to job academic advertisements per year, concluding that academic positions in biological anthropology are barely at sustainable levels.
+This echos findings from other fields, for example in biomedical sciences where there is one tenure-track position in the US for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
+PhD graduates who do not go directly into a tenure-track faculty position after graduation often go into a series of short-term appointments as part-time or limited contract instructors, also known as adjunct or contingent faculty.
+Rising numbers of contingent faculty has long been a concern in US higher education [@Trevithick2010].
+Data from the American Association of University Professors' 2021–22 faculty survey indicate that more than 60% of faculty positions in US universities were held by non-tenure track full time or part time contingent faculty members [@colby2022annual].
+Contingent faculty positions are precarious because they provide low or no health and retirement benefits, little or nothing in the way of professional development, and lower salaries relative to tenure-track positions.
+PhD graduates who have a sequence of short-term academic positions are often disadvantaged financially, due to low compensation combined with expensive frequent relocation, socially due to isolation from family and community, and professionally due to being ineligible for many decision-making roles at the universities they work at [@platzerAcademicPrecarityAmerican2018].
+
+Studies of who holds tenure-track faculty positions in archaeology at US universities, shows that success in job applications is strongly determined by where applicants get their PhD from.
+Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing].
+Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
+This network analysis further found that programs with large endowments and with faculty who hold prestigious awards and are widely cited by other scholars produce the majority of tenured and tenure-track faculty in anthropology.
+This pattern of inequality is not unique to archaeology [@wapmanQuantifyingHierarchyDynamics2022a], but these studies draw attention to the paradox of a discipline that has many members committed to fighting social inequalities, and yet continues to reproduce its own systematic inequalities and hierarchies through hiring practices that favour an elite minority of graduates with prestigious institutional affiliations.
+
+One way that some hiring committees are tackling these biases towards prestigious institutions is by providing detailed instructions to applicants on how to prepare their application materials so that the hiring committee can focus on evaluating and ranking the specific accomplishments of candidates across common categories, rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program.
+This has resulted in job ads that are often highly prescriptive in the types of documents that applicants should submit.
+For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements about their teaching and research record and plans.
+In a comparison of job ads from 1999–2000 and 2019–2020 published in _Anthropology News_ @HelloTristesTropes2021 noticed a trend toward increasingly asking for more materials from applicants.
+They found that twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy.
+In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported two job ads with particularly extreme requirements, requesting nine and thirteen documents from applicants.
+This high burden on applicants disproportionately favors people with more time and financial resources to prepare the required materials.
+
+A notable requirement that has emerged in recent years is a diversity statement, where the applicant describes their knowledge of, prior contributions to, and future goals for advancing diversity, equity, inclusion.
+This requirement has been much-debated largely due to an experiment in 2016-2022 at several University of California campuses that used diversity statements as the first cut for selecting candidates for tenure-track faculty positions [@soucek2021diversity].
+Only candidates that scored highly on their diversity statements would have the rest of their application evaluated.
+This experiment generated intense and widespread public debate about merits and risks of requiring and using diversity statements.
+These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted a variety of approaches to evaluating diversity, including some universities dropping requirements for diversity statements in job applications [@guidenAreDiversityStatements].
+
+# Methods
+
+Our primary data source is the Archaeology Academic Jobs Wiki.
+Originating in 2007, this is a set of freely accessible web pages that anyone can edit (anonymously or with a free user account) hosted by Fandom, a for-profit company.
+The Archaeology pages are part of the Academic Jobs Wiki, which coordinates similar collaboratively-edited resources for around 40 academic disciplines.
+The coordinators and contributors are nearly all anonymous or pseudonymous.
+Typically contributors copy and paste the text of job ads from other sources, such as the _Chronicle of Higher Education_, _Higher Ed Jobs_, and university websites, into the wiki, collecting ads originally posted in numerous different locations.
+Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position, such as a tally of how many people have applied, the dates of events such as requests for more materials, interviews, offer made, rejection notices, etc.
+Contributors also edit the page to ask and answer questions about the positions and the application process.
+These comments make the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market.
+Because of its reputation for aggregating ads from diverse sources and rapidly-updated information that is not available elsewhere, the Academic Jobs Wiki has a large community of users that keep it updated and accurate has become an authoritative data source for studies of hiring trends in academia [e.g.
+@Passalacqua_2018; @musial2018five] and a widely recommended resource [e.g.
+@lightfoot2021preparing].
+
+For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023 we read the text and recorded into a Google form the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methods foci on the position.
+The topical focus is what we understood as the primary intellectual focus of the position.
+The geographic focus is the region of the world that the ideal candidate has scholarly expertise on.
+The methods focus is the data-generating sub-field of archaeology that is mentioned in the ad.
+We recorded the type and number of documents requested in each ad (e.g.
+cover letter, CV, statements on research, teaching, diversity, syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad.
+
+After completing primary data collection, we studied the topical, geographic, and methods foci of each ad and collaboratively and manually reduced the variation in the raw data into 10-15 categories appearing in 20 (for topics and geography) or 10 (for methods) or more job ads to simplify analysis and visualisation.
+This means that some topics, such as gender (mentioned in `r n_jobs_mentioning_gender` ads) do not appear in our results because of their rarity in the job ads.
+Full details of the category reduction, showing the mapping between phrases in the job ads and our categories, are in our Supplementary Materials.
+Our topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
+Our methods categories were: Archaeobotany, Archaeometry, Bioarchaeology, Ceramic analysis, Computational and Digital archaeology, Geoarchaeology, Landscape analysis, Lithic analysis, Material culture analysis, and Zooarchaeology.
+Ads could have multiple or none of these three foci, and some of the foci overlap.
+Some topics include geographic regions because this is how they are typically understood by archaeologists.
+For example Mesoamerican archaeology is understood to refer to a specific time period and geographic region.
+Similarly, digital archaeology we recorded as both a method (when a job ad has a clearly distinct topical focus, such as historic archaeology) and a topic (when there is no other topics mentioned in the job ad).
+While these overlaps can make the data challenging to interpret, in our view it reflects the complex realities of  how search committees express their needs in searching for new faculty, and is insightful in how it reveals intersections between different foci.
+
+The entire R code (R Core Team, 2021) and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx to enable re-use of materials and improve reproducibility and transparency (Marwick, 2017).
+All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
+The code is released under the MIT license, the data as CC-0, and the figures as CC-BY, to enable maximum re-use.
+
+# Results
+
+We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023, and focus our analysis on the 431 ads for positions at US universities.
 @fig-show-basic-plots shows the count of ads for each year.
@@
 # how many jobs ads now?
-# sum(uni_state_country_tally_us$n) # 433, 78% of the total
+# sum(uni_state_country_tally_us$n) # 431, 78% of the total
 
@@
 No institutions in Montana posted a job ad during this period.
+The implication here is that job-seekers who are able to relocate to populous areas will have more employment options.
 
@@
 We recorded all geographic regions mentioned in the text where the successful applicant should have expertise and be research active.
-Our analysis focuses on those locations mentioned in 20 or more ads Overall, American locations dominate.
+Our analysis focuses on those locations mentioned in 20 or more ads.
+Overall, American locations dominate.
 Panel A of @fig-show-geo-trends shows that a single region of the US, the Southwest, occurs in more job ads than every other part of the world except for the Mediterranean.
@@
 
+```{r}
+schools_that_dropped_ds_req <-     
+  jobdata %>% 
+  dplyr::select(
+    year_ad_posted,
+    name_of_hiring_university, 
+    documents_requested_diversity_statement) %>% 
+  mutate(ds = case_when(
+    documents_requested_diversity_statement == "Not requested in the job ad" ~ FALSE,
+    documents_requested_diversity_statement != "Not requested in the job ad" ~ TRUE
+  )) %>% 
+    group_by(name_of_hiring_university) %>%
+    mutate(year = parse_number(str_sub(year_ad_posted, 1, 4))) %>% 
+    ungroup() %>% 
+    arrange(name_of_hiring_university, desc(year))  %>% 
+    group_by(name_of_hiring_university) %>% 
+    filter(n() >= 2) %>% 
+    select(name_of_hiring_university,
+           year,
+           documents_requested_diversity_statement) %>% 
+    mutate(documents_requested_diversity_statement = 
+             ifelse(documents_requested_diversity_statement == "One", "1", "0"),
+           documents_requested_diversity_statement = parse_number(documents_requested_diversity_statement)) %>% 
+    pivot_wider(names_from = year,
+                values_from = documents_requested_diversity_statement,
+                values_fn = ~sum(.x, na.rm = TRUE)) %>% 
+    filter(`2022` == "0") %>% 
+    mutate(prev_req_ds = rowSums(across(`2015`:`2019`), na.rm = TRUE)) %>% 
+    select(name_of_hiring_university, 
+           `2022`,
+           prev_req_ds) %>% 
+    arrange(desc(prev_req_ds))
+
+n_schools_that_dropped_ds_req <- 
+  schools_that_dropped_ds_req %>% 
+  filter(prev_req_ds == 1) %>% 
+  nrow()
+
+```
+
 Our data on the requirements for applicants support prior findings that application complexity has gradually increased over time.
@@
 Universities that previously required a diversity statement dropped that requirement after 2022 [@guidenAreDiversityStatements].
-In our data we observed one univeristy .
-The decline in recent years for names of recommenders in the initial application may be a response to criticisms of the burden on the applicant of preparing numerous complex job applications [e.g.
+In our data we observed a reversal of the diversity statement requirement at `r n_schools_that_dropped_ds_req` universities.
+Those schools posted job ads prior 2022 that did require a diversity statement and also posted an ad in 2022 that did not require one.
+
+One bright spot for applicants is the decline in recent years for requests for names of recommenders in the initial application.
+This may be a response to recent criticisms of the burden on the applicant of preparing numerous complex job applications [e.g.
 @dennisWorstAnthroJob2022].
@@
 We also recognize the possibility that the Academic Jobs Wiki might not capture all available positions or might over-represent certain types of institutions.
+Future work should evaluate this by comparing entries on the Wiki to other sources of information about the academic job market, such as the American Anthropological Association's AnthroGuide or data collected directly from universities.
 

### 2024-11-15 analysis/paper/article.qmd
confirm focus on US job ads only

@@
     theme(legend.position = c(0.5, 0.85)) +
-  guides(x = "none")
+  guides(x = "none") +
+  ylim(0,100)
+
+```
+
+```{r}
+# get US job ads only 
+state_to_st <- function(x){
+       c(state.abb, 'DC')[match(x, c(state.name, 'District of Columbia'))]
+}
+
+state_name_and_abb <-
+enframe(state.name, value = 'state_name') %>%
+       mutate(state_abbr = state_to_st(state_name))
+
+jobdata_us_only <- 
+jobdata %>% 
+  mutate(state_country = regmatches(name_of_hiring_university,
+                                    gregexpr( "(?<=\\().+?(?=\\))",
+                                              name_of_hiring_university,
+                                              perl = T))) %>%
+  unnest(state_country) %>% 
+  filter(state_country %in% state.abb)
+
+fig_prop_by_job_title_per_year_us <- 
+jobdata_us_only %>%
+  group_by(year_ad_posted) %>%
+  count(job_title_simple) %>%
+  mutate(prop = n / sum(n)) %>%
+  ggplot() +
+  aes(year_ad_posted,
+      n,
+      group = job_title_simple,
+      fill = job_title_simple) +
+ # geom_line(linewidth = 2) +
+  geom_col() +
+ theme_minimal(base_size = base_size) +
+  xlab("") +
+  ylab("Number of\ntenure track job ads") +
+  scale_fill_brewer(palette = "Dark2") +
+  guides(fill = guide_legend(nrow=2,
+                               byrow=TRUE,
+                               "Job title",
+                                keywidth=0.1,
+                 keyheight=0.1,
+                 default.unit="inch")) +
+    theme(legend.position = c(0.5, 0.85)) +
+  guides(x = "none") +
+  ylim(0,100)
+
+total_number_of_us_ads_in_our_sample <- nrow(jobdata_us_only)
+
+# all of our analysis is on the US jobs ads only
+jobdata <- jobdata_us_only
 
@@
 
-We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023, and focus our analysis on the 431 ads for positions at US universities.
+We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023, and focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities.
 @fig-show-basic-plots shows the count of ads for each year.
@@
 # sum(uni_state_country_tally$n) # 550 all of them
-
-state_to_st <- function(x){
-       c(state.abb, 'DC')[match(x, c(state.name, 'District of Columbia'))]
-}
-
-state_name_and_abb <-
-enframe(state.name, value = 'state_name') %>%
-       mutate(state_abbr = state_to_st(state_name))
 
@@
 library(fiftystater)
+# install.packages("devtools")
+# devtools::install_github("wmurphyrd/fiftystater")
 library(tidyverse)

### 2024-12-20 analysis/paper/article.qmd
In which Jess experiences the horrors of Quarto and GitHub

@@
 
+```{=html}
 <!-- Instructors to authors for AAQ
@@
 -->
+```
 
@@
 
-The aim of this paper is to explore the demand-side of the academic job market for archaeologists in the United States.
+The aim of this paper is to TEACH JESS HOW TO USE QUARTO AND GITHUB, UNFORTUNATELY FOR JESS.
 We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.
@@
 For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements about their teaching and research record and plans.
-In a comparison of job ads from 1999–2000 and 2019–2020 published in _Anthropology News_ @HelloTristesTropes2021 noticed a trend toward increasingly asking for more materials from applicants.
+In a comparison of job ads from 1999–2000 and 2019–2020 published in *Anthropology News* @HelloTristesTropes2021 noticed a trend toward increasingly asking for more materials from applicants.
 They found that twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy.
@@
 The coordinators and contributors are nearly all anonymous or pseudonymous.
-Typically contributors copy and paste the text of job ads from other sources, such as the _Chronicle of Higher Education_, _Higher Ed Jobs_, and university websites, into the wiki, collecting ads originally posted in numerous different locations.
+Typically contributors copy and paste the text of job ads from other sources, such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, into the wiki, collecting ads originally posted in numerous different locations.
 Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position, such as a tally of how many people have applied, the dates of events such as requests for more materials, interviews, offer made, rejection notices, etc.

### 2024-12-20 analysis/paper/article.qmd
In which the horrors include being uncertain about who exactly I am

--- 2024-12-20 analysis/paper/article.qmd
@@
+gitcreds::gitcreds_set()
+library(usethis)
+use_git_config(user.name = "Jess Beck")
 ---
@@
 The aim of this paper is to TEACH JESS HOW TO USE QUARTO AND GITHUB, UNFORTUNATELY FOR JESS.
-We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.
+<!-- It would be nice if you could make comments in the visual mode and not just in source mode, unfortunate --> We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.
 

### 2024-12-20 analysis/paper/article.qmd
stash

--- 2024-12-20 analysis/paper/article.qmd
@@
 
-The aim of this paper is to TEACH JESS HOW TO USE QUARTO AND GITHUB, UNFORTUNATELY FOR JESS.
+The aim of this paper is to explore the demand-side of the academic job market for archaeologists in the United States.
 We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.
@@
 Studies of who holds tenure-track faculty positions in archaeology at US universities, shows that success in job applications is strongly determined by where applicants get their PhD from.
-Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing].
+<!-- here is a comment in the middle --> Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing].
 Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].

### 2024-12-20 analysis/paper/article.qmd
resolve merge conflict
clean up copy-paste, test html comment in visual model

--- 2024-12-20 analysis/paper/article.qmd
@@
-gitcreds::gitcreds_set()
-library(usethis)
-use_git_config(user.name = "Jess Beck")
 ---
@@
 <!-- It would be nice if you could make comments in the visual mode and not just in source mode, unfortunate --> We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.
+<!-- yes we can make and view comments in visual mode, my mistake to think it was not possible -->
 

### 2025-01-08 analysis/paper/article.qmd

@@
 
-The aim of this paper is to TEACH JESS HOW TO USE QUARTO AND GITHUB, UNFORTUNATELY FOR JESS.
-<!-- It would be nice if you could make comments in the visual mode and not just in source mode, unfortunate --> We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.
-<!-- yes we can make and view comments in visual mode, my mistake to think it was not possible -->
+<!--Ben, please note that for now I am keeping track of my references.
+I have bolded them in text below; I think I will want to learn how to use Zotero and incorporate the citations using your arcane Lovecraftian ritual language (BibTex?), so that for our next paper I am capable of participating in the entire workflow with less supervision-->
+
+Attend any workshop, conference, or panel that includes early career researchers and the conversation will be steered inexorably towards the academic job market: who is hiring, who has attained a tenure-track position, and who is out of luck this season.
+This obsessive focus on career trajectories is warranted, particularly within the discipline of anthropology.
+Between 1985 and 2014, the number of PhDs produced by American anthropology departments increased by 70% (**Speakman et al., 2018a**), but the number of tenure-track positions advertised has stagnated or declined*.* As of2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions (**AAUP 2022**); today, 71% of faculty in the US are non-tenure-track(**Culver and Kezar 2022***).* The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors (**Beck, under review**), including decreasing federal support for higher education, concomitant increases in student debt (**Gusterson 2017**), a pronounced shift in university investment from faculty to administration (Graeber 2018: 162–163), and the growing privatization and market orientation of scientific research (**Mirowski 2011**).
+Whatever the cause, research within anthropology has consistently shown that at most 20% of PhDs in the discipline are obtaining jobs in American anthropology departments (**Speakman et al.
+2018b**), while the top-ranking programmes are placing fewer than one in three graduates in tenure-track jobs (**Mackie and Rockwell 2023**).
+
+These cultural and economic shifts are also well encapsulated by anecdata; any conversation with a junior scholar will reveal the impact of these larger professional transformations on personal and professional trajectories.
+The first author on this paper has applied to five academic positions over the course of his career, while the last author has applied to one hundred.
+A twenty-fold increase in the number of applications required to obtain a workable long-term academic position may seem preposterous, but these changing requirements for applicants are familiar to anyone who has obtained their PhDs since the Great Recession.
+This competitive landscape may have deeper roots than the last economic crisis; writing more than three decades ago, Rabinow pinpointed “an awareness… that standards have changed during the last thirty years and the quantitative and qualitative demands for entry into the system are immeasurably higher now” (**1992: 66**).
+
+These higher demands include increased overall productivity, higher mobility, sometimes requiring bothnational and international moves, more time spent in precarious short-term contracts, and the investment of a greater proportion of time and energy in applying for long-term positions.
+In their survey of early career researchers in European archaeology, for example, Brami et al.
+found that 52% of respondents who had obtained their PhD at least two years before the survey were on fixed-term contracts, while only 9% were in permanent positions (**2023: Figure 5**).
+In their words, “the queue of eager post-docs hoping for a long-term appointment is getting longer, as is the average time spent in the ‘queue’” (**2023: 239**).
+Ribeiro and Giamakis emphasize that this demographic comprises an academic precariat that is essential to the functioning of universities, but consists of a workforce on short-term contracts perpetually “stranded between employment and unemployment” (**2023:10**).
+
+Early career scholars with faculty ambitions are thus stuck in a double-bind, forced to funnel their energy into seemingly endless applications for which they receive little to no feedback and which remain invisible on their CVs.
+As Dennis et al.
+underscore, early career researchers are “supplicants”: “There is no way for applicants to point out that the job ads are taking too much time out of the scholarly community’s collective time bank” (**2022: 1**).
+Indeed, acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs to performing well in Zoom interviews and teaching demonstrations.
+These courses often provide valuable training in the hidden curriculum of the academy, but that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the “market” as a historically contingent set of cultural practices.
+As Rabinow maintains, “if we want ethical considerations to play a central role in the articulation of truth and power—and I think we do—then bringing such considerations into view is the necessary first step towards recognizing who we are today and setting out on the road to a better place" (**1992: 71**) **.**
+
+We argue that to understand the dynamics of hiring in academic archaeology, we must start at the beginning, with the job ad itself.
+There is an abundance of recent work on who does and does not get hired into tenure-track positions in anthropology and archaeology (Kawa 2019; Mackie and Rockwell 2023), but less scholarship on how those hires unfold.
+Research on hiring dynamics has amply demonstrated that *where* you get your PhD has an enormous impact on if and where you get a tenure-track position (Kawa et al.
+2019; Mackie and Rockwell 2023), but how about *what* you study?
+Are there topical, methodological, or regional specialties that are in consistent demand, or is there continual flux generated by intellectual fads, economic incentives, and political attention?
+Are job applications more demanding now than they were in the past, or is this a specious perception generated by the sheer volume of applications necessary to obtain a long-term academic position?
+
+To answer these questions, this paper explores the demand-side of the academic job market for archaeologists in the United States.
+Our study had two aims: first, to determine if disciplinary trends could be discerned in the topical, geographic, and methodological foci of the positions advertised over a ten-year period, and second, to investigate how requirements for applicants have changed over time.
 

### 2025-01-08 analysis/paper/article.qmd
First round of Jess edits to body of the paper.

@@
 abstract: |
-  Academic careers are frequently sought by archaeology graduate students.
+  Many archaeology graduate students pursue advanced degrees in the hope of landing an academic career.
 Job listing websites often serve as the first place for these students when seeking academic positions.
@@
 Using data from the community-edited Academic Jobs Wiki for Archaeology, we examine changes in the academic job market over time.
-We studied the text of 431 job ads posted from 2013-2023.
-Our analysis focuses on shifts in archaeological topics, methods and geographic regions mentioned in job ads.
-We investigate whether the burden on applicants has changed over time: do institutions request more information and documents from applicants at the initial stages of application, compared to a decade ago?
+We analyzed the text of 431 job ads posted from 2013--2023.
+Our analysis focuses on shifts in archaeological topics, methods and geographic regions specified in archaeological job ads.
+We also investigate whether the labor burden for applicants has changed over time: do institutions request more information and documents from applicants at the initial stages of application, compared to a decade ago?
 We assess the influence of socio-political factors on the changing focus of research topics in the field.
@@
 #| fig-cap: "A: total number of job ads posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank.
-B: Ratio of tenure-track to non-tenure-track positions over time."
+B: Ratio of tenure-track to non-tenure-track positions over time.The proportion of the bar above the red line reflects the number of non-tenure track positions available relative to tenure-track positions."
+#JB: I have added the last sentence above because I have not seen this kind of figure in an archaeology paper, so I think we should explain it.
+Will reduce friction for our readers.Please confirm that I am actually correctly explaining how the figure works, however.
+#JB: I would be eternally grateful if you would convert the hyphens between the timespans on the x-axis into en dashes 😬
 
@@
             "fig-panel-per-year.png"))
+
 ```
@@
 # Introduction
+
+```{=html}
+<!--Comments on Abstract and Keywords: 
+1. Is “jobs listing website” the appropriate category here? I often hear these referred to as “jobs boards”—that might be a better fit?
+2.
+Structurally, the second research domain is phrased as a question.
+For consistency, it would be good to provide a similar structure for the first domain, e.g.
+“Our analysis focuses on the XXX—have the topics, methods, or geographic regions specified in archaeological job ads changed over time?”
+3. Suggested keywords: archaeology, hiring, academia, job ads-->
+```
 
@@
 
-The academic job market is a source of uncertainty, especially for job seekers, as they look for ads that are well-matched to their niche of skills and experience that have taken years to cultivate and refine.
-Hiring committees also face uncertainty as they strive to equitably sift through sometimes hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realise their visions of their program and the discipline.
-For many academics there is a perception that the last two to three decades have seen the job market become increasingly difficult for job seekers.
-In many fields it is common to hear that there are fewer jobs available relative to the number of PhD graduates, higher numbers of short-term temporary positions relative to secure tenure-track positions, and increasingly complex application requirements with specific statements (e.g.
-teaching, research, and diversity) that need to be tailed for each application, making the process of applying for jobs a full-time job of its own.
+The academic job market is a source of uncertainty, for both job seekers and hiring committees.
+Job seekers must search for ads that are well-matched to the skill sets and professional experiences that they have taken years to cultivate.
+Hiring committeesmust sift through tens to hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realize<!--We should go with American spelling, no?
+With your Australian inclinations and my growing incorporation of Irish orthography into my writing, this paper is gonna be a mess...--> their visions of their program and the discipline.
+Finding the right candidate for the position—or the right position for the candidate—is akin to finding a needle in a haystack.
+
+There is also a perception among many scholars that the last three decades have seen the academic job market become increasingly competitive.
+Across the social sciences and humanities, there are fewer jobs available relative to the number of PhD graduates and higher numbers of short-term positions relative to permanent positions (**Bessner and Brenes 2021; Brami et al., 2023; Kawa et al.
+2019; Kelsky 2015; Mackie and Rockwell 2023**).
+This market saturation has arisen in tandem with increasingly complex application requirements.
+Many job ads now solicit specific documents in *addition* to the cover letter and CV, such as teaching, research, and diversity statements.
+These additional documents must be tailored for each application, making the process of applying for jobs a full-time job of its own.
 
 In anthropology in the US the number of doctoral anthropology graduates has increased by about 70% over the past 30 years, but the number of new faculty positions has not increased proportionally [@speakman2018choosing].
+New faculty positions have dwindled, in part due to the removal of the of the Age Discrimination in Employment Act (ADEA) exemption in 1994 which prohibited mandatory retirement ages in higher education (**Earle and Delpo Kulow 2015**).
+When combined with the institution of tenure in American academia, the ADEA exemption allowed [faculty to stay in their posts for as long as they liked] the median age for faculty in the U.S.
+now ranks among the highest for all professions (**Citation**).
+In tandem with the gradual de-investment in American higher education since the **XXX (Mirowski**) and the aftershocks of the 2008 recession, the ADEA exemption has cultivated an environment where new lines are few and far between.
 Among biological anthropologists, @Passalacqua_2018 found a ratio of 0.81 PhDs to job academic advertisements per year, concluding that academic positions in biological anthropology are barely at sustainable levels.
-This echos findings from other fields, for example in biomedical sciences where there is one tenure-track position in the US for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
+This echoes findings from other fields.
+In biomedical sciences, for example, there is one tenure-track position in the US for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
+**Citation of Larson et al.
+2014**.
+
 PhD graduates who do not go directly into a tenure-track faculty position after graduation often go into a series of short-term appointments as part-time or limited contract instructors, also known as adjunct or contingent faculty.
@@
 
-Studies of who holds tenure-track faculty positions in archaeology at US universities, shows that success in job applications is strongly determined by where applicants get their PhD from.
+Success in tenure-track job applications in archaeology is strongly determined by where applicants get their PhDs.
 Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing].
 Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
-This network analysis further found that programs with large endowments and with faculty who hold prestigious awards and are widely cited by other scholars produce the majority of tenured and tenure-track faculty in anthropology.
-This pattern of inequality is not unique to archaeology [@wapmanQuantifyingHierarchyDynamics2022a], but these studies draw attention to the paradox of a discipline that has many members committed to fighting social inequalities, and yet continues to reproduce its own systematic inequalities and hierarchies through hiring practices that favour an elite minority of graduates with prestigious institutional affiliations.
-
-One way that some hiring committees are tackling these biases towards prestigious institutions is by providing detailed instructions to applicants on how to prepare their application materials so that the hiring committee can focus on evaluating and ranking the specific accomplishments of candidates across common categories, rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program.
-This has resulted in job ads that are often highly prescriptive in the types of documents that applicants should submit.
-For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements about their teaching and research record and plans.
-In a comparison of job ads from 1999–2000 and 2019–2020 published in *Anthropology News* @HelloTristesTropes2021 noticed a trend toward increasingly asking for more materials from applicants.
-They found that twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy.
-In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported two job ads with particularly extreme requirements, requesting nine and thirteen documents from applicants.
+This network analysis showedthat programs with large endowments and with faculty who hold prestigious awards and are widely cited by other scholars produce the majority of tenured and tenure-track faculty in anthropology.
+Hiring bias predicated on the prestige of specific institutions and programs is not unique to anthropology.
+Targeted studies of the disciplines of sociology, communication, and operations research, and industrial systems engineering show similar trends  (**Barnett et al., 2010; Castillo et al., 2018; Feeley and Tutzauer 2022; Nevin 2019**).
+**Broader analyses comparing X, Y and Z, and analyses while a recent analysis of XXX showed Z (Clauset et al., 2015;** [@wapmanQuantifyingHierarchyDynamics2022a]).
+Within anthropology, however, such studies draw attention to the paradox of a discipline committed to fighting social inequalities continuing to reproduce its own systematic inequalities through hiring practices that favor an elite minority of graduates with prestigious institutional affiliations.
+
+One way that some hiring committees are tackling these biases towards prestigious institutions is by providing detailed instructions to applicants on how to prepare their application materials.
+In theory, detailed applications will allow the hiring committee to focus on evaluating the accomplishments of candidates across common categories, rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program.
+This push for equity has resulted in job ads that are often highly prescriptive in the types of documents that applicants should submit.
+For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements detailing their previous and future contributions to teaching, research, and diversity within their discipline.
+In a comparison of job ads from 1999–2000 and 2019–2020 published in *Anthropology News* @HelloTristesTropes2021 noticed an increase in the number of materials requested from applicants.
+For example, twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy.
+In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported two job ads with particularly extreme requirements<!--Are you opposed to naming and shaming these institutions, which were Oberlin College and Grinnell College?
+I applied for one of these positions, and have no issue referring to them by name...-->, which respectively requested nine and thirteen documents from applicants.
 This high burden on applicants disproportionately favors people with more time and financial resources to prepare the required materials.
@@
 A notable requirement that has emerged in recent years is a diversity statement, where the applicant describes their knowledge of, prior contributions to, and future goals for advancing diversity, equity, inclusion.
-This requirement has been much-debated largely due to an experiment in 2016-2022 at several University of California campuses that used diversity statements as the first cut for selecting candidates for tenure-track faculty positions [@soucek2021diversity].
+This requirement has been much-debated largely due to an experiment in 2016–2022 at several University of California campuses that used diversity statements as the first cut for selecting candidates for tenure-track faculty positions [@soucek2021diversity].
 Only candidates that scored highly on their diversity statements would have the rest of their application evaluated.
-This experiment generated intense and widespread public debate about merits and risks of requiring and using diversity statements.
-These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted a variety of approaches to evaluating diversity, including some universities dropping requirements for diversity statements in job applications [@guidenAreDiversityStatements].
+This experiment generated intense and widespread public debate about the merits and risks of requiring and using diversity statements.
+These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted in a variety of approaches to evaluating diversity, even leading some universities to entirely omit requirements for diversity statements in job applications [@guidenAreDiversityStatements].
+
+**While attention to equity in job market materials is commendable, work on hiring practices suggests ZZZ.
+For example, Warren 2019 has shown XXX.
+Rivera, in a study of hiring practices at YYY, has likewise demonstrated ZZZ.
+Somewhere in this section I want to discuss the results from sociology, specifically the Warren 2019 piece and the Rivera 2017 piece on hiring practices.**
+
+**Cite Rogge 1976.**
 
@@
 
-Our primary data source is the Archaeology Academic Jobs Wiki.
+Our primary data source is the Archaeology Academic Jobs Wiki<!--Check if Kelsky has anything too say about this.
+I also think there have been some articles about the pros and cons of the poli sci or Econ versions of this.-->.
 Originating in 2007, this is a set of freely accessible web pages that anyone can edit (anonymously or with a free user account) hosted by Fandom, a for-profit company.
@@
 The coordinators and contributors are nearly all anonymous or pseudonymous.
-Typically contributors copy and paste the text of job ads from other sources, such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, into the wiki, collecting ads originally posted in numerous different locations.
+Typically, contributors copy and paste the text of job ads from other sources, such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, into the wiki, collecting <!--consolidating?
+centralizing?--> ads originally posted in numerous different locations.
 Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position, such as a tally of how many people have applied, the dates of events such as requests for more materials, interviews, offer made, rejection notices, etc.
@@
 
-For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023 we read the text and recorded into a Google form the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methods foci on the position.
-The topical focus is what we understood as the primary intellectual focus of the position.
-The geographic focus is the region of the world that the ideal candidate has scholarly expertise on.
+For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023, we read the text and recorded the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methods foci on the position into a Google form.
+The topical focus is what we understood as the primary intellectual focus of the position—**examples of topical foci included X,Y, and Z**.
+The geographic focus is the region of the world that the ideal candidate has scholarly expertise on, **for example, X, Y, or Z**.
 The methods focus is the data-generating sub-field of archaeology that is mentioned in the ad.
-We recorded the type and number of documents requested in each ad (e.g.
-cover letter, CV, statements on research, teaching, diversity, syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad.
+**Examples of methods used in this study include X, Y, or Z.** <!--It might be helpful to make this into a summary table with examples of all foci generated by the study, and then one showing the simplified 10–15 topics.
+I understand this is in the supplementary tables already.--> We also recorded the type and number of documents requested in each ad (e.g.
+cover letter, CV, statements on research, teaching, diversity<!-- Were these all relatively easy to recognise?
+I’ve applied to places like Penn State, for example, that do a stupid combined research/teaching/diversity statement, just to make applicants do extra work, seemingly.-->, syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad.
 
@@
 Full details of the category reduction, showing the mapping between phrases in the job ads and our categories, are in our Supplementary Materials.
-Our topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
+<!--
+I think it would be much easier for our readers to parse this info as a table, unless you have any objection.
+This would also help us rein in our word count after all my contributions-->Our topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
 Our methods categories were: Archaeobotany, Archaeometry, Bioarchaeology, Ceramic analysis, Computational and Digital archaeology, Geoarchaeology, Landscape analysis, Lithic analysis, Material culture analysis, and Zooarchaeology.
@@
 Some topics include geographic regions because this is how they are typically understood by archaeologists.
-For example Mesoamerican archaeology is understood to refer to a specific time period and geographic region.
-Similarly, digital archaeology we recorded as both a method (when a job ad has a clearly distinct topical focus, such as historic archaeology) and a topic (when there is no other topics mentioned in the job ad).
-While these overlaps can make the data challenging to interpret, in our view it reflects the complex realities of how search committees express their needs in searching for new faculty, and is insightful in how it reveals intersections between different foci.
+For example, Mesoamerican archaeology is understood to refer to a specific time period and geographic region.
+Similarly, we recorded digital archaeology we recorded as both a method (when a job ad had a clearly distinct topical focus, such as historic archaeology) and a topic (when there were no other topics mentioned in the job ad).
+While these categorical overlaps can make the data challenging to interpret, in our view this approach reflects the complex realities of how search committees express their needs in searching for new faculty.
+Acknowledgement of overlaps also produces new insights into hiring dynamics through revealing intersections between different foci.
 
@@
 
-We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023, and focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities.
+We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023.
+We focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities because <!--Ben,we should provide the rationale for this, I don't remember what it was-->.
 @fig-show-basic-plots shows the count of ads for each year.
-Assistant Professor jobs are consistently the most common title and rank, and open rank or full professor are the least frequent.
-The ratio of tenure-track to non-tenure track positions is generally well above one.
-Only 2013-2014 had more non-tenure track positions than tenure track, which was followed by an upward trend peaking at 2018-2019 and the declining again into the present.
+Assistant Professor jobs are consistently the most common title and rank of positions advertised, while open rank or full professor are the least frequent.
+The ratio of tenure-track to non-tenure track positions is generally well above one; in other words, this data set is dominated by tenure-track positions .
+Only academic year 2013-2014 had more non-tenure track positions than tenure-track positions <!--I know that this is a pedantic complaint, but we should be consistent about whether or not we hyphenate tenure-track-->, which was followed by an upward trend peaking at 2018-2019 and then declining again into the present.
 
@@
 #| fig-cap: "A: Frequency of hiring institution by Carnegie classification.
-B: Inset shows map of the United States showing the count of tenure-track job ads posted by all insititutions in each state during 2013-2023"
+B: Inset shows map of the United States showing the count of tenure-track job ads posted by all insititutions in each state during 2013 -- 2023"
+#JB: So many positions in New York! Why don’t I have a job in New York?!
+# JB: This figure is great. We should make sure it works for colour-blind readers.
 
@@
 After California, the states that posted the most ads during 2013-2023 include New York, Texas, and Pennsylvania, and Florida.
-These top five states also correspond to the top five most populous US states, indicating that rates of hiring is approximately proportional to population density.
+These top five states also correspond to the top five most populous US states, suggesting that rates of hiring are approximately proportional to population density.
 These top five states for job ads are also the top five states with highest number of degrees awarded in Anthropology (U.S.
@@
 No institutions in Montana posted a job ad during this period.
-The implication here is that job-seekers who are able to relocate to populous areas will have more employment options.
+The implication here is that job-seekers who are able to relocate to populous areas will have more employment options.<!--Yes, nicely put-->
 
@@
 
+#JB: Interesting that Near East is on the up and up. You know what I am going to say about the timespans in the x-axis…
+
 # show figure from PNG file
@@
 Panel A of @fig-show-geo-trends shows that a single region of the US, the Southwest, occurs in more job ads than every other part of the world except for the Mediterranean.
-The Southwest includes Arizona and New Mexico, with portions of California, Colorado, Nevada, Oklahoma, Texas, and Utah.
+<!-- Someone in another field recently asked me what archaeologists wear to conferences, and I said “bolo ties”.
+This is why.-->The Southwest includes Arizona and New Mexico, with portions of California, Colorado, Nevada, Oklahoma, Texas, and Utah.
 It is archaeologically significant as the home of the Ancestral Pueblo, Hohokam, and Mogollon peoples who practiced irrigation agriculture and lived in relatively large settlements, compared to other regions of the US.
-The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, also in relatively large numbers.
+The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, also in relatively large numbers (**Summary citation(s), preferably including an Indigenous archaeologist from the region)**.
 The Mediterranean is prominent because it is the location that is often mentioned in job ads focused on classical archaeology (i.e.
@@
 
-Demand for jobs focusing on the Americas is generally high over time, with a peak in 2019-2020, and a decrease since then.
+Demand for jobs focusing on the Americas has been generally high over the past decade, with a peak in 2019-2020, and a subsequent decrease.
 Demand for jobs focusing on Africa was very low until 2019-2020, peaking in 2020-2021.
-The proportion of ads with a geographic focus on the Mediterranean has varied substantially, peaking at 2016 and at its lowest in 2019, showing an inverse pattern of the Americas.
-Asia and India, the Near East and Europe are consistently rare as a geographical focus in job ads.
-Asia and India, Africa and the Americas appear correlated with each other, while the Near East and Mediterranean are correlated in an opposite trend.
+The proportion of ads with a geographic focus on the Mediterranean has varied substantially, peaking in 2016 and experiencing a nadir in 2019, showing an inverse pattern to the Americas.
+Asia and India, the Near East, and Europe rarely occur as a geographical focus in job ads at US institutions.
+Asia and India, Africa, and the Americas appear correlated with each other, while the Near East and Mediterranean are correlated in an opposite trend.
 
@@
 Landscape archaeology, encompassing GIS and remote sensing, has remained prominent compared to other methods [@fig-show-metho-trends].
-Methods focused on a specific element of the archaeological record, such as Lithic analysis, Zooarchaeology and Ceramics are among the least frequently mentioned in job ads.
-Instead we see the more popular methods here are ones that are relevant to multiple elements of the archaeological record (e.g.
+<!--Might be worth exploring whether this is a skillset more frequently sought after in CRM and/or other industries—is this a transferable skill thing?--> Methods focused on a specific element of the archaeological record, such as Lithic analysis, Zooarchaeology and Ceramics are among the least frequently mentioned in job ads.
+Instead, the more popular methods are those that are relevant to multiple elements of the archaeological record (e.g.
 Archaeobotany encompasses macroscopic and microscopic plant remains; Bioarchaeology may include skeletal analysis, isotopes, proteins, etc.).
@@
 Computational and digital archaeology is the second most represented method, showing an overall increasing trend, particularly since 2020-2021.
-Archaeobotany shows a strong cyclical trend, with a rise and fall and rise again over our study period.
+<!--Congratulations--> Archaeobotany shows a strong cyclical trend, rising, falling, and then rising again over our study period.
 Archaeometry and Geoarchaeology have maintained a relatively low but steady presence in job ads, peaking in 2017-2018 and 2018-2019 and declining thereafter.
@@
 
+#JB: Please. The x-axis. My fragile mental state.
+
+#JB: Given that not all job ads are categorised by all foci, can we provide an overall n for Figure 3 for and 5, either in the captions or in the figures themselves?
+I assume that 20% of our “methods” sample is not the same as 20% of our “topics” sample.
+
 # show figure from PNG file
@@
 
-The most frequently mentioned topic in the job ads in Environmental archaeology [@fig-show-topi-trends], which is our category for phrases found in the text of ads such as human-environmental dynamics, interaction between humans and their environments, environmental change, climate change, historical ecology, ecological knowledge,human ecology, and ecological systems.
-Public archaeology is the second most frequent topic overall, this includes phrases such as cultural resource management, cultural heritage, heritage studies, museum studies, human rights, community engaged, historic preservation; social justice, community-based, repatriation, and community-engaged archaeology.
-The least frequent topics are Pleistocene archaeology, e.g.
-human origins; hunter-gatherer archaeology, and digital archaeology.
+The most frequently mentioned topic in the job ads is Environmental archaeology [@fig-show-topi-trends].
+This category encompasses such phrases as human-environmental dynamics, interaction between humans and their environments, environmental change, climate change, historical ecology, ecological knowledge,human ecology, and ecological systems.
+Public archaeology is the second most frequent topic overall, which included phrases such as cultural resource management, cultural heritage, heritage studies, museum studies, human rights, community engaged, historic preservation, social justice, community-based, repatriation, and community-engaged archaeology.
+The least common topics in our sample are Pleistocene archaeology, e.g.
+human origins, hunter-gatherer archaeology, and digital archaeology.
 
 In the years 2019-2020 and 2020-2021 there are striking changes in the popularity of topics in job ads.
-Indigenous and historical archaeology (which includes archaeology of enslaved people) became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
+Indigenous and historical archaeology (which includes archaeology of enslaved people <!--African diaspora archaeology?
+We should be careful to use the appropriate language here-->) became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
 Conversely, archaeological science, which was popular during 2012-2017, was rarely mentioned in job ads during 2019-2021.
-Ancient Europe and the Mediterranean, which was consistently one of the least frequently mentioned topics, virtually disappeared from job ads during 2019-2021.
-Mentions of complex societies in ads consistently decrease over the study period, while mentions of public archaeology consistently increase.
-The topic of Environmental Archaeology remains consistently high over time.
+Ancient Europe and the Mediterranean, which was an infrequently mentioned topic for the entire study period, virtually disappeared from job ads during 2019-2021.
+Mentions of complex societies in ads decrease at a steady rate, while mentions of public archaeology consistently increase over the study period.
+The topic of Environmental Archaeology remains consistently <!--We should be careful of the word consistent as we are relying on it a bit heavily throughout the text.
+Here, could we replace with "popular" or "stable"?-->high over time.
 
@@
 
-Job ads are frequently rich in topics, that is, they mention multiple topics in a single call for applications, more often than they mention multiple methods or geographic locations.
+In our sample, job ads were more topically rich than geographically or methodologically rich.
+That is, ads were more likely to mention multiple topics than they were to mention multiple methods or geographic locations.
 A Kruskal-Wallis test indicated significantly higher richness in topics compared to richness of geographic locations or methods in job ads (χ2 (df = `r unname(kw_test_richness$df)`, N = `r nrow(kw_test_richness_input)`) = `r round(unname(kw_test_richness$statistic), 2)`, p = `r expSup(unname(kw_test_richness$p.value))`).
@@
 Complex societies and environmental archaeology were frequently found in the same ads.
-Biological archaeology, archaeological science, and evolutionary archaeology are another cluster of topics that frequently occur together.
-Other topics are relatively isolated, for example, Pleistocene archaeology and digital archaeology rarely occur with other topics.
+Biological archaeology, archaeological science, and evolutionary archaeology are another cluster of topics that frequently co-occur.
+Other topics are relatively isolated.
+For example, Pleistocene archaeology and digital archaeology rarely occur with other topics.
 
@@
 Over our ten year study period there have been substantial changes in the instructions to applicants in terms of the type and number of documents that are requested by the search committees [@fig-requirements-over-time].
-A cover letter and CV are less requested in more recent years, especially after 2018.
+A cover letter and CV are less requested in more recent years <!--So curious as to what these jobs DO request.-->, especially after 2018.
 The requirement for a diversity statement is rare until 2019-2020, peaks around 2020-2021, then decreases towards the present.
-Requests for names of recommenders (either zero or three names, rarely only two names) reaches a maximum during 2019-2020 then decreases towards the present.
+Requests for names of recommenders (either zero or three names, rarely only two names) reaches a maximum during 2019-2020, then also decline for the remainder of the study period.
 The requirement for a research statement and teaching statement increases after 2015-2016, and becomes more frequent in job ads in more recent years.
@@
 Conversely, full professor positions have minimal requirements.
-The documents that are most frequently requested across the three ranks are cover letter and CV.
+The documents that are most frequently requested across all three ranks are cover letter and CV.
 
@@
 
-Our results reveal some of these shifts, providing insights into the dynamics of topical, methods, and geographic foci over time.
-While these dynamics are deeply felt by people actively seeking faculty jobs, we believe this is the first time they have been quantified at such a large scale.
-Trends in job ads reflect broader shifts in intellectual and practical priorities about archaeology, undergraduate education, and the process of hiring professors.
+Our results reveal some of these shifts <!--What shifts?
+It seems like there is an introductory sentence missing here.-->, providing insights into the dynamics of topical, methods, and geographic foci over time.
+While these dynamics are familiar to scholars actively seeking faculty jobs, we believe this is the first time they have been quantified at such a large scale within archaeology.
+Trends in job ads reflect broader shifts in intellectual and practical priorities concerning archaeology, undergraduate education, and the process of hiring professors.
 The demand for archaeology faculty, indicated by the total number of tenure-track jobs, maybe explained by a variety of processes.
-Overall we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-14.
-This is at odds with prior work finding that contingent positions make up more than half of faculty positions.
-This discrepancy may be due to advertising for short-term positions typically having much more limited circulation than advertising for tenure-track jobs, and thus less likely to be added to the Academic Jobs Wiki.
+Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-14.
+This is at odds with previous work that has shown that contingent positions make up more than half of faculty positions.
+**MORE DETAIL HERE.** This discrepancy may be due to the more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
 Many of these short-term positions are not advertised nationally, but only through local email lists and are filled by people close to the hiring department, such as recently graduated students.
-The downward trend in tenure-track positions during 2013-2019 maybe related to decline in undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a].
+It is thus likely that non-tenure-track jobs are less frequently added to the Academic Jobs Wiki.
+The downward trend in tenure-track positions during 2013-2019 may be related to declining undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a].
 The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus.
-
-In addition to a dip in job ads, our results show that 2019-2021 was a major inflection point in the popularity of topics and regions in job ads.
-Indigenous and historical archaeology and the Americas strongly increase in popularity at this time, while archaeological science, complex societies, the Mediterranean and Near East and show declines.
-Similarly, the number of job ads with a geographic focus on the Americas and Africa peaks during this time.
-These shifts in the topics and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices.
-COVID-19 negatively impacted Black, American Indian and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US.
-The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic with the murder of George Floyd in Minneapolis by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
+**(Citations).
+“75.2 per cent of respondents consider their career to have been negatively affected by the pandemic, whether through lost opportunities, illness, unemployment, or negative impact on mental health (Figure 6)” (Brami et al., 2023: 233).**
+
+The 2019-2021 period was also a major inflection point in the popularity of specific topics and regions in job ads.
+Calls for positions incorporating Indigenous and historical archaeology and archaeology of the Americas became far more frequent at this time, while archaeological science, complex societies, the Mediterranean and Near East and showed declines in popularity.
+Similarly, the number of job ads with a geographic focus on the Americas and Africa peaks during 2019–2021.
+These shifts in the topics and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices **(Citations)**.
+COVID-19 negatively impacted Black, American Indian and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US **(Citations)**.
+The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic and the murder of George Floyd in Minneapolis by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
 Mass protests objecting to Floyd's murder generated widespread concern about racial inequities and stimulated a broad interest in addressing systemic racial injustice.
-Our data suggest that US universities participated in this movement by adjusting their hiring plans to prioritize recruiting archaeologists working on topics relevant to Black and Indigenous communities.
+
+Our data suggest that US universities participated in this movement by adjusting their hiring plans to prioritize recruiting archaeologists working on topics relevant to Black and Indigenous communities <!--
+We should be thoughtful about where we locate agency here.
+Are institutions doing this, or departments?
+Likely it is some combination of both, but a lot of these calls are coming from departments.
+A suggestion for an avenue for future research might be to see how these trends play out at a larger scale across the social sciences and humanities, or even STEM fields.
+Aaron Clauset's next project...-->.
 Many universities may have hoped to hire Black and Indigenous archaeologists as part of their effort to tackle systemic racism.
-However as it is illegal in the US for universities use race or ethnicity as a primary factor in hiring (due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity), universities appear to have tailored the content of the job ads to focus on topics where they expect Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
+However as it is illegal in the US for universities use race or ethnicity as a primary factor in hiring (due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity), universities appear to have tailored the content of their job ads to focus on topics where they expect Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
+**Some commentary on whether and if Black/Indigenous researchers are more numerous in these fields, citing Maria Franklin, Sonya Atalay, Aja, Rachel Watson, etc.**
 
 This striking change in topics and geographic foci of 2019-2021 occurred against a backdrop of several longer-term trends.
-We found a long-term decline in topics such as the archaeology of complex societies and archaeological science, Mesoamerica and South America as a geographic focus, and methods relating to landscape archaeology.
+Over the course of the study period, we document a gradual decline in topics such as the archaeology of complex societies and archaeological science, the goegraphic foci of Mesoamerica and South America, and methods relating to landscape archaeology.
 These trends are harder to explain as we cannot link their start to a moment like COVID-19.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism.
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism **(Citations)**.
 This might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for compositional analysis.
-Other relevant factors may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation, and an increasing emphasis on more job-oriented degrees that prepare students for working in cultural heritage management rather than advanced research.
+Other relevant factors may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation.
+We also anticipate a growing emphasis on more job-oriented degrees that prepare students for working in cultural heritage management or contract archaeology **(Bardolph, Henry, and White Under Review)**.
 A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience, which may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
+<!--Alternate hypothesis: The people who study ancient mesoamerican states NEVER RETIRE.
+See: Michael Smith.
+-->
 
@@
 
-Our data on the requirements for applicants support prior findings that application complexity has gradually increased over time.
+Our data on the requirements for applicants support prior findings that the complexity of applications—and concomitantly, the labor required to apply for a tenure-track job—has gradually increased over time.
 This trend is especially pronounced for Assistant Professor positions, which make more demands on applicants than Associate and Full Professor positions.
-We saw a growing demand for research and teaching statements, consistent with previous work.
+<!--Eat the old-->.
+Consistent with results from other studies **(Citations)**, our study documented a growing demand for research and teaching statements.
 Demand for diversity statements shows a unique trajectory, peaking in 2020-2021 and declining into the present.
 This may relate to the intersecting concerns emerging during the COVID-19 pandemic, which may have declined in job ads as the most urgent time of the pandemic moves into the past.
-Another factor here may be the debates surrounding the experiment with diversity statements in hiring at some UC campuses during 2016-2022.
+<!-- I don't understand the second half of this sentence-->.Another factor here may be the debates surrounding the experiment with diversity statements in hiring at some UC campuses during 2016-2022.
 Universities that previously required a diversity statement dropped that requirement after 2022 [@guidenAreDiversityStatements].
@@
 In recognition of this burden, not only on applicants but also on colleagues writing letters of recommendation over and over to support applicants, many hiring committees now follow the recommendations of @dennisWorstAnthroJob2022 in only requesting names and letters of recommendation at later stages of the hiring process, if at all.
-
-A key limitation of our research is that it does not include an analysis of the academic profiles of those who were eventually hired from the job ads.
+**A line here about AAA recommendations to not ask for letters up front.**
+
+A key limitation of our research is that it does not include an analysis of the academic profiles of those who were eventually hired based on this sample of job ads.
 Our results reveal collective aspirations for the future of the field, but only from those who are securely employed in it as tenured professors serving on hiring committees at US universities.
-Our results do not show how these aspirations worked out in the topical and geographic foci of the people who were eventually hired for the positions studied here.
+Our results do not show how these aspirations worked out in the topical and geographic foci of the people who were eventually hired for these positions .
 Future work should consider interviewing faculty hired during our study period to match up people to the ads they responded to.
-If the successful applicants can be identified, then the match between the details of the job ad and the applicant's research can be analysed.
+If the successful applicants can be identified, then the match between the details of the job ad and the applicant's research can be analyzed.
 Then we can assess how effective the hiring process is for driving change in the discipline by setting the topics of courses that will be taught in undergraduate and graduate curricula, and research that will be supported by universities.
@@
 
-Our results show that job ads are a highly nuanced record of disciplinary history, showing shifting trends in topics, methods, and geographic foci, with higher chronological resolution than many other sources, such as undergraduate textbooks [@lyman2010american].
+Our results show that job ads offer a nuanced record of disciplinary history, documenting shifting trends in topics, methods, and geographic foci, with higher chronological resolution than many other sources, such as undergraduate textbooks [@lyman2010american].
 The short periods of time spanning the rise and fall of some of the foci we observed has implications for prospective graduate students.
-An topic that is growing in popularity as a student begins their PhD may have peaked and be in decline before they graduate.
+A topic that is growing in popularity as a student begins their PhD may have peaked and be in decline before they graduate.
 Methods seem to have a much lower frequency of change in popularity, relative to topical and geographic foci.
@@
 Jobs ads also demonstrate how archaeologists draw on current events and politics to prioritize topics, regions and methods in their hiring plans.
-This is consistent with prior findings, such @wilk1985ancient's analysis of publications that found trends in explanations of Mayan collapse during the 1960s-70 that correspond with major contemporary American political and social issues.
+This is consistent with prior findings, such @wilk1985ancient's analysis of publications that found trends in explanations of Mayan collapse during the 1960s-1970s that correspond with major contemporary American political and social issues.
 This responsiveness to current events reveals that many archaeologists accept that our narratives of the past are entangled in the social and political milieu in which we work.
@@
 
+We thank NO ONE.
+
 <!-- The following line inserts a page break  -->

### 2025-01-10 analysis/paper/article.qmd
Minor formatting edits.

@@
 This obsessive focus on career trajectories is warranted, particularly within the discipline of anthropology.
-Between 1985 and 2014, the number of PhDs produced by American anthropology departments increased by 70% (**Speakman et al., 2018a**), but the number of tenure-track positions advertised has stagnated or declined*.* As of2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions (**AAUP 2022**); today, 71% of faculty in the US are non-tenure-track(**Culver and Kezar 2022***).* The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors (**Beck, under review**), including decreasing federal support for higher education, concomitant increases in student debt (**Gusterson 2017**), a pronounced shift in university investment from faculty to administration (Graeber 2018: 162–163), and the growing privatization and market orientation of scientific research (**Mirowski 2011**).
+Between 1985 and 2014, the number of PhDs produced by American anthropology departments increased by 70% (**Speakman et al., 2018a**), but the number of tenure-track positions advertised has stagnated or declined*.* As of2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions (**AAUP 2022**); today, 71% of faculty in the US are non-tenure-track(**Culver and Kezar 2022***).* The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors (**Beck, under review**), including decreasing federal support for higher education, concomitant increases in student debt (**Gusterson 2017**), a pronounced shift in university investment from faculty to administration (**Graeber 2018: 162–163**), and the growing privatization and market orientation of scientific research (**Mirowski 2011**).
 Whatever the cause, research within anthropology has consistently shown that at most 20% of PhDs in the discipline are obtaining jobs in American anthropology departments (**Speakman et al.
@@
 A twenty-fold increase in the number of applications required to obtain a workable long-term academic position may seem preposterous, but these changing requirements for applicants are familiar to anyone who has obtained their PhDs since the Great Recession.
-This competitive landscape may have deeper roots than the last economic crisis; writing more than three decades ago, Rabinow pinpointed “an awareness… that standards have changed during the last thirty years and the quantitative and qualitative demands for entry into the system are immeasurably higher now” (**1992: 66**).
+This competitive landscape has deeper roots than the last economic crisis; writing more than three decades ago, Rabinow pinpointed “an awareness… that standards have changed during the last thirty years and the quantitative and qualitative demands for entry into the system are immeasurably higher now” (**1992: 66**).
 
@@
 We argue that to understand the dynamics of hiring in academic archaeology, we must start at the beginning, with the job ad itself.
-There is an abundance of recent work on who does and does not get hired into tenure-track positions in anthropology and archaeology (Kawa 2019; Mackie and Rockwell 2023), but less scholarship on how those hires unfold.
-Research on hiring dynamics has amply demonstrated that *where* you get your PhD has an enormous impact on if and where you get a tenure-track position (Kawa et al.
-2019; Mackie and Rockwell 2023), but how about *what* you study?
+There is an abundance of recent work on who does and does not get hired into tenure-track positions in anthropology and archaeology (**Kawa 2019; Mackie and Rockwell 2023**), but less scholarship on how those hires unfold.
+Research on hiring dynamics has amply demonstrated that *where* you get your PhD has an enormous impact on if and where you get a tenure-track position (**Kawa et al.
+2019; Mackie and Rockwell 2023**), but how about *what* you study?
 Are there topical, methodological, or regional specialties that are in consistent demand, or is there continual flux generated by intellectual fads, economic incentives, and political attention?
@@
 Job seekers must search for ads that are well-matched to the skill sets and professional experiences that they have taken years to cultivate.
-Hiring committeesmust sift through tens to hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realize<!--We should go with American spelling, no?
-With your Australian inclinations and my growing incorporation of Irish orthography into my writing, this paper is gonna be a mess...--> their visions of their program and the discipline.
+Hiring committeesmust sift through tens to hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realize their visions of their program and the discipline.
 Finding the right candidate for the position—or the right position for the candidate—is akin to finding a needle in a haystack.
@@
 now ranks among the highest for all professions (**Citation**).
-In tandem with the gradual de-investment in American higher education since the **XXX (Mirowski**) and the aftershocks of the 2008 recession, the ADEA exemption has cultivated an environment where new lines are few and far between.
+In tandem with the gradual de-investment in American higher education since the **XXX (Mirowski 2011**) and the aftershocks of the 2008 recession, the ADEA exemption has cultivated an environment where new lines are few and far between.
 Among biological anthropologists, @Passalacqua_2018 found a ratio of 0.81 PhDs to job academic advertisements per year, concluding that academic positions in biological anthropology are barely at sustainable levels.

### 2025-01-12 analysis/paper/article.qmd
BM responding to JB's comments

@@
 We analyzed the text of 431 job ads posted from 2013--2023.
-Our analysis focuses on shifts in archaeological topics, methods and geographic regions specified in archaeological job ads.
+Our analysis addresses the question of how have archaeological topics, methods, and geographic regions specified in archaeological job ads shifted over time?
 We also investigate whether the labor burden for applicants has changed over time: do institutions request more information and documents from applicants at the initial stages of application, compared to a decade ago?
@@
 keywords: |
-  keyword 1; keyword 2; keyword 3
+  archaeology; hiring; academia; job ads 
 highlights: |
@@
 
-<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->
-
-<!-- With the following code you can access and display values from the yml header above. -->
+```{=html}
+<!--Comments on Abstract and Keywords: 
+1.
+Is “jobs listing website” the appropriate category here?
+I often hear these referred to as “jobs boards”—that might be a better fit?
+BM: I agree that 'job board' is in common use, but I prefer 'jobs listing website' because it's more precise and explicit about the medium and contents
+
+2.
+Structurally, the second research domain is phrased as a question.
+For consistency, it would be good to provide a similar structure for the first domain, e.g.
+“Our analysis focuses on the XXX—have the topics, methods, or geographic regions specified in archaeological job ads changed over time?” BM: agreed, I updated the abstract
+
+3. Suggested keywords: archaeology, hiring, academia, job ads BM: added, you can also freely add these kinds of things directly, no worries -->
+```
 
@@
   xlab("") +
-  ylab("Number of\ntenure track job ads") +
+  ylab("Number of\ntenure-track job ads") +
   scale_fill_brewer(palette = "Dark2") +
@@
   # add line break for prettier plotting
-  mutate(year_ad_posted_break = str_replace(year, "-", "-\n"))
+  mutate(year_ad_posted_break = str_replace(year, "-", 
+                                            "–\n")) # en-dash
 
@@
            size = 3) +
-  labs(y = "Ratio of tenure-track\nto non-tenure track and other",
+  labs(y = "Ratio of tenure-track\nto non-tenure-track and other",
        x = "") +
@@
    #guides(x = "none")
-
 ```
@@
 #| fig-cap: "A: total number of job ads posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank.
-B: Ratio of tenure-track to non-tenure-track positions over time.The proportion of the bar above the red line reflects the number of non-tenure track positions available relative to tenure-track positions."
+B: Ratio of tenure-track to non-tenure-track positions over time.
+The red line indicates a 1:1 ratio of tenure-track to non-tenure-track positions, bars taller than that line indicate more tenure-track than non-tenure-track positions in that year"
 #JB: I have added the last sentence above because I have not seen this kind of figure in an archaeology paper, so I think we should explain it.
 Will reduce friction for our readers.Please confirm that I am actually correctly explaining how the figure works, however.
-#JB: I would be eternally grateful if you would convert the hyphens between the timespans on the x-axis into en dashes 😬
+BM: I've edited the caption further to clarify how to read that red line.
+#JB: I would be eternally grateful if you would convert the hyphens between the timespans on the x-axis into en dashes 😬 BM: done, you may commence eternal gratitude :)
 
@@
 
-```{=html}
-<!--Comments on Abstract and Keywords: 
-1. Is “jobs listing website” the appropriate category here? I often hear these referred to as “jobs boards”—that might be a better fit?
-2.
-Structurally, the second research domain is phrased as a question.
-For consistency, it would be good to provide a similar structure for the first domain, e.g.
-“Our analysis focuses on the XXX—have the topics, methods, or geographic regions specified in archaeological job ads changed over time?”
-3. Suggested keywords: archaeology, hiring, academia, job ads-->
-```
-
 <!--Ben, please note that for now I am keeping track of my references.
-I have bolded them in text below; I think I will want to learn how to use Zotero and incorporate the citations using your arcane Lovecraftian ritual language (BibTex?), so that for our next paper I am capable of participating in the entire workflow with less supervision-->
+I have bolded them in text below; I think I will want to learn how to use Zotero and incorporate the citations using your arcane Lovecraftian ritual language (BibTex?), so that for our next paper I am capable of participating in the entire workflow with less supervision.
+BM: ok great!
+Yes, BibTeX, I believe you will love it.
+-->
 
@@
 This obsessive focus on career trajectories is warranted, particularly within the discipline of anthropology.
-Between 1985 and 2014, the number of PhDs produced by American anthropology departments increased by 70% (**Speakman et al., 2018a**), but the number of tenure-track positions advertised has stagnated or declined*.* As of2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions (**AAUP 2022**); today, 71% of faculty in the US are non-tenure-track(**Culver and Kezar 2022***).* The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors (**Beck, under review**), including decreasing federal support for higher education, concomitant increases in student debt (**Gusterson 2017**), a pronounced shift in university investment from faculty to administration (**Graeber 2018: 162–163**), and the growing privatization and market orientation of scientific research (**Mirowski 2011**).
-Whatever the cause, research within anthropology has consistently shown that at most 20% of PhDs in the discipline are obtaining jobs in American anthropology departments (**Speakman et al.
-2018b**), while the top-ranking programmes are placing fewer than one in three graduates in tenure-track jobs (**Mackie and Rockwell 2023**).
+As of 2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions (**AAUP 2022**); today, 71% of faculty in the US are non-tenure-track(**Culver and Kezar 2022***).* The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors (**Beck, under review**), including decreasing federal support for higher education, concomitant increases in student debt (**Gusterson 2017**), a pronounced shift in university investment from faculty to administration (**Graeber 2018: 162–163**), and the growing privatization and market orientation of scientific research (**Mirowski 2011**).
 
@@
 The first author on this paper has applied to five academic positions over the course of his career, while the last author has applied to one hundred.
-A twenty-fold increase in the number of applications required to obtain a workable long-term academic position may seem preposterous, but these changing requirements for applicants are familiar to anyone who has obtained their PhDs since the Great Recession.
-This competitive landscape has deeper roots than the last economic crisis; writing more than three decades ago, Rabinow pinpointed “an awareness… that standards have changed during the last thirty years and the quantitative and qualitative demands for entry into the system are immeasurably higher now” (**1992: 66**).
+A twenty-fold increase in the number of applications required to obtain a workable long-term academic position may seem preposterous, but these changing requirements for applicants are familiar to anyone who has obtained their PhDs since the Great Recession of 2008.
+This competitive landscape has deeper roots than the last economic crisis; writing more than three decades ago, Rabinow pinpointed "an awareness...
+that standards have changed during the last thirty years and the quantitative and qualitative demands for entry into the system are immeasurably higher now" (**1992: 66**).
 
@@
 found that 52% of respondents who had obtained their PhD at least two years before the survey were on fixed-term contracts, while only 9% were in permanent positions (**2023: Figure 5**).
-In their words, “the queue of eager post-docs hoping for a long-term appointment is getting longer, as is the average time spent in the ‘queue’” (**2023: 239**).
-Ribeiro and Giamakis emphasize that this demographic comprises an academic precariat that is essential to the functioning of universities, but consists of a workforce on short-term contracts perpetually “stranded between employment and unemployment” (**2023:10**).
+In their words, "the queue of eager post-docs hoping for a long-term appointment is getting longer, as is the average time spent in the 'queue'" (**2023: 239**).
+Ribeiro and Giamakis emphasize that this demographic comprises an academic precariat that is essential to the functioning of universities, but consists of a workforce on short-term contracts perpetually "stranded between employment and unemployment" (**2023:10**).
 
@@
 As Dennis et al.
-underscore, early career researchers are “supplicants”: “There is no way for applicants to point out that the job ads are taking too much time out of the scholarly community’s collective time bank” (**2022: 1**).
-Indeed, acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs to performing well in Zoom interviews and teaching demonstrations.
-These courses often provide valuable training in the hidden curriculum of the academy, but that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the “market” as a historically contingent set of cultural practices.
-As Rabinow maintains, “if we want ethical considerations to play a central role in the articulation of truth and power—and I think we do—then bringing such considerations into view is the necessary first step towards recognizing who we are today and setting out on the road to a better place" (**1992: 71**) **.**
+underscore, early career researchers are "supplicants": "There is no way for applicants to point out that the job ads are taking too much time out of the scholarly community’s collective time bank" (**2022: 1**).
+Indeed, acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs to performing well in online video interviews and teaching demonstrations.
+These courses often provide valuable training in the hidden curriculum of the academy, but that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the "market" as a historically contingent set of cultural practices.
+As Rabinow maintains, "if we want ethical considerations to play a central role in the articulation of truth and power—and I think we do—then bringing such considerations into view is the necessary first step towards recognizing who we are today and setting out on the road to a better place" (**1992: 71**) **.**
 
@@
 
+<!-- BM: I think we have a little repetition in the Intro and Background sections, particulary in the mentions of data from papers by Speakman, Kawa et al.
+and Mackie & Rockwell.
+To reduce duplication, I've removed most of that from the intro, and  kept/shifted it into the background so we have a logical progression of general to more specific observations about the job market -->
+
 The academic job market is a source of uncertainty, for both job seekers and hiring committees.
 Job seekers must search for ads that are well-matched to the skill sets and professional experiences that they have taken years to cultivate.
-Hiring committeesmust sift through tens to hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realize their visions of their program and the discipline.
+Hiring committees must sift through tens to hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realize their visions of their program and the discipline.
 Finding the right candidate for the position—or the right position for the candidate—is akin to finding a needle in a haystack.
@@
 Rising numbers of contingent faculty has long been a concern in US higher education [@Trevithick2010].
-Data from the American Association of University Professors' 2021–22 faculty survey indicate that more than 60% of faculty positions in US universities were held by non-tenure track full time or part time contingent faculty members [@colby2022annual].
+Data from the American Association of University Professors' 2021–22 faculty survey indicate that more than 60% of faculty positions in US universities were held by non-tenure-track full time or part time contingent faculty members [@colby2022annual].
 Contingent faculty positions are precarious because they provide low or no health and retirement benefits, little or nothing in the way of professional development, and lower salaries relative to tenure-track positions.
@@
 Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing].
+Top-ranking programmes are placing fewer than one in three graduates in tenure-track jobs [@mackieMarketShareAccounting2023].
 Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
-This network analysis showedthat programs with large endowments and with faculty who hold prestigious awards and are widely cited by other scholars produce the majority of tenured and tenure-track faculty in anthropology.
+This network analysis showed that programs with large endowments and with faculty who hold prestigious awards and are widely cited by other scholars produce the majority of tenured and tenure-track faculty in anthropology.
 Hiring bias predicated on the prestige of specific institutions and programs is not unique to anthropology.
@@
 For example, twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy.
-In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported two job ads with particularly extreme requirements<!--Are you opposed to naming and shaming these institutions, which were Oberlin College and Grinnell College?
-I applied for one of these positions, and have no issue referring to them by name...-->, which respectively requested nine and thirteen documents from applicants.
+In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported job ads from Oberlin College and Grinnell College with particularly extreme requirements <!--Are you opposed to naming and shaming these institutions, which were Oberlin College and Grinnell College?
+I applied for one of these positions, and have no issue referring to them by name...
+BM: nope, I've added their names -->, which respectively requested nine and thirteen documents from applicants.
 This high burden on applicants disproportionately favors people with more time and financial resources to prepare the required materials.
@@
 Rivera, in a study of hiring practices at YYY, has likewise demonstrated ZZZ.
-Somewhere in this section I want to discuss the results from sociology, specifically the Warren 2019 piece and the Rivera 2017 piece on hiring practices.**
+Somewhere in this section I want to discuss the results from sociology, specifically the Warren 2019 piece and the Rivera 2017 piece on hiring practices.** <!-- ok, sounds good!
+-->
 
@@
 Our primary data source is the Archaeology Academic Jobs Wiki<!--Check if Kelsky has anything too say about this.
-I also think there have been some articles about the pros and cons of the poli sci or Econ versions of this.-->.
+I also think there have been some articles about the pros and cons of the poli sci or Econ versions of this.
+BM: I did have a search of her blog where there are ~20 mentions, all pretty neutral, her most interesting comment is to notes that it originated in the year of the Great Recession -->.
 Originating in 2007, this is a set of freely accessible web pages that anyone can edit (anonymously or with a free user account) hosted by Fandom, a for-profit company.
@@
 The coordinators and contributors are nearly all anonymous or pseudonymous.
-Typically, contributors copy and paste the text of job ads from other sources, such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, into the wiki, collecting <!--consolidating?
-centralizing?--> ads originally posted in numerous different locations.
-Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position, such as a tally of how many people have applied, the dates of events such as requests for more materials, interviews, offer made, rejection notices, etc.
+Typically, contributors copy and paste all or some of the text of job ads from other sources, such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, into the wiki, collecting <!--consolidating?
+centralizing?
+BM: I prefer the shortest and simplest word that will do the job --> ads originally posted in numerous different locations.
+Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position.
+These edits result in annotations such as a tally of how many people have applied, the dates of events such as requests for more materials, interviews, offer made, rejection notices, etc.
 Contributors also edit the page to ask and answer questions about the positions and the application process.
@@
 For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023, we read the text and recorded the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methods foci on the position into a Google form.
-The topical focus is what we understood as the primary intellectual focus of the position—**examples of topical foci included X,Y, and Z**.
-The geographic focus is the region of the world that the ideal candidate has scholarly expertise on, **for example, X, Y, or Z**.
+The topical focus is what we understood as the primary intellectual focus of the position—examples of topical foci included environmental archaeology, public archaeology, and North American archaeology.
+The geographic focus is the region of the world that the ideal candidate has scholarly expertise on, for example, Southwest US, Mediterranean, or Asia and India.
 The methods focus is the data-generating sub-field of archaeology that is mentioned in the ad.
-**Examples of methods used in this study include X, Y, or Z.** <!--It might be helpful to make this into a summary table with examples of all foci generated by the study, and then one showing the simplified 10–15 topics.
-I understand this is in the supplementary tables already.--> We also recorded the type and number of documents requested in each ad (e.g.
-cover letter, CV, statements on research, teaching, diversity<!-- Were these all relatively easy to recognise?
-I’ve applied to places like Penn State, for example, that do a stupid combined research/teaching/diversity statement, just to make applicants do extra work, seemingly.-->, syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad.
-
-After completing primary data collection, we studied the topical, geographic, and methods foci of each ad and collaboratively and manually reduced the variation in the raw data into 10-15 categories appearing in 20 (for topics and geography) or 10 (for methods) or more job ads to simplify analysis and visualisation.
+Examples of methods used in this study include archaeobotany, lithic analysis, or zooarchaeology.
+<!--It might be helpful to make this into a summary table with examples of all foci generated by the study, and then one showing the simplified 10–15 topics.
+I understand this is in the supplementary tables already.
+BM: I've added in the 'for example' you requested, but feels very repetitive to give examples in this paragraph, and then the full set of topics etc.
+in the next paragraph, and add then add a table as well.
+It's like saying "we found 5000 lithics, one random one was 45 mm long, another random piece was 12 mm wide, etc." readers don't expect that in the manuscript  --> We also recorded the type and number of documents requested in each ad (e.g.
+cover letter, CV, statements on research, teaching, diversity <!-- Were these all relatively easy to recognise?
+I’ve applied to places like Penn State, for example, that do a stupid combined research/teaching/diversity statement, just to make applicants do extra work, seemingly.
+BM: we treated these kinds of combo statements as separate requests, I don't know how many there were, I'm ok to omit that detail -->, syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad, if any.
+
+After completing primary data collection, we studied the topical, geographic, and methods text of each ad and collaboratively and manually reduced the variation in the raw data into 10-15 categories appearing in 20 (for topics and geography) or 10 (for methods) or more job ads to simplify analysis and visualisation.
 This means that some topics, such as gender (mentioned in `r n_jobs_mentioning_gender` ads) do not appear in our results because of their rarity in the job ads.
 Full details of the category reduction, showing the mapping between phrases in the job ads and our categories, are in our Supplementary Materials.
-<!--
-I think it would be much easier for our readers to parse this info as a table, unless you have any objection.
-This would also help us rein in our word count after all my contributions-->Our topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
+<!-- I think it would be much easier for our readers to parse this info as a table, unless you have any objection.
+This would also help us rein in our word count after all my contributions BM: I think it's best in the Supplement, these will be too distracting in the main text.
+We do unpack the most important categories in detail in results section.
+Let's see what the peer reviewers say --> Our final, reduced topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
 Our methods categories were: Archaeobotany, Archaeometry, Bioarchaeology, Ceramic analysis, Computational and Digital archaeology, Geoarchaeology, Landscape analysis, Lithic analysis, Material culture analysis, and Zooarchaeology.
@@
 We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023.
-We focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities because <!--Ben,we should provide the rationale for this, I don't remember what it was-->.
+We focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities <!--Ben, if we should provide the rationale for this, I don't remember what it was BM: I also don't recall, I don't see any justification for the US focus in Speakman et al or Kawa et al.
+and I can't think of a one that is either unverifiable or doesn't require heaps of additional work to validate.
+Leave it to the peer reviewers?
+-->.
 @fig-show-basic-plots shows the count of ads for each year.
 Assistant Professor jobs are consistently the most common title and rank of positions advertised, while open rank or full professor are the least frequent.
-The ratio of tenure-track to non-tenure track positions is generally well above one; in other words, this data set is dominated by tenure-track positions .
-Only academic year 2013-2014 had more non-tenure track positions than tenure-track positions <!--I know that this is a pedantic complaint, but we should be consistent about whether or not we hyphenate tenure-track-->, which was followed by an upward trend peaking at 2018-2019 and then declining again into the present.
+The ratio of tenure-track to non-tenure-track positions is generally well above one; in other words, this data set is dominated by tenure-track positions .
+Only academic year 2013-2014 had more non-tenure-track positions than tenure-track positions <!--I know that this is a pedantic complaint, but we should be consistent about whether or not we hyphenate tenure-track BM: I've edited every instance so they have hyphens.
+How do you feel about the double hyphenated 'non-tenure-track'?
+-->, which was followed by an upward trend peaking at 2018-2019 and then declining again into the present.
 
@@
 No institutions in Montana posted a job ad during this period.
-The implication here is that job-seekers who are able to relocate to populous areas will have more employment options.<!--Yes, nicely put-->
+The implication here is that job-seekers who are able to relocate to populous areas will have more employment options.
+<!--Yes, nicely put BM: thanks!
+-->
 
@@
   # add line break for prettier plotting
-  mutate(year_ad_posted_break = str_replace(year_ad_posted, "-", "-\n"))
+  mutate(year_ad_posted_break = str_replace(year_ad_posted, "-", 
+                                            "–\n")) # en-dash
 
@@
 
-#JB: Interesting that Near East is on the up and up. You know what I am going to say about the timespans in the x-axis…
+#JB: Interesting that Near East is on the up and up. You know what I am going to say about the timespans in the x-axis… BM: replaced hyphens with en-dash
 
@@
 <!-- Someone in another field recently asked me what archaeologists wear to conferences, and I said “bolo ties”.
-This is why.-->The Southwest includes Arizona and New Mexico, with portions of California, Colorado, Nevada, Oklahoma, Texas, and Utah.
+This is why.
+BM: haha, yes, exactly!
+-->The Southwest includes Arizona and New Mexico, with portions of California, Colorado, Nevada, Oklahoma, Texas, and Utah.
 It is archaeologically significant as the home of the Ancestral Pueblo, Hohokam, and Mogollon peoples who practiced irrigation agriculture and lived in relatively large settlements, compared to other regions of the US.
-The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, also in relatively large numbers (**Summary citation(s), preferably including an Indigenous archaeologist from the region)**.
+The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, also in relatively large numbers [@griffin2000native] <!-- BM: added requested citation: JB: Summary citation(s), preferably including an Indigenous archaeologist from the region -->.
 The Mediterranean is prominent because it is the location that is often mentioned in job ads focused on classical archaeology (i.e.
@@
 Landscape archaeology, encompassing GIS and remote sensing, has remained prominent compared to other methods [@fig-show-metho-trends].
-<!--Might be worth exploring whether this is a skillset more frequently sought after in CRM and/or other industries—is this a transferable skill thing?--> Methods focused on a specific element of the archaeological record, such as Lithic analysis, Zooarchaeology and Ceramics are among the least frequently mentioned in job ads.
+<!--Might be worth exploring whether this is a skillset more frequently sought after in CRM and/or other industries—is this a transferable skill thing?
+BM: yes, good point, added something on that --> The popularity of this suite of methods may reflect its importance as a skill in demand by employers in the Cultural Research Management sector.
+@morganReadyNotArchaeological2023 found that almost one-third of 599 jobs ads posted by CRM employers sought candidates with experience in GIS methods.
+Methods focused on a specific element of the archaeological record, such as Lithic analysis, Zooarchaeology and Ceramics are among the least frequently mentioned in job ads.
 Instead, the more popular methods are those that are relevant to multiple elements of the archaeological record (e.g.
@@
 Computational and digital archaeology is the second most represented method, showing an overall increasing trend, particularly since 2020-2021.
-<!--Congratulations--> Archaeobotany shows a strong cyclical trend, rising, falling, and then rising again over our study period.
+<!--Congratulations BM: haha, would you like to borrow my crystal ball?--> Archaeobotany shows a strong cyclical trend, rising, falling, and then rising again over our study period.
 Archaeometry and Geoarchaeology have maintained a relatively low but steady presence in job ads, peaking in 2017-2018 and 2018-2019 and declining thereafter.
@@
 
-#JB: Please. The x-axis. My fragile mental state.
+#JB: Please. The x-axis. My fragile mental state. BM: en-dashes added!
 
@@
 I assume that 20% of our “methods” sample is not the same as 20% of our “topics” sample.
+BM: what would the n be here, number of individual topics or methods recorded in total?
 
@@
 human origins, hunter-gatherer archaeology, and digital archaeology.
+<!-- BM: See, here is a nice unpacking of some of the key categories -->
 
 In the years 2019-2020 and 2020-2021 there are striking changes in the popularity of topics in job ads.
-Indigenous and historical archaeology (which includes archaeology of enslaved people <!--African diaspora archaeology?
-We should be careful to use the appropriate language here-->) became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
+Indigenous and historical archaeology (which includes archaeology of the African diaspora and enslaved people <!--African diaspora archaeology?
+We should be careful to use the appropriate language here BM: yes, ok, added that, but many jobs ads literally mention slavery and enslaved people, so I want to make it clear that slavery is a notable topic in American archaeology, not something to sanitize and ignore under the gloss of historical archaeology generally being about white people, a personal obsession, I guess.
+Did you know there were African ex-slaves on the first fleet to Australia?
+-->) became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
 Conversely, archaeological science, which was popular during 2012-2017, was rarely mentioned in job ads during 2019-2021.
@@
 Mentions of complex societies in ads decrease at a steady rate, while mentions of public archaeology consistently increase over the study period.
-The topic of Environmental Archaeology remains consistently <!--We should be careful of the word consistent as we are relying on it a bit heavily throughout the text.
-Here, could we replace with "popular" or "stable"?-->high over time.
+The topic of Environmental Archaeology remains <!--We should be careful of the word consistent as we are relying on it a bit heavily throughout the text.
+Here, could we replace with "popular" or "stable"?
+BM: yes, good point, seems to work if I just delete it -->high over time.
 
@@
   # add line break for prettier plotting
-  mutate(year_ad_posted_break = str_replace(year_ad_posted, "-", "-\n"))
+  mutate(year_ad_posted_break = str_replace(year_ad_posted, "-", 
+                                           "–\n")) # en-dash
 
@@
 Over our ten year study period there have been substantial changes in the instructions to applicants in terms of the type and number of documents that are requested by the search committees [@fig-requirements-over-time].
-A cover letter and CV are less requested in more recent years <!--So curious as to what these jobs DO request.-->, especially after 2018.
+Requests for cover letter and CV decline in more recent years, perhaps reflecting their status as a generic and expected component to submit without needing to be specifically requested <!--So curious as to what these jobs DO request.
+BM: I'm not confident this is a real or important trend, but I did look into the schools that don't ask for these, there are mostly lesser-known places like Texas Tech University and East Tennessee State University.
+I think there might be a few other factors here, for example the job ads on the wiki might have that generic text edited out because it's expected by everyone.
+Difficult to test this because the links the interfolio, etc original source text for most ads are broken now, I did check a dozen of them just now.
+In any case, I don't think any further work on it would be of much relevance to our main arguments -->.
 The requirement for a diversity statement is rare until 2019-2020, peaks around 2020-2021, then decreases towards the present.
@@
 
-Our results reveal some of these shifts <!--What shifts?
-It seems like there is an introductory sentence missing here.-->, providing insights into the dynamics of topical, methods, and geographic foci over time.
+Our study of a decade of tenure-track job ads in archaeology in the US reveals diverse dynamics in the demand for specialisations in topics, methods and geographic focus, and in the instructions to applicants to these jobs <!--What shifts?
+It seems like there is an introductory sentence missing here.
+BM: ok, I tried to expand it a bit -->.
 While these dynamics are familiar to scholars actively seeking faculty jobs, we believe this is the first time they have been quantified at such a large scale within archaeology.
@@
 Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-14.
-This is at odds with previous work that has shown that contingent positions make up more than half of faculty positions.
-**MORE DETAIL HERE.** This discrepancy may be due to the more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
+This is at odds with previous work that has shown that contingent positions make up more than half of faculty positions [@colby2022annual].
+**MORE DETAIL HERE.** <!-- BM: what do you want more detail on?
+The previous work is Colby 2022, discussed in the background section already.
+I've added the citation to that here also--> This discrepancy may be due to the more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
 Many of these short-term positions are not advertised nationally, but only through local email lists and are filled by people close to the hiring department, such as recently graduated students.
-It is thus likely that non-tenure-track jobs are less frequently added to the Academic Jobs Wiki.
+One possible explanation for our results is that non-tenure-track jobs are less frequently added to the Academic Jobs Wiki.
 The downward trend in tenure-track positions during 2013-2019 may be related to declining undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a].
-The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus.
-**(Citations).
-“75.2 per cent of respondents consider their career to have been negatively affected by the pandemic, whether through lost opportunities, illness, unemployment, or negative impact on mental health (Figure 6)” (Brami et al., 2023: 233).**
+The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus [@woolstonJuniorResearchersHit2020].
+A survey of early career researchers in archaeology captures the impact of this dip, with three-quarters of respondents experiencing negative impacts on their careers due to the pandemic [@bramiPrecariousFutureReflections2023] <!-- JB: “75.2 per cent of respondents consider their career to have been negatively affected by the pandemic, whether through lost opportunities, illness, unemployment, or negative impact on mental health (Figure 6)” (Brami et al., 2023: 233).
+BM: I put a sentence in for this for you, but I think it's mostly a supply-side detail, and our focus is demand-side, so I don't feel it's highly relevant here and don't want too much detail on it -->
 
@@
 Similarly, the number of job ads with a geographic focus on the Americas and Africa peaks during 2019–2021.
-These shifts in the topics and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices **(Citations)**.
-COVID-19 negatively impacted Black, American Indian and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US **(Citations)**.
+These shifts in the topics and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices [@dunivinBlackLivesMatter2022; @franklinFutureNowArchaeology2020; @flewellenFutureArchaeologyAntiracist2021; @lalukArchaeologySocialJustice2022].
+COVID-19 negatively impacted Black, American Indian and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US [@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021] <!-- JB: (Citations) BM: added -->.
 The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic and the murder of George Floyd in Minneapolis by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
@@
 
-Our data suggest that US universities participated in this movement by adjusting their hiring plans to prioritize recruiting archaeologists working on topics relevant to Black and Indigenous communities <!--
+Our data suggest that archaeology faculty at US universities participated in this movement by adjusting their hiring plans to prioritize recruiting archaeologists working on topics relevant to Black and Indigenous communities <!--
 We should be thoughtful about where we locate agency here.
@@
 A suggestion for an avenue for future research might be to see how these trends play out at a larger scale across the social sciences and humanities, or even STEM fields.
-Aaron Clauset's next project...-->.
+Aaron Clauset's next project...
+BM: ok, added 'archaeology faculty' to clarify that it's people in departments making these decisions.
+-->.
 Many universities may have hoped to hire Black and Indigenous archaeologists as part of their effort to tackle systemic racism.
 However as it is illegal in the US for universities use race or ethnicity as a primary factor in hiring (due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity), universities appear to have tailored the content of their job ads to focus on topics where they expect Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
-**Some commentary on whether and if Black/Indigenous researchers are more numerous in these fields, citing Maria Franklin, Sonya Atalay, Aja, Rachel Watson, etc.**
+<!-- Some commentary on whether and if Black/Indigenous researchers are more numerous in these fields, citing Maria Franklin, Sonya Atalay, Aja, Rachel Watson, etc.
+BM: I've looked in these articles and not found any relevant data.
+Also, I think this is going beyond the scope of our study -->
 
@@
 Over the course of the study period, we document a gradual decline in topics such as the archaeology of complex societies and archaeological science, the goegraphic foci of Mesoamerica and South America, and methods relating to landscape archaeology.
-These trends are harder to explain as we cannot link their start to a moment like COVID-19.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism **(Citations)**.
-This might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for compositional analysis.
-Other relevant factors may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation.
-We also anticipate a growing emphasis on more job-oriented degrees that prepare students for working in cultural heritage management or contract archaeology **(Bardolph, Henry, and White Under Review)**.
-A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience, which may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
+These trends are harder to explain as we cannot link their origins to a moment like COVID-19.
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological] <!-- JB: (Citations) BM: I added some, not my strong area, however -->.
+This might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
+Other factors relevant to this decline may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation.
+<!-- We also anticipate a growing emphasis on more job-oriented degrees that prepare students for working in cultural heritage management or contract archaeology (Bardolph, Henry, and White Under Review).
+BM: this sentence seems to me to be out of sequence here.
+Why are we doing anticipation here in this middle of this para explaining some past trends?
+--> A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience, which may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
 <!--Alternate hypothesis: The people who study ancient mesoamerican states NEVER RETIRE.
 See: Michael Smith.
+BM: heh!
 -->
@@
 <!--Eat the old-->.
-Consistent with results from other studies **(Citations)**, our study documented a growing demand for research and teaching statements.
+Consistent with results from other studies \[\@HelloTristesTropes2021\], our study documented a growing demand for research and teaching statements.
 Demand for diversity statements shows a unique trajectory, peaking in 2020-2021 and declining into the present.
-This may relate to the intersecting concerns emerging during the COVID-19 pandemic, which may have declined in job ads as the most urgent time of the pandemic moves into the past.
-<!-- I don't understand the second half of this sentence-->.Another factor here may be the debates surrounding the experiment with diversity statements in hiring at some UC campuses during 2016-2022.
+This may relate to the intersecting concerns emerging during the COVID-19 pandemic, which may have reached a peak in 2020-2021 and then declined over time, resulting in dropping requests for diversity statements, as the most urgent period of the pandemic moves into the past.
+<!-- I don't understand the second half of this sentence BM: edited in an attempt to clarify -->.
+Another factor here may be the debates surrounding the experiment with diversity statements in hiring at some UC campuses during 2016-2022.
 Universities that previously required a diversity statement dropped that requirement after 2022 [@guidenAreDiversityStatements].
@@
 In recognition of this burden, not only on applicants but also on colleagues writing letters of recommendation over and over to support applicants, many hiring committees now follow the recommendations of @dennisWorstAnthroJob2022 in only requesting names and letters of recommendation at later stages of the hiring process, if at all.
-**A line here about AAA recommendations to not ask for letters up front.**
+Showing sensitivity to this burden, in 2020 the American Anthropological Association issued guidance to academic departments that letters of recommendation should not be requested in the initial application, but only from short-listed candidates [@mallinLetsMakeAcademic2020] <!-- A line here about AAA recommendations to not ask for letters up front.
+BM: I looked for this statement and could only find broken links, so not sure if they are still into it.
+Added something anyway.
+-->
 
 A key limitation of our research is that it does not include an analysis of the academic profiles of those who were eventually hired based on this sample of job ads.
-Our results reveal collective aspirations for the future of the field, but only from those who are securely employed in it as tenured professors serving on hiring committees at US universities.
+Our results reveal collective aspirations for the future of the field, but only from those writing the job ads, typically faculty who are securely employed as tenured professors serving on hiring committees at US universities.
 Our results do not show how these aspirations worked out in the topical and geographic foci of the people who were eventually hired for these positions .
@@
 
-We thank NO ONE.
+We thank NO ONE. <!-- haha! so generous of you! -->
 

### 2025-01-20 analysis/paper/article.qmd
JB Round 1 minor edits and responses to BM comments, 20 Jan 2025.

@@
 
-<!--Ben, please note that for now I am keeping track of my references.
-I have bolded them in text below; I think I will want to learn how to use Zotero and incorporate the citations using your arcane Lovecraftian ritual language (BibTex?), so that for our next paper I am capable of participating in the entire workflow with less supervision.
-BM: ok great!
-Yes, BibTeX, I believe you will love it.
--->
-
 Attend any workshop, conference, or panel that includes early career researchers and the conversation will be steered inexorably towards the academic job market: who is hiring, who has attained a tenure-track position, and who is out of luck this season.
@@
 and Mackie & Rockwell.
-To reduce duplication, I've removed most of that from the intro, and  kept/shifted it into the background so we have a logical progression of general to more specific observations about the job market -->
+To reduce duplication, I've removed most of that from the intro, and  kept/shifted it into the background so we have a logical progression of general to more specific observations about the job market.
+JB: Duly noted, thanks for doing the legwork on that.
+-->
 
@@
 New faculty positions have dwindled, in part due to the removal of the of the Age Discrimination in Employment Act (ADEA) exemption in 1994 which prohibited mandatory retirement ages in higher education (**Earle and Delpo Kulow 2015**).
-When combined with the institution of tenure in American academia, the ADEA exemption allowed [faculty to stay in their posts for as long as they liked] the median age for faculty in the U.S.
+When combined with the institution of tenure in American academia, the ADEA exemption allowed faculty to stay in their posts for as long as they liked.
+The median age for faculty in the U.S.
 now ranks among the highest for all professions (**Citation**).
-In tandem with the gradual de-investment in American higher education since the **XXX (Mirowski 2011**) and the aftershocks of the 2008 recession, the ADEA exemption has cultivated an environment where new lines are few and far between.
+In tandem with the gradual de-investment in American higher education since 1980s **(Mirowski 2011**) and the aftershocks of the 2008 recession, the ADEA exemption has cultivated an environment where new lines are few and far between.
 Among biological anthropologists, @Passalacqua_2018 found a ratio of 0.81 PhDs to job academic advertisements per year, concluding that academic positions in biological anthropology are barely at sustainable levels.
 This echoes findings from other fields.
-In biomedical sciences, for example, there is one tenure-track position in the US for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
-**Citation of Larson et al.
-2014**.
+In biomedical sciences, for example, there is one tenure-track position in the US for approximately every 6.3 PhD graduates \[@Ghaffarzadegan_2015\].
+In engineering, Larson et al., have calculated that providing jobs for even 50% of PhD graduates would require the field to expand at an "improbable" rate of 14% per year (**2014: 747**).
+As they emphasize, "...
+the system in many places is saturated, far beyond capacity to absorb new PhDs in academia at the rates they are being produced" (**2014: 749**).
 
@@
 Data from the American Association of University Professors' 2021–22 faculty survey indicate that more than 60% of faculty positions in US universities were held by non-tenure-track full time or part time contingent faculty members [@colby2022annual].
-Contingent faculty positions are precarious because they provide low or no health and retirement benefits, little or nothing in the way of professional development, and lower salaries relative to tenure-track positions.
-PhD graduates who have a sequence of short-term academic positions are often disadvantaged financially, due to low compensation combined with expensive frequent relocation, socially due to isolation from family and community, and professionally due to being ineligible for many decision-making roles at the universities they work at [@platzerAcademicPrecarityAmerican2018].
+Contingent faculty positions are precarious because they provide low or no health and retirement benefits, limited opportunities for professional development, and lower salaries relative to tenure-track positions.
+PhD graduates who have a sequence of short-term academic positions are often disadvantaged financially, due to low compensation combined with expensive frequent relocation, socially due to isolation from family and community, and professionally due to being ineligible for many decision-making roles at the universities at which they work [@platzerAcademicPrecarityAmerican2018].
+<!--Well, this is depressingly familiar.
+-->
 
@@
 Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing].
-Top-ranking programmes are placing fewer than one in three graduates in tenure-track jobs [@mackieMarketShareAccounting2023].
+Top-ranking programs are placing fewer than one in three graduates in tenure-track jobs [@mackieMarketShareAccounting2023].
 Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
-This network analysis showed that programs with large endowments and with faculty who hold prestigious awards and are widely cited by other scholars produce the majority of tenured and tenure-track faculty in anthropology.
+This network analysis showed that programs with large endowments, with faculty who hold prestigious awards and are widely cited by other scholars, produce the majority of tenured and tenure-track faculty in anthropology.
 Hiring bias predicated on the prestige of specific institutions and programs is not unique to anthropology.
@@
 
-One way that some hiring committees are tackling these biases towards prestigious institutions is by providing detailed instructions to applicants on how to prepare their application materials.
+One way that some hiring committees are tackling these prestige biases is by providing detailed instructions to applicants on how to prepare their application materials.
 In theory, detailed applications will allow the hiring committee to focus on evaluating the accomplishments of candidates across common categories, rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program.
@@
 For example, twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy.
-In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported job ads from Oberlin College and Grinnell College with particularly extreme requirements <!--Are you opposed to naming and shaming these institutions, which were Oberlin College and Grinnell College?
-I applied for one of these positions, and have no issue referring to them by name...
-BM: nope, I've added their names -->, which respectively requested nine and thirteen documents from applicants.
+In their review of the 'worst job ads of 2021' @dennisWorstAnthroJob2022 reported job ads from Oberlin College and Grinnell College with particularly extreme requirements, which respectively requested nine and thirteen documents from applicants.
 This high burden on applicants disproportionately favors people with more time and financial resources to prepare the required materials.
 
-A notable requirement that has emerged in recent years is a diversity statement, where the applicant describes their knowledge of, prior contributions to, and future goals for advancing diversity, equity, inclusion.
+A notable requirement that has emerged in recent years is a diversity statement, where the applicant describes their knowledge of, prior contributions to, and future goals for advancing diversity, equity, and inclusion.
 This requirement has been much-debated largely due to an experiment in 2016–2022 at several University of California campuses that used diversity statements as the first cut for selecting candidates for tenure-track faculty positions [@soucek2021diversity].
@@
 These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted in a variety of approaches to evaluating diversity, even leading some universities to entirely omit requirements for diversity statements in job applications [@guidenAreDiversityStatements].
+<!--The Harvard PhD student staying with me right now told me that Harvard just dropped their diversity statement requirement, if we want a specific example.
+Not sure it's necessary, though.
+-->
 
@@
 centralizing?
-BM: I prefer the shortest and simplest word that will do the job --> ads originally posted in numerous different locations.
+BM: I prefer the shortest and simplest word that will do the job.
+JB: How very Strunk & White of you.
+We'll leave it as collective.--> ads originally posted in numerous different locations.
 Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position.
@@
 It's like saying "we found 5000 lithics, one random one was 45 mm long, another random piece was 12 mm wide, etc." readers don't expect that in the manuscript  --> We also recorded the type and number of documents requested in each ad (e.g.
-cover letter, CV, statements on research, teaching, diversity <!-- Were these all relatively easy to recognise?
-I’ve applied to places like Penn State, for example, that do a stupid combined research/teaching/diversity statement, just to make applicants do extra work, seemingly.
-BM: we treated these kinds of combo statements as separate requests, I don't know how many there were, I'm ok to omit that detail -->, syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad, if any.
+cover letter, CV, statements on research, teaching, diversity , syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad, if any.
 
@@
 We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023.
-We focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities <!--Ben, if we should provide the rationale for this, I don't remember what it was BM: I also don't recall, I don't see any justification for the US focus in Speakman et al or Kawa et al.
-and I can't think of a one that is either unverifiable or doesn't require heaps of additional work to validate.
-Leave it to the peer reviewers?
--->.
+We focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities .
 @fig-show-basic-plots shows the count of ads for each year.
@@
 The ratio of tenure-track to non-tenure-track positions is generally well above one; in other words, this data set is dominated by tenure-track positions .
-Only academic year 2013-2014 had more non-tenure-track positions than tenure-track positions <!--I know that this is a pedantic complaint, but we should be consistent about whether or not we hyphenate tenure-track BM: I've edited every instance so they have hyphens.
-How do you feel about the double hyphenated 'non-tenure-track'?
--->, which was followed by an upward trend peaking at 2018-2019 and then declining again into the present.
+Only academic year 2013-2014 had more non-tenure-track positions than tenure-track positions , which was followed by an upward trend peaking at 2018-2019 and then declining again into the present.
 
@@
 The implication here is that job-seekers who are able to relocate to populous areas will have more employment options.
-<!--Yes, nicely put BM: thanks!
--->
 
@@
 Panel A of @fig-show-geo-trends shows that a single region of the US, the Southwest, occurs in more job ads than every other part of the world except for the Mediterranean.
-<!-- Someone in another field recently asked me what archaeologists wear to conferences, and I said “bolo ties”.
-This is why.
-BM: haha, yes, exactly!
--->The Southwest includes Arizona and New Mexico, with portions of California, Colorado, Nevada, Oklahoma, Texas, and Utah.
+The Southwest includes Arizona and New Mexico, with portions of California, Colorado, Nevada, Oklahoma, Texas, and Utah.
 It is archaeologically significant as the home of the Ancestral Pueblo, Hohokam, and Mogollon peoples who practiced irrigation agriculture and lived in relatively large settlements, compared to other regions of the US.
-The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, also in relatively large numbers [@griffin2000native] <!-- BM: added requested citation: JB: Summary citation(s), preferably including an Indigenous archaeologist from the region -->.
+The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, also in relatively large numbers [@griffin2000native] <!-- BM: added requested citation: JB: Summary citation(s), preferably including an Indigenous archaeologist from the region.
+JB: Thanks Ben, though this was a note for myself, not you :/-->.
 The Mediterranean is prominent because it is the location that is often mentioned in job ads focused on classical archaeology (i.e.
@@
 Landscape archaeology, encompassing GIS and remote sensing, has remained prominent compared to other methods [@fig-show-metho-trends].
-<!--Might be worth exploring whether this is a skillset more frequently sought after in CRM and/or other industries—is this a transferable skill thing?
-BM: yes, good point, added something on that --> The popularity of this suite of methods may reflect its importance as a skill in demand by employers in the Cultural Research Management sector.
+The popularity of this suite of methods may reflect its importance as a skill in demand by employers in the Cultural Research Management sector.
 @morganReadyNotArchaeological2023 found that almost one-third of 599 jobs ads posted by CRM employers sought candidates with experience in GIS methods.
@@
 Computational and digital archaeology is the second most represented method, showing an overall increasing trend, particularly since 2020-2021.
-<!--Congratulations BM: haha, would you like to borrow my crystal ball?--> Archaeobotany shows a strong cyclical trend, rising, falling, and then rising again over our study period.
+Archaeobotany shows a strong cyclical trend, rising, falling, and then rising again over our study period.
 Archaeometry and Geoarchaeology have maintained a relatively low but steady presence in job ads, peaking in 2017-2018 and 2018-2019 and declining thereafter.
@@
 human origins, hunter-gatherer archaeology, and digital archaeology.
-<!-- BM: See, here is a nice unpacking of some of the key categories -->
 
@@
 Mentions of complex societies in ads decrease at a steady rate, while mentions of public archaeology consistently increase over the study period.
-The topic of Environmental Archaeology remains <!--We should be careful of the word consistent as we are relying on it a bit heavily throughout the text.
-Here, could we replace with "popular" or "stable"?
-BM: yes, good point, seems to work if I just delete it -->high over time.
+The topic of Environmental Archaeology remains high over time.
 
@@
 Over our ten year study period there have been substantial changes in the instructions to applicants in terms of the type and number of documents that are requested by the search committees [@fig-requirements-over-time].
-Requests for cover letter and CV decline in more recent years, perhaps reflecting their status as a generic and expected component to submit without needing to be specifically requested <!--So curious as to what these jobs DO request.
-BM: I'm not confident this is a real or important trend, but I did look into the schools that don't ask for these, there are mostly lesser-known places like Texas Tech University and East Tennessee State University.
-I think there might be a few other factors here, for example the job ads on the wiki might have that generic text edited out because it's expected by everyone.
-Difficult to test this because the links the interfolio, etc original source text for most ads are broken now, I did check a dozen of them just now.
-In any case, I don't think any further work on it would be of much relevance to our main arguments -->.
+Requests for cover letter and CV decline in more recent years, perhaps reflecting their status as a generic and expected component to submit without needing to be specifically requested.
 The requirement for a diversity statement is rare until 2019-2020, peaks around 2020-2021, then decreases towards the present.
@@
 It seems like there is an introductory sentence missing here.
-BM: ok, I tried to expand it a bit -->.
+BM: ok, I tried to expand it a bit.
+JB: Yes, this flows now.
+Thx.
+-->.
 While these dynamics are familiar to scholars actively seeking faculty jobs, we believe this is the first time they have been quantified at such a large scale within archaeology.
@@
 This is at odds with previous work that has shown that contingent positions make up more than half of faculty positions [@colby2022annual].
-**MORE DETAIL HERE.** <!-- BM: what do you want more detail on?
-The previous work is Colby 2022, discussed in the background section already.
-I've added the citation to that here also--> This discrepancy may be due to the more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
+**MORE DETAIL HERE FROM JB.** This discrepancy may be due to the more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
 Many of these short-term positions are not advertised nationally, but only through local email lists and are filled by people close to the hiring department, such as recently graduated students.
@@
 The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus [@woolstonJuniorResearchersHit2020].
-A survey of early career researchers in archaeology captures the impact of this dip, with three-quarters of respondents experiencing negative impacts on their careers due to the pandemic [@bramiPrecariousFutureReflections2023] <!-- JB: “75.2 per cent of respondents consider their career to have been negatively affected by the pandemic, whether through lost opportunities, illness, unemployment, or negative impact on mental health (Figure 6)” (Brami et al., 2023: 233).
-BM: I put a sentence in for this for you, but I think it's mostly a supply-side detail, and our focus is demand-side, so I don't feel it's highly relevant here and don't want too much detail on it -->
+A survey of early career researchers in archaeology captures the impact of this dip, with three-quarters of respondents experiencing negative impacts on their careers due to the pandemic [@bramiPrecariousFutureReflections2023]
 
@@
 These shifts in the topics and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices [@dunivinBlackLivesMatter2022; @franklinFutureNowArchaeology2020; @flewellenFutureArchaeologyAntiracist2021; @lalukArchaeologySocialJustice2022].
-COVID-19 negatively impacted Black, American Indian and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US [@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021] <!-- JB: (Citations) BM: added -->.
+COVID-19 negatively impacted Black, American Indian and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US [@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021].
 The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic and the murder of George Floyd in Minneapolis by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
@@
 
-Our data suggest that archaeology faculty at US universities participated in this movement by adjusting their hiring plans to prioritize recruiting archaeologists working on topics relevant to Black and Indigenous communities <!--
-We should be thoughtful about where we locate agency here.
-Are institutions doing this, or departments?
-Likely it is some combination of both, but a lot of these calls are coming from departments.
-A suggestion for an avenue for future research might be to see how these trends play out at a larger scale across the social sciences and humanities, or even STEM fields.
-Aaron Clauset's next project...
-BM: ok, added 'archaeology faculty' to clarify that it's people in departments making these decisions.
--->.
+Our data suggest that archaeology faculty at US universities participated in this movement by adjusting their hiring plans to prioritize recruiting archaeologists working on topics relevant to Black and Indigenous communities.
 Many universities may have hoped to hire Black and Indigenous archaeologists as part of their effort to tackle systemic racism.
@@
 This trend is especially pronounced for Assistant Professor positions, which make more demands on applicants than Associate and Full Professor positions.
-<!--Eat the old-->.
 Consistent with results from other studies \[\@HelloTristesTropes2021\], our study documented a growing demand for research and teaching statements.
@@
 This may relate to the intersecting concerns emerging during the COVID-19 pandemic, which may have reached a peak in 2020-2021 and then declined over time, resulting in dropping requests for diversity statements, as the most urgent period of the pandemic moves into the past.
-<!-- I don't understand the second half of this sentence BM: edited in an attempt to clarify -->.
 Another factor here may be the debates surrounding the experiment with diversity statements in hiring at some UC campuses during 2016-2022.
@@
 Added something anyway.
--->
+JB: I am finding it here under Job Board Policies: https://employers.americananthro.org/static-page/10285/job-board-policies/?-->
 
@@
 
-We thank NO ONE. <!-- haha! so generous of you! -->
+We thank NO ONE. <!-- haha! so generous of you! JB: We should probably add some people here, no? -->
 

### 2025-01-20 analysis/paper/article.qmd
Second round of JB minor edits on 20 January 2025

@@
 This echoes findings from other fields.
-In biomedical sciences, for example, there is one tenure-track position in the US for approximately every 6.3 PhD graduates \[@Ghaffarzadegan_2015\].
+In biomedical sciences, for example, there is one tenure-track position in the US for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
 In engineering, Larson et al., have calculated that providing jobs for even 50% of PhD graduates would require the field to expand at an "improbable" rate of 14% per year (**2014: 747**).
@@
 PhD graduates who have a sequence of short-term academic positions are often disadvantaged financially, due to low compensation combined with expensive frequent relocation, socially due to isolation from family and community, and professionally due to being ineligible for many decision-making roles at the universities at which they work [@platzerAcademicPrecarityAmerican2018].
-<!--Well, this is depressingly familiar.
--->
 
@@
 These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted in a variety of approaches to evaluating diversity, even leading some universities to entirely omit requirements for diversity statements in job applications [@guidenAreDiversityStatements].
-<!--The Harvard PhD student staying with me right now told me that Harvard just dropped their diversity statement requirement, if we want a specific example.
-Not sure it's necessary, though.
--->
 
@@
 
-Our primary data source is the Archaeology Academic Jobs Wiki<!--Check if Kelsky has anything too say about this.
-I also think there have been some articles about the pros and cons of the poli sci or Econ versions of this.
-BM: I did have a search of her blog where there are ~20 mentions, all pretty neutral, her most interesting comment is to notes that it originated in the year of the Great Recession -->.
+Our primary data source is the Archaeology Academic Jobs Wiki.
 Originating in 2007, this is a set of freely accessible web pages that anyone can edit (anonymously or with a free user account) hosted by Fandom, a for-profit company.
@@
 Examples of methods used in this study include archaeobotany, lithic analysis, or zooarchaeology.
-<!--It might be helpful to make this into a summary table with examples of all foci generated by the study, and then one showing the simplified 10–15 topics.
-I understand this is in the supplementary tables already.
-BM: I've added in the 'for example' you requested, but feels very repetitive to give examples in this paragraph, and then the full set of topics etc.
-in the next paragraph, and add then add a table as well.
-It's like saying "we found 5000 lithics, one random one was 45 mm long, another random piece was 12 mm wide, etc." readers don't expect that in the manuscript  --> We also recorded the type and number of documents requested in each ad (e.g.
+We also recorded the type and number of documents requested in each ad (e.g.
 cover letter, CV, statements on research, teaching, diversity , syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad, if any.
@@
 Full details of the category reduction, showing the mapping between phrases in the job ads and our categories, are in our Supplementary Materials.
-<!-- I think it would be much easier for our readers to parse this info as a table, unless you have any objection.
-This would also help us rein in our word count after all my contributions BM: I think it's best in the Supplement, these will be too distracting in the main text.
-We do unpack the most important categories in detail in results section.
-Let's see what the peer reviewers say --> Our final, reduced topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
+Our final, reduced topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
 Our methods categories were: Archaeobotany, Archaeometry, Bioarchaeology, Ceramic analysis, Computational and Digital archaeology, Geoarchaeology, Landscape analysis, Lithic analysis, Material culture analysis, and Zooarchaeology.
@@
 In the years 2019-2020 and 2020-2021 there are striking changes in the popularity of topics in job ads.
-Indigenous and historical archaeology (which includes archaeology of the African diaspora and enslaved people <!--African diaspora archaeology?
-We should be careful to use the appropriate language here BM: yes, ok, added that, but many jobs ads literally mention slavery and enslaved people, so I want to make it clear that slavery is a notable topic in American archaeology, not something to sanitize and ignore under the gloss of historical archaeology generally being about white people, a personal obsession, I guess.
-Did you know there were African ex-slaves on the first fleet to Australia?
--->) became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
+Indigenous and historical archaeology (which includes archaeology of the African diaspora and enslaved people became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
 Conversely, archaeological science, which was popular during 2012-2017, was rarely mentioned in job ads during 2019-2021.
@@
 Other factors relevant to this decline may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation.
-<!-- We also anticipate a growing emphasis on more job-oriented degrees that prepare students for working in cultural heritage management or contract archaeology (Bardolph, Henry, and White Under Review).
-BM: this sentence seems to me to be out of sequence here.
-Why are we doing anticipation here in this middle of this para explaining some past trends?
---> A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience, which may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
-<!--Alternate hypothesis: The people who study ancient mesoamerican states NEVER RETIRE.
-See: Michael Smith.
-BM: heh!
--->
+A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience, which may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
 
@@
 In recognition of this burden, not only on applicants but also on colleagues writing letters of recommendation over and over to support applicants, many hiring committees now follow the recommendations of @dennisWorstAnthroJob2022 in only requesting names and letters of recommendation at later stages of the hiring process, if at all.
-Showing sensitivity to this burden, in 2020 the American Anthropological Association issued guidance to academic departments that letters of recommendation should not be requested in the initial application, but only from short-listed candidates [@mallinLetsMakeAcademic2020] <!-- A line here about AAA recommendations to not ask for letters up front.
-BM: I looked for this statement and could only find broken links, so not sure if they are still into it.
-Added something anyway.
-JB: I am finding it here under Job Board Policies: https://employers.americananthro.org/static-page/10285/job-board-policies/?-->
+Showing sensitivity to this burden, in 2020 the American Anthropological Association issued guidance to academic departments that letters of recommendation should not be requested in the initial application, but only from short-listed candidates (**AAA Job Board Policies Citation)** [@mallinLetsMakeAcademic2020]
 

### 2025-01-21 analysis/paper/article.qmd
JB edits (final, hopefully, excepting references) on 21 January 2025

@@
 Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
-This network analysis showed that programs with large endowments, with faculty who hold prestigious awards and are widely cited by other scholars, produce the majority of tenured and tenure-track faculty in anthropology.
+This network analysis showed that programs with large endowments, with faculty who hold prestigious awards and are widely cited by other scholars, produce the majority of tenured and tenure-track faculty.
 Hiring bias predicated on the prestige of specific institutions and programs is not unique to anthropology.
-Targeted studies of the disciplines of sociology, communication, and operations research, and industrial systems engineering show similar trends (**Barnett et al., 2010; Castillo et al., 2018; Feeley and Tutzauer 2022; Nevin 2019**).
-**Broader analyses comparing X, Y and Z, and analyses while a recent analysis of XXX showed Z (Clauset et al., 2015;** [@wapmanQuantifyingHierarchyDynamics2022a]).
+Targeted studies of the disciplines of sociology, communication, and operations research, and industrial systems engineering show similar dynamics (**Barnett et al., 2010; Castillo, Myers and Chen, 2018; Feeley and Tutzauer 2022; Nevin 2019**).
+Broader analyses comparing computer science, business, and history likewise reveal \[parallel\] patterns among disparate disciplines across the humanities, social sciences, and STEM fields, in which "across disciplines, prestige hierarchies make the most accurate predictions of faculty placement" (Clauset et al., 2015:4).
+The most recent comprehensive meta-analysis of faculty placement dynamics in the United States, which examined employment and doctoral education of all tenure-track faculty at PhD-granting universities, underscored the outsized impact of institutional prestige: 80% of faculty trained in the US came from just 20.4% of institutions [@wapmanQuantifyingHierarchyDynamics2022a].
 Within anthropology, however, such studies draw attention to the paradox of a discipline committed to fighting social inequalities continuing to reproduce its own systematic inequalities through hiring practices that favor an elite minority of graduates with prestigious institutional affiliations.
@@
 These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted in a variety of approaches to evaluating diversity, even leading some universities to entirely omit requirements for diversity statements in job applications [@guidenAreDiversityStatements].
-
-**While attention to equity in job market materials is commendable, work on hiring practices suggests ZZZ.
-For example, Warren 2019 has shown XXX.
-Rivera, in a study of hiring practices at YYY, has likewise demonstrated ZZZ.
-Somewhere in this section I want to discuss the results from sociology, specifically the Warren 2019 piece and the Rivera 2017 piece on hiring practices.** <!-- ok, sounds good!
--->
-
-**Cite Rogge 1976.**
 
@@
 Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-14.
-This is at odds with previous work that has shown that contingent positions make up more than half of faculty positions [@colby2022annual].
-**MORE DETAIL HERE FROM JB.** This discrepancy may be due to the more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
+This is at odds with previous work that has shown that contingent positions make up more than half of faculty positions [@colby2022annual], and that these numbers are increasing between 2004 and 2019, there was a 36% increase in the replacement of tenure-track positions by contingent positions (**AAUP 2022**).
+This discrepancy in our data set may be due to the more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
 Many of these short-term positions are not advertised nationally, but only through local email lists and are filled by people close to the hiring department, such as recently graduated students.
@@
 However as it is illegal in the US for universities use race or ethnicity as a primary factor in hiring (due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity), universities appear to have tailored the content of their job ads to focus on topics where they expect Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
-<!-- Some commentary on whether and if Black/Indigenous researchers are more numerous in these fields, citing Maria Franklin, Sonya Atalay, Aja, Rachel Watson, etc.
-BM: I've looked in these articles and not found any relevant data.
-Also, I think this is going beyond the scope of our study -->
-
 This striking change in topics and geographic foci of 2019-2021 occurred against a backdrop of several longer-term trends.
@@
 These trends are harder to explain as we cannot link their origins to a moment like COVID-19.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological] <!-- JB: (Citations) BM: I added some, not my strong area, however -->.
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological] <!-- JB: (Citations) BM: I added some, not my strong area, however.
+JB: Do you think we could just cite Lars Fogelin's arch theory book here?-->.
 This might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
@@
 This trend is especially pronounced for Assistant Professor positions, which make more demands on applicants than Associate and Full Professor positions.
-Consistent with results from other studies \[\@HelloTristesTropes2021\], our study documented a growing demand for research and teaching statements.
+Consistent with results from other studies \[\@HelloTristesTropes2021\] <!-- JB: Ben, this does not seem to be "hyperlinked" the way other references are, just flagging.
+-->, our study documented a growing demand for research and teaching statements.
 Demand for diversity statements shows a unique trajectory, peaking in 2020-2021 and declining into the present.
@@
 
-We thank NO ONE. <!-- haha! so generous of you! JB: We should probably add some people here, no? -->
+<!--JB: We should probably add some people here, no? -->
 

### 2025-01-25 analysis/paper/article.qmd
Jess' additions to the reference list and final round of edits!

@@
 abstract: |
-  Many archaeology graduate students pursue advanced degrees in the hope of landing an academic career.
-Job listing websites often serve as the first place for these students when seeking academic positions.
+  Many archaeology graduate students pursue advanced degrees in the hope of undertaking an academic career.
+Job listing websites often serve as the first port-of-call for students  seeking academic positions.
 We examined tenure-track job advertisements over the past decade to gain insights into the academic job market for archaeologists.
-Using data from the community-edited Academic Jobs Wiki for Archaeology, we examine changes in the academic job market over time.
+Using data from the communit-edited Academic Jobs Wiki for Archaeology, we examine changes in the academic job market over time.
 We analyzed the text of 431 job ads posted from 2013--2023.
-Our analysis addresses the question of how have archaeological topics, methods, and geographic regions specified in archaeological job ads shifted over time?
+Our analysis addresses the question of how  archaeological topics, methods, and geographic regions specified in archaeological job ads shifted over time.
 We also investigate whether the labor burden for applicants has changed over time: do institutions request more information and documents from applicants at the initial stages of application, compared to a decade ago?
 We assess the influence of socio-political factors on the changing focus of research topics in the field.
-This research aims to assist current and future archaeology students and graduates in better understanding the job market and the requirements of employers, thereby aiding them in effectively preparing for their applications for positions in archaeology.
+This research aims to assist current and future archaeology students, graduates, and advisors to better understand the dynamics of the academic job market and the requirements of employers.
 keywords: |
@@
 Attend any workshop, conference, or panel that includes early career researchers and the conversation will be steered inexorably towards the academic job market: who is hiring, who has attained a tenure-track position, and who is out of luck this season.
-This obsessive focus on career trajectories is warranted, particularly within the discipline of anthropology.
-As of 2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions (**AAUP 2022**); today, 71% of faculty in the US are non-tenure-track(**Culver and Kezar 2022***).* The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors (**Beck, under review**), including decreasing federal support for higher education, concomitant increases in student debt (**Gusterson 2017**), a pronounced shift in university investment from faculty to administration (**Graeber 2018: 162–163**), and the growing privatization and market orientation of scientific research (**Mirowski 2011**).
+This obsessive focus on career trajectories is warranted.
+As of 2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions [@the20222022]; today, 71% of faculty in the US are non-tenure-track [@culver] .
+The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors [@beck], including decreasing federal support for higher education, concomitant increases in student debt [@gusterson2017], a pronounced shift in university investment from faculty to administration [@graeber2018, pp.
+162–163], and the growing privatization and market orientation of scientific research [@mirowski2011].
 
@@
 This competitive landscape has deeper roots than the last economic crisis; writing more than three decades ago, Rabinow pinpointed "an awareness...
-that standards have changed during the last thirty years and the quantitative and qualitative demands for entry into the system are immeasurably higher now" (**1992: 66**).
-
-These higher demands include increased overall productivity, higher mobility, sometimes requiring bothnational and international moves, more time spent in precarious short-term contracts, and the investment of a greater proportion of time and energy in applying for long-term positions.
+that standards have changed during the last thirty years and the quantitative and qualitative demands for entry into the system are immeasurably higher now" [-@rabinow1992, pp.66].
+
+These higher demands include increased overall productivity, higher national and international mobility, more time spent in precarious short-term contracts, and the investment of a greater amount of time and energy in applying for long-term positions.
 In their survey of early career researchers in European archaeology, for example, Brami et al.
-found that 52% of respondents who had obtained their PhD at least two years before the survey were on fixed-term contracts, while only 9% were in permanent positions (**2023: Figure 5**).
-In their words, "the queue of eager post-docs hoping for a long-term appointment is getting longer, as is the average time spent in the 'queue'" (**2023: 239**).
-Ribeiro and Giamakis emphasize that this demographic comprises an academic precariat that is essential to the functioning of universities, but consists of a workforce on short-term contracts perpetually "stranded between employment and unemployment" (**2023:10**).
-
-Early career scholars with faculty ambitions are thus stuck in a double-bind, forced to funnel their energy into seemingly endless applications for which they receive little to no feedback and which remain invisible on their CVs.
+found that 52% of respondents who had obtained their PhD at least two years before the survey were on fixed-term contracts, while only 9% were in permanent positions [-@bramiPrecariousFutureReflections2023, Figure 5].
+In their words, "the queue of eager post-docs hoping for a long-term appointment is getting longer, as is the average time spent in the 'queue'" [-@bramiPrecariousFutureReflections2023,239].
+Ribeiro and Giamakis emphasize that this demographic comprises an academic precariat that is essential to the functioning of universities, but consists of a workforce perpetually "stranded between employment and unemployment" [@ribeiro2023, pp.10].
+
+Junior scholars with faculty ambitions are thus stuck in a double-bind, forced to funnel their energy into seemingly endless applications for which they receive little to no feedback and which remain invisible on their CVs.
 As Dennis et al.
-underscore, early career researchers are "supplicants": "There is no way for applicants to point out that the job ads are taking too much time out of the scholarly community’s collective time bank" (**2022: 1**).
-Indeed, acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs to performing well in online video interviews and teaching demonstrations.
-These courses often provide valuable training in the hidden curriculum of the academy, but that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the "market" as a historically contingent set of cultural practices.
-As Rabinow maintains, "if we want ethical considerations to play a central role in the articulation of truth and power—and I think we do—then bringing such considerations into view is the necessary first step towards recognizing who we are today and setting out on the road to a better place" (**1992: 71**) **.**
+underscore, early career researchers are "supplicants": "There is no way for applicants to point out that the job ads are taking too much time out of the scholarly community’s collective time bank" [@dennisWorstAnthroJob2022, pp.1].
+Indeed, acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs, to performing well in long-list interviews and prepare teaching demonstrations.
+These courses can provide valuable training in the hidden curriculum of the academy, but that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the "market" as a historically contingent set of cultural practices.
+As Rabinow maintains, "if we want ethical considerations to play a central role in the articulation of truth and power—and I think we do—then bringing such considerations into view is the necessary first step towards recognizing who we are today and setting out on the road to a better place" [-@rabinow1992, pp.71] **.**
 
 We argue that to understand the dynamics of hiring in academic archaeology, we must start at the beginning, with the job ad itself.
-There is an abundance of recent work on who does and does not get hired into tenure-track positions in anthropology and archaeology (**Kawa 2019; Mackie and Rockwell 2023**), but less scholarship on how those hires unfold.
-Research on hiring dynamics has amply demonstrated that *where* you get your PhD has an enormous impact on if and where you get a tenure-track position (**Kawa et al.
-2019; Mackie and Rockwell 2023**), but how about *what* you study?
+There is an abundance of recent work on who does and does not get hired into tenure-track positions in anthropology and archaeology, but less scholarship on how those hires unfold.
+Research on hiring dynamics has amply demonstrated that *where* you get your PhD has an enormous impact on if and where you get a tenure-track position [@kawaSocialNetworkUS2019; @mackieMarketShareAccounting2023], but how about *what* you study?
 Are there topical, methodological, or regional specialties that are in consistent demand, or is there continual flux generated by intellectual fads, economic incentives, and political attention?
@@
 Job seekers must search for ads that are well-matched to the skill sets and professional experiences that they have taken years to cultivate.
-Hiring committees must sift through tens to hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realize their visions of their program and the discipline.
+Hiring committees must sift through tens to hundreds of applicants in the hope of finding a candidate that meets all of their needs and helps to realize their visions for their program and for the discipline.
 Finding the right candidate for the position—or the right position for the candidate—is akin to finding a needle in a haystack.
 
-There is also a perception among many scholars that the last three decades have seen the academic job market become increasingly competitive.
-Across the social sciences and humanities, there are fewer jobs available relative to the number of PhD graduates and higher numbers of short-term positions relative to permanent positions (**Bessner and Brenes 2021; Brami et al., 2023; Kawa et al.
-2019; Kelsky 2015; Mackie and Rockwell 2023**).
+Among many scholars, there is also a perception that the last three decades have seen the academic job market become increasingly competitive.
+Across the social sciences and humanities, there are fewer jobs available relative to the number of PhD graduates, and higher numbers of short-term positions relative to permanent positions[@bessner2021; @bramiPrecariousFutureReflections2023; @kawaSocialNetworkUS2019; @kelsky2015; @mackieMarketShareAccounting2023].
 This market saturation has arisen in tandem with increasingly complex application requirements.
@@
 
-In anthropology in the US the number of doctoral anthropology graduates has increased by about 70% over the past 30 years, but the number of new faculty positions has not increased proportionally [@speakman2018choosing].
-New faculty positions have dwindled, in part due to the removal of the of the Age Discrimination in Employment Act (ADEA) exemption in 1994 which prohibited mandatory retirement ages in higher education (**Earle and Delpo Kulow 2015**).
-When combined with the institution of tenure in American academia, the ADEA exemption allowed faculty to stay in their posts for as long as they liked.
+In American anthropology, the number of doctoral anthropology graduates has increased by about 70% over the past 30 years, but the number of new faculty positions has not increased proportionally [@speakman2018choosing].
+New faculty positions have dwindled, in part due to the removal of the of the Age Discrimination in Employment Act (ADEA) exemption in 1994 which prohibited mandatory retirement ages in higher education [@earle2014].
+When combined with the institution of tenure, the ADEA exemption allowed faculty to stay in their posts for as long as they liked.
 The median age for faculty in the U.S.
-now ranks among the highest for all professions (**Citation**).
-In tandem with the gradual de-investment in American higher education since 1980s **(Mirowski 2011**) and the aftershocks of the 2008 recession, the ADEA exemption has cultivated an environment where new lines are few and far between.
-Among biological anthropologists, @Passalacqua_2018 found a ratio of 0.81 PhDs to job academic advertisements per year, concluding that academic positions in biological anthropology are barely at sustainable levels.
+now ranks among the highest for all professions [@kaskie2016].
+In tandem with the gradual de-investment in American higher education since 1980s [@mirowski2011] and the aftershocks of the 2008 recession, the ADEA exemption has cultivated an environment where new lines are few and far between.
+Among biological anthropologists, for example, @Passalacqua_2018 found a ratio of 0.81 PhDs to job academic advertisements per year, concluding that academic positions in biological anthropology are barely at sustainable levels.
 This echoes findings from other fields.
-In biomedical sciences, for example, there is one tenure-track position in the US for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
-In engineering, Larson et al., have calculated that providing jobs for even 50% of PhD graduates would require the field to expand at an "improbable" rate of 14% per year (**2014: 747**).
+In biomedical sciences, there is one tenure-track position in the US for approximately every 6.3 PhD graduates \[@Ghaffarzadegan_2015\].
+In engineering, Larson et al., have calculated that providing jobs for even 50% of PhD graduates would require the field to expand at an "improbable" rate of 14% per year [-@larson2014, pp.747].
 As they emphasize, "...
-the system in many places is saturated, far beyond capacity to absorb new PhDs in academia at the rates they are being produced" (**2014: 749**).
+the system in many places is saturated, far beyond capacity to absorb new PhDs in academia at the rates they are being produced" [@larson2014, pp.749].
 
 PhD graduates who do not go directly into a tenure-track faculty position after graduation often go into a series of short-term appointments as part-time or limited contract instructors, also known as adjunct or contingent faculty.
-Rising numbers of contingent faculty has long been a concern in US higher education [@Trevithick2010].
+The rising number of contingent faculty has long been a concern in US higher education [@Trevithick2010].
 Data from the American Association of University Professors' 2021–22 faculty survey indicate that more than 60% of faculty positions in US universities were held by non-tenure-track full time or part time contingent faculty members [@colby2022annual].
 Contingent faculty positions are precarious because they provide low or no health and retirement benefits, limited opportunities for professional development, and lower salaries relative to tenure-track positions.
-PhD graduates who have a sequence of short-term academic positions are often disadvantaged financially, due to low compensation combined with expensive frequent relocation, socially due to isolation from family and community, and professionally due to being ineligible for many decision-making roles at the universities at which they work [@platzerAcademicPrecarityAmerican2018].
+PhD graduates who have a sequence of short-term academic positions are often disadvantaged financially due to low compensation combined with expensive frequent relocation, socially due to isolation from family and community, and professionally due to being ineligible for many decision-making roles at the universities at which they work [@platzerAcademicPrecarityAmerican2018].
 
@@
 Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
-This network analysis showed that programs with large endowments, with faculty who hold prestigious awards and are widely cited by other scholars, produce the majority of tenured and tenure-track faculty.
+This network analysis showed that programs with large endowments and faculty who hold prestigious awards and are widely cited by other scholars, produce the majority of tenured and tenure-track faculty.
+
 Hiring bias predicated on the prestige of specific institutions and programs is not unique to anthropology.
-Targeted studies of the disciplines of sociology, communication, and operations research, and industrial systems engineering show similar dynamics (**Barnett et al., 2010; Castillo, Myers and Chen, 2018; Feeley and Tutzauer 2022; Nevin 2019**).
-Broader analyses comparing computer science, business, and history likewise reveal \[parallel\] patterns among disparate disciplines across the humanities, social sciences, and STEM fields, in which "across disciplines, prestige hierarchies make the most accurate predictions of faculty placement" (Clauset et al., 2015:4).
-The most recent comprehensive meta-analysis of faculty placement dynamics in the United States, which examined employment and doctoral education of all tenure-track faculty at PhD-granting universities, underscored the outsized impact of institutional prestige: 80% of faculty trained in the US came from just 20.4% of institutions [@wapmanQuantifyingHierarchyDynamics2022a].
+Targeted studies of the disciplines of sociology, communication, and operations research, and industrial systems engineering show similar dynamics [@barnett2010; @castillo; @feeley2021; @nevin2019].
+Broader analyses comparing computer science, business, and history likewise comparable patterns among disparate disciplines across the humanities, social sciences, and STEM fields; "across disciplines, prestige hierarchies make the most accurate predictions of faculty placement" \[@clauset2015, pp.4\].
+The most recent comprehensive meta-analysis of faculty placement dynamics in the United States, which examined employment and doctoral education of all tenure-track faculty at PhD-granting universities, underscored the outsize impact of institutional prestige: 80% of faculty trained in the US came from just 20.4% of institutions \[@wapmanQuantifyingHierarchyDynamics2022a\].
 Within anthropology, however, such studies draw attention to the paradox of a discipline committed to fighting social inequalities continuing to reproduce its own systematic inequalities through hiring practices that favor an elite minority of graduates with prestigious institutional affiliations.
@@
 One way that some hiring committees are tackling these prestige biases is by providing detailed instructions to applicants on how to prepare their application materials.
-In theory, detailed applications will allow the hiring committee to focus on evaluating the accomplishments of candidates across common categories, rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program.
+In theory, detailed applications will allow the hiring committee to focus on evaluating the accomplishments of candidates across common categories rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program.
 This push for equity has resulted in job ads that are often highly prescriptive in the types of documents that applicants should submit.
-For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements detailing their previous and future contributions to teaching, research, and diversity within their discipline.
+For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements detailing their previous and future contributions to teaching, research, and diversity .
 In a comparison of job ads from 1999–2000 and 2019–2020 published in *Anthropology News* @HelloTristesTropes2021 noticed an increase in the number of materials requested from applicants.
@@
 The coordinators and contributors are nearly all anonymous or pseudonymous.
-Typically, contributors copy and paste all or some of the text of job ads from other sources, such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, into the wiki, collecting <!--consolidating?
+Typically, contributors copy and paste all or some of the text of job ads into the wiki, from sources such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, collecting <!--consolidating?
 centralizing?
@@
 These comments make the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market.
-Because of its reputation for aggregating ads from diverse sources and rapidly-updated information that is not available elsewhere, the Academic Jobs Wiki has a large community of users that keep it updated and accurate has become an authoritative data source for studies of hiring trends in academia [e.g.
-@Passalacqua_2018; @musial2018five] and a widely recommended resource [e.g.
+Because of its reputation for aggregating ads from diverse sources and rapidly-updated information that is not available elsewhere, the Academic Jobs Wiki has become an authoritative data source for studies of hiring trends in academia [e.g.
+@Passalacqua_2018; @musial2018five] and a widely recommended resource for applicants [e.g.
 @lightfoot2021preparing].
 
-For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023, we read the text and recorded the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methods foci on the position into a Google form.
-The topical focus is what we understood as the primary intellectual focus of the position—examples of topical foci included environmental archaeology, public archaeology, and North American archaeology.
-The geographic focus is the region of the world that the ideal candidate has scholarly expertise on, for example, Southwest US, Mediterranean, or Asia and India.
-The methods focus is the data-generating sub-field of archaeology that is mentioned in the ad.
-Examples of methods used in this study include archaeobotany, lithic analysis, or zooarchaeology.
+For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023, we read the text and recorded the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methodological foci of the position into a Google form.
+The topical focus is what we understood as the intellectual core of the position—examples of topical foci included environmental archaeology, public archaeology, and North American archaeology.
+The geographic focus is the region of the world about which the ideal candidate has scholarly expertise , for example, Southwest US, Mediterranean, or Asia and India.
+The methods focus is the data-generating sub-field of archaeology mentioned in the ad.
+Examples of methods used in this study include archaeobotany, lithic analysis, and zooarchaeology.
 We also recorded the type and number of documents requested in each ad (e.g.
-cover letter, CV, statements on research, teaching, diversity , syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad, if any.
-
-After completing primary data collection, we studied the topical, geographic, and methods text of each ad and collaboratively and manually reduced the variation in the raw data into 10-15 categories appearing in 20 (for topics and geography) or 10 (for methods) or more job ads to simplify analysis and visualisation.
+cover letter, CV, statements on research, teaching, diversity , syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad.
+
+After completing primary data collection, we studied the topical, geographic, and methods text of each ad and collaboratively and manually reduced the variation in the raw data into 10-15 categories appearing in 20 (for topics and geography) or 10 (for methods) or more job ads to simplify analysis and visualization.
 This means that some topics, such as gender (mentioned in `r n_jobs_mentioning_gender` ads) do not appear in our results because of their rarity in the job ads.
-Full details of the category reduction, showing the mapping between phrases in the job ads and our categories, are in our Supplementary Materials.
-Our final, reduced topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
+Full details of the category reduction, showing the mapping between phrases in the job ads and our categories, can be found in our Supplementary Materials.
+Our final topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology <!-- Not to be a pain in the ass, but here you capitalize both "Indigenous" and "Historical" and later you only capitalize "Indigenous".
+Can you pick one strategy and do a quick command F to make sure we are being consistent?
+-->, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
 Our methods categories were: Archaeobotany, Archaeometry, Bioarchaeology, Ceramic analysis, Computational and Digital archaeology, Geoarchaeology, Landscape analysis, Lithic analysis, Material culture analysis, and Zooarchaeology.
@@
 For example, Mesoamerican archaeology is understood to refer to a specific time period and geographic region.
-Similarly, we recorded digital archaeology we recorded as both a method (when a job ad had a clearly distinct topical focus, such as historic archaeology) and a topic (when there were no other topics mentioned in the job ad).
-While these categorical overlaps can make the data challenging to interpret, in our view this approach reflects the complex realities of how search committees express their needs in searching for new faculty.
+Similarly, we recorded digital archaeology as both a method (when a job ad had a clearly distinct topical focus, such as historic archaeology) and a topic (when there were no other topics mentioned in the job ad).
+While these categorical overlaps can make the data challenging to interpret, in our view this approach reflects the complex realities of how search committees express their needs when searching for new faculty.
 Acknowledgement of overlaps also produces new insights into hiring dynamics through revealing intersections between different foci.
 
-The entire R code (R Core Team, 2021) and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx to enable re-use of materials and improve reproducibility and transparency (Marwick, 2017).
+The entire R code (R Core Team, 2021) <!-- This citation, and Marwick 2017 below, do not seem to be input in the the BibTex style...
+--> and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx to enable re-use of materials and improve reproducibility and transparency (Marwick, 2017).
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
@@
 Assistant Professor jobs are consistently the most common title and rank of positions advertised, while open rank or full professor are the least frequent.
-The ratio of tenure-track to non-tenure-track positions is generally well above one; in other words, this data set is dominated by tenure-track positions .
-Only academic year 2013-2014 had more non-tenure-track positions than tenure-track positions , which was followed by an upward trend peaking at 2018-2019 and then declining again into the present.
+The ratio of tenure-track to non-tenure-track positions is generally well above one; in other words, this data set is dominated by tenure-track positions.
+Only academic year 2013-2014 had more non-tenure-track positions than tenure-track positions, which was followed by an upward trend peaking at 2018-2019 and then declining again into the present.
 
@@
 
-Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of institutions according to their Carnegie Classification, which is a framework for classifying US colleges and universities according to the types of degrees awarded, levels of activity such as research, and topical foci.
-Doctoral universities with high and very high research activity are by far the most active with hiring archaeology faculty.
+Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of institutions according to their Carnegie Classification, a framework for classifying US colleges and universities according to the types of degrees awarded, levels of activity such as research, and topical foci.
+Doctoral universities with high and very high research activity are by far the most active institutions hiring archaeology faculty.
 Associate's colleges, also known as community colleges, rarely post job ads for archaeology faculty.
@@
 California posted almost twice as many job ads as the next most active states.
-After California, the states that posted the most ads during 2013-2023 include New York, Texas, and Pennsylvania, and Florida.
-These top five states also correspond to the top five most populous US states, suggesting that rates of hiring are approximately proportional to population density.
-These top five states for job ads are also the top five states with highest number of degrees awarded in Anthropology (U.S.
-Department of Education nd).
+After California, the states that posted the most ads during 2013-2023 include New York, Texas, Pennsylvania, and Florida.
+These top five states correspond to the top five most populous US states, suggesting that rates of hiring are approximately proportional to population density.
+The top five states for job ads are also the five states with highest number of degrees awarded in Anthropology (U.S.
+Department of Education nd)<!--Need BibTex citation here -->.
 Similarly, the lowest counts of job ads were observed in states with the lowest populations: North Dakota, South Dakota, Alaska, and Nebraska.
@@
 The Southwest includes Arizona and New Mexico, with portions of California, Colorado, Nevada, Oklahoma, Texas, and Utah.
-It is archaeologically significant as the home of the Ancestral Pueblo, Hohokam, and Mogollon peoples who practiced irrigation agriculture and lived in relatively large settlements, compared to other regions of the US.
-The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, also in relatively large numbers [@griffin2000native] <!-- BM: added requested citation: JB: Summary citation(s), preferably including an Indigenous archaeologist from the region.
+It is archaeologically significant as the home of the Ancestral Pueblo, Hohokam, and Mogollon peoples, who practiced irrigation agriculture and lived in relatively large settlements compared to other regions of the US.
+The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, groups who had similarly high population densities[@griffin2000native] <!-- BM: added requested citation: JB: Summary citation(s), preferably including an Indigenous archaeologist from the region.
 JB: Thanks Ben, though this was a note for myself, not you :/-->.
 The Mediterranean is prominent because it is the location that is often mentioned in job ads focused on classical archaeology (i.e.
-archaeology of Bronze and Iron Age Italy and Greece).
-
-Demand for jobs focusing on the Americas has been generally high over the past decade, with a peak in 2019-2020, and a subsequent decrease.
+archaeology of Bronze Age and Iron Age Italy and Greece).
+
+Demand for jobs focusing on the Americas has been generally high over the past decade, with a peak in 2019-2020 and a subsequent decrease.
 Demand for jobs focusing on Africa was very low until 2019-2020, peaking in 2020-2021.
@@
 Asia and India, the Near East, and Europe rarely occur as a geographical focus in job ads at US institutions.
-Asia and India, Africa, and the Americas appear correlated with each other, while the Near East and Mediterranean are correlated in an opposite trend.
+Asia and India, Africa, and the Americas appear correlated with each other, while the Near East and Mediterranean are correlated in an opposite trend <!-- "are correlated in an opposite trend" is clunky.
+Can we say "show an inverse correlation", or something like that?
+-->.
 
@@
 Landscape archaeology, encompassing GIS and remote sensing, has remained prominent compared to other methods [@fig-show-metho-trends].
-The popularity of this suite of methods may reflect its importance as a skill in demand by employers in the Cultural Research Management sector.
+The popularity of this suite of methods may reflect its importance as a skill in demand by employers in the Cultural Research Management (CRM) sector.
 @morganReadyNotArchaeological2023 found that almost one-third of 599 jobs ads posted by CRM employers sought candidates with experience in GIS methods.
 Methods focused on a specific element of the archaeological record, such as Lithic analysis, Zooarchaeology and Ceramics are among the least frequently mentioned in job ads.
-Instead, the more popular methods are those that are relevant to multiple elements of the archaeological record (e.g.
+Instead, more popular methods are those that are relevant to multiple elements of the archaeological record (e.g.
 Archaeobotany encompasses macroscopic and microscopic plant remains; Bioarchaeology may include skeletal analysis, isotopes, proteins, etc.).
@@
 The most frequently mentioned topic in the job ads is Environmental archaeology [@fig-show-topi-trends].
-This category encompasses such phrases as human-environmental dynamics, interaction between humans and their environments, environmental change, climate change, historical ecology, ecological knowledge,human ecology, and ecological systems.
-Public archaeology is the second most frequent topic overall, which included phrases such as cultural resource management, cultural heritage, heritage studies, museum studies, human rights, community engaged, historic preservation, social justice, community-based, repatriation, and community-engaged archaeology.
-The least common topics in our sample are Pleistocene archaeology, e.g.
-human origins, hunter-gatherer archaeology, and digital archaeology.
-
-In the years 2019-2020 and 2020-2021 there are striking changes in the popularity of topics in job ads.
-Indigenous and historical archaeology (which includes archaeology of the African diaspora and enslaved people became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
+This category encompasses such phrases as human-environmental dynamics, interaction between humans and their environments, environmental change, climate change, historical ecology, ecological knowledge, human ecology, and ecological systems.
+Public archaeology is the second most frequent topic overall; this category included phrases such as cultural resource management, cultural heritage, heritage studies, museum studies, human rights, community engaged, historic preservation, social justice, community-based, repatriation, and community-engaged archaeology.
+The least common topics in our sample are Pleistocene archaeology (e.g.
+human origins, hunter-gatherer archaeology) and digital archaeology.
+
+The years 2019-2020 and 2020-2021 show striking changes in the popularity of topics in job ads.
+Indigenous and historical archaeology (which includes archaeology of the African diaspora and enslaved people) became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
 Conversely, archaeological science, which was popular during 2012-2017, was rarely mentioned in job ads during 2019-2021.
-Ancient Europe and the Mediterranean, which was an infrequently mentioned topic for the entire study period, virtually disappeared from job ads during 2019-2021.
+Ancient Europe and the Mediterranean, an infrequently mentioned topic for the entire study period, virtually disappeared from job ads during 2019-2021.
 Mentions of complex societies in ads decrease at a steady rate, while mentions of public archaeology consistently increase over the study period.
@@
 Complex societies and environmental archaeology were frequently found in the same ads.
-Biological archaeology, archaeological science, and evolutionary archaeology are another cluster of topics that frequently co-occur.
+Biological archaeology <!-- What is biological archaeology?
+Do you mean "bioarchaeology"?
+Are you getting the name of my specialty wrong?
+Ben!-->, Archaeological science, and Evolutionary archaeology are another cluster of topics that frequently co-occur.
 Other topics are relatively isolated.
@@
 
-Our study of a decade of tenure-track job ads in archaeology in the US reveals diverse dynamics in the demand for specialisations in topics, methods and geographic focus, and in the instructions to applicants to these jobs <!--What shifts?
+Our study of a decade of tenure-track job ads in archaeology in the US reveals diverse dynamics in the demand for specialisations in topics, methods and geographic focus, and in the instructions to applicants.
+<!--What shifts?
 It seems like there is an introductory sentence missing here.
@@
 Thx.
--->.
-While these dynamics are familiar to scholars actively seeking faculty jobs, we believe this is the first time they have been quantified at such a large scale within archaeology.
+--> While these dynamics are familiar to scholars actively seeking faculty positions, we believe this is the first time they have been quantified at such a large scale within archaeology.
 Trends in job ads reflect broader shifts in intellectual and practical priorities concerning archaeology, undergraduate education, and the process of hiring professors.
-The demand for archaeology faculty, indicated by the total number of tenure-track jobs, maybe explained by a variety of processes.
-Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-14.
-This is at odds with previous work that has shown that contingent positions make up more than half of faculty positions [@colby2022annual], and that these numbers are increasing between 2004 and 2019, there was a 36% increase in the replacement of tenure-track positions by contingent positions (**AAUP 2022**).
+The demand for archaeology faculty, indicated by the total number of tenure-track jobs, maybe affected by a variety of factors.
+Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-2014.
+This is at odds with previous work that has shown that contingent positions make up more than half of faculty positions [@colby2022annual], and that these numbers are increasing— between 2004 and 2019, for example, there was a 36% increase in the replacement of tenure-track positions by contingent positions [@the20222022].
 This discrepancy in our data set may be due to the more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
@@
 One possible explanation for our results is that non-tenure-track jobs are less frequently added to the Academic Jobs Wiki.
-The downward trend in tenure-track positions during 2013-2019 may be related to declining undergraduate enrollment in anthropology fields since 2013 [@crambChangingProfileTenureTrack2022a].
+The downward trend in tenure-track positions during 2013-2019 may be related to declining undergraduate enrollment in anthropology since 2013 [@crambChangingProfileTenureTrack2022a].
 The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus [@woolstonJuniorResearchersHit2020].
@@
 The 2019-2021 period was also a major inflection point in the popularity of specific topics and regions in job ads.
-Calls for positions incorporating Indigenous and historical archaeology and archaeology of the Americas became far more frequent at this time, while archaeological science, complex societies, the Mediterranean and Near East and showed declines in popularity.
+Calls for positions incorporating Indigenous and historical archaeology and archaeology of the Americas became far more frequent at this time, while archaeological science, complex societies, the Mediterranean and Near East and showed declines in popularity <!--Should some of these words be capitalized since they are topics?-->.
 Similarly, the number of job ads with a geographic focus on the Americas and Africa peaks during 2019–2021.
-These shifts in the topics and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices [@dunivinBlackLivesMatter2022; @franklinFutureNowArchaeology2020; @flewellenFutureArchaeologyAntiracist2021; @lalukArchaeologySocialJustice2022].
-COVID-19 negatively impacted Black, American Indian and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US [@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021].
+These shifts in the topical and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices [@dunivinBlackLivesMatter2022; @franklinFutureNowArchaeology2020; @flewellenFutureArchaeologyAntiracist2021; @lalukArchaeologySocialJustice2022].
+COVID-19 negatively impacted Black, American Indian <!-- Native American?
+--> and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US [@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021].
 The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic and the murder of George Floyd in Minneapolis by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
@@
 Many universities may have hoped to hire Black and Indigenous archaeologists as part of their effort to tackle systemic racism.
-However as it is illegal in the US for universities use race or ethnicity as a primary factor in hiring (due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity), universities appear to have tailored the content of their job ads to focus on topics where they expect Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
+Due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity, however,it is illegal in the US for universities use race or ethnicity as a primary factor in hiring.
+As a result, universities appear to have tailored the content of their job ads to focus on topics where they expect Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
+
 This striking change in topics and geographic foci of 2019-2021 occurred against a backdrop of several longer-term trends.
-Over the course of the study period, we document a gradual decline in topics such as the archaeology of complex societies and archaeological science, the goegraphic foci of Mesoamerica and South America, and methods relating to landscape archaeology.
-These trends are harder to explain as we cannot link their origins to a moment like COVID-19.
+Over the course of the study period, we document a gradual decline in topics such as the archaeology of complex societies and archaeological science, the geographic foci of Mesoamerica and South America, and methods relating to landscape archaeology.
+These trends are harder to explain as we cannot link their origins to a historical event like the COVID-19 pandemic.
 We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological] <!-- JB: (Citations) BM: I added some, not my strong area, however.
-JB: Do you think we could just cite Lars Fogelin's arch theory book here?-->.
-This might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
+JB: Do you think we could just cite Lars Fogelin's arch theory book here?
+https://www.academia.edu/109310616/An_Unauthorized_Companion_to_American_Archaeological_Theory_PDF-->.
+This theoretical trend might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
 Other factors relevant to this decline may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation.
-A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past, increased interest in climate change, environmental sustainability, and resilience, which may have shifted attention away from the study of monumental architecture, elite societies, political hierarchies and state systems.
+A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past and an increased concern with climate change, environmental sustainability, and resilience, shifting attention away from the study of monumental architecture, elite societies, political hierarchies, and state systems.
 
@@
 
-Our data on the requirements for applicants support prior findings that the complexity of applications—and concomitantly, the labor required to apply for a tenure-track job—has gradually increased over time.
+Our data on the requirements for applicants support prior findings that the complexity of applications—and concomitantly, the labor required to apply for tenure-track jobs—has gradually increased over time.
 This trend is especially pronounced for Assistant Professor positions, which make more demands on applicants than Associate and Full Professor positions.
-Consistent with results from other studies \[\@HelloTristesTropes2021\] <!-- JB: Ben, this does not seem to be "hyperlinked" the way other references are, just flagging.
--->, our study documented a growing demand for research and teaching statements.
+Consistent with results from other studies [@HelloTristesTropes2021], our project documented a growing demand for research and teaching statements.
 Demand for diversity statements shows a unique trajectory, peaking in 2020-2021 and declining into the present.
-This may relate to the intersecting concerns emerging during the COVID-19 pandemic, which may have reached a peak in 2020-2021 and then declined over time, resulting in dropping requests for diversity statements, as the most urgent period of the pandemic moves into the past.
+This may relate to the intersecting concerns <!-- interescting concerns about what?
+This statement will be stronger if it is tied to a specific argument.
+About positionality?
+About race and identity?
+About race and social class?
+--> emerging during the COVID-19 pandemic, which seem to have reached a peak in 2020-2021 and then declined over time, resulting in jettisoning requests for diversity statements as the most urgent period of the pandemic moves into the past.
 Another factor here may be the debates surrounding the experiment with diversity statements in hiring at some UC campuses during 2016-2022.
@@
 In recognition of this burden, not only on applicants but also on colleagues writing letters of recommendation over and over to support applicants, many hiring committees now follow the recommendations of @dennisWorstAnthroJob2022 in only requesting names and letters of recommendation at later stages of the hiring process, if at all.
-Showing sensitivity to this burden, in 2020 the American Anthropological Association issued guidance to academic departments that letters of recommendation should not be requested in the initial application, but only from short-listed candidates (**AAA Job Board Policies Citation)** [@mallinLetsMakeAcademic2020]
+Showing sensitivity to this burden, in 2020 the American Anthropological Association issued guidance to academic departments that letters of recommendation should not be requested in the initial application, but should only be required from short-listed candidates[@jobboar2020; @mallinLetsMakeAcademic2020].
 
 A key limitation of our research is that it does not include an analysis of the academic profiles of those who were eventually hired based on this sample of job ads.
-Our results reveal collective aspirations for the future of the field, but only from those writing the job ads, typically faculty who are securely employed as tenured professors serving on hiring committees at US universities.
-Our results do not show how these aspirations worked out in the topical and geographic foci of the people who were eventually hired for these positions .
-Future work should consider interviewing faculty hired during our study period to match up people to the ads they responded to.
-If the successful applicants can be identified, then the match between the details of the job ad and the applicant's research can be analyzed.
-Then we can assess how effective the hiring process is for driving change in the discipline by setting the topics of courses that will be taught in undergraduate and graduate curricula, and research that will be supported by universities.
+Our results reveal collective aspirations for the future of the field, but only from those writing the job ads, typically faculty who are securely employed as tenured professors serving on hiring committees at US colleges and universities.
+Our results do not show how these aspirations worked out in the topical and geographic foci of the people who were actually hired for these positions.
+Future work should consider interviewing faculty hired during our study period to match up scholars witgh the ads to which they applied.
+If the successful applicants can be identified, then we could analyze the fit between the details of the job ad and the applicant's research.
+Such an analysis would allow us to assess how effective job ads are for driving change in the discipline by setting the topics of courses that will be taught in undergraduate and graduate curricula, and research that will be supported by universities.
 We also recognize the possibility that the Academic Jobs Wiki might not capture all available positions or might over-represent certain types of institutions.
-Future work should evaluate this by comparing entries on the Wiki to other sources of information about the academic job market, such as the American Anthropological Association's AnthroGuide or data collected directly from universities.
+Future work should evaluate this by comparing entries on the Wiki to other sources of information about the academic job market, such as the American Anthropological Association's AnthroGuide, or data collected directly from universities.
 
@@
 Our results show that job ads offer a nuanced record of disciplinary history, documenting shifting trends in topics, methods, and geographic foci, with higher chronological resolution than many other sources, such as undergraduate textbooks [@lyman2010american].
-The short periods of time spanning the rise and fall of some of the foci we observed has implications for prospective graduate students.
+The short periods of time spanning the rise and fall of some of the foci we observed have implications for prospective graduate students.
 A topic that is growing in popularity as a student begins their PhD may have peaked and be in decline before they graduate.
-Methods seem to have a much lower frequency of change in popularity, relative to topical and geographic foci.
-One possible implication of this is that a graduate student who has invested in developing technical expertise in a method during their studies, in addition to a topic and region, might be less exposed to the whims of the job market than a student without a distinct area of technical expertise.
-
-Jobs ads also demonstrate how archaeologists draw on current events and politics to prioritize topics, regions and methods in their hiring plans.
-This is consistent with prior findings, such @wilk1985ancient's analysis of publications that found trends in explanations of Mayan collapse during the 1960s-1970s that correspond with major contemporary American political and social issues.
-This responsiveness to current events reveals that many archaeologists accept that our narratives of the past are entangled in the social and political milieu in which we work.
+Methods seem to have a much lower frequency of change in popularity relative to topical and geographic foci.
+One possible implication of this finding is that a graduate student who has invested in developing technical expertise in a method during their studies, in addition to a topic and region, might be less vulnerable to the vagaries of the job market than a student without a distinct area of technical expertise.
+
+Jobs ads also demonstrate how archaeologists draw on current events and politics to prioritize topics, regions, and methods in their hiring plans.
+This result is consistent with prior findings, such @wilk1985ancient's analysis of publications that identified trends in the explanations of Mayan <!-- Based on some pedantic friends, I think the adjective here may be Maya; I would check --> collapse during the 1960s-1970s that corresponded with major contemporary American political and social issues.
+This continual responsiveness to current events demonstrates that many archaeologists accept that our narratives of the past are entangled in the social and political milieu in which we work.
 Furthermore, it shows an active effort by archaeologists to take control of interpreting the past.

### 2025-01-25 analysis/paper/article.qmd
review some comments

@@
 
-```{=html}
-<!--Comments on Abstract and Keywords: 
-1.
-Is “jobs listing website” the appropriate category here?
-I often hear these referred to as “jobs boards”—that might be a better fit?
-BM: I agree that 'job board' is in common use, but I prefer 'jobs listing website' because it's more precise and explicit about the medium and contents
-
-2.
-Structurally, the second research domain is phrased as a question.
-For consistency, it would be good to provide a similar structure for the first domain, e.g.
-“Our analysis focuses on the XXX—have the topics, methods, or geographic regions specified in archaeological job ads changed over time?” BM: agreed, I updated the abstract
-
-3. Suggested keywords: archaeology, hiring, academia, job ads BM: added, you can also freely add these kinds of things directly, no worries -->
-```
-
 Keywords: `r rmarkdown::metadata$keywords`
@@
   str_replace("2021-22",  "2021-2022")
- # fix for 2021-22 DONE!
- # fix for 2023 DONE!
 
@@
 jobdata <- jobdata_us_only
-
 ```
@@
 The red line indicates a 1:1 ratio of tenure-track to non-tenure-track positions, bars taller than that line indicate more tenure-track than non-tenure-track positions in that year"
-#JB: I have added the last sentence above because I have not seen this kind of figure in an archaeology paper, so I think we should explain it.
-Will reduce friction for our readers.Please confirm that I am actually correctly explaining how the figure works, however.
-BM: I've edited the caption further to clarify how to read that red line.
-#JB: I would be eternally grateful if you would convert the hyphens between the timespans on the x-axis into en dashes 😬 BM: done, you may commence eternal gratitude :)
 
@@
 These courses can provide valuable training in the hidden curriculum of the academy, but that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the "market" as a historically contingent set of cultural practices.
-As Rabinow maintains, "if we want ethical considerations to play a central role in the articulation of truth and power—and I think we do—then bringing such considerations into view is the necessary first step towards recognizing who we are today and setting out on the road to a better place" [-@rabinow1992, pp.71] **.**
+As Rabinow maintains, "if we want ethical considerations to play a central role in the articulation of truth and power—and I think we do—then bringing such considerations into view is the necessary first step towards recognizing who we are today and setting out on the road to a better place" [-@rabinow1992, pp.71].
 
@@
 # Background
-
-<!-- BM: I think we have a little repetition in the Intro and Background sections, particulary in the mentions of data from papers by Speakman, Kawa et al.
-and Mackie & Rockwell.
-To reduce duplication, I've removed most of that from the intro, and  kept/shifted it into the background so we have a logical progression of general to more specific observations about the job market.
-JB: Duly noted, thanks for doing the legwork on that.
--->
 
@@
 This echoes findings from other fields.
-In biomedical sciences, there is one tenure-track position in the US for approximately every 6.3 PhD graduates \[@Ghaffarzadegan_2015\].
+In biomedical sciences, there is one tenure-track position in the US for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
 In engineering, Larson et al., have calculated that providing jobs for even 50% of PhD graduates would require the field to expand at an "improbable" rate of 14% per year [-@larson2014, pp.747].
@@
 Targeted studies of the disciplines of sociology, communication, and operations research, and industrial systems engineering show similar dynamics [@barnett2010; @castillo; @feeley2021; @nevin2019].
-Broader analyses comparing computer science, business, and history likewise comparable patterns among disparate disciplines across the humanities, social sciences, and STEM fields; "across disciplines, prestige hierarchies make the most accurate predictions of faculty placement" \[@clauset2015, pp.4\].
-The most recent comprehensive meta-analysis of faculty placement dynamics in the United States, which examined employment and doctoral education of all tenure-track faculty at PhD-granting universities, underscored the outsize impact of institutional prestige: 80% of faculty trained in the US came from just 20.4% of institutions \[@wapmanQuantifyingHierarchyDynamics2022a\].
+Broader analyses comparing computer science, business, and history likewise comparable patterns among disparate disciplines across the humanities, social sciences, and STEM fields; "across disciplines, prestige hierarchies make the most accurate predictions of faculty placement" [@clauset2015, pp.4].
+The most recent comprehensive meta-analysis of faculty placement dynamics in the United States, which examined employment and doctoral education of all tenure-track faculty at PhD-granting universities, underscored the outsize impact of institutional prestige: 80% of faculty trained in the US came from just 20.4% of institutions [@wapmanQuantifyingHierarchyDynamics2022a].
 Within anthropology, however, such studies draw attention to the paradox of a discipline committed to fighting social inequalities continuing to reproduce its own systematic inequalities through hiring practices that favor an elite minority of graduates with prestigious institutional affiliations.
@@
   filter(!is.na(basic2021)) %>% 
-  nrow() # 245, 290, 300, 313, 404, 430 / 433 for US unis
+  nrow()
 
@@
 B: Inset shows map of the United States showing the count of tenure-track job ads posted by all insititutions in each state during 2013 -- 2023"
-#JB: So many positions in New York! Why don’t I have a job in New York?!
-# JB: This figure is great. We should make sure it works for colour-blind readers.
 
@@
 }
-
-# BM TODO: check for job ads that have a location, but we're not getting it
 
@@
 
-# what about within the US
-# how many times each location mentioned?
+# what about within the US, how many times each location mentioned?
 p <- 
@@
 
-#JB: Interesting that Near East is on the up and up. You know what I am going to say about the timespans in the x-axis… BM: replaced hyphens with en-dash
-
 # show figure from PNG file
@@
 It is archaeologically significant as the home of the Ancestral Pueblo, Hohokam, and Mogollon peoples, who practiced irrigation agriculture and lived in relatively large settlements compared to other regions of the US.
-The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, groups who had similarly high population densities[@griffin2000native] <!-- BM: added requested citation: JB: Summary citation(s), preferably including an Indigenous archaeologist from the region.
-JB: Thanks Ben, though this was a note for myself, not you :/-->.
+The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, groups who had similarly high population densities [@griffin2000native].
 The Mediterranean is prominent because it is the location that is often mentioned in job ads focused on classical archaeology (i.e.
@@
 Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
-
-#JB: Please. The x-axis. My fragile mental state. BM: en-dashes added!
-
-#JB: Given that not all job ads are categorised by all foci, can we provide an overall n for Figure 3 for and 5, either in the captions or in the figures themselves?
-I assume that 20% of our “methods” sample is not the same as 20% of our “topics” sample.
-BM: what would the n be here, number of individual topics or methods recorded in total?
 
@@
 Our study of a decade of tenure-track job ads in archaeology in the US reveals diverse dynamics in the demand for specialisations in topics, methods and geographic focus, and in the instructions to applicants.
-<!--What shifts?
-It seems like there is an introductory sentence missing here.
-BM: ok, I tried to expand it a bit.
-JB: Yes, this flows now.
-Thx.
---> While these dynamics are familiar to scholars actively seeking faculty positions, we believe this is the first time they have been quantified at such a large scale within archaeology.
+While these dynamics are familiar to scholars actively seeking faculty positions, we believe this is the first time they have been quantified at such a large scale within archaeology.
 Trends in job ads reflect broader shifts in intellectual and practical priorities concerning archaeology, undergraduate education, and the process of hiring professors.
@@
 These shifts in the topical and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices [@dunivinBlackLivesMatter2022; @franklinFutureNowArchaeology2020; @flewellenFutureArchaeologyAntiracist2021; @lalukArchaeologySocialJustice2022].
-COVID-19 negatively impacted Black, American Indian <!-- Native American?
---> and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US [@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021].
+COVID-19 negatively impacted Black, Indigenous American, and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US [@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021].
 The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic and the murder of George Floyd in Minneapolis by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
@@
 JB: Do you think we could just cite Lars Fogelin's arch theory book here?
-https://www.academia.edu/109310616/An_Unauthorized_Companion_to_American_Archaeological_Theory_PDF-->.
+https://www.academia.edu/109310616/An_Unauthorized_Companion_to_American_Archaeological_Theory_PDF -->.
 This theoretical trend might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
@@
 Demand for diversity statements shows a unique trajectory, peaking in 2020-2021 and declining into the present.
-This may relate to the intersecting concerns <!-- interescting concerns about what?
-This statement will be stronger if it is tied to a specific argument.
-About positionality?
-About race and identity?
-About race and social class?
---> emerging during the COVID-19 pandemic, which seem to have reached a peak in 2020-2021 and then declined over time, resulting in jettisoning requests for diversity statements as the most urgent period of the pandemic moves into the past.
+This may relate to the intersecting concerns about race, identity and class inequalities emerging during the COVID-19 pandemic, which seem to have reached a peak in 2020-2021 and then declined over time, resulting in jettisoning requests for diversity statements as the most urgent period of the pandemic moves into the past.
 Another factor here may be the debates surrounding the experiment with diversity statements in hiring at some UC campuses during 2016-2022.
@@
 Jobs ads also demonstrate how archaeologists draw on current events and politics to prioritize topics, regions, and methods in their hiring plans.
-This result is consistent with prior findings, such @wilk1985ancient's analysis of publications that identified trends in the explanations of Mayan <!-- Based on some pedantic friends, I think the adjective here may be Maya; I would check --> collapse during the 1960s-1970s that corresponded with major contemporary American political and social issues.
+This result is consistent with prior findings, such @wilk1985ancient's analysis of publications that identified trends in the explanations of the Maya collapse during the 1960s-1970s that corresponded with major contemporary American political and social issues.
 This continual responsiveness to current events demonstrates that many archaeologists accept that our narratives of the past are entangled in the social and political milieu in which we work.
@@
 
-<!--JB: We should probably add some people here, no? -->
+Thanks to the anonymous contributors to the jobs Wiki
 

### 2025-01-28 analysis/paper/article.qmd
resolve merge conflict

--- 2025-01-25 analysis/paper/article.qmd
@@
 bibliography: references.bib
-csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
+csl: "../templates/society-for-american-archaeology.csl" # Insert path for the bib-style
 abstract: |
@@
   archaeology; hiring; academia; job ads 
-highlights: |
-  These are the highlights. 
 ---
@@
 Keywords: `r rmarkdown::metadata$keywords`
-
-Highlights: `r rmarkdown::metadata$highlights`
 
@@
 
-```{r}
-#| label: fig-show-basic-plots
-#| fig-cap: "A: total number of job ads posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank.
-B: Ratio of tenure-track to non-tenure-track positions over time.
-The red line indicates a 1:1 ratio of tenure-track to non-tenure-track positions, bars taller than that line indicate more tenure-track than non-tenure-track positions in that year"
-
-# show figure from PNG file
-knitr::include_graphics(here("analysis",
-            "figures", 
-            "fig-panel-per-year.png"))
-
-```
-
 # Introduction
@@
 This competitive landscape has deeper roots than the last economic crisis; writing more than three decades ago, Rabinow pinpointed "an awareness...
-that standards have changed during the last thirty years and the quantitative and qualitative demands for entry into the system are immeasurably higher now" [-@rabinow1992, pp.66].
+that standards have changed during the last thirty years and the quantitative and qualitative demands for entry into the system are immeasurably higher now" [-@rabinow1992, pp.
+66].
 
@@
 In their words, "the queue of eager post-docs hoping for a long-term appointment is getting longer, as is the average time spent in the 'queue'" [-@bramiPrecariousFutureReflections2023,239].
-Ribeiro and Giamakis emphasize that this demographic comprises an academic precariat that is essential to the functioning of universities, but consists of a workforce perpetually "stranded between employment and unemployment" [@ribeiro2023, pp.10].
+Ribeiro and Giamakis emphasize that this demographic comprises an academic precariat that is essential to the functioning of universities, but consists of a workforce perpetually "stranded between employment and unemployment" [@ribeiro2023, pp.
+10].
 
@@
 As Dennis et al.
-underscore, early career researchers are "supplicants": "There is no way for applicants to point out that the job ads are taking too much time out of the scholarly community’s collective time bank" [@dennisWorstAnthroJob2022, pp.1].
+underscore, early career researchers are "supplicants": "There is no way for applicants to point out that the job ads are taking too much time out of the scholarly community’s collective time bank" [@dennisWorstAnthroJob2022, pp.
+1].
 Indeed, acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs, to performing well in long-list interviews and prepare teaching demonstrations.
 These courses can provide valuable training in the hidden curriculum of the academy, but that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the "market" as a historically contingent set of cultural practices.
-As Rabinow maintains, "if we want ethical considerations to play a central role in the articulation of truth and power—and I think we do—then bringing such considerations into view is the necessary first step towards recognizing who we are today and setting out on the road to a better place" [-@rabinow1992, pp.71].
+As Rabinow maintains, "if we want ethical considerations to play a central role in the articulation of truth and power—and I think we do—then bringing such considerations into view is the necessary first step towards recognizing who we are today and setting out on the road to a better place" [-@rabinow1992, pp.
+71].
 
@@
 Top-ranking programs are placing fewer than one in three graduates in tenure-track jobs [@mackieMarketShareAccounting2023].
-Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
+Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 53 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
 This network analysis showed that programs with large endowments and faculty who hold prestigious awards and are widely cited by other scholars, produce the majority of tenured and tenure-track faculty.
@@
 Hiring bias predicated on the prestige of specific institutions and programs is not unique to anthropology.
-Targeted studies of the disciplines of sociology, communication, and operations research, and industrial systems engineering show similar dynamics [@barnett2010; @castillo; @feeley2021; @nevin2019].
-Broader analyses comparing computer science, business, and history likewise comparable patterns among disparate disciplines across the humanities, social sciences, and STEM fields; "across disciplines, prestige hierarchies make the most accurate predictions of faculty placement" [@clauset2015, pp.4].
-The most recent comprehensive meta-analysis of faculty placement dynamics in the United States, which examined employment and doctoral education of all tenure-track faculty at PhD-granting universities, underscored the outsize impact of institutional prestige: 80% of faculty trained in the US came from just 20.4% of institutions [@wapmanQuantifyingHierarchyDynamics2022a].
+Targeted studies of sociology, communication, and operations research, and industrial systems engineering show similar dynamics [@barnett2010; @castillo; @feeley2021; @nevin2019].
+Broader analyses comparing computer science, business, and history likewise comparable patterns among disparate disciplines across the humanities, social sciences, and STEM fields; "across disciplines, prestige hierarchies make the most accurate predictions of faculty placement" [@clauset2015, pp.
+4].
+The most recent comprehensive meta-analysis of faculty placement dynamics in the United States, which examined employment and doctoral education of all tenure-track faculty at PhD-granting universities, underscored the outsize impact of institutional prestige: 80% of faculty trained in the US came from just 20% of institutions [@wapmanQuantifyingHierarchyDynamics2022a].
 Within anthropology, however, such studies draw attention to the paradox of a discipline committed to fighting social inequalities continuing to reproduce its own systematic inequalities through hiring practices that favor an elite minority of graduates with prestigious institutional affiliations.
@@
 For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements detailing their previous and future contributions to teaching, research, and diversity .
-In a comparison of job ads from 1999–2000 and 2019–2020 published in *Anthropology News* @HelloTristesTropes2021 noticed an increase in the number of materials requested from applicants.
+In a comparison of job ads from 1999–2000 and 2019–2020 published in *Anthropology News*, @HelloTristesTropes2021 noticed an increase in the number of materials requested from applicants.
 For example, twice as many 2019–2020 job ads requested writing samples compared to 1999–2000, and nearly four times as many requested statements of teaching philosophy.
@@
 A notable requirement that has emerged in recent years is a diversity statement, where the applicant describes their knowledge of, prior contributions to, and future goals for advancing diversity, equity, and inclusion.
-This requirement has been much-debated largely due to an experiment in 2016–2022 at several University of California campuses that used diversity statements as the first cut for selecting candidates for tenure-track faculty positions [@soucek2021diversity].
+This requirement has been much-debated largely due to an experiment during 2016–2022 at several University of California campuses that used diversity statements as the first cut for selecting candidates for tenure-track faculty positions [@soucek2021diversity].
 Only candidates that scored highly on their diversity statements would have the rest of their application evaluated.
@@
 The coordinators and contributors are nearly all anonymous or pseudonymous.
-Typically, contributors copy and paste all or some of the text of job ads into the wiki, from sources such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, collecting <!--consolidating?
-centralizing?
-BM: I prefer the shortest and simplest word that will do the job.
-JB: How very Strunk & White of you.
-We'll leave it as collective.--> ads originally posted in numerous different locations.
+Typically, contributors copy and paste all or some of the text of job ads into the wiki, from sources such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, collecting ads originally posted in numerous different locations.
 Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position.
@@
 The topical focus is what we understood as the intellectual core of the position—examples of topical foci included environmental archaeology, public archaeology, and North American archaeology.
-The geographic focus is the region of the world about which the ideal candidate has scholarly expertise , for example, Southwest US, Mediterranean, or Asia and India.
+The geographic focus is the region of the world about which the ideal candidate has scholarly expertise, for example, Southwest US, Mediterranean, or Asia and India.
 The methods focus is the data-generating sub-field of archaeology mentioned in the ad.
@@
 
-After completing primary data collection, we studied the topical, geographic, and methods text of each ad and collaboratively and manually reduced the variation in the raw data into 10-15 categories appearing in 20 (for topics and geography) or 10 (for methods) or more job ads to simplify analysis and visualization.
+After completing primary data collection, we studied the topical, geographic, and methods text of each ad.
+Following the approach of @ryan2003techniques, we collaboratively and manually reduced the variation in the raw data into 10-15 categories appearing in at least 20 (for topics and geography) or 10 (for methods) job ads to simplify analysis and visualization.
 This means that some topics, such as gender (mentioned in `r n_jobs_mentioning_gender` ads) do not appear in our results because of their rarity in the job ads.
-Full details of the category reduction, showing the mapping between phrases in the job ads and our categories, can be found in our Supplementary Materials.
-Our final topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Biological anthropology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and Historical archaeology <!-- Not to be a pain in the ass, but here you capitalize both "Indigenous" and "Historical" and later you only capitalize "Indigenous".
-Can you pick one strategy and do a quick command F to make sure we are being consistent?
--->, North Mesoamerican Archaeology, Pleistocene archaeology, and Public archaeology Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
+Full details of the category reduction, showing the mapping between exact phrases found in the job ads and the categories we used for our analysis, can be found in our Supplementary Materials.
+Our final topic categories were: American archaeology, Ancient Europe and Mediterranean, Archaeological science, Archaeological theory, Bioarchaeology, Complex societies, Digital archaeology, Environmental archaeology, Evolutionary anthropology, Indigenous and historical archaeology, North Mesoamerican archaeology, Pleistocene archaeology, and Public archaeology.
+Our geographic categories were: Africa, Americas, Asia & India, Canada & Arctic, Europe, Mediterranean, Meso- & South America, Near East, Oceania, Midwest US, Northeastern US, Southeast US, Southwest US, and Western US.
 Our methods categories were: Archaeobotany, Archaeometry, Bioarchaeology, Ceramic analysis, Computational and Digital archaeology, Geoarchaeology, Landscape analysis, Lithic analysis, Material culture analysis, and Zooarchaeology.
-Ads could have multiple or none of these three foci, and some of the foci overlap.
+
+Individual ads could be recorded to have multiple or none of these topical, geographic, and methods foci, and some of the foci overlap.
 Some topics include geographic regions because this is how they are typically understood by archaeologists.
-For example, Mesoamerican archaeology is understood to refer to a specific time period and geographic region.
-Similarly, we recorded digital archaeology as both a method (when a job ad had a clearly distinct topical focus, such as historic archaeology) and a topic (when there were no other topics mentioned in the job ad).
-While these categorical overlaps can make the data challenging to interpret, in our view this approach reflects the complex realities of how search committees express their needs when searching for new faculty.
+For example, Mesoamerican archaeology is understood to refer to a specific time period _and_ a specific geographic region.
+Similarly, we recorded digital archaeology as both a method (when a job ad also had a clearly distinct topical focus, such as historic archaeology) and a topic (when there were no other topics mentioned in the job ad).
+While this polythetic approach results in categorical overlaps that can make the data challenging to interpret [@kuckartz2014qualitative], in our view this reflects the complex realities of how search committees express their needs when searching for new faculty.
 Acknowledgement of overlaps also produces new insights into hiring dynamics through revealing intersections between different foci.
 
-The entire R code (R Core Team, 2021) <!-- This citation, and Marwick 2017 below, do not seem to be input in the the BibTex style...
---> and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx to enable re-use of materials and improve reproducibility and transparency (Marwick, 2017).
+The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
@@
 We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023.
-We focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities .
+We focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities.
 @fig-show-basic-plots shows the count of ads for each year.
@@
 Only academic year 2013-2014 had more non-tenure-track positions than tenure-track positions, which was followed by an upward trend peaking at 2018-2019 and then declining again into the present.
+
+```{r}
+#| label: fig-show-basic-plots
+#| fig-cap: "A: total number of job ads posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank.
+B: Ratio of tenure-track to non-tenure-track positions over time.
+The red line indicates a 1:1 ratio of tenure-track to non-tenure-track positions, bars taller than that line indicate more tenure-track than non-tenure-track positions in that year"
+
+# show figure from PNG file
+knitr::include_graphics(here("analysis",
+            "figures", 
+            "fig-panel-per-year.png"))
+
+```
 
@@
 
-Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of institutions according to their Carnegie Classification, a framework for classifying US colleges and universities according to the types of degrees awarded, levels of activity such as research, and topical foci.
+Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of institutions according to their Carnegie Classification, a framework for classifying US colleges and universities according to the types of degrees awarded, levels of activity such as research, and topical foci [@shulman2001carnegie].
 Doctoral universities with high and very high research activity are by far the most active institutions hiring archaeology faculty.
@@
 These top five states correspond to the top five most populous US states, suggesting that rates of hiring are approximately proportional to population density.
-The top five states for job ads are also the five states with highest number of degrees awarded in Anthropology (U.S.
-Department of Education nd)<!--Need BibTex citation here -->.
+The top five states for job ads are also the five states with highest number of degrees awarded in Anthropology [@IPEDSNCES].
 Similarly, the lowest counts of job ads were observed in states with the lowest populations: North Dakota, South Dakota, Alaska, and Nebraska.
@@
 #| fig-cap: "A: Frequency of locations mentioned in the text of the job ads.
-B: Popularity of locations in job ads over time.
-Individual data points are show, overlain by a locally weighted regression line for each location to indicate temporal trends."
+B: Popularity of locations in job ads over time for locations that appear in 20 or more ads.
+Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
 
@@
 Asia and India, the Near East, and Europe rarely occur as a geographical focus in job ads at US institutions.
-Asia and India, Africa, and the Americas appear correlated with each other, while the Near East and Mediterranean are correlated in an opposite trend <!-- "are correlated in an opposite trend" is clunky.
-Can we say "show an inverse correlation", or something like that?
--->.
+Asia and India, Africa, and the Americas appear correlated with each other, while the Near East and Mediterranean are inversely correlated in an opposite trend.
 
@@
 #| fig-cap: "A: Frequency of methods mentioned in the text of the job ads.
-B: Popularity of methods in job ads over time.
-Individual data points are show, overlain by a locally weighted regression line for each location to indicate temporal trends."
+B: Popularity of methods in job ads over time for methods that appear in 10 or more ads.
+Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
 
@@
 Instead, more popular methods are those that are relevant to multiple elements of the archaeological record (e.g.
-Archaeobotany encompasses macroscopic and microscopic plant remains; Bioarchaeology may include skeletal analysis, isotopes, proteins, etc.).
+Archaeobotany encompasses macroscopic and microscopic plant remains; Bioarchaeology often includes skeletal analysis, isotopes, proteins, etc.).
 
@@
                       "Mesoamerican Archaeology",
-                      "Biological anthropology",
+                      "Bioarchaeology",
                       "Archaeological theory",
@@
 
-In our sample, job ads were more topically rich than geographically or methodologically rich.
+In our sample job ads were more topically rich than geographically or methodologically rich.
 That is, ads were more likely to mention multiple topics than they were to mention multiple methods or geographic locations.
-A Kruskal-Wallis test indicated significantly higher richness in topics compared to richness of geographic locations or methods in job ads (χ2 (df = `r unname(kw_test_richness$df)`, N = `r nrow(kw_test_richness_input)`) = `r round(unname(kw_test_richness$statistic), 2)`, p = `r expSup(unname(kw_test_richness$p.value))`).
+A Kruskal-Wallis test indicated significantly higher richness in topics compared to richness of geographic locations or methods in job ads (χ2 (df = `r unname(kw_test_richness$parameter)`, N = `r nrow(kw_test_richness_input)`) = `r round(unname(kw_test_richness$statistic), 2)`, p = `r expSup(unname(kw_test_richness$p.value))`).
 @fig-show-cooc shows topic co-occurrences in our sample.
@@
 Complex societies and environmental archaeology were frequently found in the same ads.
-Biological archaeology <!-- What is biological archaeology?
-Do you mean "bioarchaeology"?
-Are you getting the name of my specialty wrong?
-Ben!-->, Archaeological science, and Evolutionary archaeology are another cluster of topics that frequently co-occur.
+Bioarchaeology, Archaeological science, and Evolutionary archaeology are another cluster of topics that frequently co-occur.
 Other topics are relatively isolated.
@@
 The 2019-2021 period was also a major inflection point in the popularity of specific topics and regions in job ads.
-Calls for positions incorporating Indigenous and historical archaeology and archaeology of the Americas became far more frequent at this time, while archaeological science, complex societies, the Mediterranean and Near East and showed declines in popularity <!--Should some of these words be capitalized since they are topics?-->.
+Calls for positions incorporating Indigenous and historical archaeology and archaeology of the Americas became far more frequent at this time, while Archaeological science, Complex societies, the Mediterranean and Near East and showed declines in popularity.
 Similarly, the number of job ads with a geographic focus on the Americas and Africa peaks during 2019–2021.
@@
 These trends are harder to explain as we cannot link their origins to a historical event like the COVID-19 pandemic.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological] <!-- JB: (Citations) BM: I added some, not my strong area, however.
-JB: Do you think we could just cite Lars Fogelin's arch theory book here?
-https://www.academia.edu/109310616/An_Unauthorized_Companion_to_American_Archaeological_Theory_PDF -->.
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019].
 This theoretical trend might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
@@
     filter(`2022` == "0") %>% 
-    mutate(prev_req_ds = rowSums(across(`2015`:`2019`), na.rm = TRUE)) %>% 
+   rowwise() %>% 
+    mutate(prev_req_ds = sum(across(c(`2015`, `2016`, `2017`, `2018`, `2019`, `2020`, `2021`)), 
+                                 na.rm = TRUE)) %>% 
     select(name_of_hiring_university, 
@@
 In recognition of this burden, not only on applicants but also on colleagues writing letters of recommendation over and over to support applicants, many hiring committees now follow the recommendations of @dennisWorstAnthroJob2022 in only requesting names and letters of recommendation at later stages of the hiring process, if at all.
-Showing sensitivity to this burden, in 2020 the American Anthropological Association issued guidance to academic departments that letters of recommendation should not be requested in the initial application, but should only be required from short-listed candidates[@jobboar2020; @mallinLetsMakeAcademic2020].
+Showing sensitivity to this burden, in 2020 the American Anthropological Association issued guidance to academic departments that letters of recommendation should not be requested in the initial application, but should only be required from short-listed candidates[@jobboar2020; @Youngling2020].
 
@@
 
-Thanks to the anonymous contributors to the jobs Wiki
+Thanks to the anonymous contributors to the Academic Jobs Wiki, without whose efforts to collect and organise hundreds of job ads, this paper would not exist.
+To those of who contributed to the wiki but don't have a tenure-track job, we hope this paper shows your efforts were not wasted, but were a contribution to more accurate expectation-setting for future generations of archaeologists.
+
+# Data Availability Statement
+
+The data that support the findings of this study are openly available in Zenodo at http://doi.org/xxx/xxx.
 
@@
 
-# References
+# References cited
 

### 2025-02-04 analysis/paper/article.qmd
add renv

@@
 
+Tenure-track positions are a type of academic job that typically involve a probationary period (e.g.
+5-7 years) culminating in an evaluation of research, teaching, and service, that if successful, results in permanent employment, job security and academic freedom.
 Attend any workshop, conference, or panel that includes early career researchers and the conversation will be steered inexorably towards the academic job market: who is hiring, who has attained a tenure-track position, and who is out of luck this season.
 This obsessive focus on career trajectories is warranted.
-As of 2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions [@the20222022]; today, 71% of faculty in the US are non-tenure-track [@culver] .
+As of 2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions [@the20222022]; today, 71% of faculty in the US are non-tenure-track [@culver].
 The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors [@beck], including decreasing federal support for higher education, concomitant increases in student debt [@gusterson2017], a pronounced shift in university investment from faculty to administration [@graeber2018, pp.
@@
 
+# Funding Statement
+
+This research received no specific grant funding from any funding agency, commercial or not-for-profit sectors.
+
+# Competing Interests Statement
+
+Competing interests: The authors declare none
+
+# Supplementary materials
+
+Our supplementary materials are available at http://doi.org/xxx/xxx.
+
 <!-- The following line inserts a page break  -->

### 2025-02-04 analysis/paper/article.qmd
add peer review link to draft Zenodo repo (manually uploaded from GitHub)

@@
 
-The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx [ for peer reviewers the link is https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug ] to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
@@
 The data that support the findings of this study are openly available in Zenodo at http://doi.org/xxx/xxx.
+For peer review the link is: https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug
 
@@
 Our supplementary materials are available at http://doi.org/xxx/xxx.
+For peer review the link is: https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug
 

### 2025-04-26 analysis/paper/article.qmd
Submitted to American Antiquity, after manual edits to cover page of docx
minor edits in response to peer reviews

@@
 We examined tenure-track job advertisements over the past decade to gain insights into the academic job market for archaeologists.
-Using data from the communit-edited Academic Jobs Wiki for Archaeology, we examine changes in the academic job market over time.
+Using data from the community-edited Academic Jobs Wiki for Archaeology, we examine changes in the academic job market over time.
 We analyzed the text of 431 job ads posted from 2013--2023.
@@
 # explore over time
-fig_prop_by_job_title_per_year <- 
+prop_by_job_title_per_year_tbl <- 
 jobdata %>%
@@
   count(job_title_simple) %>%
-  mutate(prop = n / sum(n)) %>%
-  ggplot() +
+  mutate(prop = n / sum(n))
+
+prop_by_job_title_per_year_tbl_wide <- 
+prop_by_job_title_per_year_tbl %>% 
+  select(-prop) %>% 
+  pivot_wider(names_from = year_ad_posted,
+              values_from = n) %>% 
+  mutate(across(everything(), .fns = ~replace_na(.,0)))
+
+fig_prop_by_job_title_per_year <- 
+  ggplot(prop_by_job_title_per_year_tbl) +
   aes(year_ad_posted,
@@
 
-# Introduction
-
+<!-- Do not use “INTRODUCTION” or “ABSTRACT” as headings. -->
+
+Attend any workshop, conference, or panel that includes early career researchers and the conversation will be steered inexorably towards the academic job market: who is hiring, who has attained a tenure-track position, and who is out of luck this season.
 Tenure-track positions are a type of academic job that typically involve a probationary period (e.g.
 5-7 years) culminating in an evaluation of research, teaching, and service, that if successful, results in permanent employment, job security and academic freedom.
-Attend any workshop, conference, or panel that includes early career researchers and the conversation will be steered inexorably towards the academic job market: who is hiring, who has attained a tenure-track position, and who is out of luck this season.
 This obsessive focus on career trajectories is warranted.
@@
 1].
-Indeed, acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs, to performing well in long-list interviews and prepare teaching demonstrations.
-These courses can provide valuable training in the hidden curriculum of the academy, but that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the "market" as a historically contingent set of cultural practices.
+Indeed, acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs, to performing well in long-list interviews and preparing teaching demonstrations.
+These courses can provide valuable training in the hidden curriculum of the academy, but the fact that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the "market" as a historically contingent set of cultural practices.
 As Rabinow maintains, "if we want ethical considerations to play a central role in the articulation of truth and power—and I think we do—then bringing such considerations into view is the necessary first step towards recognizing who we are today and setting out on the road to a better place" [-@rabinow1992, pp.
@@
 Hiring bias predicated on the prestige of specific institutions and programs is not unique to anthropology.
-Targeted studies of sociology, communication, and operations research, and industrial systems engineering show similar dynamics [@barnett2010; @castillo; @feeley2021; @nevin2019].
+Targeted studies of sociology, communication, operations research, and industrial systems engineering show similar dynamics [@barnett2010; @castillo; @feeley2021; @nevin2019].
 Broader analyses comparing computer science, business, and history likewise comparable patterns among disparate disciplines across the humanities, social sciences, and STEM fields; "across disciplines, prestige hierarchies make the most accurate predictions of faculty placement" [@clauset2015, pp.
@@
 This push for equity has resulted in job ads that are often highly prescriptive in the types of documents that applicants should submit.
-For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements detailing their previous and future contributions to teaching, research, and diversity .
+For example, in addition to a cover letter and CV, job ads in many fields now require applicants to submit short statements detailing their previous and future contributions to teaching, research, and diversity.
 In a comparison of job ads from 1999–2000 and 2019–2020 published in *Anthropology News*, @HelloTristesTropes2021 noticed an increase in the number of materials requested from applicants.
@@
                 bg.r = 0.1),
-            force = 0,
-            force_pull = 100,
-            size = 3)+
+           # force = 0,
+           force_pull = 2,
+            size = 5)+
   scale_x_continuous(breaks = NULL) +

### 2025-04-27 analysis/paper/sketches.R
exploring wiki life history

@@ -252,10 +252,312 @@ ggsave(here("analysis",
        dpi = 900) # make the image nice and crisp))
 
 
+#------------------------------------------------------
+# scrape wiki page to test claims about'dead by 2018'
+
+# starting from : ratio-tt-to-non-tt
+
+base_url_to_2019 <- "https://academicjobs.fandom.com/wiki/Archaeology_Jobs_"
+base_url_after_2020 <- "https://academicjobs.fandom.com/wiki/Archaeology_"
+
+# starts at 2010-2011
+# fix for 2021-22
+# base UR
+
+years_to_2019 <- map_chr(2012:2019, ~str_glue('{.x}-{.x +1}'))
+years_after_2020 <- map_chr(2020:2022, ~str_glue('{.x}-{.x +1}'))
+# though it seems to start at 2007-8: https://academicjobs.fandom.com/wiki/Archaeology_07-08
+
+# make a set of URLs for each page for each year
+urls_for_each_year <- c(str_glue('{base_url_to_2019}{years_to_2019}'),
+                        str_glue('{base_url_after_2020}{years_after_2020}')) %>%
+  str_replace("2021-2022", "2021-22")
+
+# get the history of edits page "?action=history&offset=&limit=10000"
+
+edits_page <- "?action=history&offset=&limit=10000"
+
+edits_urls_for_each_year <-
+  str_glue('{urls_for_each_year}',
+           '{edits_page}' )
+
+library(rvest)
+
+# make a table of key edit variables
+edits_urls_for_each_year_lst_tbl <-
+map(edits_urls_for_each_year,
+    ~{
+      # get webpage
+      pge <- read_html(.x)
+
+      # extract key variables for each edit
+      # into a table
+      tibble(
+      edit_date =  pge %>%
+        html_elements(".mw-changeslist-date") %>%
+        html_text2(),
+      edit_name =  pge %>%
+        html_elements(".mw-userlink bdi") %>%
+        html_text2(),
+      edit_size =  pge %>%
+        html_elements(".mw-plusminus-neg , .mw-plusminus-null , .mw-plusminus-pos") %>%
+        html_text2()
+      )
+    }
+)
+
+# handle getting the edit comment which tells us which section
+# of the page was edited, it's blank for many edits, but there
+# might be some interesting patterns in there
+edits_urls_for_each_year_lst_section <-
+  map(edits_urls_for_each_year,
+      ~read_html(.x) %>%
+        html_elements('#pagehistory li') %>%
+        html_text2()
+  )
+
[…]

### 2025-04-28 analysis/paper/sketches.R
more wiki exploration

@@ -375,6 +375,28 @@ bind_rows(edits_urls_for_each_year_lst_tbl_with_comments,
 
 # visualisations ----------------------------------
 
+ggplot(edits_for_each_year_tbl) +
+  aes(month(edit_month,
+            label = TRUE,
+            abbr = FALSE)) +
+  geom_bar() +
+  theme_minimal() +
+  facet_wrap(~ edit_year_only,
+             ncol = 1) +
+  labs(x = "", y = "Number of edits") +
+  theme(axis.text.x = element_text(angle = 90,
+                                   vjust = 0.5))
+
+ggplot(edits_for_each_year_tbl) +
+  aes(edit_month) +
+  geom_bar() +
+  theme_minimal() +
+  facet_wrap(~ edit_year_only,
+             ncol = 1) +
+  labs(x = "", y = "Number of edits") +
+  theme(axis.text.x = element_text(angle = 90,
+                                   vjust = 0.5))
+
 # Number of edits per year
 p_edits_per_year <-
 ggplot(edits_for_each_year_tbl) +
@@ -402,6 +424,19 @@ ggplot() +
   labs(x = "", y = "Number of editors\n(distinct usernames or IP addresses)") +
   theme(axis.text.x = element_text(angle = 90,
                                    vjust = 0.5))
+
+# number of edits per editor
+edits_for_each_year_tbl %>%
+  group_by(edit_name) %>%
+  tally(sort = TRUE) %>%
+  ggplot() +
+  aes(n) +
+  geom_histogram() +
+  scale_x_log10() +
+  scale_y_log10() +
+  labs(x = "Number of edits",
+       y = "Number of editors") +
+  theme_minimal()
 
 # Editor activity over time
 top_editors_per_year <-
@@ -442,6 +477,15 @@ edits_for_each_year_tbl %>%
                                          units = "days") / (365.25)) %>%
   mutate(first_edit_year = year(first_edit),
          last_edit_year =  year(last_edit))
+
+#  year of every editor's first edit
+ggplot(edits_for_each_year_tbl_span) +
+ aes(first_edit_year) +
+  geom_bar() +
+  theme_minimal() +
+  labs(x = "",
+       y = "Number of editors making their first edit")
+
 
 # Most editors are only active for less that one year
 p_editor_life_distr <-

### 2025-05-01 analysis/paper/article.qmd
increase text size and improve number placement on Fig 2 map

@@
 
+# save this as SVG, then edit by hand to position the numbers on the
+# map nicely, then export that hand-edited SVG as a png to show in the
+# paper 
 ggsave(here("analysis",
             "figures", 
-            "fig-map-and-carnegie-classification.png"),
+            "fig-map-and-carnegie-classification.svg"),
        plot = p_map_c,
@@
 
-# show figure from PNG file
+# show figure from PNG file, this is after hand edits to place the
+# numbers on the map nicely
 knitr::include_graphics(here("analysis",
             "figures",
-            "fig-map-and-carnegie-classification.png"))
+            "fig-map-and-carnegie-classification-hand-edit.png"))
 ```

### 2025-05-01 analysis/paper/sketches.R

@@ -255,6 +255,8 @@ ggsave(here("analysis",
 #------------------------------------------------------
 # scrape wiki page to test claims about'dead by 2018'
 
+library(tidyverse)
+
 # starting from : ratio-tt-to-non-tt
 
 base_url_to_2019 <- "https://academicjobs.fandom.com/wiki/Archaeology_Jobs_"
@@ -376,6 +378,16 @@ bind_rows(edits_urls_for_each_year_lst_tbl_with_comments,
 # visualisations ----------------------------------
 
 ggplot(edits_for_each_year_tbl) +
+  aes(as_date(edit_month)) +
+  geom_bar() +
+  theme_minimal() +
+  labs(x = "", y = "Number of edits") +
+  scale_x_date(date_breaks = "1 year",
+               date_labels = "%Y") +
+  theme(axis.text.x = element_text(angle = 90,
+                                   vjust = 0.5))
+
+ggplot(edits_for_each_year_tbl) +
   aes(month(edit_month,
             label = TRUE,
             abbr = FALSE)) +
@@ -392,12 +404,13 @@ ggplot(edits_for_each_year_tbl) +
   geom_bar() +
   theme_minimal() +
   facet_wrap(~ edit_year_only,
-             ncol = 1) +
+             ncol = 1,
+             scales = "free") +
   labs(x = "", y = "Number of edits") +
   theme(axis.text.x = element_text(angle = 90,
                                    vjust = 0.5))
 
-# Number of edits per year
+# Number of ediyear_ad_posted# Number of edits per year
 p_edits_per_year <-
 ggplot(edits_for_each_year_tbl) +
   aes(as_date(edit_year)) +

### 2025-05-01 analysis/paper/article.qmd
deleted plot and text on job ad requirements by position. R2#20

@@
 Requests for course descriptions, syllabus samples, teaching evaluations, transcripts and writing samples are consistently low over time (not shown here).
-
-```{r}
-#| label: requirements-by-position
-
-# do the requirements differ for associate positions 
-jobdata_requirements_by_rank <- 
-jobdata %>% 
-  mutate(position_title = case_when(
-    str_detect(title_of_position_tenure_track_jobs_only, 
-               "associate") ~ "associate",
-    str_detect(title_of_position_tenure_track_jobs_only, 
-               "assistant") ~ "assistant",
-    str_detect(title_of_position_tenure_track_jobs_only, 
-               "full") ~ "full"))  %>% 
-  select(position_title,
-         starts_with("documents_requested")) %>% 
-  pivot_longer(-position_title) %>% 
-  mutate(value = case_when(
-    value == "Not requested in the job ad" ~ 0,
-    value == "One" ~ 1,
-    value == "Two (e.g. two syllabi)" ~ 2,
-    value == "Three" ~ 3,
-    .default = 0
-  )) %>% 
-  filter(!is.na(position_title))   %>% 
-  mutate(position_title = str_to_title(position_title)) %>% 
-    # trim names a bit
-  mutate(name = str_remove(name, "documents_requested_")) %>% 
-  mutate(name = str_replace_all(name, "_", " "))
-
-jobdata_requirements_by_rank_means <- 
-  jobdata_requirements_by_rank %>% 
-  group_by(position_title,
-           name) %>% 
-  summarise(mean = mean(value))
-
-p_req_rank <- 
-ggplot() +
-    geom_jitter(data = jobdata_requirements_by_rank,
-                  aes(position_title, 
-                      value),
-                height = 0.05,
-                alpha = 0.1) +
-    geom_point(data = jobdata_requirements_by_rank_means,
-               aes(position_title,
-                   mean),
-               size = 4,
-               colour = "red") +
-  facet_wrap( ~ name,
-              scales = "free_y") +
-   scale_y_continuous(breaks = integer_breaks()) +
-  xlab("") +
-  ylab("Number requested in job ad") +
-  theme_minimal()
-
-ggsave(filename = here("analysis",
-            "figures", 
-            "fig-requirements-by-rank.png"),
-       plot = p_req_rank,
-       bg ="white",
-       h = 6, # experiment with h and w to get the right size and proportion 
-       w = 8,
-       units = "in",
-       dpi = 900) # make the image nice and crisp
-```
-
-```{r}
-#| label: fig-requirements-by-position
-#| fig-cap: "Variation in application requirements by academic position.
-Each point represents an individual job ad.
-The points have been jittered to avoid overlap.
-The red point is the mean value for each group."
-
-# show figure from PNG file
-knitr::include_graphics(here("analysis",
-            "figures",
-            "fig-requirements-by-rank.png"))
-```
-
-Instructions to applicants vary by the rank of the position advertised [@fig-requirements-by-position].
-Assistant professor positions tend to have the most complex instructions with many different types of documents required.
-Conversely, full professor positions have minimal requirements.
-The documents that are most frequently requested across all three ranks are cover letter and CV.
 

### 2025-05-01 analysis/paper/article.qmd
add table showing breakdown of jobs by type, rank, tenure status

@@
 library(ggbeeswarm)
+library(flextable)
 # This CSV file was downloaded from our data sheet here
@@
 @fig-show-basic-plots shows the count of ads for each year.
+@tbl-show-basic-counts shows the breakdown by different job types, ranks, and tenure status.
 Assistant Professor jobs are consistently the most common title and rank of positions advertised, while open rank or full professor are the least frequent.
@@
 
+```
+
+```{r}
+#| label: tbl-show-basic-counts
+#| tbl-cap: "Breakdown of counts of job ads by rank and tenure status.
+Values in parentheses are for the United States only.
+TT = tenure-track, NTT = non-tenure-track.
+Not all ads include unambiguous information about tenure status, so the sum of TT and NTT does not equal the sum of all jobs in all ranks."
+
+#---------------------------------------------
+# want a table to show values for fig 1
+
+prop_by_job_title_per_year_tbl_wide <-
+prop_by_job_title_per_year_tbl %>%
+  select(-prop) %>%
+  pivot_wider(names_from = job_title_simple,
+              values_from = n)
+
+prop_by_job_title_per_year_us_only_tbl <-
+jobdata_us_only %>%
+  group_by(year_ad_posted) %>%
+  count(job_title_simple) %>%
+  mutate(prop = n / sum(n)) %>%
+  select(-prop) %>%
+  pivot_wider(names_from = job_title_simple,
+              values_from = n)
+
+# combine tables with total (US) numbers
+
+# 1. Join the two tables by year, adding suffixes to distinguish columns
+combined_data <- left_join(
+  prop_by_job_title_per_year_tbl_wide,
+  prop_by_job_title_per_year_us_only_tbl,
+  by = "year_ad_posted",
+  suffix = c("_wide", "_us") # Suffixes for columns from each table
+)
+
+# 2. Identify the original value column names (excluding the join key)
+value_cols <- setdiff(names(prop_by_job_title_per_year_tbl_wide), "year_ad_posted")
+
+# 3. Mutate to create the combined columns
+#    Iterate through the original value column names
+for (col_name in value_cols) {
+  # Construct the names of the columns from the joined table
+  col_wide_name <- paste0(col_name, "_wide")
+  col_us_name <- paste0(col_name, "_us")
+
+  # Use mutate to create a new column (overwriting the original name)
+  # with the combined string "Value_Wide (Value_US)"
+  # Handle NA values by replacing them with a placeholder like "-"
+  combined_data <- combined_data %>%
+    mutate(
+      !!col_name := paste0(
+        # Coalesce NA to "-" before pasting
+        coalesce(as.character(!!sym(col_wide_name)), "-"),
+        " (",
+        coalesce(as.character(!!sym(col_us_name)), "-"),
+        ")"
+      )
+    )
+}
+
+# 4. Select only the year column and the newly created combined columns
+final_combined_table <- combined_data %>%
+  select(year_ad_posted, all_of(value_cols))
+
+# let's add TT and NTT columns to that
+count_of_tt_jobs_per_year_from_our_form_tbl <- 
+count_of_tt_jobs_per_year_from_our_form %>%
+  select(year_ad_posted = year,
+         TT = n_tt_jobs,
+         NTT = n_non_tt_jobs) %>%
+  left_join(final_combined_table) %>%
+  relocate(TT, NTT, .after = last_col()) %>%
+  relocate(`Full Professor`, .after = `Associate Professor`) %>%
+  rename( `Year ad posted` = year_ad_posted) %>%
+  # get numeric year, sort, then remove it
+  mutate(year_num = parse_number(str_sub(`Year ad posted`, 6L))) %>%
+  arrange(year_num) %>%
+  select(-year_num)
+
+ft <- flextable(count_of_tt_jobs_per_year_from_our_form_tbl)
+
+# Apply formatting for a clean, professional look similar to booktabs
+ft <- ft %>%
+  # Apply a theme with minimal lines, good header/footer separation
+  theme_booktabs() %>% 
+  # Adjust column widths to fit
+  autofit() %>% fit_to_width(7) %>% 
+  # Left-align all columns (adjust if needed)
+  align(align = "left", part = "all") %>% 
+  # Center numeric columns (adjust selector)
+  align(j = ~ . -year_ad_posted, align = "center", part = "all") %>% 
+  # Add some vertical padding
+  padding(padding.top = 4, padding.bottom = 4, part = "all")
+
+# Print the flextable object - it will render as a table in Word
+ft
 ```

### 2025-05-01 analysis/paper/sketches.R

@@ -616,7 +616,76 @@ plot_grid(
                     labels = c("C", "F"))
           )
 
+#---------------------------------------------
+# want a table to show values for fig 1
 
+prop_by_job_title_per_year_tbl_wide <-
+prop_by_job_title_per_year_tbl %>%
+  select(-prop) %>%
+  pivot_wider(names_from = job_title_simple,
+              values_from = n)
+
+prop_by_job_title_per_year_us_only_tbl <-
+jobdata_us_only %>%
+  group_by(year_ad_posted) %>%
+  count(job_title_simple) %>%
+  mutate(prop = n / sum(n)) %>%
+  select(-prop) %>%
+  pivot_wider(names_from = job_title_simple,
+              values_from = n)
+
+# combine tables with total (US) numbers
+
+# 1. Join the two tables by year, adding suffixes to distinguish columns
+combined_data <- left_join(
+  prop_by_job_title_per_year_tbl_wide,
+  prop_by_job_title_per_year_us_only_tbl,
+  by = "year_ad_posted",
+  suffix = c("_wide", "_us") # Suffixes for columns from each table
+)
+
+# 2. Identify the original value column names (excluding the join key)
+value_cols <- setdiff(names(prop_by_job_title_per_year_tbl_wide), "year_ad_posted")
+
+# 3. Mutate to create the combined columns
+#    Iterate through the original value column names
+for (col_name in value_cols) {
+  # Construct the names of the columns from the joined table
+  col_wide_name <- paste0(col_name, "_wide")
+  col_us_name <- paste0(col_name, "_us")
+
+  # Use mutate to create a new column (overwriting the original name)
+  # with the combined string "Value_Wide (Value_US)"
+  # Handle NA values by replacing them with a placeholder like "-"
+  combined_data <- combined_data %>%
+    mutate(
+      !!col_name := paste0(
+        # Coalesce NA to "-" before pasting
+        coalesce(as.character(!!sym(col_wide_name)), "-"),
+        " (",
+        coalesce(as.character(!!sym(col_us_name)), "-"),
+        ")"
+      )
+    )
+}
+
+# 4. Select only the year column and the newly created combined columns
+final_combined_table <- combined_data %>%
+  select(year_ad_posted, all_of(value_cols))
+
+# let's add TT and NTT columns to that
+count_of_tt_jobs_per_year_from_our_form %>%
[…]

### 2025-05-02 analysis/paper/article.qmd
updating section on background to the wiki

@@
 library(flextable)
+library(rvest)
 # This CSV file was downloaded from our data sheet here
@@
 # time the quarto document is rendered. So this code block is
-# eval: false to avoid repeated and unnesecary scraping
+# eval: false to avoid repeated and unnecessary scraping
 
@@
 
+```{r}
+#| eval: false
+
+#------------------------------------------------------
+# scrape wiki page to describe editing dyanmics,
+# save the results so we don't do this often
+
+# starting from : ratio-tt-to-non-tt above
+# get the history of edits page "?action=history&offset=&limit=10000"
+
+edits_page <- "?action=history&offset=&limit=10000"
+
+edits_urls_for_each_year <-
+  str_glue('{urls_for_each_year}',
+           '{edits_page}' )
+
+library(rvest)
+
+# scrape from the webpages and make a table of key edit variables
+# this takes a few minutes
+edits_urls_for_each_year_lst_tbl <-
+map(edits_urls_for_each_year,
+    ~{
+      # get webpage
+      pge <- read_html(.x)
+
+      # extract key variables for each edit
+      # into a table
+      tibble(
+      edit_date =  pge %>%
+        html_elements(".mw-changeslist-date") %>%
+        html_text2(),
+      edit_name =  pge %>%
+        html_elements(".mw-userlink bdi") %>%
+        html_text2(),
+      edit_size =  pge %>%
+        html_elements(".mw-plusminus-neg , .mw-plusminus-null , .mw-plusminus-pos") %>%
+        html_text2()
+      )
+    }
+)
+
+# handle getting the edit comment which tells us which section
+# of the page was edited, it's blank for many edits, but there
+# might be some interesting patterns in there
+edits_urls_for_each_year_lst_section <-
+  map(edits_urls_for_each_year,
+      ~read_html(.x) %>%
+        html_elements('#pagehistory li') %>%
+        html_text2()
+  )
+
+edits_urls_for_each_year_lst_section_tbl <-
+map(edits_urls_for_each_year_lst_section,
+    ~{
+      # Extract time and date
+      edit_date <-
+        stringr::str_extract(.x, "\\d{2}:\\d{2}, \\d{1,2} [A-Za-z]+ \\d{4}")
+
+      edit_size <- str_extract(.x, "bytes\\s([+-]?\\d+)") |>
+         str_remove("bytes\\s")
+
+      # Extract text between → and undo
+      edit_comment <- stringr::str_extract(.x, "→(.*?) undo")
+      edit_comment <- stringr::str_remove_all(edit_comment, "→| undo")  # clean up
+
+      # Combine into a data frame
+      tibble(edit_date,
+             edit_size,
+             edit_comment)
+    }
+)
+
+# join together key variables and edit comments
+edits_urls_for_each_year_lst_tbl_with_comments <-
+map2(edits_urls_for_each_year_lst_tbl,
+     edits_urls_for_each_year_lst_section_tbl,
+    ~ left_join(.x, .y,
+                relationship = "many-to-many") %>%
+      # deduplicate
+      distinct()
+)
+
+names(edits_urls_for_each_year_lst_tbl_with_comments) <-
+  c(years_to_2019,
+    years_after_2020)
+
+# combine list into one big data frame
+edits_for_each_year_tbl <-
+bind_rows(edits_urls_for_each_year_lst_tbl_with_comments,
+          .id = "year_ad_posted") %>%
+  mutate(
+    edit_size_num = parse_number(str_replace(edit_size, "−", "-")),
+    edit_time = parse_date_time(edit_date,
+                                 orders = "HM, d B Y"),
+    # round down to first day of month
+    edit_month = floor_date(edit_time,
+                            unit = "month"),
+    edit_year = floor_date(edit_month,
+                           unit = "year"),
+    edit_year_only = year(edit_month),
+    job_market_year = parse_number(str_sub(year_ad_posted,
+                              -4L))
+  ) %>%
+  # delete edits that were made after the job market year
+  filter(edit_year_only == job_market_year | edit_year == (job_market_year -1))
+
+write_csv(edits_for_each_year_tbl,
+          here::here("analysis/data/raw_data/edits_for_each_year_tbl.csv"))
+
+```
+
+```{r}
+#| label: fig-panel-wiki-characteristics
+#| fig-cap: "A: Total number of edits to each Wiki page (additions and deletions) per year.
+B: Number of unique editors, as represented by distinct usernames or IP addresses (recorded for editors without usernames), active per year.
+C: Active periods of individual editors, the black lines connect the date of the first and last edit of all editors that were active for  three months or more.
+D: Typical size of each edit per year, either addition or removal of text, one word is roughly 4-6 Bytes.
+E: Distribution of the numbers of edits per job ad for each year.
+F: Histogram showing that the majority of editors are only active on the site for less than one year."
+
+edits_for_each_year_tbl <- read_csv(
+          here::here("analysis/data/raw_data/edits_for_each_year_tbl.csv"))
+
+# visualisations ----------------------------------
+
+# Number of edits per year
+p_edits_per_year <-
+ggplot(edits_for_each_year_tbl) +
+  aes(as_date(edit_year)) +
+  geom_bar() +
+  theme_minimal() +
+  scale_x_date(date_breaks = "1 year",
+               date_labels = "%Y") +
+  labs(x = "", y = "Number of edits") +
+  theme(axis.text.x = element_text(angle = 90,
+                                   vjust = 0.5))
+
+# Number of unique editors per year
+p_editors_per_year <-
+edits_for_each_year_tbl %>%
+  group_by(edit_year) %>%
+  summarise(n_distinct_editors = n_distinct(edit_name)) %>%
+ggplot() +
+  aes(as_date(edit_year),
+      n_distinct_editors) +
+  geom_col() +
+  theme_minimal() +
+  scale_x_date(date_breaks = "1 year",
+               date_labels = "%Y") +
+  labs(x = "", y = "Number of editors\n(distinct usernames or IP addresses)") +
+  theme(axis.text.x = element_text(angle = 90,
+                                   vjust = 0.5))
+
+# Editor activity over time
+
+# life span of an editor, from first edit to last edit
+edits_for_each_year_tbl_span <-
+edits_for_each_year_tbl %>%
+  group_by(edit_name) %>%
+  summarise(first_edit = min(edit_month),
+            last_edit =  max(edit_month),
+            edit_span = last_edit - first_edit,
+            edit_span_years = as.numeric(edit_span,
+                                         units = "days") / (365.25)) %>%
+  mutate(first_edit_year = year(first_edit),
+         last_edit_year =  year(last_edit))
+
+# Most editors are only active for less that one year
+p_editor_life_distr <-
+ggplot(edits_for_each_year_tbl_span) +
+  aes(edit_span_years) +
+  geom_histogram() +
+  theme_minimal() +
+  scale_y_log10() +
+  labs(x = "Years editing", y = "Number of editors")
+
+# how many active for less than one year?
+edits_for_each_year_tbl_span_tbl <- 
+edits_for_each_year_tbl_span %>% 
+  group_by(edit_span_years) %>% 
+  summarise(n = n()) %>% 
+  mutate(perc = n / sum(n) * 100)
+
+edits_for_each_year_tbl_span_tbl_less_than_one_year <-
+edits_for_each_year_tbl_span_tbl %>% 
+  mutate(sum_less_than_one = if_else(cumsum(edit_span_years) < 1, n, 0)) %>% pull(sum_less_than_one) %>% sum()
+
+edits_for_each_year_tbl_span_tbl_less_than_one_year_perc <- 
+  round(
+    edits_for_each_year_tbl_span_tbl_less_than_one_year /  sum(edits_for_each_year_tbl_span_tbl$n) * 100, 0)
+
+# for those editors active for more than three months:
+p_editor_life_distr_year <-
+edits_for_each_year_tbl_span %>%
+ filter(edit_span_years > 0.25) %>%
+ggplot() +
+  aes(as_date(first_edit_year),
+      reorder(edit_name, first_edit_year)) +
+  geom_segment(aes(x =   first_edit,
+                   xend = last_edit),
+               linewidth = 1,
+               lineend = "butt") +
+  theme_minimal() +
+  labs(x = "", y = "Individual editors\n(usernames and IP addresses redacted)") +
+  theme(axis.text.y=element_blank())
+
+# size of edits per year
+p_edit_size_per_year <-
+edits_for_each_year_tbl %>%
+  group_by(edit_year) %>%
+  summarise(edit_size_num = sum(edit_size_num)) %>%
+  ggplot() +
+  aes(as_date(edit_year),
+      edit_size_num) +
+  geom_col() +
+  theme_minimal() +
+  scale_x_date(date_breaks = "1 year",
+               date_labels = "%Y") +
+  labs(x = "", y = "Sum of edits (Bytes)") +
+  theme(axis.text.x = element_text(angle = 90,
+                                   vjust = 0.5))
+
+# typical edit size per year
+p_median_edit_size_per_year <-
+edits_for_each_year_tbl %>%
+  group_by(edit_year) %>%
+  # median, because sometimes there are 1-2 very large
+  # page reformatting edits
+  summarise(edit_size_num = median(edit_size_num)) %>%
+  ggplot() +
+  aes(as_date(edit_year),
+      edit_size_num) +
+  geom_col() +
+  theme_minimal() +
+  scale_x_date(date_breaks = "1 year",
+               date_labels = "%Y") +
+  labs(x = "", y = "Median edit size (Bytes)") +
+  theme(axis.text.x = element_text(angle = 90,
+                                   vjust = 0.5))
+
+# which sections get the most edits?
+section_edit_tally_tbl <-
+edits_for_each_year_tbl %>%
+  mutate(edit_comment = str_squish(edit_comment)) %>%
+  filter(!is.na(edit_comment)) %>%
+  # remove all caps text at the end of the position name, those are status changes
+  # in the same position
+  mutate(edit_comment = gsub("\\s+[A-Z]+(?:\\s+[A-Z]+)*$", "", edit_comment)) %>%
+  filter(!edit_comment %in% c("Current Users",
+                              "DISCUSSION, RUMORS, SPECULATION",
+                              "DISCUSSION, RUMORS AND SPECULATION",
+                              "General Discussion, Rumors, and Speculation",
+                              "Rejection Etiquette",
+                              "Current Users Edit",
+                              "Description",
+                              "RESEARCH",
+                              "TENURE",
+                              "NON-TENURE-TRACK",
+                              "ASSISTANT",
+                              "TENURE-TRACK OR TENURED / FULL-TIME",
+                              "Tenure-Track or Tenured / Full-time Position",
+                              "Posting Names of Job-Getters",
+                              "VISITING POSITIONS / LIMITED TERM APPOINTMENTS / POSTDOCS")) %>%
+  filter(!str_detect(edit_comment, "Updated:|Wiki")) %>%
+  group_by(edit_year,
+           edit_comment) %>%
+  tally(sort = TRUE)
+
+# edits per section for each year
+library(ggbeeswarm)
+p_edit_number_per_entry_per_year <-
+ggplot(section_edit_tally_tbl) +
+  aes(as_date(edit_year),
+      group = as_date(edit_year),
+      n) +
+  geom_boxplot() +
+  theme_minimal() +
+  scale_x_date(date_breaks = "1 year",
+               date_labels = "%Y") +
+  labs(x = "", y = "Number of edits per job ad") +
+  theme(axis.text.x = element_text(angle = 90,
+                                   vjust = 0.5))
+
+library(cowplot)
+plot_grid(
+  plot_grid(p_edits_per_year,
+          p_editors_per_year,
+          p_median_edit_size_per_year,
+          p_edit_number_per_entry_per_year,
+          labels = c("A", "B", "D", "E")),
+          plot_grid(p_editor_life_distr_year,
+                    p_editor_life_distr,
+                    rel_widths = c(2),
+                    ncol = 1,
+                    labels = c("C", "F"))
+          )
+```
+
+Because there is no single editorial authority, collaboratively-edited resources such as the Academic Jobs Wiki can be highly variable in the amount and type of activity over time.
+It is important to characterise this activity to assess the reliability of the content is as a source of information about the job market.
+We can get some insights into the dynamics of editing activity from the MediaWiki technology used by Academic Jobs Wiki that creates a public record of every act of adding or removing text to any page in the Wiki [@barrett2008mediawiki].
+@fig-panel-wiki-characteristics shows the temporal trends in editing of the Academic Jobs Wiki.
+The number of edits and editors roughly halved from 2013 to 2017, and while the number of edits remained relatively steady after 2017, the number of editors continued to gradually decline.
+The vast majority of editors are active on the site for less than a year (n = r` edits_for_each_year_tbl_span_tbl_less_than_one_year`, `r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%), perhaps reflecting a single season of job-searching.
+Editors with a longer history likely come from two groups, job seekers participating in multiple seasons of job applications, and faculty on search committees posting multiple job ads.
+
+```{r}
+#| label: fig-panel-wiki-self-report-numbers
+#| fig-cap: "Number of users self-reporting in each category of 'current users' per year"
+
+# get the dat
+# 2021-22 and 2022-2023
+current_user_tbl <-
+map(
+urls_for_each_year[1:9],
+~.x %>%
+  read_html() %>%
+  html_elements("p+ table td") %>%
+  html_text2()
+)
+
+current_user_tbl_2021_22 <-
+    urls_for_each_year[10]  %>%
+      read_html() %>%
+      html_elements("table:nth-child(1086) td") %>%
+      html_text2()
+
+current_user_tbl_2022_2023 <-
+    urls_for_each_year[11] %>%
+      read_html() %>%
+      html_elements("h2+ .fandom-table td") %>%
+      html_text2()
+
+current_user_tbl_all <- append(current_user_tbl,
+                        c(list(current_user_tbl_2021_22),
+                          list(current_user_tbl_2022_2023)
+                        ))
+
+names(current_user_tbl_all) <- urls_for_each_year
+
+names(current_user_tbl_all)[names(current_user_tbl_all) == "https://academicjobs.fandom.com/wiki/Archaeology_2021-22"] <- "https://academicjobs.fandom.com/wiki/Archaeology_2021-2022"
+
+wiki_data_list <-
+map(current_user_tbl_all,
+    ~.x %>% .[1:40] %>%
+    stringi::stri_remove_empty(.x))
+
+# Use map_dfr to iterate through the list and its names, binding results
+tidy_table <- map_dfr(wiki_data_list, ~{
+  # Process one vector (.x) at a time
+  tibble(raw = .x) %>%
+    # Create a lagged column to easily access the previous element
+    mutate(
+      previous_raw = lag(raw),
+      # Attempt to parse the current value as a number
+      # suppressWarnings handles cases where it's clearly not numeric
+      # !is.na checks if parsing was successful
+      is_count = !is.na(suppressWarnings(readr::parse_number(raw))),
+      # Check if the *previous* value looks like a number
+      prev_is_count = !is.na(suppressWarnings(readr::parse_number(previous_raw)))
+    ) %>%
+    # Keep rows where:
+    # 1. The current row IS a count (is_count == TRUE)
+    # 2. The previous row exists (previous_raw is not NA)
+    # 3. The previous row IS NOT a count (prev_is_count == FALSE)
+    #    This ensures we grab the status text, not a preceding number
+    filter(is_count & !is.na(previous_raw) & !prev_is_count) %>%
+    # Select and rename the relevant columns, converting count to numeric
+    transmute(
+      status = previous_raw,
+      count = readr::parse_number(raw) # Convert count string to number
+    )
+}, .id = "url") %>% # .id creates a column 'url' from the list names
+  # Extract the year from the URL
+  mutate(
+    year = str_extract(basename(url), "\\d{4}-\\d{4}"), # Extracts YYYY-YYYY pattern
+    .before = status # Place the year column before status
+  ) %>%
+  select(year, status, count) # Keep only the final desired columns
+
+#  the resulting tidy table
+tidy_table_to_plot <- 
+tidy_table %>%
+  filter(!status %in% c("I'm a lurker",
+                        "I have a current but term-limited faculty job",
+                        "I have some other teaching, research, or non-academic gig",
+                        "Nope (independent scholar, on sabbatical from adjuncting, etc.)")) %>%
+  mutate(year = str_replace(year, "-", "-\n"))
+
+# summary over time
+how_many_people_use_the_wiki <- 
+tidy_table_to_plot %>% 
+filter(status == "How many people use the wiki?")
+
+how_many_people_use_the_wiki_perc_start_end <- round(
+(how_many_people_use_the_wiki$count[1] - how_many_people_use_the_wiki$count[nrow(how_many_people_use_the_wiki)]) / how_many_people_use_the_wiki$count[1] * 100, 0)
+
+# proportion of search committee members
+search_committee_members_tbl <- 
+tidy_table_to_plot %>% 
+  filter(status %in% c( "How many people use the wiki?",
+                       "I'm a search committee member")) %>% 
+  pivot_wider(names_from = status,
+              values_from = count) %>% 
+  mutate(perc = round(`I'm a search committee member` / 
+          `How many people use the wiki?` * 100, 0)) %>% 
+  drop_na()
+
+search_committee_members_perc_start <- 
+  search_committee_members_tbl$perc[1] 
+search_committee_members_perc_end <- 
+  search_committee_members_tbl$perc[nrow(search_committee_members_tbl)]
+
+  ggplot(tidy_table_to_plot) +
+  aes(year,
+      count) +
+  geom_col() +
+  theme_minimal() +
+  facet_wrap(~ status,
+             scales = "free") +
+  xlab("") +
+  theme(axis.text.x = element_text(size = 6))
+
+tidy_table_to_plot %>% 
+filter(status != "How many people use the wiki?") %>% 
+ggplot() +
+  aes(year,
+      count,
+      fill = status) +
+  geom_col() +
+  theme_minimal() +
+  xlab("") +
+  scale_fill_brewer(type = "qual") +
+  theme(axis.text.x = element_text(size = 6),
+        legend.position = c(.7,.7))
+```
+
+There seems to be a shift over time in the balance of editors in these two groups, as indicated by the number of self-reporting users over time.
+@fig-panel-wiki-self-report-numbers shows the data found on each page in a section labeled 'current users', where users can volunteer to update a table tracking the status of users.
+These data indicate that the number of self-reporting users, from the values recording at 'How many people use the wiki?', declined by `r how_many_people_use_the_wiki_perc_start_end`% during the study period.
+At the same time, the proportion of users self-identifying as "search committee members" increased, from under `r search_committee_members_perc_start`% in 2014-2018 to over `r search_committee_members_perc_end`% from 2020 onward.
+This suggests that over time fewer job applicants have been contributing job postings and sharing information about their experiences, and search committee members were increasingly posting job advertisements themselves.
+Overall the Archaeology Academic Jobs Wiki is best understood not as a comprehensive archive of all jobs posted during the study period, but a biased sample of the job ads that were of most interest to archaeologists applying for jobs in a given year.
+It represents the jobs that early career applicants were most focused on applying for, and getting information on.
+
 For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023, we read the text and recorded the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methodological foci of the position into a Google form.
@@
 Values in parentheses are for the United States only.
-TT = tenure-track, NTT = non-tenure-track.
+TT = tenure-track, US-only, NTT = non-tenure-track, US and elsewhere.
 Not all ads include unambiguous information about tenure status, so the sum of TT and NTT does not equal the sum of all jobs in all ranks."

### 2025-05-02 analysis/paper/sketches.R

@@ -687,6 +687,110 @@ count_of_tt_jobs_per_year_from_our_form %>%
   arrange(year_num) %>%
   select(-year_num)
 
+#---------------------------------------------------------------
+# can we get that table of 'current users'
+
+# 2021-22 and 2022-2023
+current_user_tbl <-
+map(
+urls_for_each_year[1:9],
+~.x %>%
+  read_html() %>%
+  html_elements("p+ table td") %>%
+  html_text2()
+)
+
+current_user_tbl_2021_22 <-
+    urls_for_each_year[10]  %>%
+      read_html() %>%
+      html_elements("table:nth-child(1086) td") %>%
+      html_text2()
+
+current_user_tbl_2022_2023 <-
+    urls_for_each_year[11] %>%
+      read_html() %>%
+      html_elements("h2+ .fandom-table td") %>%
+      html_text2()
+
+current_user_tbl_all <- append(current_user_tbl,
+                        c(list(current_user_tbl_2021_22),
+                          list(current_user_tbl_2022_2023)
+                        ))
+
+names(current_user_tbl_all) <- urls_for_each_year
+
+names(current_user_tbl_all)[names(current_user_tbl_all) == "https://academicjobs.fandom.com/wiki/Archaeology_2021-22"] <- "https://academicjobs.fandom.com/wiki/Archaeology_2021-2022"
+
+wiki_data_list <-
+map(current_user_tbl_all,
+    ~.x %>% .[1:40] %>%
+    stringi::stri_remove_empty(.x))
+
+# Use map_dfr to iterate through the list and its names, binding results
+tidy_table <- map_dfr(wiki_data_list, ~{
+  # Process one vector (.x) at a time
+  tibble(raw = .x) %>%
+    # Create a lagged column to easily access the previous element
+    mutate(
+      previous_raw = lag(raw),
+      # Attempt to parse the current value as a number
+      # suppressWarnings handles cases where it's clearly not numeric
+      # !is.na checks if parsing was successful
+      is_count = !is.na(suppressWarnings(readr::parse_number(raw))),
+      # Check if the *previous* value looks like a number
+      prev_is_count = !is.na(suppressWarnings(readr::parse_number(previous_raw)))
+    ) %>%
+    # Keep rows where:
+    # 1. The current row IS a count (is_count == TRUE)
+    # 2. The previous row exists (previous_raw is not NA)
+    # 3. The previous row IS NOT a count (prev_is_count == FALSE)
[…]

### 2025-05-02 analysis/paper/article.qmd
add plot showing self-reported user category counts, with color palette from "The Thing"

@@
   fig-path: "../figures/"
-  fig-dpi: 900
+  fig-dpi: 1000
 filters:
@@
 ```{r}
-#| label: fig-panel-wiki-characteristics
-#| fig-cap: "A: Total number of edits to each Wiki page (additions and deletions) per year.
-B: Number of unique editors, as represented by distinct usernames or IP addresses (recorded for editors without usernames), active per year.
-C: Active periods of individual editors, the black lines connect the date of the first and last edit of all editors that were active for  three months or more.
-D: Typical size of each edit per year, either addition or removal of text, one word is roughly 4-6 Bytes.
-E: Distribution of the numbers of edits per job ad for each year.
-F: Histogram showing that the majority of editors are only active on the site for less than one year."
 
@@
 # visualisations ----------------------------------
+small_base_size <- 4
 
@@
   geom_bar() +
-  theme_minimal() +
+  theme_minimal(base_size = small_base_size) +
   scale_x_date(date_breaks = "1 year",
@@
   geom_col() +
-  theme_minimal() +
+  theme_minimal(base_size = small_base_size) +
   scale_x_date(date_breaks = "1 year",
@@
   geom_histogram() +
-  theme_minimal() +
+  theme_minimal(base_size = small_base_size) +
   scale_y_log10() +
@@
                lineend = "butt") +
-  theme_minimal() +
+  theme_minimal(base_size = small_base_size) +
   labs(x = "", y = "Individual editors\n(usernames and IP addresses redacted)") +
@@
   geom_boxplot() +
-  theme_minimal() +
+  theme_minimal(base_size = small_base_size) +
   scale_x_date(date_breaks = "1 year",
@@
                                    vjust = 0.5))
-
-library(cowplot)
-plot_grid(
-  plot_grid(p_edits_per_year,
-          p_editors_per_year,
-          p_median_edit_size_per_year,
-          p_edit_number_per_entry_per_year,
-          labels = c("A", "B", "D", "E")),
-          plot_grid(p_editor_life_distr_year,
-                    p_editor_life_distr,
-                    rel_widths = c(2),
-                    ncol = 1,
-                    labels = c("C", "F"))
-          )
-```
-
-Because there is no single editorial authority, collaboratively-edited resources such as the Academic Jobs Wiki can be highly variable in the amount and type of activity over time.
-It is important to characterise this activity to assess the reliability of the content is as a source of information about the job market.
-We can get some insights into the dynamics of editing activity from the MediaWiki technology used by Academic Jobs Wiki that creates a public record of every act of adding or removing text to any page in the Wiki [@barrett2008mediawiki].
-@fig-panel-wiki-characteristics shows the temporal trends in editing of the Academic Jobs Wiki.
-The number of edits and editors roughly halved from 2013 to 2017, and while the number of edits remained relatively steady after 2017, the number of editors continued to gradually decline.
-The vast majority of editors are active on the site for less than a year (n = r` edits_for_each_year_tbl_span_tbl_less_than_one_year`, `r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%), perhaps reflecting a single season of job-searching.
-Editors with a longer history likely come from two groups, job seekers participating in multiple seasons of job applications, and faculty on search committees posting multiple job ads.
+```
 
@@
                         "Nope (independent scholar, on sabbatical from adjuncting, etc.)")) %>%
-  mutate(year = str_replace(year, "-", "-\n"))
+  mutate(year = str_sub(year, 6L))
 
@@
 
-  ggplot(tidy_table_to_plot) +
-  aes(year,
-      count) +
-  geom_col() +
-  theme_minimal() +
-  facet_wrap(~ status,
-             scales = "free") +
-  xlab("") +
-  theme(axis.text.x = element_text(size = 6))
-
+# "The Thing" Inspired Palette (8 Colors)
+the_thing_palette <- c(
+  "#6488AC", # 1. Icy Blue
+  "#1A2A3A", # 2. Deep Shadow Blue/Black
+  "#E8702A", # 3. Flamethrower Orange
+  "#9D2323", # 4. Blood Red/Creature Tissue
+  "#8B7355", # 5. Outpost Brown
+  "#E1E1E1", # 6. Snowy Off-White/Grey
+  "#5F7C6A", # 7. Parka Green/Drab
+  "#D9A443"  # 8. Warning Yellow/Orange (Muted)
+)
+
+how_many_people_use_the_wiki <- 
 tidy_table_to_plot %>% 
   filter(status != "How many people use the wiki?") %>% 
+  mutate(status = str_wrap(status, 26)) %>% 
 ggplot() +
@@
   geom_col() +
-  theme_minimal() +
+  theme_minimal(base_size = small_base_size) +
   xlab("") +
-  scale_fill_brewer(type = "qual") +
-  theme(axis.text.x = element_text(size = 6),
-        legend.position = c(.7,.7))
-```
+  scale_fill_manual(values = the_thing_palette) +
+  theme(legend.position = c(.7,.8),
+        legend.key.size = unit(0.25,"line"),
+        legend.text = element_text(margin = margin(r = 10, 
+                                                   t = 2,
+                                                   b = 2,
+                                                   unit = "pt"))) +
+  guides(fill=guide_legend(nrow=3,
+                           byrow=TRUE),
+         )
+```
+
+```{r}
+#| label: fig-panel-wiki-characteristics
+#| fig-cap: "A: Total number of edits to each Wiki page (additions and deletions) per year.
+B: Number of unique editors, as represented by distinct usernames or IP addresses (recorded for editors without usernames), active per year.
+C: Active periods of individual editors, the black lines connect the date of the first and last edit of all editors that were active for three months or more.
+D: Typical size of each edit per year, either addition or removal of text, one word is roughly 4-6 Bytes.
+E: Distribution of the numbers of edits per job ad for each year.
+F: Histogram showing that the majority of editors are only active on the site for less than one year.
+G: Breakdown of self-reported Wiki page user categories for each year."
+#| fig-height: 5
+
+library(cowplot)
+plot_grid(
+ plot_grid(
+  plot_grid(p_edits_per_year,
+          p_editors_per_year,
+          p_median_edit_size_per_year,
+          p_edit_number_per_entry_per_year,
+          labels = c("A", "B", "D", "E"),
+          label_size = small_base_size),
+          plot_grid(p_editor_life_distr_year,
+                    p_editor_life_distr,
+                    rel_widths = c(2),
+                    ncol = 1,
+                    labels = c("C", "F"),
+                    label_size = small_base_size)),
+  plot_grid(how_many_people_use_the_wiki,
+            labels = "G",
+            label_size = small_base_size),
+  rel_heights = c(2,1),
+  ncol = 1)
+```
+
+Because there is no single editorial authority, collaboratively-edited resources such as the Academic Jobs Wiki can be highly variable in the amount and type of activity over time.
+It is important to characterise this activity to assess the reliability of the content is as a source of information about the job market.
+We can get some insights into the dynamics of editing activity from the MediaWiki technology used by Academic Jobs Wiki that creates a public record of every act of adding or removing text to any page in the Wiki [@barrett2008mediawiki].
+@fig-panel-wiki-characteristics shows the temporal trends in editing of the Academic Jobs Wiki.
+The number of edits and editors roughly halved from 2013 to 2017, and while the number of edits remained relatively steady after 2017, the number of editors continued to gradually decline.
+The vast majority of editors are active on the site for less than a year (n = r` edits_for_each_year_tbl_span_tbl_less_than_one_year`, `r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%), perhaps reflecting a single season of job-searching.
+Editors with a longer history likely come from two groups, job seekers participating in multiple seasons of job applications, and faculty on search committees posting multiple job ads.
 
 There seems to be a shift over time in the balance of editors in these two groups, as indicated by the number of self-reporting users over time.
-@fig-panel-wiki-self-report-numbers shows the data found on each page in a section labeled 'current users', where users can volunteer to update a table tracking the status of users.
-These data indicate that the number of self-reporting users, from the values recording at 'How many people use the wiki?', declined by `r how_many_people_use_the_wiki_perc_start_end`% during the study period.
+Panel G of @fig-panel-wiki-characteristics shows the data found on each page in a section labeled 'current users', where users can volunteer to update a table tracking the status of users.
+These data indicate that the total number of self-reporting users, from the values recording at 'How many people use the wiki?', declined by `r how_many_people_use_the_wiki_perc_start_end`% during the study period.
 At the same time, the proportion of users self-identifying as "search committee members" increased, from under `r search_committee_members_perc_start`% in 2014-2018 to over `r search_committee_members_perc_end`% from 2020 onward.
@@
 Overall the Archaeology Academic Jobs Wiki is best understood not as a comprehensive archive of all jobs posted during the study period, but a biased sample of the job ads that were of most interest to archaeologists applying for jobs in a given year.
-It represents the jobs that early career applicants were most focused on applying for, and getting information on.
+It represents the jobs that early career applicants were most focused on applying for, and sharing information on.
 

### 2025-05-02 analysis/paper/sketches.R

@@ -792,6 +792,9 @@ discussion_section <-
 
 names(discussion_section) <- urls_for_each_year
 
+#----------------------------------------------
+# can we geolocate IP addresses?
+
 
 
 

### 2025-05-02 analysis/paper/article.qmd
replace plot of changing requirements over time to match reviewer's suggestion

@@
 plot_grid(
- fig_prop_by_job_title_per_year + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
+ fig_prop_by_job_title_per_year_us + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
   fig_ratio_tt_2_ntt_jobs_per_year + theme(plot.margin = unit(c(0, 0, 0, 0), "cm")),
@@
                               -4L))
-  ) %>%
+  ) # %>%
   # delete edits that were made after the job market year
-  filter(edit_year_only == job_market_year | edit_year == (job_market_year -1))
+  # filter(edit_year_only == job_market_year | edit_year == (job_market_year -1))
 
@@
 p_edits_per_year <-
-ggplot(edits_for_each_year_tbl) +
+  edits_for_each_year_tbl %>% 
+  filter(between(edit_year,
+                 as.POSIXct("2012-01-01 0:00:00"), 
+                 as.POSIXct("2023-01-01 0:00:00"))) %>% 
+ggplot() +
   aes(as_date(edit_year)) +
@@
 edits_for_each_year_tbl %>%
+    filter(between(edit_year,
+                 as.POSIXct("2012-01-01 0:00:00"), 
+                 as.POSIXct("2023-01-01 0:00:00"))) %>% 
   group_by(edit_year) %>%
@@
 edits_for_each_year_tbl %>%
+    filter(between(edit_year,
+                 as.POSIXct("2012-01-01 0:00:00"), 
+                 as.POSIXct("2023-01-01 0:00:00"))) %>% 
   group_by(edit_name) %>%
@@
                    xend = last_edit),
-               linewidth = 1,
+               linewidth = 0.1,
                lineend = "butt") +
@@
   labs(x = "", y = "Individual editors\n(usernames and IP addresses redacted)") +
-  theme(axis.text.y=element_blank())
+  theme(axis.text.y=element_blank(),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(),)
 
@@
 p_edit_number_per_entry_per_year <-
-ggplot(section_edit_tally_tbl) +
+section_edit_tally_tbl %>% 
+    filter(between(edit_year,
+                 as.POSIXct("2012-01-01 0:00:00"), 
+                 as.POSIXct("2023-01-01 0:00:00"))) %>% 
+  ggplot() + 
   aes(as_date(edit_year),
@@
       n) +
-  geom_boxplot() +
+  geom_boxplot(outlier.size = 0.5) +
   theme_minimal(base_size = small_base_size) +
@@
         legend.text = element_text(margin = margin(r = 10, 
+                                                   l = 2,
                                                    t = 2,
@@
 F: Histogram showing that the majority of editors are only active on the site for less than one year.
-G: Breakdown of self-reported Wiki page user categories for each year."
+G: Breakdown of self-reported Wiki user categories for each year."
 #| fig-height: 5
@@
 
+```{r}
+# geolocate users by IP address
+# devtools::install_github("ironholds/rgeolocate")
+
+library(rgeolocate)
+file <- system.file("extdata","GeoLite2-Country.mmdb", package = "rgeolocate")
+results <- maxmind(unique(edits_for_each_year_tbl$edit_name),
+                   file,
+                   "country_code")
+
+n_all_user_ids <- length(unique(edits_for_each_year_tbl$edit_name))
+
+results_ip_addresses <-
+  results %>%
+  drop_na()
+
+n_all_user_ips <-  nrow(results_ip_addresses)
+
+# percentage of users that we can geolocate
+per_users_we_can_geolocate <- 
+  round(n_all_user_ips / n_all_user_ids * 100, 0)
+
+# tally and percentage
+results_ip_addresses_tbl <- 
+results_ip_addresses %>%
+  group_by(country_code) %>%
+  tally(sort = TRUE) %>%
+  mutate(perc = n / sum(n) * 100)
+
+results_ip_addresses_tbl_us_scl <- 
+  results_ip_addresses_tbl %>% 
+  filter(country_code == "US") %>% 
+  pull(perc) %>% 
+  round(0)
+
+```
+
 Because there is no single editorial authority, collaboratively-edited resources such as the Academic Jobs Wiki can be highly variable in the amount and type of activity over time.
@@
 @fig-panel-wiki-characteristics shows the temporal trends in editing of the Academic Jobs Wiki.
-The number of edits and editors roughly halved from 2013 to 2017, and while the number of edits remained relatively steady after 2017, the number of editors continued to gradually decline.
-The vast majority of editors are active on the site for less than a year (n = r` edits_for_each_year_tbl_span_tbl_less_than_one_year`, `r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%), perhaps reflecting a single season of job-searching.
+Every edit includes the author's identity recorded as either an IP address (a numerical address that includes information about the location of the user's computer) or a username.
+Of the `r n_all_user_ids` unique editors in our sample, `r per_users_we_can_geolocate`% used IP addresses rather than usernames, and geolocation analysis shows that `r results_ip_addresses_tbl_us_scl`% of IP addresses were based in the US, indicating the Wiki is primarily used by archaeologists in the US.
+The number of edits and editors roughly halved from 2013 to 2017, and continued to gradually decline.
+Despite this trend, the size of individual edits slightly increases over time, and the number of edits per job ad is relatively constant over time.
+The majority of editors are active on the site for less than a year (n = r` edits_for_each_year_tbl_span_tbl_less_than_one_year`, `r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%), perhaps reflecting a single season of job-searching.
 Editors with a longer history likely come from two groups, job seekers participating in multiple seasons of job applications, and faculty on search committees posting multiple job ads.
@@
 #| label: fig-show-basic-plots
-#| fig-cap: "A: total number of job ads posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank.
+#| fig-cap: "A: total number of job ads from US institutions posted to the Academic Jobs Wiki for Archaeology in each year, with coloured sections showing the proportion of jobs by title and rank.
 B: Ratio of tenure-track to non-tenure-track positions over time.
@@
 p_req <- 
-ggplot(jobdata_requirements_means) +
+jobdata_requirements %>%
+  mutate(name = case_when(
+    name == "cover letter" ~ "Cover letter",
+    name == "cv" ~ "CV",
+    name == "research statement" ~ "Research statement",
+    name == "teaching statement" ~ "Teaching statement",
+    name == "diversity statement" ~ "Diversity statement",
+    name == "names of recommenders" ~ "Names of recommenders")) %>%
+  ggplot() +
   aes(year_ad_posted_break,
-      mean_n,
-      group = name) +
-  geom_smooth(linewidth = 2,
-              colour = "black") +
-  geom_jitter(data = jobdata_requirements,
-             aes(year_ad_posted_break, 
-                 value),
-             alpha = 0.1,
-             height = 0.2,
-             width =  0.1) +
+      fill = factor(value),
+      group = factor(value)) +
+  geom_bar(stat="count",
+           position = "fill") +
   facet_wrap(~name,
@@
   xlab("") +
-  ylab("Number requested in job ad") +
+  ylab("Proportion of all job ads in a year") +
   scale_y_continuous(breaks = integer_breaks()) +
-  theme_minimal(base_size = 14) +
+  theme_minimal(base_size = base_size) +
+  scale_fill_brewer(palette = "Dark2") +
+  guides(fill = guide_legend("Count")) +
   theme(axis.text.x = element_text(size = 8),
-        strip.text = element_text( size = 16))
+        strip.text = element_text( size = 14))
 
@@
 #| label: fig-requirements-over-time
-#| fig-cap: "Changing requirements in job ads over time.
-Each point represents an individual job ad.
-The points have been jittered to avoid overlap.
-The black line is a locally weighted regression line to identify temporal trends.
-The grey region is the 95% confidence interval for the regression line."
+#| fig-cap: "Changing requirements in job ads over time. 'Count' refers to the number of each item required."
 
@@
 Over our ten year study period there have been substantial changes in the instructions to applicants in terms of the type and number of documents that are requested by the search committees [@fig-requirements-over-time].
-Requests for cover letter and CV decline in more recent years, perhaps reflecting their status as a generic and expected component to submit without needing to be specifically requested.
+Requests for cover letter and CV decline slightly in more recent years, perhaps reflecting their status as a generic and expected component to submit without needing to be specifically requested.
 The requirement for a diversity statement is rare until 2019-2020, peaks around 2020-2021, then decreases towards the present.

### 2025-05-02 analysis/paper/sketches.R

@@ -779,21 +779,161 @@ tidy_table %>%
 #-----------------------------------------------
 # can we get 'discussion' section
 
-xpath_selector <- "//h2[span/@id='DISCUSSION,_RUMORS,_SPECULATION']/following-sibling::*[count(preceding-sibling::h2[span/@id='DISCUSSION,_RUMORS,_SPECULATION']) = 1]"
+urls_for_each_year_xpath_selector_tbl <-
+tibble(urls_for_each_year = urls_for_each_year,
+       xpath_selector = c("//h2[span/@id='General_Discussion.2C_Rumors.2C_and_Speculation']/following-sibling::*[count(preceding-sibling::h2[span/@id='General_Discussion.2C_Rumors.2C_and_Speculation']) = 1]", # 2012-2013
+                          "//h2[span/@id='DISCUSSION,_RUMORS_AND_SPECULATION']/following-sibling::*[count(preceding-sibling::h2[span/@id='DISCUSSION,_RUMORS_AND_SPECULATION']) = 1]", # 2013-2014
+                          "//h2[span/@id='DISCUSSION.2C_RUMORS.2C_SPECULATION']/following-sibling::*[count(preceding-sibling::h2[span/@id='DISCUSSION.2C_RUMORS.2C_SPECULATION']) = 1]", # 2014-2015
+                          "//h2[span/@id='DISCUSSION,_RUMORS,_SPECULATION']/following-sibling::*[count(preceding-sibling::h2[span/@id='DISCUSSION,_RUMORS,_SPECULATION']) = 1]", # 2015-2016
+                          "//h2[span/@id='DISCUSSION,_RUMORS,_SPECULATION']/following-sibling::*[count(preceding-sibling::h2[span/@id='DISCUSSION,_RUMORS,_SPECULATION']) = 1]", # 2016-2017
+                          "//h2[span/@id='DISCUSSION,_RUMORS,_SPECULATION']/following-sibling::*[count(preceding-sibling::h2[span/@id='DISCUSSION,_RUMORS,_SPECULATION']) = 1]", # 2017-2018
+                          "//h2[span/@id='DISCUSSION,_RUMORS,_SPECULATION']/following-sibling::*[count(preceding-sibling::h2[span/@id='DISCUSSION,_RUMORS,_SPECULATION']) = 1]", # 2018-2019
[…]

### 2025-05-08 analysis/paper/article.qmd

@@
 We focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities.
-@fig-show-basic-plots shows the count of ads for each year.
+@fig-show-basic-plots shows the count of ads for each year, where year refers to the year the job ad was posted.
 @tbl-show-basic-counts shows the breakdown by different job types, ranks, and tenure status.
@@
 The demand for archaeology faculty, indicated by the total number of tenure-track jobs, maybe affected by a variety of factors.
+
 Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-2014.
 This is at odds with previous work that has shown that contingent positions make up more than half of faculty positions [@colby2022annual], and that these numbers are increasing— between 2004 and 2019, for example, there was a 36% increase in the replacement of tenure-track positions by contingent positions [@the20222022].
-This discrepancy in our data set may be due to the more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
+This discrepancy in our data set may be due to several factors.
+The more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
 Many of these short-term positions are not advertised nationally, but only through local email lists and are filled by people close to the hiring department, such as recently graduated students.
-One possible explanation for our results is that non-tenure-track jobs are less frequently added to the Academic Jobs Wiki.
+Another factor is bias in our data, with most Wiki users likely seeking a tenure-track job, non-tenure-track jobs are may have been less frequently added to the Academic Jobs Wiki because they were peripheral to the goal of most users.
+This bias limits the reliability of our results on the ratio of tenure-track to non-tenure-track.
+
 The downward trend in tenure-track positions during 2013-2019 may be related to declining undergraduate enrollment in anthropology since 2013 [@crambChangingProfileTenureTrack2022a].

### 2025-05-12 analysis/paper/article.qmd
Minor edits to text

@@
 As Dennis et al.
-underscore, early career researchers are "supplicants": "There is no way for applicants to point out that the job ads are taking too much time out of the scholarly community’s collective time bank" [@dennisWorstAnthroJob2022, pp.
-1].
-Indeed, acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs, to performing well in long-list interviews and preparing teaching demonstrations.
+underscore, early career researchers are "supplicants"—"There is no way for applicants to point out that the job ads are taking too much time out of the scholarly community’s collective time bank" [@dennisWorstAnthroJob2022, pp.
+1].Acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs to performing well in long-list interviews and preparing teaching demonstrations.
 These courses can provide valuable training in the hidden curriculum of the academy, but the fact that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the "market" as a historically contingent set of cultural practices.
@@
 Contingent faculty positions are precarious because they provide low or no health and retirement benefits, limited opportunities for professional development, and lower salaries relative to tenure-track positions.
-PhD graduates who have a sequence of short-term academic positions are often disadvantaged financially due to low compensation combined with expensive frequent relocation, socially due to isolation from family and community, and professionally due to being ineligible for many decision-making roles at the universities at which they work [@platzerAcademicPrecarityAmerican2018].
+PhD graduates who have a sequence of short-term academic positions are often disadvantaged financially due to low compensation combined with expensive and frequent relocation, restrictive socially due to isolation from family and community, and limited professionally due to being ineligible for many decision-making roles at the universities at which they work [@platzerAcademicPrecarityAmerican2018].
 
@@
 Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 53 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
-This network analysis showed that programs with large endowments and faculty who hold prestigious awards and are widely cited by other scholars, produce the majority of tenured and tenure-track faculty.
+This network analysis showed that programs with large endowments and widely cited faculty who hold prestigious awards produce the majority of tenured and tenure-track faculty.
 
@@
 Targeted studies of sociology, communication, operations research, and industrial systems engineering show similar dynamics [@barnett2010; @castillo; @feeley2021; @nevin2019].
-Broader analyses comparing computer science, business, and history likewise comparable patterns among disparate disciplines across the humanities, social sciences, and STEM fields; "across disciplines, prestige hierarchies make the most accurate predictions of faculty placement" [@clauset2015, pp.
+Broader analyses comparing computer science, business, and history identify comparable patterns among disparate disciplines across the humanities, social sciences, and STEM field—"across disciplines, prestige hierarchies make the most accurate predictions of faculty placement" [@clauset2015, pp.
 4].
 The most recent comprehensive meta-analysis of faculty placement dynamics in the United States, which examined employment and doctoral education of all tenure-track faculty at PhD-granting universities, underscored the outsize impact of institutional prestige: 80% of faculty trained in the US came from just 20% of institutions [@wapmanQuantifyingHierarchyDynamics2022a].
-Within anthropology, however, such studies draw attention to the paradox of a discipline committed to fighting social inequalities continuing to reproduce its own systematic inequalities through hiring practices that favor an elite minority of graduates with prestigious institutional affiliations.
-
-One way that some hiring committees are tackling these prestige biases is by providing detailed instructions to applicants on how to prepare their application materials.
+Within anthropology, such studies draw attention to a central paradox—a discipline purportedly committed to fighting social inequalities continues to reproduce its own systemic inequalities through hiring practices that favor an elite minority of candidates with prestigious affiliations.
+
+One way that some hiring committees are tackling prestige biases is by providing detailed instructions to applicants on how to prepare their application materials.
 In theory, detailed applications will allow the hiring committee to focus on evaluating the accomplishments of candidates across common categories rather than ranking based on prestige signals in a CV, such as a name of the applicant's graduate program.
@@
 These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted in a variety of approaches to evaluating diversity, even leading some universities to entirely omit requirements for diversity statements in job applications [@guidenAreDiversityStatements].
+**MORE DETAILS ON THIS HERE—JB WILL ADD.**
 
@@
 Our primary data source is the Archaeology Academic Jobs Wiki.
-Originating in 2007, this is a set of freely accessible web pages that anyone can edit (anonymously or with a free user account) hosted by Fandom, a for-profit company.
+Originating in 2007, the wiki is a set of freely accessible web pages that anyone can edit anonymously or with a free user account.
+The site is hosted by Fandom, a for-profit company.
 The Archaeology pages are part of the Academic Jobs Wiki, which coordinates similar collaboratively-edited resources for around 40 academic disciplines.
 The coordinators and contributors are nearly all anonymous or pseudonymous.
-Typically, contributors copy and paste all or some of the text of job ads into the wiki, from sources such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites, collecting ads originally posted in numerous different locations.
+Typically, contributors copy and paste all or some of the text of job ads into the wiki, from a variety of sources such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites.
 Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position.
@@
 Contributors also edit the page to ask and answer questions about the positions and the application process.
-These comments make the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market.
+These comments have made <!--If we are a bit more slippery about our tenses, we don't need to argue that the wiki is STILL crucial, esp.
+if late we will discuss its decline--> the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market.
 Because of its reputation for aggregating ads from diverse sources and rapidly-updated information that is not available elsewhere, the Academic Jobs Wiki has become an authoritative data source for studies of hiring trends in academia [e.g.
@@
 Despite this trend, the size of individual edits slightly increases over time, and the number of edits per job ad is relatively constant over time.
-The majority of editors are active on the site for less than a year (n = r` edits_for_each_year_tbl_span_tbl_less_than_one_year`, `r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%), perhaps reflecting a single season of job-searching.
+The majority of editors are active on the site for less than a year (n = r`edits_for_each_year_tbl_span_tbl_less_than_one_year`, `r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%), perhaps reflecting a single season of job-searching.
 Editors with a longer history likely come from two groups, job seekers participating in multiple seasons of job applications, and faculty on search committees posting multiple job ads.
@@
 Some topics include geographic regions because this is how they are typically understood by archaeologists.
-For example, Mesoamerican archaeology is understood to refer to a specific time period _and_ a specific geographic region.
+For example, Mesoamerican archaeology is understood to refer to a specific time period *and* a specific geographic region.
 Similarly, we recorded digital archaeology as both a method (when a job ad also had a clearly distinct topical focus, such as historic archaeology) and a topic (when there were no other topics mentioned in the job ad).
@@
 
-The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx [ for peer reviewers the link is https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug ] to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx \[ for peer reviewers the link is https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug \] to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.

### 2025-05-13 analysis/paper/article.qmd
Some more substantial edits, including a few new paragraphs of text and some queries for BM.

@@
 
+<!--JB: Given that we have expanded our section on the wiki by several paragraphs, it might be useful for us to divide the methods into two subsections: (1) Dynamics of the wiki, and (2) Scraping the wiki and categorizing ads.
+Feel free to pick better titles, but I think this would provide more guidance for our readers so they do not get lost in the welter of details about the wiki.
+Lmk what you think -->
+
 Our primary data source is the Archaeology Academic Jobs Wiki.
@@
 Contributors also edit the page to ask and answer questions about the positions and the application process.
-These comments have made <!--If we are a bit more slippery about our tenses, we don't need to argue that the wiki is STILL crucial, esp.
+These comments have made <!--JB: If we are a bit more slippery about our tenses, we don't need to argue that the wiki is STILL crucial, esp.
 if late we will discuss its decline--> the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market.
@@
 
-Because there is no single editorial authority, collaboratively-edited resources such as the Academic Jobs Wiki can be highly variable in the amount and type of activity over time.
+Collaboratively-edited resources such as the Academic Jobs Wiki can be highly variable in the amount and type of editorial activity over time.
 It is important to characterise this activity to assess the reliability of the content is as a source of information about the job market.
-We can get some insights into the dynamics of editing activity from the MediaWiki technology used by Academic Jobs Wiki that creates a public record of every act of adding or removing text to any page in the Wiki [@barrett2008mediawiki].
-@fig-panel-wiki-characteristics shows the temporal trends in editing of the Academic Jobs Wiki.
-Every edit includes the author's identity recorded as either an IP address (a numerical address that includes information about the location of the user's computer) or a username.
-Of the `r n_all_user_ids` unique editors in our sample, `r per_users_we_can_geolocate`% used IP addresses rather than usernames, and geolocation analysis shows that `r results_ip_addresses_tbl_us_scl`% of IP addresses were based in the US, indicating the Wiki is primarily used by archaeologists in the US.
-The number of edits and editors roughly halved from 2013 to 2017, and continued to gradually decline.
-Despite this trend, the size of individual edits slightly increases over time, and the number of edits per job ad is relatively constant over time.
+The MediaWiki technology used by Academic Jobs Wiki creates a public record of every act of adding or removing text to any page in the Wiki [@barrett2008mediawiki].
+@fig-panel-wiki-characteristics, which allows us to assess the dynamics of editing activity over the ten years of our study period.
+Every edit includes the author's identity, recorded as either an IP address (a numerical address that includes information about the location of the user's computer) <!--good call being explicit here--> or a username.
+Of the `r n_all_user_ids` unique editors in our sample, `r per_users_we_can_geolocate`% used IP addresses rather than usernames, and geolocation analysis shows that `r results_ip_addresses_tbl_us_scl`% of IP addresses were based in the US, indicating the Archaeology Jobs Wiki is primarily edited and visited by users in the U.S.
+<!--JB: technically we don't know if they are archaeologists; could be people from other disciplines e.g.
+Middle Eastern Studies, history etc.-->.
+The number of edits and active editors was roughly halved from 2013 to 2017, and then continued to gradually decline.
+Despite this downward trend, the size of individual edits slightly increases over time, and the number of edits per job ad is relatively constant across our study period.
 The majority of editors are active on the site for less than a year (n = r`edits_for_each_year_tbl_span_tbl_less_than_one_year`, `r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%), perhaps reflecting a single season of job-searching.
@@
 
-There seems to be a shift over time in the balance of editors in these two groups, as indicated by the number of self-reporting users over time.
+<!--JB: Ben, this is where I'd like some of the details about min, max, etc.
+What proportion of people are on the wiki for more than a year, and how long do they stay on?
+ALSO, and I just thought of this, if ECRs are highly mobile, the same ECR might be represented by different, year-long IP addresses over their period of job seeking.
+For example, when I was on the market, I lived in Michigan, Pennsylvania, New York, England, and Massachusetts—would I not look like five distinct users who each spent 1-2 years on the wiki using our rationale here?
+This perhaps merits further discussion) -->
+
+Longer-term editors who represent job seekers active for multiple seasons reflect one of the underdiscussed realities of the academic job market in archaeology—time spent on that market.
+Previous research on professional trajectories in academic archaeology suggests that early career researchers have a limited shelf life.
+Mackie and Rockwell, for example, identify a hiring plateau that occurs around seven years after candidates defend their PhDs [-@mackieMarketShareAccounting2023:5].
+However, too little time on the market also seems to have a negative effect on placement rates—“very few graduates obtain TT employment when they are all but dissertation (ABD) or even immediately after graduation” [@mackieMarketShareAccounting2023:5].
+These patterns are not restricted to the United States.
+In Brami et al.’s [-@bramiPrecariousFutureReflections2023] survey of 419 early career researchers in European archaeology, likelihood of holding a permanent position shifted relative to  time since PhD.
+Of the respondents who were 1 year or less to 4 years post-PhD, only 3% (3/116) held permanent positions.
+These numbers increased slightly for participants who had been on the market for 5–7 years, with 14% (7/49) holding permanent positions; similarly, 18% (6/34) of respondents who were on the market for 8 years or more held permanent positions.
+
+This trajectory in archaeology—where there appears to be a temporal “sweet spot” post-PhD for obtaining permanent academic positions—dovetails with a larger meta-analysis of the career trajectories of PhDs from the humanities and humanistic social sciences.
+Main et al.
+[-@main2019] used longitudinal Graduate Education Survey (GES) data from the Andrew W.
+Mellon Foundation to analyse the career trajectories of over 5,000 PhDs from disciplines in the humanities and humanistic social sciences at three time points: six months post PhD, three years post PhD, and eight years post-PhD.
+The proportion of surveyed doctorates in tenure-track or tenured faculty positions increased with time since PhD—41% of the sample held a tenure-track position six months after defending, 64% held a tenure-track position three years after defending, and 68% held a tenure-track or tenured position eight years after defending.
+In keeping with broader patterns observed across the humanities and social sciences, the variable tenures of editors on the Archaeology Jobs Wiki may thus partially reflect the periodicity of hiring relative to job seekers' time since PhD.
+
+There seems to be a shift over time in the balance of editors who identify as job-seekers versus search committee members, as indicated by the number of self-reporting users over time.
 Panel G of @fig-panel-wiki-characteristics shows the data found on each page in a section labeled 'current users', where users can volunteer to update a table tracking the status of users.
@@
 At the same time, the proportion of users self-identifying as "search committee members" increased, from under `r search_committee_members_perc_start`% in 2014-2018 to over `r search_committee_members_perc_end`% from 2020 onward.
-This suggests that over time fewer job applicants have been contributing job postings and sharing information about their experiences, and search committee members were increasingly posting job advertisements themselves.
-Overall the Archaeology Academic Jobs Wiki is best understood not as a comprehensive archive of all jobs posted during the study period, but a biased sample of the job ads that were of most interest to archaeologists applying for jobs in a given year.
-It represents the jobs that early career applicants were most focused on applying for, and sharing information on.
+This suggests that, over time, fewer job applicants have been contributing job postings and sharing information about their experiences, and search committee members have increasingly been posting job advertisements themselves.
+Overall, the Archaeology Academic Jobs Wiki is best understood not as a comprehensive archive of all jobs posted during the study period, but a biased sample of the the jobs that early career applicants were most focused on applying for, and sharing information about.
 
@@
 We also recorded the type and number of documents requested in each ad (e.g.
-cover letter, CV, statements on research, teaching, diversity , syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad.
+cover letter, CV, statements on research, teaching, diversity, syllabi, course descriptions, writing samples, transcripts) and how many names/letters of recommenders were requested in the ad.
 
@@
 
-Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of institutions according to their Carnegie Classification, a framework for classifying US colleges and universities according to the types of degrees awarded, levels of activity such as research, and topical foci [@shulman2001carnegie].
+Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of <!-- should we specify "hiring" institutions here?--> institutions according to their Carnegie Classification, a framework for classifying US colleges and universities according to the types of degrees awarded, levels of research activity, and topical foci [@shulman2001carnegie].
 Doctoral universities with high and very high research activity are by far the most active institutions hiring archaeology faculty.
@@
 
-## Geographic trends over time in job ads
+## Geographic trends in job ads over time
 
@@
 The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, groups who had similarly high population densities [@griffin2000native].
-The Mediterranean is prominent because it is the location that is often mentioned in job ads focused on classical archaeology (i.e.
+The Mediterranean is prominent because it is the region that is often mentioned in job ads focused on classical archaeology (i.e.
 archaeology of Bronze Age and Iron Age Italy and Greece).
@@
 The proportion of ads with a geographic focus on the Mediterranean has varied substantially, peaking in 2016 and experiencing a nadir in 2019, showing an inverse pattern to the Americas.
-Asia and India, the Near East, and Europe rarely occur as a geographical focus in job ads at US institutions.
+Asia and India, the Near East, and Europe <!--crying face emoji--> rarely occur as a geographical focus in job ads at US institutions.
 Asia and India, Africa, and the Americas appear correlated with each other, while the Near East and Mediterranean are inversely correlated in an opposite trend.
 
-## Method trends over time
+## Method trends in job ads over time
 
@@
 Landscape archaeology, encompassing GIS and remote sensing, has remained prominent compared to other methods [@fig-show-metho-trends].
-The popularity of this suite of methods may reflect its importance as a skill in demand by employers in the Cultural Research Management (CRM) sector.
-@morganReadyNotArchaeological2023 found that almost one-third of 599 jobs ads posted by CRM employers sought candidates with experience in GIS methods.
+The popularity of this suite of skills may reflect its demand by employers in the Cultural Research Management (CRM) sector.
+@morganReadyNotArchaeological2023 found that almost one-third of 599 jobs ads posted by CRM employers sought candidates with experience in GIS <!--JB: We're using "methods" a lot in this paragraph, so to avoid redundancy I would recommend writing GIS skills or simply GIS-->.
 Methods focused on a specific element of the archaeological record, such as Lithic analysis, Zooarchaeology and Ceramics are among the least frequently mentioned in job ads.
@@
 
-## Topic trends over time
+## Topic trends in job ads over time
 
@@
 
-The most frequently mentioned topic in the job ads is Environmental archaeology [@fig-show-topi-trends].
+The most frequently mentioned topic in this sample of job ads is Environmental archaeology [@fig-show-topi-trends].
 This category encompasses such phrases as human-environmental dynamics, interaction between humans and their environments, environmental change, climate change, historical ecology, ecological knowledge, human ecology, and ecological systems.
@@
 
-In our sample job ads were more topically rich than geographically or methodologically rich.
+In our sample, job ads were more topically rich than geographically or methodologically rich.
 That is, ads were more likely to mention multiple topics than they were to mention multiple methods or geographic locations.
@@
 Trends in job ads reflect broader shifts in intellectual and practical priorities concerning archaeology, undergraduate education, and the process of hiring professors.
-The demand for archaeology faculty, indicated by the total number of tenure-track jobs, maybe affected by a variety of factors.
+The demand for archaeology faculty, indicated by the total number of tenure-track jobs, may be affected by a variety of factors.
 
 Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-2014.
-This is at odds with previous work that has shown that contingent positions make up more than half of faculty positions [@colby2022annual], and that these numbers are increasing— between 2004 and 2019, for example, there was a 36% increase in the replacement of tenure-track positions by contingent positions [@the20222022].
-This discrepancy in our data set may be due to several factors.
-The more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
-Many of these short-term positions are not advertised nationally, but only through local email lists and are filled by people close to the hiring department, such as recently graduated students.
-Another factor is bias in our data, with most Wiki users likely seeking a tenure-track job, non-tenure-track jobs are may have been less frequently added to the Academic Jobs Wiki because they were peripheral to the goal of most users.
-This bias limits the reliability of our results on the ratio of tenure-track to non-tenure-track.
+The wiki’s emphasis on tenure-track jobs runs counter to well-established patterns in the academic job market, wherein short-term and contingent positions are fast outpacing “the last good job\[s\] in America” [@aronowitz2001].
+Research across anthropological sub-fields has shown that the number of tenure-track positions advertised within the discipline has also declined.
+Analyzing the job ads posted in *Anthropology News* in 1999–20000 versus 2019–2020, for example, Gershon and Rachok [-@HelloTristesTropes2021] found that the number of tenure-track positions decreased by 25% between their two time slices.
+Similarly, in his examination of jobs ads posted on the Biological Anthropology Academic Job Wiki between 2010–2017, Passalacqua noted that approximately one quarter of the 474 ads posted were for adjunct or visiting staff [@Passalacqua_2018, pp.773].
+Though Passalacqua did not examine how the proportion of job types changed over time, he did underscore “a current diverging trend of decreasing academic job advertisements and increasing doctoral degrees in biological anthropology which could result in serious consequences for the discipline” [-@Passalacqua_2018, pp.773].
+Such observations have a surprising longevity.
+Rogge raised the alarm nearly half a century ago, warning of the “maladaptive” consequences of the exponential rate of disciplinary growth and impending consequences for job seekers [@rogge1976, pp.839].
+The dwindling number of permanent academic positions is not a trend unique to anthropology—as of 2019, the AAUP documented a 36% increase in contingent positions over the preceding 15 years (AAUP 2022), and as of 2022, contingent positions make up more than half of faculty positions in the United States [@colby2022annual] .
+
+This discrepancy in our data set—wherein advertisements for tenure-track positions dominate the wiki despite their increasing rarity in larger market—may be due to several factors.
+The first involves the more limited circulation of advertising for short-term positions relative to advertising for tenure-track jobs.
+Many of these short-term positions are not advertised nationally, but are instead disseminated through local email lists and are filled by people close to the hiring department, such as recently graduated students.
+A second factor is bias in our data—with most Wiki users likely seeking a tenure-track job, non-tenure-track jobs may have been less frequently added to the Academic Jobs Wiki because they were peripheral to the goal of most users.
+This bias limits the reliability of our results on the ratio of tenure-track to non-tenure-track positions.
+Tellingly, Passalacqua’s examination of a similar wiki over a seven-year span showed a similar dominance of permanent positions—291/474 (63%) were for tenure-track hires at various levels (2018: Table 2).
+This suggests that these patterns are a reflection of the wiki, rather than the market itself.
+<!--JB MAYBE ADD A NOTE HERE FROM A CRISIS PAPER, PERHAPS MELISSA'S, ABOUT DISPARITY IN TRAINING AND APPLIED ANTHROPOLOGY.
+-->
 
@@
 The 2019-2021 period was also a major inflection point in the popularity of specific topics and regions in job ads.
-Calls for positions incorporating Indigenous and historical archaeology and archaeology of the Americas became far more frequent at this time, while Archaeological science, Complex societies, the Mediterranean and Near East and showed declines in popularity.
+Calls for positions incorporating Indigenous and historical archaeology, and archaeology of the Americas, became far more frequent at this time, while Archaeological science, Complex societies, and the Mediterranean and Near East showed declines in popularity.
 Similarly, the number of job ads with a geographic focus on the Americas and Africa peaks during 2019–2021.
 These shifts in the topical and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices [@dunivinBlackLivesMatter2022; @franklinFutureNowArchaeology2020; @flewellenFutureArchaeologyAntiracist2021; @lalukArchaeologySocialJustice2022].
-COVID-19 negatively impacted Black, Indigenous American, and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US [@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021].
+COVID-19 also negatively impacted Black, Indigenous American, and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US [@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021].
 The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic and the murder of George Floyd in Minneapolis by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
@@
 Due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity, however,it is illegal in the US for universities use race or ethnicity as a primary factor in hiring.
-As a result, universities appear to have tailored the content of their job ads to focus on topics where they expect Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
+As a result, universities appear to have tailored the content of their job ads to focus on topics where they expected Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
 
 This striking change in topics and geographic foci of 2019-2021 occurred against a backdrop of several longer-term trends.
-Over the course of the study period, we document a gradual decline in topics such as the archaeology of complex societies and archaeological science, the geographic foci of Mesoamerica and South America, and methods relating to landscape archaeology.
-These trends are harder to explain as we cannot link their origins to a historical event like the COVID-19 pandemic.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019].
+Over the course of the study period, we document a gradual decline in topics such as the Archaeology of complex societies and Archaeological science, the geographic foci of Mesoamerica and South America, and methods relating to Landscape archaeology <!--There is some shifting throughout the paper as to whether we capitalize the first letter of region/methods/topics or not.
+I think we should be consistent and capitalize throughout-->.
+These trends are harder to explain as we cannot link their origins to a historical event such as the COVID-19 pandemic.
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019] <!--Should we cite Cate Frieman here?-->.
 This theoretical trend might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
@@
 
-One bright spot for applicants is the decline in recent years for requests for names of recommenders in the initial application.
+One bright spot for applicants is the decline in requests for names of recommenders in the initial application in recent years.
 This may be a response to recent criticisms of the burden on the applicant of preparing numerous complex job applications [e.g.
@@
 Our results do not show how these aspirations worked out in the topical and geographic foci of the people who were actually hired for these positions.
-Future work should consider interviewing faculty hired during our study period to match up scholars witgh the ads to which they applied.
+Future work should consider interviewing faculty hired during our study period to match up scholars with the ads to which they applied.
 If the successful applicants can be identified, then we could analyze the fit between the details of the job ad and the applicant's research.
 Such an analysis would allow us to assess how effective job ads are for driving change in the discipline by setting the topics of courses that will be taught in undergraduate and graduate curricula, and research that will be supported by universities.
-We also recognize the possibility that the Academic Jobs Wiki might not capture all available positions or might over-represent certain types of institutions.
-Future work should evaluate this by comparing entries on the Wiki to other sources of information about the academic job market, such as the American Anthropological Association's AnthroGuide, or data collected directly from universities.
+We also recognize the possibility that the Academic Jobs Wiki does not capture all available positions and likely over-represents certain types of institutions.
+Future work should evaluate these biases by comparing entries on the Wiki to other sources of information about the academic job market, such as the American Anthropological Association's AnthroGuide, or data collected directly from universities.
 
@@
 Thanks to the anonymous contributors to the Academic Jobs Wiki, without whose efforts to collect and organise hundreds of job ads, this paper would not exist.
-To those of who contributed to the wiki but don't have a tenure-track job, we hope this paper shows your efforts were not wasted, but were a contribution to more accurate expectation-setting for future generations of archaeologists.
+To those of you who contributed to the wiki but did not obtain a tenure-track job, and/or decided to pivot to another industry, we hope this paper shows your time on the wiki wa not wasted.
+Your efforts have helped contribute to more accurate expectation-setting for future generations of archaeologists.
 

### 2025-05-14 analysis/paper/article.qmd
working on JB's comments

@@
 underscore, early career researchers are "supplicants"—"There is no way for applicants to point out that the job ads are taking too much time out of the scholarly community’s collective time bank" [@dennisWorstAnthroJob2022, pp.
-1].Acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs to performing well in long-list interviews and preparing teaching demonstrations.
+1].
+Acknowledging the new reality that applying for academic positions is in and of itself a full-time job, a growing number of anthropology graduate programmes are offering professionalization courses that cover the ins and outs of everything from crafting strong cover letters and CVs to performing well in long-list interviews and preparing teaching demonstrations.
 These courses can provide valuable training in the hidden curriculum of the academy, but the fact that job applications are so complicated that they require training to navigate suggests more disciplinary attention should be paid to the "market" as a historically contingent set of cultural practices.
@@
 These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted in a variety of approaches to evaluating diversity, even leading some universities to entirely omit requirements for diversity statements in job applications [@guidenAreDiversityStatements].
-**MORE DETAILS ON THIS HERE—JB WILL ADD.**
-
-# Methods
-
-<!--JB: Given that we have expanded our section on the wiki by several paragraphs, it might be useful for us to divide the methods into two subsections: (1) Dynamics of the wiki, and (2) Scraping the wiki and categorizing ads.
-Feel free to pick better titles, but I think this would provide more guidance for our readers so they do not get lost in the welter of details about the wiki.
-Lmk what you think -->
+<!-- **MORE DETAILS ON THIS HERE—JB WILL ADD.** BM: ok, thank you!
+-->
+
+# Materials and methods
+
+## Contextualising the Archaeology Academic Jobs Wiki as a data source
 
@@
 Contributors also edit the page to ask and answer questions about the positions and the application process.
-These comments have made <!--JB: If we are a bit more slippery about our tenses, we don't need to argue that the wiki is STILL crucial, esp.
-if late we will discuss its decline--> the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market.
+These comments make the Academic Jobs Wiki a unique resource for timely and specific information for job-seekers about positions they are interested in, and one of the most important internet resources for the academic job market.
 Because of its reputation for aggregating ads from diverse sources and rapidly-updated information that is not available elsewhere, the Academic Jobs Wiki has become an authoritative data source for studies of hiring trends in academia [e.g.
@@
 @fig-panel-wiki-characteristics, which allows us to assess the dynamics of editing activity over the ten years of our study period.
-Every edit includes the author's identity, recorded as either an IP address (a numerical address that includes information about the location of the user's computer) <!--good call being explicit here--> or a username.
+Every edit includes the author's identity, recorded as either an IP address (a numerical address that includes information about the location of the user's computer) or a username.
 Of the `r n_all_user_ids` unique editors in our sample, `r per_users_we_can_geolocate`% used IP addresses rather than usernames, and geolocation analysis shows that `r results_ip_addresses_tbl_us_scl`% of IP addresses were based in the US, indicating the Archaeology Jobs Wiki is primarily edited and visited by users in the U.S.
-<!--JB: technically we don't know if they are archaeologists; could be people from other disciplines e.g.
-Middle Eastern Studies, history etc.-->.
-The number of edits and active editors was roughly halved from 2013 to 2017, and then continued to gradually decline.
+The number of edits and active editors roughly halved from 2013 to 2017, and then continued to gradually decline.
 Despite this downward trend, the size of individual edits slightly increases over time, and the number of edits per job ad is relatively constant across our study period.
@@
 Editors with a longer history likely come from two groups, job seekers participating in multiple seasons of job applications, and faculty on search committees posting multiple job ads.
-
-<!--JB: Ben, this is where I'd like some of the details about min, max, etc.
-What proportion of people are on the wiki for more than a year, and how long do they stay on?
-ALSO, and I just thought of this, if ECRs are highly mobile, the same ECR might be represented by different, year-long IP addresses over their period of job seeking.
-For example, when I was on the market, I lived in Michigan, Pennsylvania, New York, England, and Massachusetts—would I not look like five distinct users who each spent 1-2 years on the wiki using our rationale here?
-This perhaps merits further discussion) -->
+That said, inferring the number of editors from the number of IP addresses is complicated by the high mobility of job seekers.
+For example one person could be represented by several IP addresses as they relocate from one city to another while working in a series of short-term appointments such as post-doctoral fellowships and adjunct teaching positions.
+This might result in an overestimation of the number of editors on the Wiki and underestimation of the duration of their editorial activity.
+
+<!--JB: Ben, this is where I'd like some of the details about min, max, etc. What proportion of people are on the wiki for more than a year, and how long do they stay on? -->
 
@@
 
+## Methods of data collection and classification
+
 For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023, we read the text and recorded the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methodological foci of the position into a Google form.
@@
 
-The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx \[ for peer reviewers the link is https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug \] to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx [ for peer reviewers the link is https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug] to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
@@
 
-Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of <!-- should we specify "hiring" institutions here?--> institutions according to their Carnegie Classification, a framework for classifying US colleges and universities according to the types of degrees awarded, levels of research activity, and topical foci [@shulman2001carnegie].
+Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of hiring institutions according to their Carnegie Classification, a framework for classifying US colleges and universities according to the types of degrees awarded, levels of research activity, and topical foci [@shulman2001carnegie].
 Doctoral universities with high and very high research activity are by far the most active institutions hiring archaeology faculty.
@@
 The proportion of ads with a geographic focus on the Mediterranean has varied substantially, peaking in 2016 and experiencing a nadir in 2019, showing an inverse pattern to the Americas.
-Asia and India, the Near East, and Europe <!--crying face emoji--> rarely occur as a geographical focus in job ads at US institutions.
+Asia and India, the Near East, and Europe <!--crying face emoji BM: sorry!
+--> rarely occur as a geographical focus in job ads at US institutions.
 Asia and India, Africa, and the Americas appear correlated with each other, while the Near East and Mediterranean are inversely correlated in an opposite trend.
@@
 The popularity of this suite of skills may reflect its demand by employers in the Cultural Research Management (CRM) sector.
-@morganReadyNotArchaeological2023 found that almost one-third of 599 jobs ads posted by CRM employers sought candidates with experience in GIS <!--JB: We're using "methods" a lot in this paragraph, so to avoid redundancy I would recommend writing GIS skills or simply GIS-->.
+@morganReadyNotArchaeological2023 found that almost one-third of 599 jobs ads posted by CRM employers sought candidates with experience in GIS.
 Methods focused on a specific element of the archaeological record, such as Lithic analysis, Zooarchaeology and Ceramics are among the least frequently mentioned in job ads.
@@
 <!--JB MAYBE ADD A NOTE HERE FROM A CRISIS PAPER, PERHAPS MELISSA'S, ABOUT DISPARITY IN TRAINING AND APPLIED ANTHROPOLOGY.
+BM: [inserting earplugs] GO FOR IT!
 -->
@@
 These trends are harder to explain as we cannot link their origins to a historical event such as the COVID-19 pandemic.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019] <!--Should we cite Cate Frieman here?-->.
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019] <!--Should we cite Cate Frieman here?
+BM: grimace emoji -->.
 This theoretical trend might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
@@
 Thanks to the anonymous contributors to the Academic Jobs Wiki, without whose efforts to collect and organise hundreds of job ads, this paper would not exist.
-To those of you who contributed to the wiki but did not obtain a tenure-track job, and/or decided to pivot to another industry, we hope this paper shows your time on the wiki wa not wasted.
+To those of you who contributed to the wiki but did not obtain a tenure-track job, and/or decided to pivot to another industry, we hope this paper shows your time on the wiki was not wasted.
 Your efforts have helped contribute to more accurate expectation-setting for future generations of archaeologists.

### 2025-05-14 analysis/paper/article.qmd
fix a citation

@@
 Previous research on professional trajectories in academic archaeology suggests that early career researchers have a limited shelf life.
-Mackie and Rockwell, for example, identify a hiring plateau that occurs around seven years after candidates defend their PhDs [-@mackieMarketShareAccounting2023:5].
-However, too little time on the market also seems to have a negative effect on placement rates—“very few graduates obtain TT employment when they are all but dissertation (ABD) or even immediately after graduation” [@mackieMarketShareAccounting2023:5].
+Mackie and Rockwell, for example, identify a hiring plateau that occurs around seven years after candidates defend their PhDs [-@mackieMarketShareAccounting2023, p.
+5].
+However, too little time on the market also seems to have a negative effect on placement rates—"very few graduates obtain TT employment when they are all but dissertation (ABD) or even immediately after graduation" [@mackieMarketShareAccounting2023, p.
+5].
 These patterns are not restricted to the United States.

### 2025-05-15 analysis/paper/article.qmd
added more details on duration of editorial activity

@@
 p_editor_life_distr <-
-ggplot(edits_for_each_year_tbl_span) +
+  edits_for_each_year_tbl_span %>% 
+  mutate(edit_span_years = edit_span_years + 0.001) %>% 
+ggplot() +
   aes(edit_span_years) +
   geom_histogram() +
+  scale_y_continuous(trans = scales::pseudo_log_trans(sigma = 0.01),
+                     breaks = c(0, 1, 10, 100, 1000), minor_breaks = NULL) +
   theme_minimal(base_size = small_base_size) +
-  scale_y_log10() +
   labs(x = "Years editing", y = "Number of editors")
@@
 
+# how many active for less than one day?
+edits_for_each_year_tbl_span_tbl_less_than_one_day  <- 
+  edits_for_each_year_tbl_span_tbl[1, ]$n
+
+edits_for_each_year_tbl_span_tbl_less_than_one_day_perc  <- 
+  round(edits_for_each_year_tbl_span_tbl[1, ]$perc, 0)
+
+# how many active for less than one year?
 edits_for_each_year_tbl_span_tbl_less_than_one_year <-
 edits_for_each_year_tbl_span_tbl %>% 
-  mutate(sum_less_than_one = if_else(cumsum(edit_span_years) < 1, n, 0)) %>% pull(sum_less_than_one) %>% sum()
+  filter(between(edit_span_years, 0.01, 1)) %>% 
+  pull(n) %>% sum()
 
@@
     edits_for_each_year_tbl_span_tbl_less_than_one_year /  sum(edits_for_each_year_tbl_span_tbl$n) * 100, 0)
+
+# how many active for more than one year?
+edits_for_each_year_tbl_span_tbl_more_than_one_year <-
+edits_for_each_year_tbl_span_tbl %>% 
+  filter(between(edit_span_years, 1, 10)) %>% 
+  pull(n) %>% sum()
+
+edits_for_each_year_tbl_span_tbl_more_than_one_year_perc <- 
+  round(edits_for_each_year_tbl_span_tbl_more_than_one_year / sum(edits_for_each_year_tbl_span_tbl$n) * 100, 0)
+
+# average length for those more than one day
+edits_for_each_year_tbl_span_tbl_mean <- 
+edits_for_each_year_tbl_span_tbl %>% 
+  filter(between(edit_span_years, 0.01, 10)) %>% 
+  pull(n) %>% mean() %>% round(1)
 
@@
 Despite this downward trend, the size of individual edits slightly increases over time, and the number of edits per job ad is relatively constant across our study period.
-The majority of editors are active on the site for less than a year (n = r`edits_for_each_year_tbl_span_tbl_less_than_one_year`, `r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%), perhaps reflecting a single season of job-searching.
+The majority of editors are active on the site for a single day (n = `r edits_for_each_year_tbl_span_tbl_less_than_one_day`, `r edits_for_each_year_tbl_span_tbl_less_than_one_day_perc`%).
+There are `r edits_for_each_year_tbl_span_tbl_less_than_one_year` (`r edits_for_each_year_tbl_span_tbl_less_than_one_year_perc`%) editors who were active for between one day and one year.
+Only `r edits_for_each_year_tbl_span_tbl_more_than_one_year` (n = `r edits_for_each_year_tbl_span_tbl_more_than_one_year_perc`%) editors were active for more than one year, with `r max(round(edits_for_each_year_tbl_span_tbl$edit_span_years,0))` years as the longest period any single editor was active on the Wiki.
+For those active for more than one day, the average duration of editorial activity is `r edits_for_each_year_tbl_span_tbl_mean` years.
+This suggests that most of the editing activity on the Wiki reflects a single season of job-searching.
 Editors with a longer history likely come from two groups, job seekers participating in multiple seasons of job applications, and faculty on search committees posting multiple job ads.
@@
 This might result in an overestimation of the number of editors on the Wiki and underestimation of the duration of their editorial activity.
-
-<!--JB: Ben, this is where I'd like some of the details about min, max, etc. What proportion of people are on the wiki for more than a year, and how long do they stay on? -->
 
@@
 Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-2014.
-The wiki’s emphasis on tenure-track jobs runs counter to well-established patterns in the academic job market, wherein short-term and contingent positions are fast outpacing “the last good job\[s\] in America” [@aronowitz2001].
+The wiki’s emphasis on tenure-track jobs runs counter to well-established patterns in the academic job market, wherein short-term and contingent positions are fast outpacing “the last good job[s] in America” [@aronowitz2001].
 Research across anthropological sub-fields has shown that the number of tenure-track positions advertised within the discipline has also declined.

### 2025-05-15 analysis/paper/article.qmd
attempt to standardize captilisation of topics and methods throughout

@@
 Landscape archaeology, although dominant, has fluctuated over the years and has been on a downtrend since 2018-2019.
-Computational and digital archaeology is the second most represented method, showing an overall increasing trend, particularly since 2020-2021.
+Computational and Digital archaeology is the second most represented method, showing an overall increasing trend, particularly since 2020-2021.
 Archaeobotany shows a strong cyclical trend, rising, falling, and then rising again over our study period.
@@
 The least common topics in our sample are Pleistocene archaeology (e.g.
-human origins, hunter-gatherer archaeology) and digital archaeology.
+human origins, hunter-gatherer archaeology) and Digital archaeology.
 
@@
 Ancient Europe and the Mediterranean, an infrequently mentioned topic for the entire study period, virtually disappeared from job ads during 2019-2021.
-Mentions of complex societies in ads decrease at a steady rate, while mentions of public archaeology consistently increase over the study period.
+Mentions of Complex societies in ads decrease at a steady rate, while mentions of public archaeology consistently increase over the study period.
 The topic of Environmental Archaeology remains high over time.
@@
 Indigenous and historical archaeology often occurs in job ads with Public archaeology and North American archaeology.
-Complex societies and environmental archaeology were frequently found in the same ads.
+Complex societies and Environmental archaeology were frequently found in the same ads.
 Bioarchaeology, Archaeological science, and Evolutionary archaeology are another cluster of topics that frequently co-occur.
 Other topics are relatively isolated.
-For example, Pleistocene archaeology and digital archaeology rarely occur with other topics.
+For example, Pleistocene archaeology and Digital archaeology rarely occur with other topics.
 
@@
 Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-2014.
-The wiki’s emphasis on tenure-track jobs runs counter to well-established patterns in the academic job market, wherein short-term and contingent positions are fast outpacing “the last good job[s] in America” [@aronowitz2001].
+The wiki’s emphasis on tenure-track jobs runs counter to well-established patterns in the academic job market, wherein short-term and contingent positions are fast outpacing "the last good job[s] in America" [@aronowitz2001].
 Research across anthropological sub-fields has shown that the number of tenure-track positions advertised within the discipline has also declined.
@@
 Similarly, in his examination of jobs ads posted on the Biological Anthropology Academic Job Wiki between 2010–2017, Passalacqua noted that approximately one quarter of the 474 ads posted were for adjunct or visiting staff [@Passalacqua_2018, pp.773].
-Though Passalacqua did not examine how the proportion of job types changed over time, he did underscore “a current diverging trend of decreasing academic job advertisements and increasing doctoral degrees in biological anthropology which could result in serious consequences for the discipline” [-@Passalacqua_2018, pp.773].
+Though Passalacqua did not examine how the proportion of job types changed over time, he did underscore "a current diverging trend of decreasing academic job advertisements and increasing doctoral degrees in biological anthropology which could result in serious consequences for the discipline" [-@Passalacqua_2018, pp.773].
 Such observations have a surprising longevity.
-Rogge raised the alarm nearly half a century ago, warning of the “maladaptive” consequences of the exponential rate of disciplinary growth and impending consequences for job seekers [@rogge1976, pp.839].
+Rogge raised the alarm nearly half a century ago, warning of the "maladaptive" consequences of the exponential rate of disciplinary growth and impending consequences for job seekers [@rogge1976, pp.839].
 The dwindling number of permanent academic positions is not a trend unique to anthropology—as of 2019, the AAUP documented a 36% increase in contingent positions over the preceding 15 years (AAUP 2022), and as of 2022, contingent positions make up more than half of faculty positions in the United States [@colby2022annual] .
@@
 This striking change in topics and geographic foci of 2019-2021 occurred against a backdrop of several longer-term trends.
-Over the course of the study period, we document a gradual decline in topics such as the Archaeology of complex societies and Archaeological science, the geographic foci of Mesoamerica and South America, and methods relating to Landscape archaeology <!--There is some shifting throughout the paper as to whether we capitalize the first letter of region/methods/topics or not.
-I think we should be consistent and capitalize throughout-->.
+Over the course of the study period, we document a gradual decline in topics such as Complex societies and Archaeological science, the geographic foci of Mesoamerica and South America, and methods relating to Landscape archaeology.
 These trends are harder to explain as we cannot link their origins to a historical event such as the COVID-19 pandemic.

### 2025-05-22 analysis/paper/article.qmd
minor edit
More text added on valorization of tenure-track jobs in higher ed.

@@
 The coordinators and contributors are nearly all anonymous or pseudonymous.
-Typically, contributors copy and paste all or some of the text of job ads into the wiki, from a variety of sources such as the *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites.
+Typically, contributors copy and paste all or some of the text of job ads into the wiki, from a variety of sources such as *The* *Chronicle of Higher Education*, *Higher Ed Jobs*, and university websites.
 Other contributors then edit the web page to add comments below an ad to share relevant information based on their experience in applying for that position.
@@
 
-The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx [ for peer reviewers the link is https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug] to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx \[ for peer reviewers the link is https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug\] to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
@@
 The proportion of ads with a geographic focus on the Mediterranean has varied substantially, peaking in 2016 and experiencing a nadir in 2019, showing an inverse pattern to the Americas.
-Asia and India, the Near East, and Europe <!--crying face emoji BM: sorry!
---> rarely occur as a geographical focus in job ads at US institutions.
+Asia and India, the Near East, and Europe rarely occur as a geographical focus in job ads at US institutions.
 Asia and India, Africa, and the Americas appear correlated with each other, while the Near East and Mediterranean are inversely correlated in an opposite trend.
@@
 Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-2014.
-The wiki’s emphasis on tenure-track jobs runs counter to well-established patterns in the academic job market, wherein short-term and contingent positions are fast outpacing "the last good job[s] in America" [@aronowitz2001].
+The wiki’s emphasis on tenure-track jobs runs counter to well-established patterns in the academic job market, wherein short-term and contingent positions are fast outpacing "the last good job\[s\] in America" [@aronowitz2001].
 Research across anthropological sub-fields has shown that the number of tenure-track positions advertised within the discipline has also declined.
@@
 Tellingly, Passalacqua’s examination of a similar wiki over a seven-year span showed a similar dominance of permanent positions—291/474 (63%) were for tenure-track hires at various levels (2018: Table 2).
-This suggests that these patterns are a reflection of the wiki, rather than the market itself.
-<!--JB MAYBE ADD A NOTE HERE FROM A CRISIS PAPER, PERHAPS MELISSA'S, ABOUT DISPARITY IN TRAINING AND APPLIED ANTHROPOLOGY.
-BM: [inserting earplugs] GO FOR IT!
--->
-
-The downward trend in tenure-track positions during 2013-2019 may be related to declining undergraduate enrollment in anthropology since 2013 [@crambChangingProfileTenureTrack2022a].
+This congruence between subfields suggests that these patterns are a reflection of the *wiki*, rather than the market itself.
+
+It is no secret that tenure-track jobs have traditionally been considered the "gold standard" of career outcomes for PhD students in the social sciences and humanities.
+As Kelsky outlines in *The Professor Is In*, her best-selling guide for PhD students pursuing faculty careers, "...most ranking graduate programs sill consider any Ph.D.
+who doesn't land a tenure-track job a failure or an aberration...
+Graduate students absorb this value system and judge themselves harshly" [-@kelsky2015, pp.
+11].
+These rubrics of success have been common across the humanities and social sciences for over half a century.
+Evaluating the demise of the National Endowment for Humanities Funded "Program to ready PhDs for Careers in Business" (CIB), which aimed to facilitate humanities PhDs' transition to the corporate world in the late 1970s, Franczak points to the dissonance between participants' incentives and outcomes:
+
+“The biggest lesson we can take from CIB is also the most obvious.
+PhDs, especially in the humanities, want to be academics.
+The deep reservoir of adjunct or contingent faculty that elite and non-elite universities alike depend on for their courses is testament to this fact—as is the excellent scholarship so many adjuncts produce without department support.
+To pretend otherwise is disingenuous, and, as CIB shows, possibly dangerous, too" [@bessner2021, pp.
+36].
+
+The desire to pursue an academic career also characterizes more recent PhD cohorts, including those within archaeology.
+Brami et al's survey found that despite the long odds and grueling precaraity, 71% of their early career respondents wished to remain in academia [=@bramiPrecariousFutureReflections2023, pp.
+242].
+Overall, the obsessive valorization of permanent faculty positions in the face of a precipitous decline in available jobs is indicative of the potent combination of denial and delusion inacross academia.
+As Cefkin and Schwegler argue "This asymmetry between the “ideal” career path and the experiences of most graduates highlights the deep and growing chasm between the future that higher education institutions envision—and strongly incentivize—and what is actually happening" [@cefkin].
+
+Returning to our results, the downward trend in tenure-track positions during 2013-2019 may be related to declining undergraduate enrollment in anthropology since 2013 [@crambChangingProfileTenureTrack2022a].
 The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus [@woolstonJuniorResearchersHit2020].
@@
 Many universities may have hoped to hire Black and Indigenous archaeologists as part of their effort to tackle systemic racism.
-Due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity, however,it is illegal in the US for universities use race or ethnicity as a primary factor in hiring.
+Due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity, however, it is illegal in the US for universities use race or ethnicity as a primary factor in hiring.
 As a result, universities appear to have tailored the content of their job ads to focus on topics where they expected Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
@@
 These trends are harder to explain as we cannot link their origins to a historical event such as the COVID-19 pandemic.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019] <!--Should we cite Cate Frieman here?
-BM: grimace emoji -->.
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019].
 This theoretical trend might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.

### 2025-05-23 analysis/paper/article.qmd
minor edits, review JB's additions to the discussion section

@@
 Tellingly, Passalacqua’s examination of a similar wiki over a seven-year span showed a similar dominance of permanent positions—291/474 (63%) were for tenure-track hires at various levels (2018: Table 2).
-This congruence between subfields suggests that these patterns are a reflection of the *wiki*, rather than the market itself.
+This congruence between subfields suggests that these patterns are a reflection of the Wiki, rather than the market itself.
 
 It is no secret that tenure-track jobs have traditionally been considered the "gold standard" of career outcomes for PhD students in the social sciences and humanities.
-As Kelsky outlines in *The Professor Is In*, her best-selling guide for PhD students pursuing faculty careers, "...most ranking graduate programs sill consider any Ph.D.
-who doesn't land a tenure-track job a failure or an aberration...
+As Kelsky outlines in *The Professor Is In*, her influential and best-selling guide for PhD students pursuing faculty careers, "...most ranking graduate programs sill consider any PhD who doesn't land a tenure-track job a failure or an aberration...
 Graduate students absorb this value system and judge themselves harshly" [-@kelsky2015, pp.
@@
 The desire to pursue an academic career also characterizes more recent PhD cohorts, including those within archaeology.
-Brami et al's survey found that despite the long odds and grueling precaraity, 71% of their early career respondents wished to remain in academia [=@bramiPrecariousFutureReflections2023, pp.
+Brami et al's survey found that despite the long odds and grueling precaraity, 71% of their early career respondents wished to remain in academia [-@bramiPrecariousFutureReflections2023, pp.
 242].
 Overall, the obsessive valorization of permanent faculty positions in the face of a precipitous decline in available jobs is indicative of the potent combination of denial and delusion inacross academia.
-As Cefkin and Schwegler argue "This asymmetry between the “ideal” career path and the experiences of most graduates highlights the deep and growing chasm between the future that higher education institutions envision—and strongly incentivize—and what is actually happening" [@cefkin].
+As Cefkin and Schwegler argue "This asymmetry between the 'ideal' career path and the experiences of most graduates highlights the deep and growing chasm between the future that higher education institutions envision—and strongly incentivize—and what is actually happening" [@cefkin].
 
@@
 These trends are harder to explain as we cannot link their origins to a historical event such as the COVID-19 pandemic.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, relational perspectives, in combination with human subjectivity and cultural framing, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019].
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, subjective, and relational perspectives, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019].
 This theoretical trend might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.

### 2025-05-23 analysis/paper/article.qmd
Git suggests I must commit before pulling so I am doing what it tells me.

--- 2025-05-22 analysis/paper/article.qmd
@@
 The desire to pursue an academic career also characterizes more recent PhD cohorts, including those within archaeology.
-Brami et al's survey found that despite the long odds and grueling precaraity, 71% of their early career respondents wished to remain in academia [=@bramiPrecariousFutureReflections2023, pp.
+Brami et al's survey found that despite the long odds and grueling precaraity, 71% of their early career respondents wished to remain in academia [= @bramiPrecariousFutureReflections2023, pp.
 242].

### 2025-05-23 analysis/paper/article.qmd
Last commit from JB! (I hope)

--- 2025-05-23 analysis/paper/article.qmd
@@
 These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted in a variety of approaches to evaluating diversity, even leading some universities to entirely omit requirements for diversity statements in job applications [@guidenAreDiversityStatements].
-<!-- **MORE DETAILS ON THIS HERE—JB WILL ADD.** BM: ok, thank you!
--->
+Institutional attitudes to diversity are likely to remain in flux for the foreseeable future.The 2023 Supreme Court decision in Students for Fair Admissions v.
+Harvard, which effectively ended affirmative action at American colleges and universities, in tandem with the Trump administration’s 2025 targeting of DEI initiatives [@knox2025] foreshadow a shift in judicial and political currents that will reshape institutional approaches to—and even definitions of—diversity.
+The 180˚pivot in the past five years—from requiring carefully crafted diversity statements as a matter of course to sweeping policy shifts rendering such documents obsolete—indicates that national and institutional commitments to diversity may only be strategic plan deep.
 
@@
 The desire to pursue an academic career also characterizes more recent PhD cohorts, including those within archaeology.
-Brami et al's survey found that despite the long odds and grueling precaraity, 71% of their early career respondents wished to remain in academia [-@bramiPrecariousFutureReflections2023, pp.
+Brami et al's survey found that despite the long odds and professional precarity, 71% of their early career respondents wished to remain in academia [= @bramiPrecariousFutureReflections2023, pp.
 242].
-Overall, the obsessive valorization of permanent faculty positions in the face of a precipitous decline in available jobs is indicative of the potent combination of denial and delusion inacross academia.
-As Cefkin and Schwegler argue "This asymmetry between the 'ideal' career path and the experiences of most graduates highlights the deep and growing chasm between the future that higher education institutions envision—and strongly incentivize—and what is actually happening" [@cefkin].
+Overall, the obsessive valorization of permanent faculty positions in the face of a precipitous decline in available jobs is indicative of the potent combination of denial and delusion across academia.
+As Cefkin and Schwegler argue "This asymmetry between the “ideal” career path and the experiences of most graduates highlights the deep and growing chasm between the future that higher education institutions envision—and strongly incentivize—and what is actually happening" [@cefkin].
 
@@
 The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus [@woolstonJuniorResearchersHit2020].
-A survey of early career researchers in archaeology captures the impact of this dip, with three-quarters of respondents experiencing negative impacts on their careers due to the pandemic [@bramiPrecariousFutureReflections2023]
+A survey of early career researchers in archaeology captures the impact of this dip, with three-quarters of respondents experiencing negative impacts on their careers due to the pandemic [@bramiPrecariousFutureReflections2023].
 
@@
 These trends are harder to explain as we cannot link their origins to a historical event such as the COVID-19 pandemic.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, subjective, and relational perspectives, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019].
-This theoretical trend might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, subjective, and relational perspectives, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019].This theoretical trend might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
 Other factors relevant to this decline may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation.
@@
 
-Our results show that job ads offer a nuanced record of disciplinary history, documenting shifting trends in topics, methods, and geographic foci, with higher chronological resolution than many other sources, such as undergraduate textbooks [@lyman2010american].
-The short periods of time spanning the rise and fall of some of the foci we observed have implications for prospective graduate students.
+The data introduced in this paper were initially collected and analyzed in spring and summer 2024 and the manuscript was drafted in autumn 2024.
+Between submitting the manuscript and receiving our initial round of revisions in April 2025, the landscape of U.S.
+higher education has shifted dramatically.
+The first hundred days of the new administration focused on instilling a climate of “economic precarity and legal uncertainty” across the sector through a cthuluesque strategy that included withholding federal funding for colleges and universities, decimating the Department of Education, revoking international student visas, and attacking Diversity Equity, and Inclusion (DEI) initiatives as “unconstitutional” [@knox2025].
+This “flood the zone” approach [@broadwater2025] shows no signs of abating; in April 2025, the president signed an executive order to overhaul the higher education accreditation system [@thewhitehouse2025], signifying the administration’s continued commitment to overhauling higher education remains intact.
+
+Such challenges to the autonomy and self-governance in higher education are particularly remarkable given that many architects of the current administration hold degrees from the very institutions they are attempting to dismantle, from Donald Trump (Wharton School of Business), to J.D.
+Vance (Yale Law School), Chris Rufo (Georgetown), Stephen Miller (Duke University), and Peter Hegseth (Harvard Kennedy School).
+Reviewing the profiles of the board of trustees for the Heritage Foundation—the conservative think tank which generated the Project 2025 report that has been used as a blueprint for feeding US higher education “into a woodchipper” (Kunder 2025)—is also illustrative.
+Two thirds of the trustees on the Heritage Foundation board have advanced degrees, while one third have doctoral degrees from institutions such as Harvard, Oxford, the University of Colorado, and the University of Texas.
+Lindsey Burke, who authored the Project 2025 chapter on the Department of Education, which includes, holds a PhD from George Mason University.
+Such career trajectories are a testament to the elite chiasmus that decries the merits and value of higher education while leveraging the credentials granted by elite institutions for professional gain.
+As Ho notes of the parallel systems of credentialing that undergird the ideology of “smartness” amongst Wall Street investment banks, playing the role of “master of the universe” requires both “especially strong doses of self-confidence and institutional legitimation” [-@ho2009, pp.
+41].
+Even archaeology is susceptible to the power and pull of such prestige hierarchies, which influence everything from hiring networks [@speakman2018; @kawaSocialNetworkUS2019] to publishing decisions [@beck2021] to research design[@wobst1983].
+
+To echo Paul Farmer, “all this is both interesting and horrible” [-@farmer2004, pp.316].
+For the purposes of our paper, however, this sociopolitical tumult provokes a suite of new questions about academic futures.
+Many archaeologists now have urgent concerns about what topics are likely to be “safe” and lead to secure ongoing employment considering the current political and funding landscape.
+Will states with stronger cultural and environmental regulations offer more stability in employment, and if so, which states are likely to be the safest?
+These are legitimate and crucial questions, but our examination of the past decade of hiring dynamics in academic archaeology cannot predict the future, especially in a mercurial political and economic landscape that is experiencing tectonic shifts in bureaucracy, organization, and funding on a near-weekly basis.
+Our study does, however, point to at least four findings that have implications for the next ten years of hiring, despite the turbulent political climate of 2024–2025.
+
+First, the economic uncertainty induced by the recent increases in tariffs is likely to have both short-term and long-term implications that slow economic growth [@schneider2025], effects which will seep into the higher education sector.
+The last two periods of major economic upheaval in the United States—the Great Recession and the COVID-19 pandemic—had pronounced and negative consequences for colleges and universities.
+These consequences were visible even within academic archaeology.
+Speakman et al., 2018, for example, show the precipitous decline in the proportion of PhDs hired into tenure-track anthropology departments after the last recession (see the steep decline in Figure 5 between 2007 and 2009), resulting in the “increased use of \[non-tenure track faculty\] by higher education institutions, and more anthropology doctorates entering the workforce”[@speakman2018, pp.12].
+As our results demonstrate, during academic year 2020–2021, only 12 tenure-track jobs in archaeology were posted on the wiki, a 70% decrease relative to the mean of 39 per year posted during our survey period (Table 1).
+As Knox and Alonso [-@knox2025] highlight, hiring freezes and program cuts are already coming into effect just four months into the new administration.
+Given past patterns, it is likely that tariffs will result in tenure-track positions in academic archaeology becoming even more scarce, and faculty and students should prepare for this.
+Whether this signals a death knell for the discipline or an opportunity for re-envisioning and reconfiguring the field [@cefkin] is up for debate.
+
+Second, our results show that hiring in academic archaeology is responsive to larger cultural and political shifts.
+The increased popularity of Indigenous and historical archaeology from 2019–2021 reflects the prominence and impact of movements such as Black Lives Matter and a then-growing interest in conversations about race and national history.
+The continual responsiveness to current events revealed by our analysis of archaeology job ads demonstrates that our narratives of the past are entangled in the social and political milieu in which we work.
+This finding is consistent with previous work—@trigger1984 outlined how archaeological research programs are shaped by national histories, while @wilk1985ancient, @blakey1983, @meltzer1983, and @soffer1983 have interrogated the entanglement of identity, ideology, national history, politics, and archaeology across space and time.
+Our study of job ads shows an active effort by archaeologists to take control of interpreting the past, offering an example of how disciplinary choices can shape archaeological research programs.
+That said, recent reports of National Science Foundation funding reviews triggered by a litany of keywords \[@johnson2025\]—including “cultural relevance,” “institutional,” “historically,” “socioeconomic,” and “systemic”—suggest the likelihood of impending topical shifts in archaeological research, as much an individual response to top-down funding pressures \[@wobst1983\] as a disciplinary response to the current historical moment.
+Third, the high chronological resolution of our analysis relative to previous studies \[@lyman2010american\] shows that the rise and fall of some of the foci we observed relatively span short periods of time.
+This limits the predictive power of our results and has implications for prospective graduate students and their mentors.
 A topic that is growing in popularity as a student begins their PhD may have peaked and be in decline before they graduate.
 Methods seem to have a much lower frequency of change in popularity relative to topical and geographic foci.
-One possible implication of this finding is that a graduate student who has invested in developing technical expertise in a method during their studies, in addition to a topic and region, might be less vulnerable to the vagaries of the job market than a student without a distinct area of technical expertise.
-
-Jobs ads also demonstrate how archaeologists draw on current events and politics to prioritize topics, regions, and methods in their hiring plans.
-This result is consistent with prior findings, such @wilk1985ancient's analysis of publications that identified trends in the explanations of the Maya collapse during the 1960s-1970s that corresponded with major contemporary American political and social issues.
-This continual responsiveness to current events demonstrates that many archaeologists accept that our narratives of the past are entangled in the social and political milieu in which we work.
-Furthermore, it shows an active effort by archaeologists to take control of interpreting the past.
+One implication of this finding is that a graduate student who has invested in developing technical expertise in a method during their studies, in addition to a topic and region, might be less vulnerable to the vagaries of the job market than a student without a distinct area of technical expertise.
+
+Finally, our work also reveals the market forces that shaped the early careers of the next generation of academic mentors.
+Assistant and associate professors who experienced the academic market as job seekers themselves between 2012–2023 probably anticipate that their students will need publications, external grants, teaching experience, interview preparation, extensive and tailored portfolios of job markets, and the ability to spend many years on the market before obtaining a permanent position, if they obtain one at all.
+This raises the question of whether mentors’ experiences on a market characterized by a recession and a pandemic will be enough to guide their students through the volatility of the next decade, and how to update our programs to prepare our graduates for success.
 

### 2025-05-24 analysis/paper/article.qmd
ready to resubmit after peer review

--- 2025-05-23 analysis/paper/article.qmd
@@
 We examined tenure-track job advertisements over the past decade to gain insights into the academic job market for archaeologists.
-Using data from the community-edited Academic Jobs Wiki for Archaeology, we examine changes in the academic job market over time.
-We analyzed the text of 431 job ads posted from 2013--2023.
-Our analysis addresses the question of how  archaeological topics, methods, and geographic regions specified in archaeological job ads shifted over time.
-We also investigate whether the labor burden for applicants has changed over time: do institutions request more information and documents from applicants at the initial stages of application, compared to a decade ago?
-We assess the influence of socio-political factors on the changing focus of research topics in the field.
-This research aims to assist current and future archaeology students, graduates, and advisors to better understand the dynamics of the academic job market and the requirements of employers.
+Using data from the community-edited Academic Jobs Wiki for Archaeology, we examined changes in the academic job market over time.
+We investigated the editing dynamics of the Wiki to understand its users and their biases.
+We then analyzed the text of 431 job ads posted from 2013--2023.
+Our analysis addresses the question of how  archaeological topics, methods, and geographic regions specified in archaeological job ads have shifted over time.
+We also explored whether the labor burden for applicants has changed over time: do institutions request more information and documents from applicants at the initial stages of application, compared to a decade ago?
+Finally, we assessed the influence of socio-political factors on the changing focus of research topics in the field.
+We conclude with implications for archaeology students, graduates, and advisors seeking to understand the dynamics of the academic job market and the requirements of employers.
 keywords: |
@@
 https://documents.saa.org/container/docs/default-source/doc-publications/style-guide/saa-style-guide_english_updated_2021_final08023c15928949dabd02faafb269fb1c.pdf?sfvrsn=c1f41c1b_11
+
+"The abstract should not exceed 200 words in length."
 
@@
 Tenure-track positions are a type of academic job that typically involve a probationary period (e.g.
-5-7 years) culminating in an evaluation of research, teaching, and service, that if successful, results in permanent employment, job security and academic freedom.
+5–7 years) culminating in an evaluation of research, teaching, and service, that if successful, results in permanent employment, job security and academic freedom.
 This obsessive focus on career trajectories is warranted.
-As of 2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions [@the20222022]; today, 71% of faculty in the US are non-tenure-track [@culver].
+As of 2019, 53.5% of colleges and universities had replaced tenure-eligible lines with contingent positions [@the20222022]; today, 71% of faculty in the U.S.
+are non-tenure-track [@culver].
 The erosion of permanent jobs in American higher education is linked to a complex intersection of political and economic factors [@beck], including decreasing federal support for higher education, concomitant increases in student debt [@gusterson2017], a pronounced shift in university investment from faculty to administration [@graeber2018, pp.
@@
 Among many scholars, there is also a perception that the last three decades have seen the academic job market become increasingly competitive.
-Across the social sciences and humanities, there are fewer jobs available relative to the number of PhD graduates, and higher numbers of short-term positions relative to permanent positions[@bessner2021; @bramiPrecariousFutureReflections2023; @kawaSocialNetworkUS2019; @kelsky2015; @mackieMarketShareAccounting2023].
+Across the social sciences and humanities, there are fewer jobs available relative to the number of PhD graduates, and higher numbers of short-term positions relative to permanent positions [@bessner2021; @bramiPrecariousFutureReflections2023; @kawaSocialNetworkUS2019; @kelsky2015; @mackieMarketShareAccounting2023].
 This market saturation has arisen in tandem with increasingly complex application requirements.
@@
 In tandem with the gradual de-investment in American higher education since 1980s [@mirowski2011] and the aftershocks of the 2008 recession, the ADEA exemption has cultivated an environment where new lines are few and far between.
-Among biological anthropologists, for example, @Passalacqua_2018 found a ratio of 0.81 PhDs to job academic advertisements per year, concluding that academic positions in biological anthropology are barely at sustainable levels.
+Among biological anthropologists, for example, @Passalacqua_2018 found ratios ranging from 1.21 to 0.82 PhDs to job academic advertisements per year, concluding that academic positions in biological anthropology are barely at sustainable levels.
 This echoes findings from other fields.
-In biomedical sciences, there is one tenure-track position in the US for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
+In biomedical sciences, there is one tenure-track position in the U.S.
+for approximately every 6.3 PhD graduates [@Ghaffarzadegan_2015].
 In engineering, Larson et al., have calculated that providing jobs for even 50% of PhD graduates would require the field to expand at an "improbable" rate of 14% per year [-@larson2014, pp.747].
@@
 PhD graduates who do not go directly into a tenure-track faculty position after graduation often go into a series of short-term appointments as part-time or limited contract instructors, also known as adjunct or contingent faculty.
-The rising number of contingent faculty has long been a concern in US higher education [@Trevithick2010].
-Data from the American Association of University Professors' 2021–22 faculty survey indicate that more than 60% of faculty positions in US universities were held by non-tenure-track full time or part time contingent faculty members [@colby2022annual].
+The rising number of contingent faculty has long been a concern in U.S.
+higher education [@Trevithick2010].
+Data from the American Association of University Professors' 2021–22 faculty survey indicate that more than 60% of faculty positions in U.S.
+universities were held by non-tenure-track full time or part time contingent faculty members [@colby2022annual].
 Contingent faculty positions are precarious because they provide low or no health and retirement benefits, limited opportunities for professional development, and lower salaries relative to tenure-track positions.
@@
 Success in tenure-track job applications in archaeology is strongly determined by where applicants get their PhDs.
-Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing].
+Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 U.S.
+graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing].
 Top-ranking programs are placing fewer than one in three graduates in tenure-track jobs [@mackieMarketShareAccounting2023].
-Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 53 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
+Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the U.S.
+in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 53 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019].
 This network analysis showed that programs with large endowments and widely cited faculty who hold prestigious awards produce the majority of tenured and tenure-track faculty.
@@
 4].
-The most recent comprehensive meta-analysis of faculty placement dynamics in the United States, which examined employment and doctoral education of all tenure-track faculty at PhD-granting universities, underscored the outsize impact of institutional prestige: 80% of faculty trained in the US came from just 20% of institutions [@wapmanQuantifyingHierarchyDynamics2022a].
+The most recent comprehensive meta-analysis of faculty placement dynamics in the United States, which examined employment and doctoral education of all tenure-track faculty at PhD-granting universities, underscored the outsize impact of institutional prestige: 80% of faculty trained in the U.S.
+came from just 20% of institutions [@wapmanQuantifyingHierarchyDynamics2022a].
 Within anthropology, such studies draw attention to a central paradox—a discipline purportedly committed to fighting social inequalities continues to reproduce its own systemic inequalities through hiring practices that favor an elite minority of candidates with prestigious affiliations.
@@
 These debates drew attention to the challenges of hiring faculty from underrepresented minorities, and resulted in a variety of approaches to evaluating diversity, even leading some universities to entirely omit requirements for diversity statements in job applications [@guidenAreDiversityStatements].
-<!-- **MORE DETAILS ON THIS HERE—JB WILL ADD.** BM: ok, thank you!
--->
+Although some time has passed since the UC experiment, institutional attitudes to diversity are likely to remain in flux for the foreseeable future.
+The 2023 Supreme Court decision in Students for Fair Admissions v.
+Harvard, which effectively ended affirmative action at American colleges and universities, in tandem with the Trump administration’s 2025 targeting of DEI initiatives [@knox2025] foreshadow a shift in judicial and political currents that is reshaping institutional approaches to—and even definitions of—diversity.
+The 180° pivot over the past five years—from requiring carefully crafted diversity statements as a matter of course to abrupt and sweeping policy shifts rendering such documents obsolete—indicates that national and institutional commitments to diversity may be only as deep as the latest strategic plan.
 
@@
 The MediaWiki technology used by Academic Jobs Wiki creates a public record of every act of adding or removing text to any page in the Wiki [@barrett2008mediawiki].
-@fig-panel-wiki-characteristics, which allows us to assess the dynamics of editing activity over the ten years of our study period.
+@fig-panel-wiki-characteristics, which allows us to measure the dynamics of editing activity over the ten years of our study period.
 Every edit includes the author's identity, recorded as either an IP address (a numerical address that includes information about the location of the user's computer) or a username.
-Of the `r n_all_user_ids` unique editors in our sample, `r per_users_we_can_geolocate`% used IP addresses rather than usernames, and geolocation analysis shows that `r results_ip_addresses_tbl_us_scl`% of IP addresses were based in the US, indicating the Archaeology Jobs Wiki is primarily edited and visited by users in the U.S.
+Of the `r n_all_user_ids` unique editors in our sample, `r per_users_we_can_geolocate`% used IP addresses rather than usernames, and geolocation analysis shows that `r results_ip_addresses_tbl_us_scl`% of IP addresses were based in the U.S., indicating the Archaeology Jobs Wiki is primarily edited and visited by users in the U.S.
 The number of edits and active editors roughly halved from 2013 to 2017, and then continued to gradually decline.
@@
 
-This trajectory in archaeology—where there appears to be a temporal “sweet spot” post-PhD for obtaining permanent academic positions—dovetails with a larger meta-analysis of the career trajectories of PhDs from the humanities and humanistic social sciences.
-Main et al.
-[-@main2019] used longitudinal Graduate Education Survey (GES) data from the Andrew W.
+This trajectory in archaeology—where there appears to be a temporal "sweet spot" post-PhD for obtaining permanent academic positions—dovetails with a larger meta-analysis of the career trajectories of PhDs from the humanities and humanistic social sciences.
+@main2019 used longitudinal Graduate Education Survey (GES) data from the Andrew W.
 Mellon Foundation to analyse the career trajectories of over 5,000 PhDs from disciplines in the humanities and humanistic social sciences at three time points: six months post PhD, three years post PhD, and eight years post-PhD.
 The proportion of surveyed doctorates in tenure-track or tenured faculty positions increased with time since PhD—41% of the sample held a tenure-track position six months after defending, 64% held a tenure-track position three years after defending, and 68% held a tenure-track or tenured position eight years after defending.
-In keeping with broader patterns observed across the humanities and social sciences, the variable tenures of editors on the Archaeology Jobs Wiki may thus partially reflect the periodicity of hiring relative to job seekers' time since PhD.
+In keeping with broader patterns observed across the humanities and social sciences, the variable tenures of editors on the Archaeology Jobs Wiki may partially reflect the temporal patterns in hiring relative to job seekers' time since PhD.
 
@@
 These data indicate that the total number of self-reporting users, from the values recording at 'How many people use the wiki?', declined by `r how_many_people_use_the_wiki_perc_start_end`% during the study period.
-At the same time, the proportion of users self-identifying as "search committee members" increased, from under `r search_committee_members_perc_start`% in 2014-2018 to over `r search_committee_members_perc_end`% from 2020 onward.
+At the same time, the proportion of users self-identifying as "search committee members" increased, from under `r search_committee_members_perc_start`% in 2014–2018 to over `r search_committee_members_perc_end`% from 2020 onward.
 This suggests that, over time, fewer job applicants have been contributing job postings and sharing information about their experiences, and search committee members have increasingly been posting job advertisements themselves.
-Overall, the Archaeology Academic Jobs Wiki is best understood not as a comprehensive archive of all jobs posted during the study period, but a biased sample of the the jobs that early career applicants were most focused on applying for, and sharing information about.
+Overall, the Archaeology Academic Jobs Wiki is best understood not as a comprehensive and neutral archive of all jobs posted during the study period, but a biased sample of the the jobs that early career applicants were most focused on applying for, and sharing information about.
 
@@
 
-For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013-2023, we read the text and recorded the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methodological foci of the position into a Google form.
+For each tenure-track job advertised on the Archaeology Academic Jobs Wiki during 2013–2023, we read the text and recorded the name of the hiring institution, the title of the position, and exact words and phrases from the ad about the topical, geographic, and methodological foci of the position into a Google form.
 The topical focus is what we understood as the intellectual core of the position—examples of topical foci included environmental archaeology, public archaeology, and North American archaeology.
@@
 After completing primary data collection, we studied the topical, geographic, and methods text of each ad.
-Following the approach of @ryan2003techniques, we collaboratively and manually reduced the variation in the raw data into 10-15 categories appearing in at least 20 (for topics and geography) or 10 (for methods) job ads to simplify analysis and visualization.
+Following the approach of @ryan2003techniques, we collaboratively and manually reduced the variation in the raw data into 10–15 categories appearing in at least 20 (for topics and geography) or 10 (for methods) job ads to simplify analysis and visualization.
 This means that some topics, such as gender (mentioned in `r n_jobs_mentioning_gender` ads) do not appear in our results because of their rarity in the job ads.
@@
 
-The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/xxx/xx \[ for peer reviewers the link is https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug\] to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
+The entire R code [@rlanguage] and data files used for all the analyses and visualizations contained in this paper are openly available at https://doi.org/10.5281/zenodo.14798941 to enable re-use of materials and improve reproducibility and transparency [@Marwick2017].
 All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository.
@@
 
-We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013-2023.
-We focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at US universities.
+We collected data from `r total_number_of_ads_in_our_sample` ads for tenure-track jobs in archaeology posted during 2013–2023.
+We focus our analysis here on the `r total_number_of_us_ads_in_our_sample` ads for positions at U.S.
+universities.
 @fig-show-basic-plots shows the count of ads for each year, where year refers to the year the job ad was posted.
@@
 The ratio of tenure-track to non-tenure-track positions is generally well above one; in other words, this data set is dominated by tenure-track positions.
-Only academic year 2013-2014 had more non-tenure-track positions than tenure-track positions, which was followed by an upward trend peaking at 2018-2019 and then declining again into the present.
+Only academic year 2013–2014 had more non-tenure-track positions than tenure-track positions, which was followed by an upward trend peaking at 2018–2019 and then declining again into the present.
 
@@
 
-Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of hiring institutions according to their Carnegie Classification, a framework for classifying US colleges and universities according to the types of degrees awarded, levels of research activity, and topical foci [@shulman2001carnegie].
+Panel A of @fig-show-map-of-hiring-institution-and-cc shows the frequencies of hiring institutions according to their Carnegie Classification, a framework for classifying U.S.
+colleges and universities according to the types of degrees awarded, levels of research activity, and topical foci [@shulman2001carnegie].
 Doctoral universities with high and very high research activity are by far the most active institutions hiring archaeology faculty.
@@
 California posted almost twice as many job ads as the next most active states.
-After California, the states that posted the most ads during 2013-2023 include New York, Texas, Pennsylvania, and Florida.
-These top five states correspond to the top five most populous US states, suggesting that rates of hiring are approximately proportional to population density.
+After California, the states that posted the most ads during 2013–2023 include New York, Texas, Pennsylvania, and Florida.
+These top five states correspond to the top five most populous U.S.
+states, suggesting that rates of hiring are approximately proportional to population density.
 The top five states for job ads are also the five states with highest number of degrees awarded in Anthropology [@IPEDSNCES].
@@
 The Southwest includes Arizona and New Mexico, with portions of California, Colorado, Nevada, Oklahoma, Texas, and Utah.
-It is archaeologically significant as the home of the Ancestral Pueblo, Hohokam, and Mogollon peoples, who practiced irrigation agriculture and lived in relatively large settlements compared to other regions of the US.
+It is archaeologically significant as the home of the Ancestral Pueblo, Hohokam, and Mogollon peoples, who practiced irrigation agriculture and lived in relatively large settlements compared to other regions of the U.S.
 The area was later occupied by the Navajo, Ute, Southern Paiute, Hopi and Zuni, groups who had similarly high population densities [@griffin2000native].
@@
 
-Demand for jobs focusing on the Americas has been generally high over the past decade, with a peak in 2019-2020 and a subsequent decrease.
-Demand for jobs focusing on Africa was very low until 2019-2020, peaking in 2020-2021.
+Demand for jobs focusing on the Americas has been generally high over the past decade, with a peak in 2019–2020 and a subsequent decrease.
+Demand for jobs focusing on Africa was very low until 2019–2020, peaking in 2020–2021.
 The proportion of ads with a geographic focus on the Mediterranean has varied substantially, peaking in 2016 and experiencing a nadir in 2019, showing an inverse pattern to the Americas.
-Asia and India, the Near East, and Europe rarely occur as a geographical focus in job ads at US institutions.
+Asia and India, the Near East, and Europe rarely occur as a geographical focus in job ads at U.S.
+institutions.
 Asia and India, Africa, and the Americas appear correlated with each other, while the Near East and Mediterranean are inversely correlated in an opposite trend.
@@
 
-Landscape archaeology, although dominant, has fluctuated over the years and has been on a downtrend since 2018-2019.
-Computational and Digital archaeology is the second most represented method, showing an overall increasing trend, particularly since 2020-2021.
+Landscape archaeology, although dominant, has fluctuated over the years and has been on a downtrend since 2018–2019.
+Computational and Digital archaeology is the second most represented method, showing an overall increasing trend, particularly since 2020–2021.
 Archaeobotany shows a strong cyclical trend, rising, falling, and then rising again over our study period.
-Archaeometry and Geoarchaeology have maintained a relatively low but steady presence in job ads, peaking in 2017-2018 and 2018-2019 and declining thereafter.
-Lithic analysis and Zooarchaeology are also mentioned relatively infrequently in job ads and show an inverse correlation with each other after 2018-2019.
+Archaeometry and Geoarchaeology have maintained a relatively low but steady presence in job ads, peaking in 2017–2018 and 2018–2019 and declining thereafter.
+Lithic analysis and Zooarchaeology are also mentioned relatively infrequently in job ads and show an inverse correlation with each other after 2018–2019.
 
@@
 
-The years 2019-2020 and 2020-2021 show striking changes in the popularity of topics in job ads.
-Indigenous and historical archaeology (which includes archaeology of the African diaspora and enslaved people) became the most popular topic at this time, rising from being one of the least popular topics from 2012-2017.
-Conversely, archaeological science, which was popular during 2012-2017, was rarely mentioned in job ads during 2019-2021.
-Ancient Europe and the Mediterranean, an infrequently mentioned topic for the entire study period, virtually disappeared from job ads during 2019-2021.
+The years 2019–2020 and 2020–2021 show striking changes in the popularity of topics in job ads.
+Indigenous and historical archaeology (which includes archaeology of the African diaspora and enslaved people) became the most popular topic at this time, rising from being one of the least popular topics from 2012–2017.
+Conversely, archaeological science, which was popular during 2012–2017, was rarely mentioned in job ads during 2019–2021.
+Ancient Europe and the Mediterranean, an infrequently mentioned topic for the entire study period, virtually disappeared from job ads during 2019–2021.
 Mentions of Complex societies in ads decrease at a steady rate, while mentions of public archaeology consistently increase over the study period.
@@
 #| label: fig-show-cooc
-#| fig-cap: "Heatmap of topic co-occurrance in job ads."
+#| fig-cap: "Heatmap of topic co-occurrence in job ads."
 
@@
 Requests for cover letter and CV decline slightly in more recent years, perhaps reflecting their status as a generic and expected component to submit without needing to be specifically requested.
-The requirement for a diversity statement is rare until 2019-2020, peaks around 2020-2021, then decreases towards the present.
-Requests for names of recommenders (either zero or three names, rarely only two names) reaches a maximum during 2019-2020, then also decline for the remainder of the study period.
-The requirement for a research statement and teaching statement increases after 2015-2016, and becomes more frequent in job ads in more recent years.
+The requirement for a diversity statement is rare until 2019–2020, peaks around 2020–2021, then decreases towards the present.
+Requests for names of recommenders (either zero or three names, rarely only two names) reaches a maximum during 2019–2020, then also decline for the remainder of the study period.
+The requirement for a research statement and teaching statement increases after 2015–2016, and becomes more frequent in job ads in more recent years.
 Requests for course descriptions, syllabus samples, teaching evaluations, transcripts and writing samples are consistently low over time (not shown here).
@@
 
-Our study of a decade of tenure-track job ads in archaeology in the US reveals diverse dynamics in the demand for specialisations in topics, methods and geographic focus, and in the instructions to applicants.
+Our study of a decade of tenure-track job ads in archaeology in the U.S.
+reveals diverse dynamics in the demand for specialisations in topics, methods and geographic focus, and in the instructions to applicants.
 While these dynamics are familiar to scholars actively seeking faculty positions, we believe this is the first time they have been quantified at such a large scale within archaeology.
@@
 
-Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013-2014.
+Overall, we found more tenure-track jobs advertised each year than non-tenure-track, with the exception of 2013–2014.
 The wiki’s emphasis on tenure-track jobs runs counter to well-established patterns in the academic job market, wherein short-term and contingent positions are fast outpacing "the last good job\[s\] in America" [@aronowitz2001].
 Research across anthropological sub-fields has shown that the number of tenure-track positions advertised within the discipline has also declined.
-Analyzing the job ads posted in *Anthropology News* in 1999–20000 versus 2019–2020, for example, Gershon and Rachok [-@HelloTristesTropes2021] found that the number of tenure-track positions decreased by 25% between their two time slices.
-Similarly, in his examination of jobs ads posted on the Biological Anthropology Academic Job Wiki between 2010–2017, Passalacqua noted that approximately one quarter of the 474 ads posted were for adjunct or visiting staff [@Passalacqua_2018, pp.773].
+Analyzing the job ads posted in *Anthropology News* in 1999–2000 versus 2019–2020, for example, Gershon and Rachok [-@HelloTristesTropes2021] found that the number of tenure-track positions decreased by 25% between their two time slices.
+Similarly, in his examination of jobs ads posted on the Biological Anthropology Academic Job Wiki between 2010–2017, Passalacqua noted that approximately one quarter of the 474 ads posted were for adjunct or visiting staff [-@Passalacqua_2018, pp.773].
 Though Passalacqua did not examine how the proportion of job types changed over time, he did underscore "a current diverging trend of decreasing academic job advertisements and increasing doctoral degrees in biological anthropology which could result in serious consequences for the discipline" [-@Passalacqua_2018, pp.773].
@@
 This bias limits the reliability of our results on the ratio of tenure-track to non-tenure-track positions.
-Tellingly, Passalacqua’s examination of a similar wiki over a seven-year span showed a similar dominance of permanent positions—291/474 (63%) were for tenure-track hires at various levels (2018: Table 2).
+Tellingly, Passalacqua's examination of a similar Wiki over a seven-year span showed a similar dominance of permanent positions—291/474 (63%) were for tenure-track hires at various levels (2018: Table 2).
 This congruence between subfields suggests that these patterns are a reflection of the Wiki, rather than the market itself.
@@
 The desire to pursue an academic career also characterizes more recent PhD cohorts, including those within archaeology.
-Brami et al's survey found that despite the long odds and grueling precaraity, 71% of their early career respondents wished to remain in academia [-@bramiPrecariousFutureReflections2023, pp.
+Brami et al's survey found that despite the long odds and professional precarity, 71% of their early career respondents wished to remain in academia [-@bramiPrecariousFutureReflections2023, pp.
 242].
-Overall, the obsessive valorization of permanent faculty positions in the face of a precipitous decline in available jobs is indicative of the potent combination of denial and delusion inacross academia.
+Overall, the obsessive valorization of permanent faculty positions in the face of a precipitous decline in available jobs is indicative of the potent combination of denial and delusion across academia.
 As Cefkin and Schwegler argue "This asymmetry between the 'ideal' career path and the experiences of most graduates highlights the deep and growing chasm between the future that higher education institutions envision—and strongly incentivize—and what is actually happening" [@cefkin].
 
-Returning to our results, the downward trend in tenure-track positions during 2013-2019 may be related to declining undergraduate enrollment in anthropology since 2013 [@crambChangingProfileTenureTrack2022a].
-The big dip during 2020-2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus [@woolstonJuniorResearchersHit2020].
-A survey of early career researchers in archaeology captures the impact of this dip, with three-quarters of respondents experiencing negative impacts on their careers due to the pandemic [@bramiPrecariousFutureReflections2023]
-
-The 2019-2021 period was also a major inflection point in the popularity of specific topics and regions in job ads.
+Returning to our results, the downward trend in tenure-track positions during 2013–2019 may be related to declining undergraduate enrollment in anthropology since 2013 [@crambChangingProfileTenureTrack2022a].
+The big dip during 2020–2021 is explained by the hiring freezes at many institutions resulting from the COVID-19 pandemic, which caused extreme disruption and uncertainty as universities focused on adapting to online instruction and assessment in an effort to minimize the spread of the virus [@woolstonJuniorResearchersHit2020].
+A survey of early career researchers in archaeology captures the impact of this dip, with three-quarters of respondents experiencing negative impacts on their careers due to the pandemic [@bramiPrecariousFutureReflections2023].
+
+The 2019–2021 period was also a major inflection point in the popularity of specific topics and regions in job ads.
 Calls for positions incorporating Indigenous and historical archaeology, and archaeology of the Americas, became far more frequent at this time, while Archaeological science, Complex societies, and the Mediterranean and Near East showed declines in popularity.
@@
 These shifts in the topical and geographic foci of job ads were likely influenced by broader cultural movements such as Black Lives Matter, protests about racial injustice, and efforts to amplify Indigenous voices [@dunivinBlackLivesMatter2022; @franklinFutureNowArchaeology2020; @flewellenFutureArchaeologyAntiracist2021; @lalukArchaeologySocialJustice2022].
-COVID-19 also negatively impacted Black, Indigenous American, and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the US [@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021].
+COVID-19 also negatively impacted Black, Indigenous American, and Hispanic communities with significantly higher infection and morbidity rates, drawing attention to racial and socio-economic inequality in the U.S.
+[@taiDisproportionateImpactCOVID192021; @mackeyRacialEthnicDisparities2021].
 The Black Lives Matter movement, dating back to 2013, intersected profoundly with the pandemic and the murder of George Floyd in Minneapolis by a police officer in May 2020, three months after the World Health Organization declared COVID-19 to be a pandemic.
@@
 
-Our data suggest that archaeology faculty at US universities participated in this movement by adjusting their hiring plans to prioritize recruiting archaeologists working on topics relevant to Black and Indigenous communities.
+Our data suggest that archaeology faculty at U.S.
+universities participated in this movement by adjusting their hiring plans to prioritize recruiting archaeologists working on topics relevant to Black and Indigenous communities.
 Many universities may have hoped to hire Black and Indigenous archaeologists as part of their effort to tackle systemic racism.
-Due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity, however, it is illegal in the US for universities use race or ethnicity as a primary factor in hiring.
+Due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity, however, it is illegal for U.S.
+universities use race or ethnicity as a primary factor in hiring.
 As a result, universities appear to have tailored the content of their job ads to focus on topics where they expected Black and Indigenous researchers to be most numerous, in an effort to ensure that these researchers would be well-represented in the pool of applicants.
 
-This striking change in topics and geographic foci of 2019-2021 occurred against a backdrop of several longer-term trends.
+This striking change in topics and geographic foci of 2019–2021 occurred against a backdrop of several longer-term trends.
 Over the course of the study period, we document a gradual decline in topics such as Complex societies and Archaeological science, the geographic foci of Mesoamerica and South America, and methods relating to Landscape archaeology.
 These trends are harder to explain as we cannot link their origins to a historical event such as the COVID-19 pandemic.
-We might speculate that a growing preference for archaeological approaches that privilege agency-driven, subjective, and relational perspectives, is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019].
+We might speculate that a growing preference for archaeological approaches that privilege agency-driven, subjective, and relational perspectives is one enduring legacy of debates in the 1980s and 90s about processualism versus post-processualism [@hodderArchaeologicalProcessIntroduction1999; @johnson2019archaeological; @Fogelin2019].
 This theoretical trend might explain why archaeological science is showing a decline, as demand for methods for analyzing artefact materiality, ontology, and power displace physical laboratory methods for technological, functional, and compositional analyses.
-Other factors relevant to this decline may include increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation.
+Other factors relevant to this decline may include the increasing difficulty of obtaining research funding to support archaeological science research, such as laboratory facilities and instrumentation.
 A decline in interest in the archaeology of complex societies may reflect several themes that intersect with broader social changes, such as growing interest in Indigenous and non-state actors in the past and an increased concern with climate change, environmental sustainability, and resilience, shifting attention away from the study of monumental architecture, elite societies, political hierarchies, and state systems.
@@
 Consistent with results from other studies [@HelloTristesTropes2021], our project documented a growing demand for research and teaching statements.
-Demand for diversity statements shows a unique trajectory, peaking in 2020-2021 and declining into the present.
-This may relate to the intersecting concerns about race, identity and class inequalities emerging during the COVID-19 pandemic, which seem to have reached a peak in 2020-2021 and then declined over time, resulting in jettisoning requests for diversity statements as the most urgent period of the pandemic moves into the past.
-Another factor here may be the debates surrounding the experiment with diversity statements in hiring at some UC campuses during 2016-2022.
-Universities that previously required a diversity statement dropped that requirement after 2022 [@guidenAreDiversityStatements].
+Demand for diversity statements shows a unique trajectory, peaking in 2020–2021 and declining into the present.
+This may relate to the intersecting concerns about race, identity and class inequalities emerging during the COVID-19 pandemic, which seem to have reached a peak in 2020–2021 and then declined over time, resulting in jettisoning requests for diversity statements as the most urgent period of the pandemic moves into the past.
+Another factor here may be the debates surrounding the experiment with diversity statements in hiring at some UC campuses during 2016–2022.
+Some universities that previously required a diversity statement dropped that requirement after 2022 [@guidenAreDiversityStatements].
 In our data we observed a reversal of the diversity statement requirement at `r n_schools_that_dropped_ds_req` universities.
@@
 A key limitation of our research is that it does not include an analysis of the academic profiles of those who were eventually hired based on this sample of job ads.
-Our results reveal collective aspirations for the future of the field, but only from those writing the job ads, typically faculty who are securely employed as tenured professors serving on hiring committees at US colleges and universities.
+Our results reveal collective aspirations for the future of the field, but only from those writing the job ads, typically faculty who are securely employed as tenured professors serving on hiring committees at U.S.
+colleges and universities.
 Our results do not show how these aspirations worked out in the topical and geographic foci of the people who were actually hired for these positions.
@@
 
-Our results show that job ads offer a nuanced record of disciplinary history, documenting shifting trends in topics, methods, and geographic foci, with higher chronological resolution than many other sources, such as undergraduate textbooks [@lyman2010american].
-The short periods of time spanning the rise and fall of some of the foci we observed have implications for prospective graduate students.
+The data introduced in this paper were initially collected and analyzed in spring and summer 2024 and the manuscript was drafted in autumn 2024.
+Between submitting the manuscript and receiving our initial round of revisions in April 2025, the landscape of U.S.
+higher education has shifted dramatically.
+The first hundred days of the new federal administration focused on instilling a climate of "economic precarity and legal uncertainty" across the sector through a cthuluesque strategy that included withholding federal funding for colleges and universities, decimating the Department of Education, revoking international student visas, and attacking Diversity Equity, and Inclusion (DEI) initiatives as "unconstitutional" [@knox2025].
+This deliberately destabilizing "flood the zone" approach [@broadwater2025] shows no signs of abating.
+In April 2025, the U.S.
+president signed an executive order to overhaul the higher education accreditation system [@thewhitehouse2025], signifying the administration’s continued commitment to align higher education institutions to its ideological agenda.
+
+Such challenges to the autonomy and self-governance in higher education are particularly remarkable given that many architects of the current administration hold degrees from the very institutions they are attempting to dismantle, from Donald Trump (Wharton School of Business), to J.D.
+Vance (Yale Law School), Chris Rufo (Georgetown), Stephen Miller (Duke University), and Peter Hegseth (Harvard Kennedy School).
+Reviewing the profiles of the board of trustees for the Heritage Foundation—the conservative think tank which generated the Project 2025 report that has been used as a blueprint for feeding U.S.
+higher education "into a woodchipper" (Kunder 2025)—is also illustrative.
+Two thirds of the trustees on the Heritage Foundation board have advanced degrees, while one third have doctoral degrees from institutions such as Harvard, Oxford, the University of Colorado, and the University of Texas.
+Lindsey Burke, who authored the Project 2025 chapter on the Department of Education, holds a PhD from George Mason University.
+Such career trajectories are a testament to the elite chiasmus that decries the merits and value of higher education while leveraging the credentials granted by elite institutions for professional gain.
+As Ho notes of the parallel systems of credentialing that undergird the ideology of "smartness" among Wall Street investment banks, playing the role of "master of the universe" requires both "especially strong doses of self-confidence and institutional legitimation" [-@ho2009, pp.
+41].
+Even archaeology is susceptible to the power and pull of such prestige hierarchies, which influence everything from hiring networks [@speakman2018; @kawaSocialNetworkUS2019] to publishing decisions [@beck2021] to research design[@wobst1983].
+
+To echo Paul Farmer, "all this is both interesting and horrible" [-@farmer2004, pp.316].
+Contextualising our job ad findings in this sociopolitical tumult provokes a suite of new questions about academic futures.
+Many archaeologists now have urgent concerns about what topics are likely to be "safe" and lead to secure ongoing employment, considering the current political and funding landscape.
+Will states with stronger cultural and environmental regulations offer more stability in employment, and if so, which states are likely to be the safest?
+These are legitimate and crucial questions, but predicting the future goes beyond the scope of our examination of the past decade of hiring dynamics in academic archaeology.
+Extrapolating future trends is inherently uncertain, especially in a mercurial political and economic landscape that is experiencing tectonic shifts in bureaucracy, organization, and funding on a near-weekly basis.
+Our study does, however, point to at least four findings that have implications for the next ten years of hiring, despite the turbulent political climate of 2024–2025.
+
+First, the macroeconomic uncertainty induced by the increased tariffs introduced in early 2025 is likely to have both short-term and long-term implications that slow economic growth [@schneider2025], with effects that will seep into the higher education sector.
+The last two periods of major economic upheaval in the U.S.—the Great Recession and the COVID-19 pandemic—had pronounced and negative consequences for colleges and universities.
+These consequences were visible even within academic archaeology.
+@speakman2018, for example, show the precipitous decline in the proportion of PhDs hired into tenure-track anthropology departments after the last recession (see the steep decline in Figure 5 between 2007 and 2009), resulting in the "increased use of \[non-tenure track faculty\] by higher education institutions, and more anthropology doctorates entering the workforce" [-@speakman2018, pp.12].
+As our results demonstrate, during academic year 2020–2021, only 12 tenure-track jobs in archaeology were posted on the Wiki, a 70% decrease relative to the mean of 39 per year posted during our survey period [@tbl-show-basic-counts].
+As @knox2025 highlight, hiring freezes and program cuts are already coming into effect just four months into the new administration.
+Given past patterns, it is likely that increased tariffs will result in tenure-track positions in academic archaeology becoming even more scarce, and faculty and students should prepare for this.
+Whether this signals a death knell for the discipline or an opportunity for re-envisioning and reconfiguring the field [@cefkin] is up for debate.
+
+Second, our results show that hiring in academic archaeology is responsive to larger cultural and political shifts.
+The increased popularity of Indigenous and historical archaeology from 2019–2021 reflects the prominence and impact of movements such as Black Lives Matter and a then-growing interest in conversations about race and national history.
+The continual responsiveness to current events revealed by our analysis of archaeology job ads demonstrates that our narratives of the past are entangled in the social and political milieu in which we work.
+This finding is consistent with previous work—@trigger1984 outlined how archaeological research programs are shaped by national histories, while @soffer1983, @blakey1983, @meltzer1983, and @wilk1985ancient have interrogated the entanglement of identity, ideology, national history, politics, and archaeology across space and time.
+Our study of job ads shows an active effort by archaeologists to take control of interpreting the past, offering an example of how disciplinary choices can shape archaeological research programs.
+That said, in early 2025 many researchers had their work disrupted by cancellations of funding from the U.S.
+National Science Foundation if their projects contained certain keywords [@Mervis], and others were told to cease activities if their projects contain certain words, including "cultural relevance," "institutional," "historically," "socioeconomic," and "systemic" [@johnson2025].
+This unprecedented keyword vetting suggests that future topical shifts in archaeological research may be as much an individual response to top-down funding pressures [@wobst1983] as a disciplinary response to the current historical moment.
+
+Third, the high chronological resolution of our analysis relative to previous studies of disciplinary trends [e.g.
+@lyman2010american] shows that the rise and fall of some of the foci we observed span relatively short periods of time.
+This limits the predictive power of our results and has implications for prospective graduate students and their mentors.
 A topic that is growing in popularity as a student begins their PhD may have peaked and be in decline before they graduate.
 Methods seem to have a much lower frequency of change in popularity relative to topical and geographic foci.
-One possible implication of this finding is that a graduate student who has invested in developing technical expertise in a method during their studies, in addition to a topic and region, might be less vulnerable to the vagaries of the job market than a student without a distinct area of technical expertise.
-
-Jobs ads also demonstrate how archaeologists draw on current events and politics to prioritize topics, regions, and methods in their hiring plans.
-This result is consistent with prior findings, such @wilk1985ancient's analysis of publications that identified trends in the explanations of the Maya collapse during the 1960s-1970s that corresponded with major contemporary American political and social issues.
-This continual responsiveness to current events demonstrates that many archaeologists accept that our narratives of the past are entangled in the social and political milieu in which we work.
-Furthermore, it shows an active effort by archaeologists to take control of interpreting the past.
+One implication of this finding is that a graduate student who has invested in developing technical expertise in a method during their studies, in addition to a topic and region, might be less vulnerable to the vagaries of the job market than a student without a distinct area of technical expertise in generating data from material records of past human behaviour.
+
+Finally, our work also reveals the market forces that shaped the early careers of the next generation of academic mentors.
+Assistant and associate professors who experienced the academic market as job seekers themselves between 2012–2023 probably anticipate that their students will need publications, external grants, teaching experience, interview preparation, extensive and tailored portfolios of job markets, and the ability to spend many years on the market before obtaining a permanent position, if they obtain one at all.
+This raises the question of whether mentors' experiences on a market characterized by a recession and a pandemic will be enough to guide their students through the volatility of the next decade, and how they should update their programs to prepare their graduates for success.
 
@@
 
-The data that support the findings of this study are openly available in Zenodo at http://doi.org/xxx/xxx.
-For peer review the link is: https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug
+The data that support the findings of this study are openly available in Zenodo at https://doi.org/10.5281/zenodo.14798941
 
@@
 
-Our supplementary materials are available at http://doi.org/xxx/xxx.
-For peer review the link is: https://zenodo.org/records/14798942?preview=1&token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6ImMyYWE2MDc3LTMxMzAtNDM1Yi1hYTJkLWMyMGMyNWVhOWMyYiIsImRhdGEiOnt9LCJyYW5kb20iOiJiMjkyMTAyN2UwNzEwNWQ2OWIyZmQ4OTkwMjU0MzY1YyJ9.slnC1BbNgZkiJjrRqmKFK9a1bpI-e9VBbjcU7L5OEeYFs3whcL3POsfVj7phDnoaB5A_6mP9wpGirEaT9Dtvug
+Our supplementary materials are available at https://doi.org/10.5281/zenodo.14798941
 

### 2025-05-24 analysis/paper/article.qmd
resubmission complete

@@
 41].
-Even archaeology is susceptible to the power and pull of such prestige hierarchies, which influence everything from hiring networks [@speakman2018; @kawaSocialNetworkUS2019] to publishing decisions [@beck2021] to research design[@wobst1983].
+Even archaeology is susceptible to the power and pull of such prestige hierarchies, which influence everything from hiring networks [@speakman2018; @kawaSocialNetworkUS2019] to publishing decisions [@beck2021] to research design [@wobst1983].
 
@@
 That said, in early 2025 many researchers had their work disrupted by cancellations of funding from the U.S.
-National Science Foundation if their projects contained certain keywords [@Mervis], and others were told to cease activities if their projects contain certain words, including "cultural relevance," "institutional," "historically," "socioeconomic," and "systemic" [@johnson2025].
-This unprecedented keyword vetting suggests that future topical shifts in archaeological research may be as much an individual response to top-down funding pressures [@wobst1983] as a disciplinary response to the current historical moment.
+National Science Foundation if their projects contained certain keywords [@Mervis], and others were told to cease activities if their projects contain keywords such as "cultural relevance," "institutional," "historically," "socioeconomic," and "systemic" [@johnson2025].
+This unprecedented keyword vetting suggests that future topical shifts in archaeological research may be as much an individual response to top-down funding pressures [cf.
+@wobst1983] as a disciplinary response to the current historical moment.
 

### 2025-07-08 README.Rmd
revisions for 'accept with revision' decision, add renv

@@
 )
-# Please put your title here to include it in the file below.
-Title <- "Title of your paper goes here"
 ```
 
-# Careers in Ruins: Academic Archaeology Job Trends From 2013 – 2023
+# Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology
 
@@
 
-This repository contains the data and code for our poster presented at the Society of American Archaeology Annual Meeting in New Orleans 17-21 April 2024:
+This repository contains the data and code for our paper:
+
+> Ben Marwick, Anne Marie Poole, Ailin Zhang, Setareh Shafizadeh, and Jess Beck (2025). Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology. _American Antiquity_
+
+An earlier version was presented as a poster at the Society of American Archaeology Annual Meeting in New Orleans 17-21 April 2024:
 
@@
 - open the `.Rproj` file in RStudio
-- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
-[DESCRIPTION](/DESCRIPTION) file).
+- run `renv::install()` to ensure you have the packages this analysis depends on 
 - finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console

### 2025-07-08 analysis/paper/article.qmd
revisions for 'accept with revision' decision, add renv

@@
 Tenure-track positions are a type of academic job that typically involve a probationary period (e.g.
-5–7 years) culminating in an evaluation of research, teaching, and service, that if successful, results in permanent employment, job security and academic freedom.
+five to seven years) culminating in an evaluation of research, teaching, and service, that if successful, results in permanent employment, job security and academic freedom.
 This obsessive focus on career trajectories is warranted.
@@
 In American anthropology, the number of doctoral anthropology graduates has increased by about 70% over the past 30 years, but the number of new faculty positions has not increased proportionally [@speakman2018choosing].
-New faculty positions have dwindled, in part due to the removal of the of the Age Discrimination in Employment Act (ADEA) exemption in 1994 which prohibited mandatory retirement ages in higher education [@earle2014].
+New faculty positions have dwindled, in part due to the removal of the Age Discrimination in Employment Act (ADEA) exemption in 1994 which prohibited mandatory retirement ages in higher education [@earle2014].
 When combined with the institution of tenure, the ADEA exemption allowed faculty to stay in their posts for as long as they liked.
@@
 Harvard, which effectively ended affirmative action at American colleges and universities, in tandem with the Trump administration’s 2025 targeting of DEI initiatives [@knox2025] foreshadow a shift in judicial and political currents that is reshaping institutional approaches to—and even definitions of—diversity.
-The 180° pivot over the past five years—from requiring carefully crafted diversity statements as a matter of course to abrupt and sweeping policy shifts rendering such documents obsolete—indicates that national and institutional commitments to diversity may be only as deep as the latest strategic plan.
+The 180 degree pivot over the past five years—from requiring carefully crafted diversity statements as a matter of course to abrupt and sweeping policy shifts rendering such documents obsolete—indicates that national and institutional commitments to diversity may be only as deep as the latest strategic plan.
 
@@
 
-## Contextualising the Archaeology Academic Jobs Wiki as a data source
+## Contextualizing the Archaeology Academic Jobs Wiki as a data source
 
@@
  filter(edit_span_years > 0.25) %>%
+  mutate(first_edit_year = as_date(first_edit_year)) %>% 
 ggplot() +
-  aes(as_date(first_edit_year),
+  aes(first_edit_year,
       reorder(edit_name, first_edit_year)) +
@@
     panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(),)
+    panel.grid.minor = element_blank())  +
+  scale_x_datetime(date_breaks = "1 year", date_labels = "%Y")
 
@@
                date_labels = "%Y") +
-  labs(x = "", y = "Sum of edits (Bytes)") +
+  labs(x = "", y = "Sum of edits (bytes)") +
   theme(axis.text.x = element_text(angle = 90,
@@
                date_labels = "%Y") +
-  labs(x = "", y = "Median edit size (Bytes)") +
+  labs(x = "", y = "Median edit size (bytes)") +
   theme(axis.text.x = element_text(angle = 90,
@@
 C: Active periods of individual editors, the black lines connect the date of the first and last edit of all editors that were active for three months or more.
-D: Typical size of each edit per year, either addition or removal of text, one word is roughly 4-6 Bytes.
+D: Typical size of each edit per year, either addition or removal of text, one word is roughly 4-6 bytes.
 E: Distribution of the numbers of edits per job ad for each year.
@@
 G: Breakdown of self-reported Wiki user categories for each year."
-#| fig-height: 5
+#| fig-height: 4
 
@@
 In Brami et al.’s [-@bramiPrecariousFutureReflections2023] survey of 419 early career researchers in European archaeology, likelihood of holding a permanent position shifted relative to time since PhD.
-Of the respondents who were 1 year or less to 4 years post-PhD, only 3% (3/116) held permanent positions.
-These numbers increased slightly for participants who had been on the market for 5–7 years, with 14% (7/49) holding permanent positions; similarly, 18% (6/34) of respondents who were on the market for 8 years or more held permanent positions.
+Of the respondents who were one year or less than four years post-PhD, only 3% (3/116) held permanent positions.
+These numbers increased slightly for participants who had been on the market for five to seven years, with 14% (7/49) holding permanent positions; similarly, 18% (6/34) of respondents who were on the market for eight years or more held permanent positions.
 
@@
 @main2019 used longitudinal Graduate Education Survey (GES) data from the Andrew W.
-Mellon Foundation to analyse the career trajectories of over 5,000 PhDs from disciplines in the humanities and humanistic social sciences at three time points: six months post PhD, three years post PhD, and eight years post-PhD.
+Mellon Foundation to analyse the career trajectories of over 5,000 PhDs from disciplines in the humanities and humanistic social sciences at three time points: six months post-PhD, three years post-PhD, and eight years post-PhD.
 The proportion of surveyed doctorates in tenure-track or tenured faculty positions increased with time since PhD—41% of the sample held a tenure-track position six months after defending, 64% held a tenure-track position three years after defending, and 68% held a tenure-track or tenured position eight years after defending.
@@
 #| fig-cap: "A: Frequency of hiring institution by Carnegie classification.
-B: Inset shows map of the United States showing the count of tenure-track job ads posted by all insititutions in each state during 2013 -- 2023"
+B: Inset shows map of the United States showing the count of tenure-track job ads posted by all institutions in each state during 2013 -- 2023"
 
@@
 
-Panel B of @fig-show-map-of-hiring-institution-and-cc shows the geographic distribution of the hiring institutions.
+The geographic distribution of hiring institutions is shown in @fig-show-map-of-hiring-institution-and-cc Panel B.
 California posted almost twice as many job ads as the next most active states.
@@
 Our study of a decade of tenure-track job ads in archaeology in the U.S.
-reveals diverse dynamics in the demand for specialisations in topics, methods and geographic focus, and in the instructions to applicants.
+reveals diverse dynamics in the demand for specializations in topics, methods and geographic focus, and in the instructions to applicants.
 While these dynamics are familiar to scholars actively seeking faculty positions, we believe this is the first time they have been quantified at such a large scale within archaeology.
@@
 Analyzing the job ads posted in *Anthropology News* in 1999–2000 versus 2019–2020, for example, Gershon and Rachok [-@HelloTristesTropes2021] found that the number of tenure-track positions decreased by 25% between their two time slices.
-Similarly, in his examination of jobs ads posted on the Biological Anthropology Academic Job Wiki between 2010–2017, Passalacqua noted that approximately one quarter of the 474 ads posted were for adjunct or visiting staff [-@Passalacqua_2018, pp.773].
+Similarly, in his examination of jobs ads posted on the Biological Anthropology Academic Jobs Wiki between 2010–2017, Passalacqua noted that approximately one quarter of the 474 ads posted were for adjunct or visiting staff [-@Passalacqua_2018, pp.773].
 Though Passalacqua did not examine how the proportion of job types changed over time, he did underscore "a current diverging trend of decreasing academic job advertisements and increasing doctoral degrees in biological anthropology which could result in serious consequences for the discipline" [-@Passalacqua_2018, pp.773].
@@
 It is no secret that tenure-track jobs have traditionally been considered the "gold standard" of career outcomes for PhD students in the social sciences and humanities.
-As Kelsky outlines in *The Professor Is In*, her influential and best-selling guide for PhD students pursuing faculty careers, "...most ranking graduate programs sill consider any PhD who doesn't land a tenure-track job a failure or an aberration...
+As Kelsky outlines in *The Professor Is In*, her influential and best-selling guide for PhD students pursuing faculty careers, "...most ranking graduate programs still consider any PhD who doesn't land a tenure-track job a failure or an aberration...
 Graduate students absorb this value system and judge themselves harshly" [-@kelsky2015, pp.
@@
 Many universities may have hoped to hire Black and Indigenous archaeologists as part of their effort to tackle systemic racism.
-Due to the 1964 Civil Rights Act which prohibits hiring based solely on race or ethnicity, however, it is illegal for U.S.
+Due to the Civil Rights Act of 1964 which prohibits hiring based solely on race or ethnicity, however, it is illegal for U.S.
 universities use race or ethnicity as a primary factor in hiring.
@@
 In recognition of this burden, not only on applicants but also on colleagues writing letters of recommendation over and over to support applicants, many hiring committees now follow the recommendations of @dennisWorstAnthroJob2022 in only requesting names and letters of recommendation at later stages of the hiring process, if at all.
-Showing sensitivity to this burden, in 2020 the American Anthropological Association issued guidance to academic departments that letters of recommendation should not be requested in the initial application, but should only be required from short-listed candidates[@jobboar2020; @Youngling2020].
+Showing sensitivity to this burden, in 2020 the American Anthropological Association issued guidance to academic departments that letters of recommendation should not be requested in the initial application, but should only be required from short-listed candidates [@jobboar2020; @Youngling2020].
 
@@
 To echo Paul Farmer, "all this is both interesting and horrible" [-@farmer2004, pp.316].
-Contextualising our job ad findings in this sociopolitical tumult provokes a suite of new questions about academic futures.
+Contextualizing our job ad findings in this sociopolitical tumult provokes a suite of new questions about academic futures.
 Many archaeologists now have urgent concerns about what topics are likely to be "safe" and lead to secure ongoing employment, considering the current political and funding landscape.

### 2025-07-08 analysis/paper/article.qmd
got the title page sorted

@@
 ---
-title: "Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology"
-author:
-  - Ben Marwick:
-      correspondence: "yes"
-      email: bmarwick@uw.edu
-      orcid: 0000-0001-7879-4531
-      institute:
-        - uw
-  - Anne Marie Poole:
-      institute: uw
-      orcid: xxx
-  - Ailin Zhang:
-      institute: uw
-      orcid: xxx
-  - Setareh Shafizadeh:
-      orcid: 0000-0003-2231-4310
-      institute: uw
-  - Jess Beck:
-      orcid: 0000-0002-7387-2307
-      institute:
-        - ucd
-institute:
-  - uw: Department of Anthropology, University of Washington, Seattle, USA
-  - ucd: School of Archaeology, University College Dublin, Ireland
-
-title-block-published: "Last updated"  
-date: now
-date-format: long
+# title: "Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology"
+# author:
+#   - Ben Marwick:
+#       correspondence: "yes"
+#       email: bmarwick@uw.edu
+#       orcid: 0000-0001-7879-4531
+#       institute:
+#         - uw
+#   - Anne Marie Poole:
+#       institute: uw
+#       orcid: xxx
+#   - Ailin Zhang:
+#       institute: uw
+#       orcid: xxx
+#   - Setareh Shafizadeh:
+#       orcid: 0000-0003-2231-4310
+#       institute: uw
+#   - Jess Beck:
+#       orcid: 0000-0002-7387-2307
+#       institute:
+#         - ucd
+# institute:
+#   - uw: Department of Anthropology, University of Washington, Seattle, USA
+#   - ucd: School of Archaeology, University College Dublin, Ireland
+#   
+# title-block-published: "Last updated"  
+# date: now
+# date-format: long
+# ---  Output Options ---
 format: 
@@
 filters:
-  - ../templates/scholarly-metadata.lua
-  - ../templates/author-info-blocks.lua
+ # - ../templates/scholarly-metadata.lua
+ # - ../templates/author-info-blocks.lua
   - ../templates/pagebreak.lua
@@
 csl: "../templates/society-for-american-archaeology.csl" # Insert path for the bib-style
-abstract: |
+reference-section-title: "References Cited" # Required title for the reference list
+---
+
+```{=html}
+<!-- Instructors to authors for AAQ
+
+https://documents.saa.org/container/docs/default-source/doc-publications/style-guide/saa-style-guide_english_updated_2021_final08023c15928949dabd02faafb269fb1c.pdf?sfvrsn=c1f41c1b_11
+
+"The abstract should not exceed 200 words in length."
+
+"AAQ recommends 10,000 words for ARTICLES... Total word count includes
+the title page, abstracts (English and second language), keywords (English and second language), text, acknowledgments, funding statements, statements of data availability and competing interests, lists of online supplemental material, figure and table captions, text of all tables (for Excel tables, copy and paste into a Word document to check the word count), notes, and the references cited section"
+
+Our notes document: https://docs.google.com/document/d/12FNfuAPIdZPBlKTEZK_HEWM8GwR_ZHyS/edit Ailin: https://docs.google.com/document/d/13OzXGFdZFraEb60ZXGwMUcRxCRKZxAQG/edit & https://docs.google.com/document/d/1NUF6dD8EnStlM6OameVvI9v6ehtf9IpWNt5SCYZgHPQ/edit
+
+Our SAA poster: Poole et al. https://osf.io/p5uzg & https://docs.google.com/presentation/d/1bljiJiE_JmywxRCaUGh4NkkcaMnQNh2uIBUcmxlp-yw/edit#slide=id.g26d90de87be_2_0
+
+Our URS23 poster: Zhang et al. https://docs.google.com/presentation/d/1u9OGBJNRilNBXAsysiY3z4Q5_qOQa4fxxzBfVlZnQYk/edit#slide=id.g2d1d7d9d51d_0_0
+
+-->
+```
+
+<!-- 
+MANUAL SAA-COMPLIANT TITLE PAGE FOR WORD OUTPUT
+
+For Word output, we create the title page manually in the document body.
+The headings below will be styled using the "Heading 1" and "Normal"
+styles from my reference docx in ../templates/template.docx
+-->
+\
+\
+
+::: {custom-style="Title Page Centered"}
+
+Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology
+
+:::
+
+<!-- Use backslashes for significant vertical spacing -->
+\
+\
+<!-- This div applies our custom centered style to the top content -->
+
+::: {custom-style="Title Page Centered not bold"}
+
+<!-- This text will use the "Title Page Centered" style (Centered, Not Bold) -->
+Ben Marwick, Anne Marie Poole, Ailin Zhang, Setareh Shafizadeh, and Jess Beck
+\
+DO NOT CITE IN ANY CONTEXT WITHOUT PERMISSION OF THE AUTHOR(S)
+\
+
+:::
+
+\
+\
+
+::: {custom-style="Authors on title page"}
+ **Ben Marwick**,  **Anne Marie Poole**,  **Ailin Zhang**,  **Setareh Shafizadeh** Department of Anthropology, University of Washington, Seattle, WA, USA (bmarwick@uw.edu, corresponding author)
+
+ **Jess Beck** School of Archaeology, University College Dublin, Ireland
+
+:::
+
+<!-- This command creates a page break after the title page. -->
+\newpage
+
 Many archaeology graduate students pursue advanced degrees in the hope of undertaking an academic career.
@@
 We conclude with implications for archaeology students, graduates, and advisors seeking to understand the dynamics of the academic job market and the requirements of employers.
-keywords: |
-  archaeology; hiring; academia; job ads 
----
-
-Keywords: `r rmarkdown::metadata$keywords`
+
+Keywords: archaeology; hiring; academia; job ads
+
+\newpage
+
+Muchos estudiantes de posgrado en arqueología cursan estudios avanzados con la esperanza de emprender una carrera académica.
+Los portales de ofertas laborales suelen ser el primer punto de referencia para quienes buscan puestos en el ámbito académico.
+Examinamos anuncios de plazas de planta (tenure-track) publicados durante la última década con el objetivo de comprender mejor el mercado laboral académico en arqueología.
+Utilizando datos del Academic Jobs Wiki for Archaeology, un recurso colaborativo editado por la comunidad, analizamos los cambios en dicho mercado a lo largo del tiempo.
+Investigamos también las dinámicas de edición del Wiki para comprender mejor a sus usuarios y los sesgos que puedan introducir.
+Posteriormente, analizamos el texto de 431 anuncios de empleo publicados entre 2013 y 2023.
+Nuestra investigación aborda cómo han variado, a lo largo del tiempo, los temas, métodos y regiones geográficas mencionados en las convocatorias para puestos académicos en arqueología.
+Asimismo, exploramos si la carga de trabajo para quienes postulan ha cambiado: ¿las instituciones solicitan hoy más documentos e información en las etapas iniciales del proceso, en comparación con hace una década?
+Finalmente, evaluamos la influencia de factores sociopolíticos en la evolución de los temas de investigación solicitados.
+Concluimos con una reflexión sobre las implicaciones de nuestros hallazgos para estudiantes, egresados y asesores que buscan comprender las dinámicas del mercado laboral académico y los requisitos de las instituciones empleadoras.
+
+Palabras clave: arqueología; contratación académica; academia; convocatorias laborales
+
+\newpage
 
 <!-- The actual document text starts here: -->
-
-```{=html}
-<!-- Instructors to authors for AAQ
-
-https://documents.saa.org/container/docs/default-source/doc-publications/style-guide/saa-style-guide_english_updated_2021_final08023c15928949dabd02faafb269fb1c.pdf?sfvrsn=c1f41c1b_11
-
-"The abstract should not exceed 200 words in length."
-
-"AAQ recommends 10,000 words for ARTICLES... Total word count includes
-the title page, abstracts (English and second language), keywords (English and second language), text, acknowledgments, funding statements, statements of data availability and competing interests, lists of online supplemental material, figure and table captions, text of all tables (for Excel tables, copy and paste into a Word document to check the word count), notes, and the references cited section"
-
-Our notes document: https://docs.google.com/document/d/12FNfuAPIdZPBlKTEZK_HEWM8GwR_ZHyS/edit Ailin: https://docs.google.com/document/d/13OzXGFdZFraEb60ZXGwMUcRxCRKZxAQG/edit & https://docs.google.com/document/d/1NUF6dD8EnStlM6OameVvI9v6ehtf9IpWNt5SCYZgHPQ/edit
-
-Our SAA poster: Poole et al. https://osf.io/p5uzg & https://docs.google.com/presentation/d/1bljiJiE_JmywxRCaUGh4NkkcaMnQNh2uIBUcmxlp-yw/edit#slide=id.g26d90de87be_2_0
-
-Our URS23 poster: Zhang et al. https://docs.google.com/presentation/d/1u9OGBJNRilNBXAsysiY3z4Q5_qOQa4fxxzBfVlZnQYk/edit#slide=id.g2d1d7d9d51d_0_0
-
--->
-```
 
@@
 #| fig-cap: "Number of users self-reporting in each category of 'current users' per year"
+#| fig-width: 3
 
@@
 G: Breakdown of self-reported Wiki user categories for each year."
-#| fig-height: 4
+#| fig-width: 3
+#| fig-height: 2
 
@@
 
-```{r}
+```{r, fig.width = 1, fig.height = 1}
 #| label: fig-show-basic-plots
@@
 The red line indicates a 1:1 ratio of tenure-track to non-tenure-track positions, bars taller than that line indicate more tenure-track than non-tenure-track positions in that year"
+#| fig-width: 1
+#| fig-height: 1
 
@@
 B: Inset shows map of the United States showing the count of tenure-track job ads posted by all institutions in each state during 2013 -- 2023"
+#| fig-width: 3
 
@@
 Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
+#| fig-width: 3
 
@@
 #| fig-cap: "Heatmap of topic co-occurrence in job ads."
+#| fig-width: 3
 
@@
 #| fig-cap: "Changing requirements in job ads over time. 'Count' refers to the number of each item required."
+#| fig-width: 3
 

### 2025-07-08 analysis/paper/article.tex
got the title page sorted

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{fontspec}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{colortbl}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{longtable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{float}
\usepackage{wrapfig}
% Load titlesec package to customize heading styles
\usepackage{titlesec}

% SAA Heading Style: Level 1 (Centered, Bold, No Number)
\titleformat{\section}{\normalfont\bfseries\filcenter}{}{0em}{}[\vspace{1em}]

% SAA Heading Style: Level 2 (Flush-left, Bold, No Number)
\titleformat{\subsection}{\normalfont\bfseries}{}{0em}{}

% SAA Heading Style: Level 3 (Flush-left, Italic, No Number)
\titleformat{\subsubsection}{\normalfont\itshape}{}{0em}{}

% Remove all section numbering
\setcounter{secnumdepth}{0}

% Customize caption format to "Figure 1." and "Table 1."
\usepackage[labelfont=bf,labelsep=period]{caption}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology},
  pdfauthor={Ben Marwick1,; Anne Marie Poole1; Ailin Zhang1; Setareh Shafizadeh1; Jess Beck2},
  pdfkeywords={archaeology; hiring; academia; job ads},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions
in Archaeology}
\author{}
\date{July 8, 2025}

\begin{document}
\maketitle
\begin{abstract}
Many archaeology graduate students pursue advanced degrees in the hope
of undertaking an academic career. Job listing websites often serve as
the first port-of-call for students seeking academic positions. We
examined tenure-track job advertisements over the past decade to gain
insights into the academic job market for archaeologists. Using data
from the community-edited Academic Jobs Wiki for Archaeology, we
examined changes in the academic job market over time. We investigated
the editing dynamics of the Wiki to understand its users and their
biases. We then analyzed the text of 431 job ads posted from 2013--2023.
Our analysis addresses the question of how archaeological topics,
methods, and geographic regions specified in archaeological job ads have
shifted over time. We also explored whether the labor burden for
applicants has changed over time: do institutions request more
information and documents from applicants at the initial stages of
application, compared to a decade ago? Finally, we assessed the
influence of socio-political factors on the changing focus of research
topics in the field. We conclude with implications for archaeology
students, graduates, and advisors seeking to understand the dynamics of
the academic job market and the requirements of employers.
\end{abstract}

\begin{titlepage}
    \centering
    \vspace*{\stretch{1.0}}
    {\huge\bfseries \thetitle\par}
    \vspace{2cm}
    {\Large \theauthor\par}
    \vspace{1cm}
    \theaffiliation
    \vspace*{\stretch{2.0}}
\end{titlepage}
\doublespacing

\setstretch{2}
\textsuperscript{1} Department of Anthropology, University of
Washington, Seattle, USA\\
\textsuperscript{2} School of Archaeology, University College Dublin,
Ireland

\textsuperscript{*} Correspondence:
\href{mailto:bmarwick@uw.edu}{Ben Marwick
\textless{}bmarwick@uw.edu\textgreater{}}

Keywords: archaeology; hiring; academia; job ads

Attend any workshop, conference, or panel that includes early career
researchers and the conversation will be steered inexorably towards the
academic job market: who is hiring, who has attained a tenure-track
position, and who is out of luck this season. Tenure-track positions are
a type of academic job that typically involve a probationary period
(e.g.~five to seven years) culminating in an evaluation of research,
teaching, and service, that if successful, results in permanent
employment, job security and academic freedom. This obsessive focus on
career trajectories is warranted. As of 2019, 53.5\% of colleges and
universities had replaced tenure-eligible lines with contingent
positions (American Association of University Professors 2022a); today,
71\% of faculty in the U.S. are non-tenure-track (Culver and Kezar
2022). The erosion of permanent jobs in American higher education is
linked to a complex intersection of political and economic factors (Beck
2025), including decreasing federal support for higher education,
concomitant increases in student debt (Gusterson 2017), a pronounced
shift in university investment from faculty to administration (Graeber
2018:162--163), and the growing privatization and market orientation of
scientific research (Mirowski 2011).

These cultural and economic shifts are also well encapsulated by
anecdata; any conversation with a junior scholar will reveal the impact
of these larger professional transformations on personal and
professional trajectories. The first author on this paper has applied to
five academic positions over the course of his career, while the last
author has applied to one hundred. A twenty-fold increase in the number
of applications required to obtain a workable long-term academic
position may seem preposterous, but these changing requirements for
applicants are familiar to anyone who has obtained their PhDs since the
Great Recession of 2008. This competitive landscape has deeper roots
than the last economic crisis; writing more than three decades ago,
Rabinow pinpointed ``an awareness\ldots{} that standards have changed
during the last thirty years and the quantitative and qualitative
demands for entry into the system are immeasurably higher now''
(1992:66).

These higher demands include increased overall productivity, higher
national and international mobility, more time spent in precarious
short-term contracts, and the investment of a greater amount of time and
energy in applying for long-term positions. In their survey of early
career researchers in European archaeology, for example, Brami et
al.~found that 52\% of respondents who had obtained their PhD at least
two years before the survey were on fixed-term contracts, while only 9\%
were in permanent positions (2023:Figure 5). In their words, ``the queue
of eager post-docs hoping for a long-term appointment is getting longer,
as is the average time spent in the `queue'\,'' (2023:239). Ribeiro and
Giamakis emphasize that this demographic comprises an academic precariat
that is essential to the functioning of universities, but consists of a
workforce perpetually ``stranded between employment and unemployment''
(Ribeiro and Giamakis 2023:10).

Junior scholars with faculty ambitions are thus stuck in a double-bind,
forced to funnel their energy into seemingly endless applications for
which they receive little to no feedback and which remain invisible on
their CVs. As Dennis et al.~underscore, early career researchers are
``supplicants''---``There is no way for applicants to point out that the
job ads are taking too much time out of the scholarly community's
collective time bank'' (Dennis et al. 2022:1). Acknowledging the new
reality that applying for academic positions is in and of itself a
full-time job, a growing number of anthropology graduate programmes are
offering professionalization courses that cover the ins and outs of
everything from crafting strong cover letters and CVs to performing well
in long-list interviews and preparing teaching demonstrations. These
courses can provide valuable training in the hidden curriculum of the
academy, but the fact that job applications are so complicated that they
require training to navigate suggests more disciplinary attention should
be paid to the ``market'' as a historically contingent set of cultural
practices. As Rabinow maintains, ``if we want ethical considerations to
play a central role in the articulation of truth and power---and I think
we do---then bringing such considerations into view is the necessary
first step towards recognizing who we are today and setting out on the
road to a better place'' (1992:71).

We argue that to understand the dynamics of hiring in academic
archaeology, we must start at the beginning, with the job ad itself.
There is an abundance of recent work on who does and does not get hired
into tenure-track positions in anthropology and archaeology, but less
scholarship on how those hires unfold. Research on hiring dynamics has
amply demonstrated that \emph{where} you get your PhD has an enormous
impact on if and where you get a tenure-track position (Kawa et al.
2019; Mackie and Rockwell 2023), but how about \emph{what} you study?
Are there topical, methodological, or regional specialties that are in
consistent demand, or is there continual flux generated by intellectual
fads, economic incentives, and political attention? Are job applications
more demanding now than they were in the past, or is this a specious
perception generated by the sheer volume of applications necessary to
obtain a long-term academic position?

To answer these questions, this paper explores the demand-side of the
academic job market for archaeologists in the United States. Our study
had two aims: first, to determine if disciplinary trends could be
discerned in the topical, geographic, and methodological foci of the
positions advertised over a ten-year period, and second, to investigate
how requirements for applicants have changed over time.

\section{Background}\label{background}

The academic job market is a source of uncertainty, for both job seekers
and hiring committees. Job seekers must search for ads that are
well-matched to the skill sets and professional experiences that they
have taken years to cultivate. Hiring committees must sift through tens
to hundreds of applicants in the hope of finding a candidate that meets
all of their needs and helps to realize their visions for their program
and for the discipline. Finding the right candidate for the
position---or the right position for the candidate---is akin to finding
a needle in a haystack.

Among many scholars, there is also a perception that the last three
decades have seen the academic job market become increasingly
competitive. Across the social sciences and humanities, there are fewer
jobs available relative to the number of PhD graduates, and higher
numbers of short-term positions relative to permanent positions (Bessner
and Brenes 2021; Brami et al. 2023; Kawa et al. 2019; Kelsky 2015;
Mackie and Rockwell 2023). This market saturation has arisen in tandem
with increasingly complex application requirements. Many job ads now
solicit specific documents in \emph{addition} to the cover letter and
CV, such as teaching, research, and diversity statements. These
additional documents must be tailored for each application, making the
process of applying for jobs a full-time job of its own.

In American anthropology, the number of doctoral anthropology graduates
has increased by about 70\% over the past 30 years, but the number of
new faculty positions has not increased proportionally (Speakman,
Hadden, Colvin, Cramb, Jones, Jones, Kling, et al. 2018). New faculty
positions have dwindled, in part due to the removal of the Age
Discrimination in Employment Act (ADEA) exemption in 1994 which
prohibited mandatory retirement ages in higher education (Earle and
DelPo Kulow 2014). When combined with the institution of tenure, the
ADEA exemption allowed faculty to stay in their posts for as long as
they liked. The median age for faculty in the U.S. now ranks among the
highest for all professions (Kaskie 2016). In tandem with the gradual
de-investment in American higher education since 1980s (Mirowski 2011)
and the aftershocks of the 2008 recession, the ADEA exemption has
cultivated an environment where new lines are few and far between. Among
biological anthropologists, for example, Passalacqua (2018) found ratios
ranging from 1.21 to 0.82 PhDs to job academic advertisements per year,
concluding that academic positions in biological anthropology are barely
at sustainable levels. This echoes findings from other fields. In
biomedical sciences, there is one tenure-track position in the U.S. for
approximately every 6.3 PhD graduates (Ghaffarzadegan et al. 2015). In
engineering, Larson et al., have calculated that providing jobs for even
50\% of PhD graduates would require the field to expand at an
``improbable'' rate of 14\% per year (2014:747). As they emphasize,
``\ldots{} the system in many places is saturated, far beyond capacity
to absorb new PhDs in academia at the rates they are being produced''
(Larson et al. 2014:749).

PhD graduates who do not go directly into a tenure-track faculty
position after graduation often go into a series of short-term
appointments as part-time or limited contract instructors, also known as
adjunct or contingent faculty. The rising number of contingent faculty
has long been a concern in U.S. higher education (Trevithick 2010). Data
from the American Association of University Professors' 2021--22 faculty
survey indicate that more than 60\% of faculty positions in U.S.
universities were held by non-tenure-track full time or part time
contingent faculty members (American Association of University
Professors 2022b). Contingent faculty positions are precarious because
they provide low or no health and retirement benefits, limited
opportunities for professional development, and lower salaries relative
to tenure-track positions. PhD graduates who have a sequence of
short-term academic positions are often disadvantaged financially due to
low compensation combined with expensive and frequent relocation,
restrictive socially due to isolation from family and community, and
limited professionally due to being ineligible for many decision-making
roles at the universities at which they work (Platzer and Allison 2018).

Success in tenure-track job applications in archaeology is strongly
determined by where applicants get their PhDs. Data from the 2014--2015
AnthroGuide publication of the American Anthropological Association
shows that just ten out of over 100 U.S. graduate programs produced over
30\% of the graduates hired into tenure-track faculty positions
(Speakman, Hadden, Colvin, Cramb, Jones, Jones, Kling, et al. 2018).
Top-ranking programs are placing fewer than one in three graduates in
tenure-track jobs (Mackie and Rockwell 2023). Similarly, a network
analysis of 1,918 faculty holding tenured or tenure-track positions at
PhD-granting anthropology programs in the U.S. in 2015 (including 506
archaeologists) found that the just fifteen graduate programs produced
53 percent of tenured and tenure-track positions (Kawa et al. 2019).
This network analysis showed that programs with large endowments and
widely cited faculty who hold prestigious awards produce the majority of
tenured and tenure-track faculty.

Hiring bias predicated on the prestige of specific institutions and
programs is not unique to anthropology. Targeted studies of sociology,
communication, operations research, and industrial systems engineering
show similar dynamics (Barnett et al. 2010; Castillo et al. 2018; Feeley
and Tutzauer 2021; Nevin 2019). Broader analyses comparing computer
science, business, and history identify comparable patterns among
disparate disciplines across the humanities, social sciences, and STEM
field---``across disciplines, prestige hierarchies make the most
accurate predictions of faculty placement'' (Clauset et al. 2015:4). The
most recent comprehensive meta-analysis of faculty placement dynamics in
the United States, which examined employment and doctoral education of
all tenure-track faculty at PhD-granting universities, underscored the
outsize impact of institutional prestige: 80\% of faculty trained in the
U.S. came from just 20\% of institutions (Wapman et al. 2022). Within
anthropology, such studies draw attention to a central paradox---a
discipline purportedly committed to fighting social inequalities
continues to reproduce its own systemic inequalities through hiring
practices that favor an elite minority of candidates with prestigious
affiliations.

One way that some hiring committees are tackling prestige biases is by
providing detailed instructions to applicants on how to prepare their
application materials. In theory, detailed applications will allow the
hiring committee to focus on evaluating the accomplishments of
candidates across common categories rather than ranking based on
prestige signals in a CV, such as a name of the applicant's graduate
program. This push for equity has resulted in job ads that are often
highly prescriptive in the types of documents that applicants should
submit. For example, in addition to a cover letter and CV, job ads in
many fields now require applicants to submit short statements detailing
their previous and future contributions to teaching, research, and
diversity. In a comparison of job ads from 1999--2000 and 2019--2020
published in \emph{Anthropology News}, Gershon and Rachok (2021) noticed
an increase in the number of materials requested from applicants. For
example, twice as many 2019--2020 job ads requested writing samples
compared to 1999--2000, and nearly four times as many requested
statements of teaching philosophy. In their review of the `worst job ads
of 2021' Dennis et al. (2022) reported job ads from Oberlin College and
Grinnell College with particularly extreme requirements, which
respectively requested nine and thirteen documents from applicants. This
high burden on applicants disproportionately favors people with more
time and financial resources to prepare the required materials.

A notable requirement that has emerged in recent years is a diversity
statement, where the applicant describes their knowledge of, prior
contributions to, and future goals for advancing diversity, equity, and
inclusion. This requirement has been much-debated largely due to an
experiment during 2016--2022 at several University of California
campuses that used diversity statements as the first cut for selecting
candidates for tenure-track faculty positions (Soucek 2022). Only
candidates that scored highly on their diversity statements would have
the rest of their application evaluated. This experiment generated
intense and widespread public debate about the merits and risks of
requiring and using diversity statements. These debates drew attention
to the challenges of hiring faculty from underrepresented minorities,
and resulted in a variety of approaches to evaluating diversity, even
leading some universities to entirely omit requirements for diversity
statements in job applications (Guiden 2024). Although some time has
passed since the UC experiment, institutional attitudes to diversity are
likely to remain in flux for the foreseeable future. The 2023 Supreme
Court decision in Students for Fair Admissions v. Harvard, which
effectively ended affirmative action at American colleges and
universities, in tandem with the Trump administration's 2025 targeting
of DEI initiatives (Knox and Alonso 2025) foreshadow a shift in judicial
and political currents that is reshaping institutional approaches
to---and even definitions of---diversity. The 180 degree pivot over the
past five years---from requiring carefully crafted diversity statements
as a matter of course to abrupt and sweeping policy shifts rendering
such documents obsolete---indicates that national and institutional
commitments to diversity may be only as deep as the latest strategic
plan.

\section{Materials and methods}\label{materials-and-methods}

\subsection{Contextualizing the Archaeology Academic Jobs Wiki as a data
source}\label{contextualizing-the-archaeology-academic-jobs-wiki-as-a-data-source}

Our primary data source is the Archaeology Academic Jobs Wiki.
Originating in 2007, the wiki is a set of freely accessible web pages
that anyone can edit anonymously or with a free user account. The site
is hosted by Fandom, a for-profit company. The Archaeology pages are
part of the Academic Jobs Wiki, which coordinates similar
collaboratively-edited resources for around 40 academic disciplines. The
coordinators and contributors are nearly all anonymous or pseudonymous.
Typically, contributors copy and paste all or some of the text of job
ads into the wiki, from a variety of sources such as \emph{The}
\emph{Chronicle of Higher Education}, \emph{Higher Ed Jobs}, and
university websites. Other contributors then edit the web page to add
comments below an ad to share relevant information based on their
experience in applying for that position. These edits result in
annotations such as a tally of how many people have applied, the dates
of events such as requests for more materials, interviews, offer made,
rejection notices, etc. Contributors also edit the page to ask and
answer questions about the positions and the application process. These
comments make the Academic Jobs Wiki a unique resource for timely and
specific information for job-seekers about positions they are interested
in, and one of the most important internet resources for the academic
job market. Because of its reputation for aggregating ads from diverse
sources and rapidly-updated information that is not available elsewhere,
the Academic Jobs Wiki has become an authoritative data source for
studies of hiring trends in academia (Musial and Holmes 2018; e.g.
Passalacqua 2018) and a widely recommended resource for applicants (e.g.
Lightfoot et al. 2021).

\begin{figure}

\centering{

\includegraphics{article_files/figure-pdf/fig-panel-wiki-characteristics-1.pdf}

}

\caption{\label{fig-panel-wiki-characteristics}A: Total number of edits
to each Wiki page (additions and deletions) per year. B: Number of
unique editors, as represented by distinct usernames or IP addresses
(recorded for editors without usernames), active per year. C: Active
periods of individual editors, the black lines connect the date of the
first and last edit of all editors that were active for three months or
more. D: Typical size of each edit per year, either addition or removal
of text, one word is roughly 4-6 bytes. E: Distribution of the numbers
of edits per job ad for each year. F: Histogram showing that the
majority of editors are only active on the site for less than one year.
G: Breakdown of self-reported Wiki user categories for each year.}

\end{figure}%

Collaboratively-edited resources such as the Academic Jobs Wiki can be
highly variable in the amount and type of editorial activity over time.
It is important to characterise this activity to assess the reliability
of the content is as a source of information about the job market. The
MediaWiki technology used by Academic Jobs Wiki creates a public record
of every act of adding or removing text to any page in the Wiki (Barrett
2008). Figure~\ref{fig-panel-wiki-characteristics}, which allows us to
measure the dynamics of editing activity over the ten years of our study
period. Every edit includes the author's identity, recorded as either an
IP address (a numerical address that includes information about the
location of the user's computer) or a username. Of the 2824 unique
editors in our sample, 91\% used IP addresses rather than usernames, and
geolocation analysis shows that 81\% of IP addresses were based in the
U.S., indicating the Archaeology Jobs Wiki is primarily edited and
visited by users in the U.S. The number of edits and active editors
roughly halved from 2013 to 2017, and then continued to gradually
decline. Despite this downward trend, the size of individual edits
slightly increases over time, and the number of edits per job ad is
relatively constant across our study period. The majority of editors are
active on the site for a single day (n = 2241, 81\%). There are 412
(15\%) editors who were active for between one day and one year. Only
113 (n = 4\%) editors were active for more than one year, with 8 years
as the longest period any single editor was active on the Wiki. For
those active for more than one day, the average duration of editorial
activity is 4.6 years. This suggests that most of the editing activity
on the Wiki reflects a single season of job-searching. Editors with a
longer history likely come from two groups, job seekers participating in
multiple seasons of job applications, and faculty on search committees
posting multiple job ads. That said, inferring the number of editors
from the number of IP addresses is complicated by the high mobility of
job seekers. For example one person could be represented by several IP
addresses as they relocate from one city to another while working in a
series of short-term appointments such as post-doctoral fellowships and
adjunct teaching positions. This might result in an overestimation of
the number of editors on the Wiki and underestimation of the duration of
their editorial activity.

Longer-term editors who represent job seekers active for multiple
seasons reflect one of the underdiscussed realities of the academic job
market in archaeology---time spent on that market. Previous research on
professional trajectories in academic archaeology suggests that early
career researchers have a limited shelf life. Mackie and Rockwell, for
example, identify a hiring plateau that occurs around seven years after
candidates defend their PhDs (2023:5). However, too little time on the
market also seems to have a negative effect on placement rates---``very
few graduates obtain TT employment when they are all but dissertation
(ABD) or even immediately after graduation'' (Mackie and Rockwell
2023:5). These patterns are not restricted to the United States. In
Brami et al.'s (2023) survey of 419 early career researchers in European
archaeology, likelihood of holding a permanent position shifted relative
to time since PhD. Of the respondents who were one year or less than
four years post-PhD, only 3\% (3/116) held permanent positions. These
numbers increased slightly for participants who had been on the market
for five to seven years, with 14\% (7/49) holding permanent positions;
similarly, 18\% (6/34) of respondents who were on the market for eight
years or more held permanent positions.

This trajectory in archaeology---where there appears to be a temporal
``sweet spot'' post-PhD for obtaining permanent academic
positions---dovetails with a larger meta-analysis of the career
trajectories of PhDs from the humanities and humanistic social sciences.
Main et al. (2019) used longitudinal Graduate Education Survey (GES)
data from the Andrew W. Mellon Foundation to analyse the career
trajectories of over 5,000 PhDs from disciplines in the humanities and
humanistic social sciences at three time points: six months post-PhD,
three years post-PhD, and eight years post-PhD. The proportion of
surveyed doctorates in tenure-track or tenured faculty positions
increased with time since PhD---41\% of the sample held a tenure-track
position six months after defending, 64\% held a tenure-track position
three years after defending, and 68\% held a tenure-track or tenured
position eight years after defending. In keeping with broader patterns
observed across the humanities and social sciences, the variable tenures
of editors on the Archaeology Jobs Wiki may partially reflect the
temporal patterns in hiring relative to job seekers' time since PhD.

There seems to be a shift over time in the balance of editors who
identify as job-seekers versus search committee members, as indicated by
the number of self-reporting users over time. Panel G of
Figure~\ref{fig-panel-wiki-characteristics} shows the data found on each
page in a section labeled `current users', where users can volunteer to
update a table tracking the status of users. These data indicate that
the total number of self-reporting users, from the values recording at
`How many people use the wiki?', declined by 82\% during the study
period. At the same time, the proportion of users self-identifying as
``search committee members'' increased, from under 5\% in 2014--2018 to
over 14\% from 2020 onward. This suggests that, over time, fewer job
applicants have been contributing job postings and sharing information
about their experiences, and search committee members have increasingly
been posting job advertisements themselves. Overall, the Archaeology
Academic Jobs Wiki is best understood not as a comprehensive and neutral
archive of all jobs posted during the study period, but a biased sample
of the the jobs that early career applicants were most focused on
applying for, and sharing information about.

\subsection{Methods of data collection and
classification}\label{methods-of-data-collection-and-classification}

For each tenure-track job advertised on the Archaeology Academic Jobs
Wiki during 2013--2023, we read the text and recorded the name of the
hiring institution, the title of the position, and exact words and
phrases from the ad about the topical, geographic, and methodological
foci of the position into a Google form. The topical focus is what we
understood as the intellectual core of the position---examples of
topical foci included environmental archaeology, public archaeology, and
North American archaeology. The geographic focus is the region of the
world about which the ideal candidate has scholarly expertise, for
example, Southwest US, Mediterranean, or Asia and India. The methods
focus is the data-generating sub-field of archaeology mentioned in the
ad. Examples of methods used in this study include archaeobotany, lithic
analysis, and zooarchaeology. We also recorded the type and number of
documents requested in each ad (e.g.~cover letter, CV, statements on
research, teaching, diversity, syllabi, course descriptions, writing
samples, transcripts) and how many names/letters of recommenders were
requested in the ad.

After completing primary data collection, we studied the topical,
geographic, and methods text of each ad. Following the approach of Ryan
and Bernard (2003), we collaboratively and manually reduced the
variation in the raw data into 10--15 categories appearing in at least
20 (for topics and geography) or 10 (for methods) job ads to simplify
analysis and visualization. This means that some topics, such as gender
(mentioned in 6 ads) do not appear in our results because of their
rarity in the job ads. Full details of the category reduction, showing
the mapping between exact phrases found in the job ads and the
categories we used for our analysis, can be found in our Supplementary
Materials. Our final topic categories were: American archaeology,
Ancient Europe and Mediterranean, Archaeological science, Archaeological
theory, Bioarchaeology, Complex societies, Digital archaeology,
Environmental archaeology, Evolutionary anthropology, Indigenous and
historical archaeology, North Mesoamerican archaeology, Pleistocene
archaeology, and Public archaeology. Our geographic categories were:
Africa, Americas, Asia \& India, Canada \& Arctic, Europe,
Mediterranean, Meso- \& South America, Near East, Oceania, Midwest US,
Northeastern US, Southeast US, Southwest US, and Western US. Our methods
categories were: Archaeobotany, Archaeometry, Bioarchaeology, Ceramic
analysis, Computational and Digital archaeology, Geoarchaeology,
Landscape analysis, Lithic analysis, Material culture analysis, and
Zooarchaeology.

Individual ads could be recorded to have multiple or none of these
topical, geographic, and methods foci, and some of the foci overlap.
Some topics include geographic regions because this is how they are
typically understood by archaeologists. For example, Mesoamerican
archaeology is understood to refer to a specific time period \emph{and}
a specific geographic region. Similarly, we recorded digital archaeology
as both a method (when a job ad also had a clearly distinct topical
focus, such as historic archaeology) and a topic (when there were no
other topics mentioned in the job ad). While this polythetic approach
results in categorical overlaps that can make the data challenging to
interpret (Kuckartz 2014), in our view this reflects the complex
realities of how search committees express their needs when searching
for new faculty. Acknowledgement of overlaps also produces new insights
into hiring dynamics through revealing intersections between different
foci.

The entire R code (R Core Team 2024) and data files used for all the
analyses and visualizations contained in this paper are openly available
at https://doi.org/10.5281/zenodo.14798941 to enable re-use of materials
and improve reproducibility and transparency (Marwick 2017). All of the
figures, tables, and statistical test results presented here can be
independently reproduced with the code and data in this repository. The
code is released under the MIT license, the data as CC-0, and the
figures as CC-BY, to enable maximum re-use.

\section{Results}\label{results}

We collected data from 547 ads for tenure-track jobs in archaeology
posted during 2013--2023. We focus our analysis here on the 431 ads for
positions at U.S. universities. Figure~\ref{fig-show-basic-plots} shows
the count of ads for each year, where year refers to the year the job ad
was posted. Table~\ref{tbl-show-basic-counts} shows the breakdown by
different job types, ranks, and tenure status. Assistant Professor jobs
are consistently the most common title and rank of positions advertised,
while open rank or full professor are the least frequent. The ratio of
tenure-track to non-tenure-track positions is generally well above one;
in other words, this data set is dominated by tenure-track positions.
Only academic year 2013--2014 had more non-tenure-track positions than
tenure-track positions, which was followed by an upward trend peaking at
2018--2019 and then declining again into the present.

\begin{figure}

\centering{

\includegraphics[width=6.3in,height=\textheight]{../figures/fig-panel-per-year.png}

}

\caption{\label{fig-show-basic-plots}A: total number of job ads from US
institutions posted to the Academic Jobs Wiki for Archaeology in each
year, with coloured sections showing the proportion of jobs by title and
rank. B: Ratio of tenure-track to non-tenure-track positions over time.
The red line indicates a 1:1 ratio of tenure-track to non-tenure-track
positions, bars taller than that line indicate more tenure-track than
non-tenure-track positions in that year}

\end{figure}%

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{2pt}

\renewcommand*{\arraystretch}{1.5}

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

\begin{longtable}[c]{|p{0.84in}|p{1.02in}|p{1.05in}|p{0.78in}|p{0.68in}|p{1.45in}|p{0.30in}|p{0.37in}}

\caption{\label{tbl-show-basic-counts}Breakdown of counts of job ads by
rank and tenure status. Values in parentheses are for the United States
only. TT = tenure-track, US-only, NTT = non-tenure-track, US and
elsewhere. Not all ads include unambiguous information about tenure
status, so the sum of TT and NTT does not equal the sum of all jobs in
all ranks.}

\tabularnewline

\ascline{1.5pt}{666666}{1-8}

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Year\ ad\ posted}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Assistant\ Professor}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Associate\ Professor}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Full\ Professor}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Open\ Rank}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Other\ (Curator,\ Director,\ etc.)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{TT}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{NTT}}}} \\

\ascline{1.5pt}{666666}{1-8}\endfirsthead

\ascline{1.5pt}{666666}{1-8}

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Year\ ad\ posted}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Assistant\ Professor}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Associate\ Professor}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Full\ Professor}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Open\ Rank}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{Other\ (Curator,\ Director,\ etc.)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{TT}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{NTT}}}} \\

\ascline{1.5pt}{666666}{1-8}\endhead

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2012-2013}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{27\ (26)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{7\ (7)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{-\ (-)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{4\ (4)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{4\ (4)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{41}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{24}}}} \\

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2013-2014}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{19\ (14)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{5\ (2)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{3\ (3)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2\ (1)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{5\ (1)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{21}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{77}}}} \\

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2014-2015}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{42\ (36)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{6\ (6)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{3\ (3)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{9\ (8)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{5\ (4)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{57}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{35}}}} \\

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2015-2016}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{44\ (39)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{5\ (4)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{1\ (1)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2\ (2)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{5\ (3)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{49}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{25}}}} \\

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2016-2017}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{35\ (27)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{4\ (4)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{1\ (1)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2\ (2)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{-\ (-)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{34}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{25}}}} \\

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2017-2018}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{30\ (25)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{7\ (7)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{-\ (-)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{3\ (3)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{8\ (6)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{41}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{20}}}} \\

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2018-2019}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{35\ (33)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{5\ (4)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{1\ (1)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{1\ (1)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{5\ (4)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{43}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{14}}}} \\

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2019-2020}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{33\ (29)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2\ (2)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{1\ (-)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{3\ (3)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2\ (1)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{34}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{20}}}} \\

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2020-2021}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{14\ (11)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{3\ (-)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{1\ (1)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{-\ (-)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2\ (-)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{12}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{19}}}} \\

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2021-2022}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{33\ (29)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2\ (1)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2\ (2)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{5\ (2)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{8\ (5)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{39}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{27}}}} \\

\multicolumn{1}{>{\centering}m{\dimexpr 0.84in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{2022-2023}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.02in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{67\ (41)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.05in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{10\ (8)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.78in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{4\ (4)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.68in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{1\ (-)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 1.45in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{19\ (6)}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.3in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{59}}}} & \multicolumn{1}{>{\centering}m{\dimexpr 0.37in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Helvetica}{126}}}} \\

\ascline{1.5pt}{666666}{1-8}

\end{longtable}

\arrayrulecolor[HTML]{000000}

\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}

\global\setlength{\tabcolsep}{\Oldtabcolsep}

\renewcommand*{\arraystretch}{1}

\subsection{Characteristics of the hiring
institutions}\label{characteristics-of-the-hiring-institutions}

\begin{figure}

\centering{

\includegraphics[width=7.2in,height=\textheight]{../figures/fig-map-and-carnegie-classification-hand-edit.png}

}

\caption{\label{fig-show-map-of-hiring-institution-and-cc}A: Frequency
of hiring institution by Carnegie classification. B: Inset shows map of
the United States showing the count of tenure-track job ads posted by
all institutions in each state during 2013 -- 2023}

\end{figure}%

Panel A of Figure~\ref{fig-show-map-of-hiring-institution-and-cc} shows
the frequencies of hiring institutions according to their Carnegie
Classification, a framework for classifying U.S. colleges and
universities according to the types of degrees awarded, levels of
research activity, and topical foci (Shulman 2001). Doctoral
universities with high and very high research activity are by far the
most active institutions hiring archaeology faculty. Associate's
colleges, also known as community colleges, rarely post job ads for
archaeology faculty.

The geographic distribution of hiring institutions is shown in
Figure~\ref{fig-show-map-of-hiring-institution-and-cc} Panel B.
California posted almost twice as many job ads as the next most active
states. After California, the states that posted the most ads during
2013--2023 include New York, Texas, Pennsylvania, and Florida. These top
five states correspond to the top five most populous U.S. states,
suggesting that rates of hiring are approximately proportional to
population density. The top five states for job ads are also the five
states with highest number of degrees awarded in Anthropology (National
Center for Education Statistics 2025). Similarly, the lowest counts of
job ads were observed in states with the lowest populations: North
Dakota, South Dakota, Alaska, and Nebraska. No institutions in Montana
posted a job ad during this period. The implication here is that
job-seekers who are able to relocate to populous areas will have more
employment options.

\subsection{Geographic trends in job ads over
time}\label{geographic-trends-in-job-ads-over-time}

\begin{figure}

\centering{

\includegraphics[width=7.2in,height=\textheight]{../figures/fig-geo-focus-by-year.png}

}

\caption{\label{fig-show-geo-trends}A: Frequency of locations mentioned
in the text of the job ads. B: Popularity of locations in job ads over
time for locations that appear in 20 or more ads. Individual data points
are shown, overlain by a locally weighted regression line for each
location to indicate temporal trends.}

\end{figure}%

We recorded all geographic regions mentioned in the text where the
successful applicant should have expertise and be research active. Our
analysis focuses on those locations mentioned in 20 or more ads.
Overall, American locations dominate. Panel A of
Figure~\ref{fig-show-geo-trends} shows that a single region of the US,
the Southwest, occurs in more job ads than every other part of the world
except for the Mediterranean. The Southwest includes Arizona and New
Mexico, with portions of California, Colorado, Nevada, Oklahoma, Texas,
and Utah. It is archaeologically significant as the home of the
Ancestral Pueblo, Hohokam, and Mogollon peoples, who practiced
irrigation agriculture and lived in relatively large settlements
compared to other regions of the U.S. The area was later occupied by the
Navajo, Ute, Southern Paiute, Hopi and Zuni, groups who had similarly
high population densities (Griffin-Pierce 2000). The Mediterranean is
prominent because it is the region that is often mentioned in job ads
focused on classical archaeology (i.e.~archaeology of Bronze Age and
Iron Age Italy and Greece).

Demand for jobs focusing on the Americas has been generally high over
the past decade, with a peak in 2019--2020 and a subsequent decrease.
Demand for jobs focusing on Africa was very low until 2019--2020,
peaking in 2020--2021. The proportion of ads with a geographic focus on
the Mediterranean has varied substantially, peaking in 2016 and
experiencing a nadir in 2019, showing an inverse pattern to the
Americas. Asia and India, the Near East, and Europe rarely occur as a
geographical focus in job ads at U.S. institutions. Asia and India,
Africa, and the Americas appear correlated with each other, while the
Near East and Mediterranean are inversely correlated in an opposite
trend.

\subsection{Method trends in job ads over
time}\label{method-trends-in-job-ads-over-time}

\begin{figure}

\centering{

\includegraphics[width=7.2in,height=\textheight]{../figures/fig-method-focus-by-year.png}

}

\caption{\label{fig-show-metho-trends}A: Frequency of methods mentioned
in the text of the job ads. B: Popularity of methods in job ads over
time for methods that appear in 10 or more ads. Individual data points
are shown, overlain by a locally weighted regression line for each
location to indicate temporal trends.}

\end{figure}%

Landscape archaeology, encompassing GIS and remote sensing, has remained
prominent compared to other methods Figure~\ref{fig-show-metho-trends}.
The popularity of this suite of skills may reflect its demand by
employers in the Cultural Research Management (CRM) sector. Morgan
(2023) found that almost one-third of 599 jobs ads posted by CRM
employers sought candidates with experience in GIS. Methods focused on a
specific element of the archaeological record, such as Lithic analysis,
Zooarchaeology and Ceramics are among the least frequently mentioned in
job ads. Instead, more popular methods are those that are relevant to
multiple elements of the archaeological record (e.g.~Archaeobotany
encompasses macroscopic and microscopic plant remains; Bioarchaeology
often includes skeletal analysis, isotopes, proteins, etc.).

Landscape archaeology, although dominant, has fluctuated over the years
and has been on a downtrend since 2018--2019. Computational and Digital
archaeology is the second most represented method, showing an overall
increasing trend, particularly since 2020--2021. Archaeobotany shows a
strong cyclical trend, rising, falling, and then rising again over our
study period. Archaeometry and Geoarchaeology have maintained a
relatively low but steady presence in job ads, peaking in 2017--2018 and
2018--2019 and declining thereafter. Lithic analysis and Zooarchaeology
are also mentioned relatively infrequently in job ads and show an
inverse correlation with each other after 2018--2019.

\subsection{Topic trends in job ads over
time}\label{topic-trends-in-job-ads-over-time}

\begin{figure}

\centering{

\includegraphics[width=7.2in,height=\textheight]{../figures/fig-topic-focus-by-year.png}

}

\caption{\label{fig-show-topi-trends}A: Frequency of topics mentioned in
the text of the job ads. B: Popularity of topics in job ads over time
for topics that appear in 20 or more ads. Individual data points are
shown, overlain by a locally weighted regression line for each location
to indicate temporal trends.}

\end{figure}%

The most frequently mentioned topic in this sample of job ads is
Environmental archaeology Figure~\ref{fig-show-topi-trends}. This
category encompasses such phrases as human-environmental dynamics,
interaction between humans and their environments, environmental change,
climate change, historical ecology, ecological knowledge, human ecology,
and ecological systems. Public archaeology is the second most frequent
topic overall; this category included phrases such as cultural resource
management, cultural heritage, heritage studies, museum studies, human
rights, community engaged, historic preservation, social justice,
community-based, repatriation, and community-engaged archaeology. The
least common topics in our sample are Pleistocene archaeology
(e.g.~human origins, hunter-gatherer archaeology) and Digital
archaeology.

The years 2019--2020 and 2020--2021 show striking changes in the
popularity of topics in job ads. Indigenous and historical archaeology
(which includes archaeology of the African diaspora and enslaved people)
became the most popular topic at this time, rising from being one of the
least popular topics from 2012--2017. Conversely, archaeological
science, which was popular during 2012--2017, was rarely mentioned in
job ads during 2019--2021. Ancient Europe and the Mediterranean, an
infrequently mentioned topic for the entire study period, virtually
disappeared from job ads during 2019--2021. Mentions of Complex
societies in ads decrease at a steady rate, while mentions of public
archaeology consistently increase over the study period. The topic of
Environmental Archaeology remains high over time.

\begin{figure}

\centering{

\includegraphics[width=7.2in,height=\textheight]{../figures/fig-topic-cooc-heatmap.png}

}

\caption{\label{fig-show-cooc}Heatmap of topic co-occurrence in job
ads.}

\end{figure}%

In our sample, job ads were more topically rich than geographically or
methodologically rich. That is, ads were more likely to mention multiple
topics than they were to mention multiple methods or geographic
locations. A Kruskal-Wallis test indicated significantly higher richness
in topics compared to richness of geographic locations or methods in job
ads (χ2 (df = 1, N = 836) = 160.42, p =
9\emph{x}10\textsuperscript{-37}). Figure~\ref{fig-show-cooc} shows
topic co-occurrences in our sample. Indigenous and historical
archaeology often occurs in job ads with Public archaeology and North
American archaeology. Complex societies and Environmental archaeology
were frequently found in the same ads. Bioarchaeology, Archaeological
science, and Evolutionary archaeology are another cluster of topics that
frequently co-occur. Other topics are relatively isolated. For example,
Pleistocene archaeology and Digital archaeology rarely occur with other
topics.

\subsection{Instructions to applicants over
time}\label{instructions-to-applicants-over-time}

\begin{figure}

\centering{

\includegraphics[width=10.8in,height=\textheight]{../figures/fig-requirements-per-year.png}

}

\caption{\label{fig-requirements-over-time}Changing requirements in job
ads over time. `Count' refers to the number of each item required.}

\end{figure}%

Over our ten year study period there have been substantial changes in
the instructions to applicants in terms of the type and number of
documents that are requested by the search committees
Figure~\ref{fig-requirements-over-time}. Requests for cover letter and
CV decline slightly in more recent years, perhaps reflecting their
status as a generic and expected component to submit without needing to
be specifically requested. The requirement for a diversity statement is
rare until 2019--2020, peaks around 2020--2021, then decreases towards
the present. Requests for names of recommenders (either zero or three
names, rarely only two names) reaches a maximum during 2019--2020, then
also decline for the remainder of the study period. The requirement for
a research statement and teaching statement increases after 2015--2016,
and becomes more frequent in job ads in more recent years. Requests for
course descriptions, syllabus samples, teaching evaluations, transcripts
and writing samples are consistently low over time (not shown here).

\section{Discussion}\label{discussion}

Our study of a decade of tenure-track job ads in archaeology in the U.S.
reveals diverse dynamics in the demand for specializations in topics,
methods and geographic focus, and in the instructions to applicants.
While these dynamics are familiar to scholars actively seeking faculty
positions, we believe this is the first time they have been quantified
at such a large scale within archaeology. Trends in job ads reflect
broader shifts in intellectual and practical priorities concerning
archaeology, undergraduate education, and the process of hiring
professors. The demand for archaeology faculty, indicated by the total
number of tenure-track jobs, may be affected by a variety of factors.

Overall, we found more tenure-track jobs advertised each year than
non-tenure-track, with the exception of 2013--2014. The wiki's emphasis
on tenure-track jobs runs counter to well-established patterns in the
academic job market, wherein short-term and contingent positions are
fast outpacing ``the last good job{[}s{]} in America'' (Aronowitz 2001).
Research across anthropological sub-fields has shown that the number of
tenure-track positions advertised within the discipline has also
declined. Analyzing the job ads posted in \emph{Anthropology News} in
1999--2000 versus 2019--2020, for example, Gershon and Rachok (2021)
found that the number of tenure-track positions decreased by 25\%
between their two time slices. Similarly, in his examination of jobs ads
posted on the Biological Anthropology Academic Jobs Wiki between
2010--2017, Passalacqua noted that approximately one quarter of the 474
ads posted were for adjunct or visiting staff (2018:773). Though
Passalacqua did not examine how the proportion of job types changed over
time, he did underscore ``a current diverging trend of decreasing
academic job advertisements and increasing doctoral degrees in
biological anthropology which could result in serious consequences for
the discipline'' (2018:773). Such observations have a surprising
longevity. Rogge raised the alarm nearly half a century ago, warning of
the ``maladaptive'' consequences of the exponential rate of disciplinary
growth and impending consequences for job seekers (Rogge 1976:839). The
dwindling number of permanent academic positions is not a trend unique
to anthropology---as of 2019, the AAUP documented a 36\% increase in
contingent positions over the preceding 15 years (AAUP 2022), and as of
2022, contingent positions make up more than half of faculty positions
in the United States (American Association of University Professors
2022b) .

This discrepancy in our data set---wherein advertisements for
tenure-track positions dominate the wiki despite their increasing rarity
in larger market---may be due to several factors. The first involves the
more limited circulation of advertising for short-term positions
relative to advertising for tenure-track jobs. Many of these short-term
positions are not advertised nationally, but are instead disseminated
through local email lists and are filled by people close to the hiring
department, such as recently graduated students. A second factor is bias
in our data---with most Wiki users likely seeking a tenure-track job,
non-tenure-track jobs may have been less frequently added to the
Academic Jobs Wiki because they were peripheral to the goal of most
users. This bias limits the reliability of our results on the ratio of
tenure-track to non-tenure-track positions. Tellingly, Passalacqua's
examination of a similar Wiki over a seven-year span showed a similar
dominance of permanent positions---291/474 (63\%) were for tenure-track
hires at various levels (2018: Table 2). This congruence between
subfields suggests that these patterns are a reflection of the Wiki,
rather than the market itself.

It is no secret that tenure-track jobs have traditionally been
considered the ``gold standard'' of career outcomes for PhD students in
the social sciences and humanities. As Kelsky outlines in \emph{The
Professor Is In}, her influential and best-selling guide for PhD
students pursuing faculty careers, ``\ldots most ranking graduate
programs still consider any PhD who doesn't land a tenure-track job a
failure or an aberration\ldots{} Graduate students absorb this value
system and judge themselves harshly'' (2015:11). These rubrics of
success have been common across the humanities and social sciences for
over half a century. Evaluating the demise of the National Endowment for
Humanities Funded ``Program to ready PhDs for Careers in Business''
(CIB), which aimed to facilitate humanities PhDs' transition to the
corporate world in the late 1970s, Franczak points to the dissonance
between participants' incentives and outcomes:

``The biggest lesson we can take from CIB is also the most obvious.
PhDs, especially in the humanities, want to be academics. The deep
reservoir of adjunct or contingent faculty that elite and non-elite
universities alike depend on for their courses is testament to this
fact---as is the excellent scholarship so many adjuncts produce without
department support. To pretend otherwise is disingenuous, and, as CIB
shows, possibly dangerous, too'' (Bessner and Brenes 2021:36).

The desire to pursue an academic career also characterizes more recent
PhD cohorts, including those within archaeology. Brami et al's survey
found that despite the long odds and professional precarity, 71\% of
their early career respondents wished to remain in academia (2023:242).
Overall, the obsessive valorization of permanent faculty positions in
the face of a precipitous decline in available jobs is indicative of the
potent combination of denial and delusion across academia. As Cefkin and
Schwegler argue ``This asymmetry between the `ideal' career path and the
experiences of most graduates highlights the deep and growing chasm
between the future that higher education institutions envision---and
strongly incentivize---and what is actually happening'' (Cefkin and
Schwegler 2024).

Returning to our results, the downward trend in tenure-track positions
during 2013--2019 may be related to declining undergraduate enrollment
in anthropology since 2013 (Cramb et al. 2022). The big dip during
2020--2021 is explained by the hiring freezes at many institutions
resulting from the COVID-19 pandemic, which caused extreme disruption
and uncertainty as universities focused on adapting to online
instruction and assessment in an effort to minimize the spread of the
virus (Woolston 2020). A survey of early career researchers in
archaeology captures the impact of this dip, with three-quarters of
respondents experiencing negative impacts on their careers due to the
pandemic (Brami et al. 2023).

The 2019--2021 period was also a major inflection point in the
popularity of specific topics and regions in job ads. Calls for
positions incorporating Indigenous and historical archaeology, and
archaeology of the Americas, became far more frequent at this time,
while Archaeological science, Complex societies, and the Mediterranean
and Near East showed declines in popularity. Similarly, the number of
job ads with a geographic focus on the Americas and Africa peaks during
2019--2021. These shifts in the topical and geographic foci of job ads
were likely influenced by broader cultural movements such as Black Lives
Matter, protests about racial injustice, and efforts to amplify
Indigenous voices (Dunivin et al. 2022; Flewellen et al. 2021; Franklin
et al. 2020; Laluk et al. 2022). COVID-19 also negatively impacted
Black, Indigenous American, and Hispanic communities with significantly
higher infection and morbidity rates, drawing attention to racial and
socio-economic inequality in the U.S. (Mackey et al. 2021; Tai et al.
2021). The Black Lives Matter movement, dating back to 2013, intersected
profoundly with the pandemic and the murder of George Floyd in
Minneapolis by a police officer in May 2020, three months after the
World Health Organization declared COVID-19 to be a pandemic. Mass
protests objecting to Floyd's murder generated widespread concern about
racial inequities and stimulated a broad interest in addressing systemic
racial injustice.

Our data suggest that archaeology faculty at U.S. universities
participated in this movement by adjusting their hiring plans to
prioritize recruiting archaeologists working on topics relevant to Black
and Indigenous communities. Many universities may have hoped to hire
Black and Indigenous archaeologists as part of their effort to tackle
systemic racism. Due to the Civil Rights Act of 1964 which prohibits
hiring based solely on race or ethnicity, however, it is illegal for
U.S. universities use race or ethnicity as a primary factor in hiring.
As a result, universities appear to have tailored the content of their
job ads to focus on topics where they expected Black and Indigenous
researchers to be most numerous, in an effort to ensure that these
researchers would be well-represented in the pool of applicants.

This striking change in topics and geographic foci of 2019--2021
occurred against a backdrop of several longer-term trends. Over the
course of the study period, we document a gradual decline in topics such
as Complex societies and Archaeological science, the geographic foci of
Mesoamerica and South America, and methods relating to Landscape
archaeology. These trends are harder to explain as we cannot link their
origins to a historical event such as the COVID-19 pandemic. We might
speculate that a growing preference for archaeological approaches that
privilege agency-driven, subjective, and relational perspectives is one
enduring legacy of debates in the 1980s and 90s about processualism
versus post-processualism (Fogelin 2019; Hodder 1999; Johnson 2019).
This theoretical trend might explain why archaeological science is
showing a decline, as demand for methods for analyzing artefact
materiality, ontology, and power displace physical laboratory methods
for technological, functional, and compositional analyses. Other factors
relevant to this decline may include the increasing difficulty of
obtaining research funding to support archaeological science research,
such as laboratory facilities and instrumentation. A decline in interest
in the archaeology of complex societies may reflect several themes that
intersect with broader social changes, such as growing interest in
Indigenous and non-state actors in the past and an increased concern
with climate change, environmental sustainability, and resilience,
shifting attention away from the study of monumental architecture, elite
societies, political hierarchies, and state systems.

Our data on the requirements for applicants support prior findings that
the complexity of applications---and concomitantly, the labor required
to apply for tenure-track jobs---has gradually increased over time. This
trend is especially pronounced for Assistant Professor positions, which
make more demands on applicants than Associate and Full Professor
positions. Consistent with results from other studies (Gershon and
Rachok 2021), our project documented a growing demand for research and
teaching statements. Demand for diversity statements shows a unique
trajectory, peaking in 2020--2021 and declining into the present. This
may relate to the intersecting concerns about race, identity and class
inequalities emerging during the COVID-19 pandemic, which seem to have
reached a peak in 2020--2021 and then declined over time, resulting in
jettisoning requests for diversity statements as the most urgent period
of the pandemic moves into the past. Another factor here may be the
debates surrounding the experiment with diversity statements in hiring
at some UC campuses during 2016--2022. Some universities that previously
required a diversity statement dropped that requirement after 2022
(Guiden 2024). In our data we observed a reversal of the diversity
statement requirement at 4 universities. Those schools posted job ads
prior 2022 that did require a diversity statement and also posted an ad
in 2022 that did not require one.

One bright spot for applicants is the decline in requests for names of
recommenders in the initial application in recent years. This may be a
response to recent criticisms of the burden on the applicant of
preparing numerous complex job applications (e.g. Dennis et al. 2022).
In recognition of this burden, not only on applicants but also on
colleagues writing letters of recommendation over and over to support
applicants, many hiring committees now follow the recommendations of
Dennis et al. (2022) in only requesting names and letters of
recommendation at later stages of the hiring process, if at all. Showing
sensitivity to this burden, in 2020 the American Anthropological
Association issued guidance to academic departments that letters of
recommendation should not be requested in the initial application, but
should only be required from short-listed candidates (American
Anthropological Association 2020; Youngling and Gershon 2020).

A key limitation of our research is that it does not include an analysis
of the academic profiles of those who were eventually hired based on
this sample of job ads. Our results reveal collective aspirations for
the future of the field, but only from those writing the job ads,
typically faculty who are securely employed as tenured professors
serving on hiring committees at U.S. colleges and universities. Our
results do not show how these aspirations worked out in the topical and
geographic foci of the people who were actually hired for these
positions. Future work should consider interviewing faculty hired during
our study period to match up scholars with the ads to which they
applied. If the successful applicants can be identified, then we could
analyze the fit between the details of the job ad and the applicant's
research. Such an analysis would allow us to assess how effective job
ads are for driving change in the discipline by setting the topics of
courses that will be taught in undergraduate and graduate curricula, and
research that will be supported by universities. We also recognize the
possibility that the Academic Jobs Wiki does not capture all available
positions and likely over-represents certain types of institutions.
Future work should evaluate these biases by comparing entries on the
Wiki to other sources of information about the academic job market, such
as the American Anthropological Association's AnthroGuide, or data
collected directly from universities.

\section{Conclusion}\label{conclusion}

The data introduced in this paper were initially collected and analyzed
in spring and summer 2024 and the manuscript was drafted in autumn 2024.
Between submitting the manuscript and receiving our initial round of
revisions in April 2025, the landscape of U.S. higher education has
shifted dramatically. The first hundred days of the new federal
administration focused on instilling a climate of ``economic precarity
and legal uncertainty'' across the sector through a cthuluesque strategy
that included withholding federal funding for colleges and universities,
decimating the Department of Education, revoking international student
visas, and attacking Diversity Equity, and Inclusion (DEI) initiatives
as ``unconstitutional'' (Knox and Alonso 2025). This deliberately
destabilizing ``flood the zone'' approach (Broadwater 2025) shows no
signs of abating. In April 2025, the U.S. president signed an executive
order to overhaul the higher education accreditation system (The White
House 2025), signifying the administration's continued commitment to
align higher education institutions to its ideological agenda.

Such challenges to the autonomy and self-governance in higher education
are particularly remarkable given that many architects of the current
administration hold degrees from the very institutions they are
attempting to dismantle, from Donald Trump (Wharton School of Business),
to J.D. Vance (Yale Law School), Chris Rufo (Georgetown), Stephen Miller
(Duke University), and Peter Hegseth (Harvard Kennedy School). Reviewing
the profiles of the board of trustees for the Heritage Foundation---the
conservative think tank which generated the Project 2025 report that has
been used as a blueprint for feeding U.S. higher education ``into a
woodchipper'' (Kunder 2025)---is also illustrative. Two thirds of the
trustees on the Heritage Foundation board have advanced degrees, while
one third have doctoral degrees from institutions such as Harvard,
Oxford, the University of Colorado, and the University of Texas. Lindsey
Burke, who authored the Project 2025 chapter on the Department of
Education, holds a PhD from George Mason University. Such career
trajectories are a testament to the elite chiasmus that decries the
merits and value of higher education while leveraging the credentials
granted by elite institutions for professional gain. As Ho notes of the
parallel systems of credentialing that undergird the ideology of
``smartness'' among Wall Street investment banks, playing the role of
``master of the universe'' requires both ``especially strong doses of
self-confidence and institutional legitimation'' (2009:41). Even
archaeology is susceptible to the power and pull of such prestige
hierarchies, which influence everything from hiring networks (Kawa et
al. 2019; Speakman, Hadden, Colvin, Cramb, Jones, Jones, Lulewicz, et
al. 2018) to publishing decisions (Beck et al. 2021) to research design
(Wobst and Keene 1983).

To echo Paul Farmer, ``all this is both interesting and horrible''
(2004:316). Contextualizing our job ad findings in this sociopolitical
tumult provokes a suite of new questions about academic futures. Many
archaeologists now have urgent concerns about what topics are likely to
be ``safe'' and lead to secure ongoing employment, considering the
current political and funding landscape. Will states with stronger
cultural and environmental regulations offer more stability in
employment, and if so, which states are likely to be the safest? These
are legitimate and crucial questions, but predicting the future goes
beyond the scope of our examination of the past decade of hiring
dynamics in academic archaeology. Extrapolating future trends is
inherently uncertain, especially in a mercurial political and economic
landscape that is experiencing tectonic shifts in bureaucracy,
organization, and funding on a near-weekly basis. Our study does,
however, point to at least four findings that have implications for the
next ten years of hiring, despite the turbulent political climate of
2024--2025.

First, the macroeconomic uncertainty induced by the increased tariffs
introduced in early 2025 is likely to have both short-term and long-term
implications that slow economic growth (Schneider 2025), with effects
that will seep into the higher education sector. The last two periods of
major economic upheaval in the U.S.---the Great Recession and the
COVID-19 pandemic---had pronounced and negative consequences for
colleges and universities. These consequences were visible even within
academic archaeology. Speakman, Hadden, Colvin, Cramb, Jones, Jones,
Lulewicz, et al. (2018), for example, show the precipitous decline in
the proportion of PhDs hired into tenure-track anthropology departments
after the last recession (see the steep decline in Figure 5 between 2007
and 2009), resulting in the ``increased use of {[}non-tenure track
faculty{]} by higher education institutions, and more anthropology
doctorates entering the workforce'' (2018:12). As our results
demonstrate, during academic year 2020--2021, only 12 tenure-track jobs
in archaeology were posted on the Wiki, a 70\% decrease relative to the
mean of 39 per year posted during our survey period
Table~\ref{tbl-show-basic-counts}. As Knox and Alonso (2025) highlight,
hiring freezes and program cuts are already coming into effect just four
months into the new administration. Given past patterns, it is likely
that increased tariffs will result in tenure-track positions in academic
archaeology becoming even more scarce, and faculty and students should
prepare for this. Whether this signals a death knell for the discipline
or an opportunity for re-envisioning and reconfiguring the field (Cefkin
and Schwegler 2024) is up for debate.

Second, our results show that hiring in academic archaeology is
responsive to larger cultural and political shifts. The increased
popularity of Indigenous and historical archaeology from 2019--2021
reflects the prominence and impact of movements such as Black Lives
Matter and a then-growing interest in conversations about race and
national history. The continual responsiveness to current events
revealed by our analysis of archaeology job ads demonstrates that our
narratives of the past are entangled in the social and political milieu
in which we work. This finding is consistent with previous
work---Trigger (1984) outlined how archaeological research programs are
shaped by national histories, while Soffer (1983), Blakey (1983),
Meltzer (1983), and Wilk (1985) have interrogated the entanglement of
identity, ideology, national history, politics, and archaeology across
space and time. Our study of job ads shows an active effort by
archaeologists to take control of interpreting the past, offering an
example of how disciplinary choices can shape archaeological research
programs. That said, in early 2025 many researchers had their work
disrupted by cancellations of funding from the U.S. National Science
Foundation if their projects contained certain keywords (Mervis 2025),
and others were told to cease activities if their projects contain
keywords such as ``cultural relevance,'' ``institutional,''
``historically,'' ``socioeconomic,'' and ``systemic'' (Johnson et al.
2025). This unprecedented keyword vetting suggests that future topical
shifts in archaeological research may be as much an individual response
to top-down funding pressures (cf. Wobst and Keene 1983) as a
disciplinary response to the current historical moment.

Third, the high chronological resolution of our analysis relative to
previous studies of disciplinary trends (e.g. Lyman 2010) shows that the
rise and fall of some of the foci we observed span relatively short
periods of time. This limits the predictive power of our results and has
implications for prospective graduate students and their mentors. A
topic that is growing in popularity as a student begins their PhD may
have peaked and be in decline before they graduate. Methods seem to have
a much lower frequency of change in popularity relative to topical and
geographic foci. One implication of this finding is that a graduate
student who has invested in developing technical expertise in a method
during their studies, in addition to a topic and region, might be less
vulnerable to the vagaries of the job market than a student without a
distinct area of technical expertise in generating data from material
records of past human behaviour.

Finally, our work also reveals the market forces that shaped the early
careers of the next generation of academic mentors. Assistant and
associate professors who experienced the academic market as job seekers
themselves between 2012--2023 probably anticipate that their students
will need publications, external grants, teaching experience, interview
preparation, extensive and tailored portfolios of job markets, and the
ability to spend many years on the market before obtaining a permanent
position, if they obtain one at all. This raises the question of whether
mentors' experiences on a market characterized by a recession and a
pandemic will be enough to guide their students through the volatility
of the next decade, and how they should update their programs to prepare
their graduates for success.

\section{Acknowledgements}\label{acknowledgements}

Thanks to the anonymous contributors to the Academic Jobs Wiki, without
whose efforts to collect and organise hundreds of job ads, this paper
would not exist. To those of you who contributed to the wiki but did not
obtain a tenure-track job, and/or decided to pivot to another industry,
we hope this paper shows your time on the wiki was not wasted. Your
efforts have helped contribute to more accurate expectation-setting for
future generations of archaeologists.

\section{Data Availability Statement}\label{data-availability-statement}

The data that support the findings of this study are openly available in
Zenodo at https://doi.org/10.5281/zenodo.14798941

\section{Funding Statement}\label{funding-statement}

This research received no specific grant funding from any funding
agency, commercial or not-for-profit sectors.

\section{Competing Interests
Statement}\label{competing-interests-statement}

Competing interests: The authors declare none

\section{Supplementary materials}\label{supplementary-materials}

Our supplementary materials are available at
https://doi.org/10.5281/zenodo.14798941

\newpage

\section{References cited}\label{references-cited}

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\bibitem[\citeproctext]{ref-jobboar2020}
\CSLBlock{American Anthropological Association}
\CSLLeftMargin{ 2020}%
\CSLRightInline{Job board policies.
\url{https://employers.americananthro.org/static-page/10285/job-board-policies/},
accessed January 26, 2025.}

\bibitem[\citeproctext]{ref-colby2022annual}
\CSLBlock{American Association of University Professors}
\CSLLeftMargin{ 2022b}%
\CSLRightInline{The annual report on the economic status of the
profession, 2021--22. American Association of University Professors.
\url{https://www.aaup.org/report/annual-report-economic-status-profession-2021-22},
accessed January 26, 2025.}

\bibitem[\citeproctext]{ref-the20222022}
\CSLLeftMargin{ 2022a }%
\CSLRightInline{The 2022 AAUP survey of tenure practices.
\url{https://www.aaup.org/report/2022-aaup-survey-tenure-practices},
accessed January 26, 2025.}

\bibitem[\citeproctext]{ref-aronowitz2001}
\CSLBlock{Aronowitz, Stanley}
\CSLLeftMargin{ 2001}%
\CSLRightInline{\emph{The last good job in america: Work and education
in the new global technoculture}. Critical perspectives series. Rowman
\& Littlefield : Distributed by National Book Network, Lanham, Md.}

\bibitem[\citeproctext]{ref-barnett2010}
\CSLBlock{Barnett, George A., James A. Danowski, Thomas Hugh Feeley, and
Jordan Stalker}
\CSLLeftMargin{ 2010}%
\CSLRightInline{Measuring quality in communication doctoral education
using network analysis of faculty-hiring patterns. \emph{Journal of
Communication} 60(2):388--411.
DOI:\href{https://doi.org/10.1111/j.1460-2466.2010.01487.x}{10.1111/j.1460-2466.2010.01487.x}.}

\bibitem[\citeproctext]{ref-barrett2008mediawiki}
\CSLBlock{Barrett, Daniel J}
\CSLLeftMargin{ 2008}%
\CSLRightInline{\emph{MediaWiki}. O'Reilly Media, Inc.}

\bibitem[\citeproctext]{ref-beck}
\CSLBlock{Beck, Jess}
\CSLLeftMargin{ 2025}%
\CSLRightInline{Stubborn architecture: The jobs crisis in academic
anthropology. In. SAR Press, in press, Santa Fe.}

\bibitem[\citeproctext]{ref-beck2021}
\CSLBlock{Beck, Jess, Erik Gjesfjeld, and Stephen Chrisomalis}
\CSLLeftMargin{ 2021}%
\CSLRightInline{Prestige or Perish: Publishing Decisions in Academic
Archaeology. \emph{American Antiquity} 86(4):669--695.
DOI:\href{https://doi.org/10.1017/aaq.2021.64}{10.1017/aaq.2021.64}.}

\bibitem[\citeproctext]{ref-bessner2021}
\CSLBlock{Bessner, Daniel, and Michael Brenes}
\CSLLeftMargin{ 2021}%
\CSLRightInline{The academic jobs crisis: A forum. \emph{Passport: The
Society for Historians of American Foreign Relations Review}
52(1):3053.}

\bibitem[\citeproctext]{ref-blakey1983}
\CSLBlock{Blakey, Michael}
\CSLLeftMargin{ 1983}%
\CSLRightInline{Socio-political bias and ideological production in
historical archaeology. In, edited by Joan M. Gero, David M. Lacy, and
Michael L. Blakey, pp. 5--16. Research reports, department of
anthropology. University of Massachusetts, Amherst.}

\bibitem[\citeproctext]{ref-bramiPrecariousFutureReflections2023}
\CSLBlock{Brami, Maxime, Stephanie Emra, Antoine Muller, Bianca
Preda-Bălănică, Benjamin Irvine, Bogdana Milić, Aldo Malagó, Katie
Meheux, and Manuel Fernández-Götz}
\CSLLeftMargin{ 2023}%
\CSLRightInline{A precarious future: Reflections from a survey of early
career researchers in archaeology. \emph{European Journal of
Archaeology} 26(2):226--250.
DOI:\href{https://doi.org/10.1017/eaa.2022.41}{10.1017/eaa.2022.41}.}

\bibitem[\citeproctext]{ref-broadwater2025}
\CSLBlock{Broadwater, Luke}
\CSLLeftMargin{ 2025}%
\CSLRightInline{\href{https://www.nytimes.com/2025/01/28/us/politics/trump-policy-blitz.html}{Trump{'}s
{`}flood the zone{'} strategy leaves opponents gasping in outrage}.
\emph{The New York Times}.}

\bibitem[\citeproctext]{ref-castillo}
\CSLBlock{Castillo, Enrique del, Adam Meyers, and Peng Chen}
\CSLLeftMargin{ 2018}%
\CSLRightInline{A Social Network Analysis of the Operations
Research/Industrial Engineering Faculty Hiring Network. \emph{preprint
on arXiv.org}.
DOI:\href{https://doi.org/10.48550/arXiv.1803.00125}{10.48550/arXiv.1803.00125}.}

\bibitem[\citeproctext]{ref-cefkin}
\CSLBlock{Cefkin, Melissa, and Tara Schwegler}
\CSLLeftMargin{ 2024}%
\CSLRightInline{\href{https://www.insidehighered.com/opinion/career-advice/2024/05/09/graduate-work-must-focus-both-academic-and-applied-opinion}{A
nonapocalyptic vision of graduate education{'}s future}. \emph{Inside
Higher Ed}.}

\bibitem[\citeproctext]{ref-clauset2015}
\CSLBlock{Clauset, Aaron, Samuel Arbesman, and Daniel B. Larremore}
\CSLLeftMargin{ 2015}%
\CSLRightInline{Systematic inequality and hierarchy in faculty hiring
networks. \emph{Science Advances} 1(1):e1400005.
DOI:\href{https://doi.org/10.1126/sciadv.1400005}{10.1126/sciadv.1400005}.}

\bibitem[\citeproctext]{ref-crambChangingProfileTenureTrack2022a}
\CSLBlock{Cramb, Justin, Brandon T. Ritchison, Carla S. Hadden, Qian
Zhang, Edgar Alarcón-Tinajero, Xianyan Chen, K. C. Jones, Travis Jones,
Katharine Napora, Matthew Veres, and Victor D. Thompson}
\CSLLeftMargin{ 2022}%
\CSLRightInline{The changing profile of tenure-track faculty in
archaeology. \emph{Advances in Archaeological Practice} 10(4):371--381.
DOI:\href{https://doi.org/10.1017/aap.2022.8}{10.1017/aap.2022.8},
accessed October 16, 2024.}

\bibitem[\citeproctext]{ref-culver}
\CSLBlock{Culver, KC, and Adrianna Kezar}
\CSLLeftMargin{ 2022}%
\CSLRightInline{\href{https://doi.org/10.17226/26405}{The impacts of
2020 on advancement of non-tenure-track and adjunct faculty}. In
\emph{Promotion, tenure, and advancement through the lens of 2020:
Proceedings of a workshop---in brief}, edited by Maria Lund Dahlberg,
pp. 16--19. National Academies of Sciences, Engineering,; Medicine.}

\bibitem[\citeproctext]{ref-dennisWorstAnthroJob2022}
\CSLBlock{Dennis, Dannah, Dada Docot, Danielle Gendron, and Ilana
Gershon}
\CSLLeftMargin{ 2022}%
\CSLRightInline{The worst of anthro job ads for 2021. \emph{American
Anthropologist} 124(4):900--905.
DOI:\href{https://doi.org/10.1111/aman.13781}{10.1111/aman.13781},
accessed October 22, 2024.}

\bibitem[\citeproctext]{ref-dunivinBlackLivesMatter2022}
\CSLBlock{Dunivin, Zackary Okun, Harry Yaojun Yan, Jelani Ince, and
Fabio Rojas}
\CSLLeftMargin{ 2022}%
\CSLRightInline{Black lives matter protests shift public discourse.
\emph{Proceedings of the National Academy of Sciences}
119(10):e2117320119.
DOI:\href{https://doi.org/10.1073/pnas.2117320119}{10.1073/pnas.2117320119}.}

\bibitem[\citeproctext]{ref-earle2014}
\CSLBlock{Earle, Beverley, and Marianne DelPo Kulow}
\CSLLeftMargin{ 2014}%
\CSLRightInline{The {"}Deeply Toxic{"} Damage Caused by the Abolition of
Mandatory Retirement and Its Collision with Tenure in Higher Education:
A Proposal for Statutory Repair. \emph{Southern California
Interdisciplinary Law Journal}:369--418.}

\bibitem[\citeproctext]{ref-farmer2004}
\CSLBlock{Farmer, Paul}
\CSLLeftMargin{ 2004}%
\CSLRightInline{An Anthropology of Structural Violence. \emph{Current
Anthropology} 45(3):305--325.
DOI:\href{https://doi.org/10.1086/382250}{10.1086/382250}.}

\bibitem[\citeproctext]{ref-feeley2021}
\CSLBlock{Feeley, Thomas Hugh, and Frank Tutzauer}
\CSLLeftMargin{ 2021}%
\CSLRightInline{The faculty hiring network for PhD-granting
communication programs. \emph{Scientometrics} 126(5):3983--4003.
DOI:\href{https://doi.org/10.1007/s11192-021-03917-y}{10.1007/s11192-021-03917-y}.}

\bibitem[\citeproctext]{ref-flewellenFutureArchaeologyAntiracist2021}
\CSLBlock{Flewellen, Ayana Omilade, Justin P. Dunnavant, Alicia Odewale,
Alexandra Jones, Tsione Wolde-Michael, Zoë Crossland, and Maria
Franklin}
\CSLLeftMargin{ 2021}%
\CSLRightInline{{``}The Future of Archaeology Is Antiracist{''}:
Archaeology in the Time of Black Lives Matter. \emph{American Antiquity}
86(2):224--243.
DOI:\href{https://doi.org/10.1017/aaq.2021.18}{10.1017/aaq.2021.18}.}

\bibitem[\citeproctext]{ref-Fogelin2019}
\CSLBlock{Fogelin, Lars}
\CSLLeftMargin{ 2019}%
\CSLRightInline{An unauthorized companion to american archaeological
theory. Self-published. \url{https://arizona.academia.edu/LarsFogelin},
accessed January 26, 2025.}

\bibitem[\citeproctext]{ref-franklinFutureNowArchaeology2020}
\CSLBlock{Franklin, Maria, Justin P. Dunnavant, Ayana Omilade Flewellen,
and Alicia Odewale}
\CSLLeftMargin{ 2020}%
\CSLRightInline{The Future is Now: Archaeology and the Eradication of
Anti-Blackness. \emph{International Journal of Historical Archaeology}
24(4):753--766.
DOI:\href{https://doi.org/10.1007/s10761-020-00577-1}{10.1007/s10761-020-00577-1}.}

\bibitem[\citeproctext]{ref-HelloTristesTropes2021}
\CSLBlock{Gershon, Ilana, and Dafna Rachok}
\CSLLeftMargin{ 2021}%
\CSLRightInline{Hello to {Tristes Tropes}. Anthropology News.
\url{https://www.anthropology-news.org/articles/hello-to-tristes-tropes/},
accessed August 2, 2022.}

\bibitem[\citeproctext]{ref-Ghaffarzadegan_2015}
\CSLBlock{Ghaffarzadegan, Navid, Joshua Hawley, Richard Larson, and Yi
Xue}
\CSLLeftMargin{ 2015}%
\CSLRightInline{A note on PhD population growth in biomedical sciences.
\emph{Systems Research and Behavioral Science} 32(3):402--405.
DOI:\href{https://doi.org/10.1002/sres.2324}{10.1002/sres.2324}.}

\bibitem[\citeproctext]{ref-graeber2018}
\CSLBlock{Graeber, David}
\CSLLeftMargin{ 2018}%
\CSLRightInline{\emph{Bullshit jobs}. Simon \& Schuster, New York.}

\bibitem[\citeproctext]{ref-griffin2000native}
\CSLBlock{Griffin-Pierce, Trudy}
\CSLLeftMargin{ 2000}%
\CSLRightInline{\emph{Native peoples of the southwest}. UNM Press.}

\bibitem[\citeproctext]{ref-guidenAreDiversityStatements}
\CSLBlock{Guiden, Mary}
\CSLLeftMargin{ 2024}%
\CSLRightInline{Are diversity statements a thing of the past? Harvard is
the latest to drop this requirement for job applicants.
\url{https://www.higheredjobs.com/Articles/articleDisplay.cfm?ID=3948},
accessed November 1, 2024.}

\bibitem[\citeproctext]{ref-gusterson2017}
\CSLBlock{Gusterson, Hugh}
\CSLLeftMargin{ 2017}%
\CSLRightInline{Homework: Toward a critical ethnography of the
university: AES presidential address, 2017. \emph{American Ethnologist}
44(3):435--450.
DOI:\href{https://doi.org/10.1111/amet.12520}{10.1111/amet.12520}.}

\bibitem[\citeproctext]{ref-ho2009}
\CSLBlock{Ho, Karen Zouwen}
\CSLLeftMargin{ 2009}%
\CSLRightInline{\emph{Liquidated : An ethnography of wall street}. Duke
University Press, Durham.}

\bibitem[\citeproctext]{ref-hodderArchaeologicalProcessIntroduction1999}
\CSLBlock{Hodder, Ian}
\CSLLeftMargin{ 1999}%
\CSLRightInline{\emph{The archaeological process: An introduction}.
Blackwell, Oxford.}

\bibitem[\citeproctext]{ref-johnson2025}
\CSLBlock{Johnson, Carolyn Y., Scott Dance, and Joel Achenbach}
\CSLLeftMargin{ 2025}%
\CSLRightInline{\href{https://www.washingtonpost.com/science/2025/02/04/national-science-foundation-trump-executive-orders-words/}{Here
are the words putting science in the crosshairs of trump{'}s orders}.
\emph{The Washington Post}.}

\bibitem[\citeproctext]{ref-johnson2019archaeological}
\CSLBlock{Johnson, Matthew}
\CSLLeftMargin{ 2019}%
\CSLRightInline{\emph{Archaeological theory: An introduction}. John
Wiley \& Sons.}

\bibitem[\citeproctext]{ref-kaskie2016}
\CSLBlock{Kaskie, Brian}
\CSLLeftMargin{ 2016}%
\CSLRightInline{The Academy Is Aging in Place: Assessing Alternatives
for Modifying Institutions of Higher Education. \emph{The
Gerontologist}:gnw001.
DOI:\href{https://doi.org/10.1093/geront/gnw001}{10.1093/geront/gnw001}.}

\bibitem[\citeproctext]{ref-kawaSocialNetworkUS2019}
\CSLBlock{Kawa, Nicholas C., José A. Clavijo Michelangeli, Jessica L.
Clark, Daniel Ginsberg, and Christopher McCarty}
\CSLLeftMargin{ 2019}%
\CSLRightInline{The {Social Network} of {US Academic Anthropology} and
{Its Inequalities}. \emph{American Anthropologist} 121(1):14--29.
DOI:\href{https://doi.org/10.1111/aman.13158}{10.1111/aman.13158},
accessed October 31, 2024.}

\bibitem[\citeproctext]{ref-kelsky2015}
\CSLBlock{Kelsky, Karen}
\CSLLeftMargin{ 2015}%
\CSLRightInline{\emph{The professor is in: The essential guide to
turning your ph.d. Into a job}. Three Rivers Press, New York.}

\bibitem[\citeproctext]{ref-knox2025}
\CSLBlock{Knox, Liam, and Johanna Alonso}
\CSLLeftMargin{ 2025}%
\CSLRightInline{\href{https://www.insidehighered.com/news/government/politics-elections/2025/04/30/how-trumps-first-100-days-transformed-higher-ed}{Trump{'}s
100-day war on higher ed}. \emph{Inside Highe Ed}.}

\bibitem[\citeproctext]{ref-kuckartz2014qualitative}
\CSLBlock{Kuckartz, Udo}
\CSLLeftMargin{ 2014}%
\CSLRightInline{\emph{Qualitative text analysis: A guide to methods,
practice and using software}. Sage.}

\bibitem[\citeproctext]{ref-lalukArchaeologySocialJustice2022}
\CSLBlock{Laluk, Nicholas C., Lindsay M. Montgomery, Rebecca Tsosie,
Christine McCleave, Rose Miron, Stephanie Russo Carroll, Joseph Aguilar,
Ashleigh Big Wolf Thompson, Peter Nelson, Jun Sunseri, Isabel Trujillo,
GeorgeAnn M. DeAntoni, Greg Castro, and Tsim D. Schneider}
\CSLLeftMargin{ 2022}%
\CSLRightInline{Archaeology and Social Justice in Native America.
\emph{American Antiquity} 87(4):659--682.
DOI:\href{https://doi.org/10.1017/aaq.2022.59}{10.1017/aaq.2022.59}.}

\bibitem[\citeproctext]{ref-larson2014}
\CSLBlock{Larson, Richard C., Navid Ghaffarzadegan, and Yi Xue}
\CSLLeftMargin{ 2014}%
\CSLRightInline{Too Many PhD Graduates or Too Few Academic Job Openings:
The Basic Reproductive Number R0 in Academia. \emph{Systems Research and
Behavioral Science} 31(6):745--750.
DOI:\href{https://doi.org/10.1002/sres.2210}{10.1002/sres.2210}.}

\bibitem[\citeproctext]{ref-lightfoot2021preparing}
\CSLBlock{Lightfoot, Elizabeth, Cynthia Franklin, and Raiza Beltran}
\CSLLeftMargin{ 2021}%
\CSLRightInline{Preparing for the academic job market: A guide for
social work doctoral students and their mentors. \emph{Journal of Social
Work Education} 57(1):153--164.}

\bibitem[\citeproctext]{ref-lyman2010american}
\CSLBlock{Lyman, R Lee}
\CSLLeftMargin{ 2010}%
\CSLRightInline{American archaeology textbooks as reflections of the
history of the discipline. \emph{North American Archaeologist}
31(1):1--25.}

\bibitem[\citeproctext]{ref-mackeyRacialEthnicDisparities2021}
\CSLBlock{Mackey, Katherine, Chelsea K. Ayers, Karli K. Kondo, Somnath
Saha, Shailesh M. Advani, Sarah Young, Hunter Spencer, Max Rusek,
Johanna Anderson, Stephanie Veazie, Mia Smith, and Devan Kansagara}
\CSLLeftMargin{ 2021}%
\CSLRightInline{Racial and ethnic disparities in
COVID-19{\textendash}related infections, hospitalizations, and deaths.
\emph{Annals of Internal Medicine} 174(3):362--373.
DOI:\href{https://doi.org/10.7326/M20-6306}{10.7326/M20-6306}.}

\bibitem[\citeproctext]{ref-mackieMarketShareAccounting2023}
\CSLBlock{Mackie, Madeline E., and Heather Rockwell}
\CSLLeftMargin{ 2023}%
\CSLRightInline{Beyond market share: Accounting for doctoral program
size in recent rates of anthropology faculty job placement. \emph{PLOS
ONE} 18(5):e0285330.
DOI:\href{https://doi.org/10.1371/journal.pone.0285330}{10.1371/journal.pone.0285330}.}

\bibitem[\citeproctext]{ref-main2019}
\CSLBlock{Main, Joyce B., Sarah Prenovitz, and Ronald G. Ehrenberg}
\CSLLeftMargin{ 2019}%
\CSLRightInline{In Pursuit of a Tenure-Track Faculty Position: Career
Progression and Satisfaction of Humanities and Social Sciences
Doctorates. \emph{The Review of Higher Education} 42(4):1309--1336.
DOI:\href{https://doi.org/10.1353/rhe.2019.0067}{10.1353/rhe.2019.0067}.}

\bibitem[\citeproctext]{ref-Marwick2017}
\CSLBlock{Marwick, Ben}
\CSLLeftMargin{ 2017}%
\CSLRightInline{Computational reproducibility in archaeological
research: Basic principles and a case study of their implementation.
\emph{Journal of Archaeological Method and Theory} 24(2):424--450.
DOI:\href{https://doi.org/10.1007/s10816-015-9272-9}{10.1007/s10816-015-9272-9}.}

\bibitem[\citeproctext]{ref-meltzer1983}
\CSLBlock{Meltzer, David}
\CSLLeftMargin{ 1983}%
\CSLRightInline{Prehistory, power and politics in the bureau of american
ethnology, 1879{\textendash}1906. In, edited by Joan M. Gero, David M.
Lacy, and Michael L. Blakey, pp. 6777. Research reports, department of
anthropology. University of Massachusetts, Amherst.}

\bibitem[\citeproctext]{ref-Mervis}
\CSLBlock{Mervis, Jeffrey}
\CSLLeftMargin{ 2025}%
\CSLRightInline{NSF's grant cuts fall heaviest on scientists from
underrepresented groups.
\url{https://www.science.org/content/article/nsf-s-grant-cuts-fall-heaviest-scientists-underrepresented-groups}.}

\bibitem[\citeproctext]{ref-mirowski2011}
\CSLBlock{Mirowski, Philip}
\CSLLeftMargin{ 2011}%
\CSLRightInline{\emph{Science-mart: Privatizing american science}.
Harvard University Press, Cambridge, Mass.}

\bibitem[\citeproctext]{ref-morganReadyNotArchaeological2023}
\CSLBlock{Morgan, Rachel}
\CSLLeftMargin{ 2023}%
\CSLRightInline{Ready or Not: An Archaeological Knowledge, Skills, and
Abilities Needs Assessment. \emph{Advances in Archaeological Practice}
11(4):371--387.
DOI:\href{https://doi.org/10.1017/aap.2023.21}{10.1017/aap.2023.21}.}

\bibitem[\citeproctext]{ref-musial2018five}
\CSLBlock{Musial, Jennifer, and Christina Holmes}
\CSLLeftMargin{ 2018}%
\CSLRightInline{Five-year study on hiring trends in gender, women's, and
feminist studies. \emph{Feminist Studies} 44(2):253--272.}

\bibitem[\citeproctext]{ref-IPEDSNCES}
\CSLBlock{National Center for Education Statistics}
\CSLLeftMargin{ 2025}%
\CSLRightInline{Integrated postsecondary education data system (IPEDS).
U.S. Department of Education. \url{https://nces.ed.gov/ipeds/}, accessed
January 25, 2025.}

\bibitem[\citeproctext]{ref-nevin2019}
\CSLBlock{Nevin, Andrew D.}
\CSLLeftMargin{ 2019}%
\CSLRightInline{Academic Hiring Networks and Institutional Prestige: A
Case Study of Canadian Sociology. \emph{Canadian Review of
Sociology/Revue canadienne de sociologie} 56(3):389--420.
DOI:\href{https://doi.org/10.1111/cars.12252}{10.1111/cars.12252}.}

\bibitem[\citeproctext]{ref-Passalacqua_2018}
\CSLBlock{Passalacqua, Nicholas V.}
\CSLLeftMargin{ 2018}%
\CSLRightInline{Are careers in biological anthropology sustainable?
\emph{American Journal of Physical Anthropology} 166(3):772--776.
DOI:\href{https://doi.org/10.1002/ajpa.23457}{10.1002/ajpa.23457}.}

\bibitem[\citeproctext]{ref-platzerAcademicPrecarityAmerican2018}
\CSLBlock{Platzer, David, and Anne Allison}
\CSLLeftMargin{ 2018}%
\CSLRightInline{Academic {Precarity} in {American Anthropology}. Society
for Cultural Anthropology.
\url{https://culanth.org/fieldsights/academic-precarity-in-american-anthropology},
accessed October 31, 2024.}

\bibitem[\citeproctext]{ref-rlanguage}
\CSLBlock{R Core Team}
\CSLLeftMargin{ 2024}%
\CSLRightInline{\emph{\href{https://www.R-project.org/}{R: A language
and environment for statistical computing}}. R Foundation for
Statistical Computing, Vienna, Austria.}

\bibitem[\citeproctext]{ref-rabinow1992}
\CSLBlock{Rabinow, Paul}
\CSLLeftMargin{ 1992}%
\CSLRightInline{For hire: Resolutely late modern. In, pp. 5971. School
of American Research Press, Santa Fe.}

\bibitem[\citeproctext]{ref-ribeiro2023}
\CSLBlock{Ribeiro, Artur, and Christos Giamakis}
\CSLLeftMargin{ 2023}%
\CSLRightInline{On Class and Elitism in Archaeology. \emph{Open
Archaeology} 9(1):20220309.
DOI:\href{https://doi.org/10.1515/opar-2022-0309}{10.1515/opar-2022-0309}.}

\bibitem[\citeproctext]{ref-rogge1976}
\CSLBlock{Rogge, A. E.}
\CSLLeftMargin{ 1976}%
\CSLRightInline{A Look at Academic Anthropology: Through a Graph Darkly.
\emph{American Anthropologist} 78(4):829--843.
DOI:\href{https://doi.org/10.1525/aa.1976.78.4.02a00070}{10.1525/aa.1976.78.4.02a00070}.}

\bibitem[\citeproctext]{ref-ryan2003techniques}
\CSLBlock{Ryan, Gery W, and H Russell Bernard}
\CSLLeftMargin{ 2003}%
\CSLRightInline{Techniques to identify themes. \emph{Field methods}
15(1):85--109.}

\bibitem[\citeproctext]{ref-schneider2025}
\CSLBlock{Schneider, Howard}
\CSLLeftMargin{ 2025}%
\CSLRightInline{\href{https://www.reuters.com/markets/us/feds-powell-weigh-amid-tariff-fray-market-drop-2025-04-04/}{Fed{'}s
powell says larger-than-expected tariffs likely to boost inflation, slow
growth}. \emph{Reuters}.}

\bibitem[\citeproctext]{ref-shulman2001carnegie}
\CSLBlock{Shulman, Lee S}
\CSLLeftMargin{ 2001}%
\CSLRightInline{\emph{The {C}arnegie classification of institutions of
higher education}. The Carnegie Foundation for the Advancement of
Teaching, California.}

\bibitem[\citeproctext]{ref-soffer1983}
\CSLBlock{Soffer, Olga}
\CSLLeftMargin{ 1983}%
\CSLRightInline{Politics of the paleolithic in the USSR: A case of
paradigms lost. In, edited by Joan M. Gero, David M. Lacy, and Michael
L. Blakey, pp. 91105. Research reports, department of anthropology.
University of Massachusetts, Amherst.}

\bibitem[\citeproctext]{ref-soucek2021diversity}
\CSLBlock{Soucek, Brian}
\CSLLeftMargin{ 2022}%
\CSLRightInline{\href{https://lawreview.law.ucdavis.edu/archives/55/4/diversity-statements.html}{Diversity
statements}. \emph{University of California Davis Law Review}
55:1989--2050.}

\bibitem[\citeproctext]{ref-speakman2018choosing}
\CSLBlock{Speakman, Robert J., Carla S. Hadden, Matthew H. Colvin,
Justin Cramb, K. C. Jones, Travis W. Jones, Corbin L Kling, Isabelle
Lulewicz, Katharine G. Napora, Katherine L. Reinberger, and others}
\CSLLeftMargin{ 2018}%
\CSLRightInline{Choosing a path to the ancient world in a modern market:
The reality of faculty jobs in archaeology. \emph{American Antiquity}
83(1):1--12.}

\bibitem[\citeproctext]{ref-speakman2018}
\CSLBlock{Speakman, Robert J., Carla S. Hadden, Matthew H. Colvin,
Justin Cramb, K. C. Jones, Travis W. Jones, Isabelle Lulewicz, Katharine
G. Napora, Katherine L. Reinberger, Brandon T. Ritchison, Alexandra R.
Edwards, and Victor D. Thompson}
\CSLLeftMargin{ 2018}%
\CSLRightInline{Market share and recent hiring trends in anthropology
faculty positions. Edited by John P. Hart. \emph{PLOS ONE}
13(9):e0202528.
DOI:\href{https://doi.org/10.1371/journal.pone.0202528}{10.1371/journal.pone.0202528}.}

\bibitem[\citeproctext]{ref-taiDisproportionateImpactCOVID192021}
\CSLBlock{Tai, Don Bambino Geno, Aditya Shah, Chyke A Doubeni, Irene G
Sia, and Mark L Wieland}
\CSLLeftMargin{ 2021}%
\CSLRightInline{The disproportionate impact of COVID-19 on racial and
ethnic minorities in the united states. \emph{Clinical Infectious
Diseases} 72(4):703--706.
DOI:\href{https://doi.org/10.1093/cid/ciaa815}{10.1093/cid/ciaa815}.}

\bibitem[\citeproctext]{ref-thewhitehouse2025}
\CSLBlock{The White House}
\CSLLeftMargin{ 2025}%
\CSLRightInline{\emph{\href{https://www.whitehouse.gov/fact-sheets/2025/04/fact-sheet-president-donald-j-trump-reforms-accreditation-to-strengthen-higher-education/}{Fact
sheet: President donald j. Trump reforms accreditation to strengthen
higher education}}.}

\bibitem[\citeproctext]{ref-Trevithick2010}
\CSLBlock{Trevithick, Alan}
\CSLLeftMargin{ 2010}%
\CSLRightInline{Anthropology and the new faculty majority: {Adjunct} and
contingent labor. \emph{Anthropology News} 51(9):4--4.
DOI:\href{https://doi.org/10.1111/j.1556-3502.2010.51904.x}{10.1111/j.1556-3502.2010.51904.x}.}

\bibitem[\citeproctext]{ref-trigger1984}
\CSLBlock{Trigger, Bruce}
\CSLLeftMargin{ 1984}%
\CSLRightInline{Alternative archaeologies: Nationalist, colonialist,
imperialist. \emph{Man} 19:355370.}

\bibitem[\citeproctext]{ref-wapmanQuantifyingHierarchyDynamics2022a}
\CSLBlock{Wapman, K. Hunter, Sam Zhang, Aaron Clauset, and Daniel B.
Larremore}
\CSLLeftMargin{ 2022}%
\CSLRightInline{Quantifying hierarchy and dynamics in {US} faculty
hiring and retention. \emph{Nature} 610(7930):120--127.
DOI:\href{https://doi.org/10.1038/s41586-022-05222-x}{10.1038/s41586-022-05222-x},
accessed November 1, 2024.}

\bibitem[\citeproctext]{ref-wilk1985ancient}
\CSLBlock{Wilk, Richard R}
\CSLLeftMargin{ 1985}%
\CSLRightInline{The ancient {M}aya and the political present.
\emph{Journal of Anthropological Research} 41(3):307--326.}

\bibitem[\citeproctext]{ref-wobst1983}
\CSLBlock{Wobst, H. Martin, and Arthur S. Keene}
\CSLLeftMargin{ 1983}%
\CSLRightInline{Archaeological explanation as political economy. In,
edited by Joan M. Gero, David M. Lacy, and Michael L. Blakey, pp. 7989.
Research reports, department of anthropology. University of
Massachusetts, Amherst.}

\bibitem[\citeproctext]{ref-woolstonJuniorResearchersHit2020}
\CSLBlock{Woolston, Chris}
\CSLLeftMargin{ 2020}%
\CSLRightInline{Junior researchers hit by coronavirus-triggered hiring
freezes. \emph{Nature} 582(7812):449--450.
DOI:\href{https://doi.org/10.1038/d41586-020-01656-3}{10.1038/d41586-020-01656-3}.}

\bibitem[\citeproctext]{ref-Youngling2020}
\CSLBlock{Youngling, Elizabeth, and Ilana Gershon}
\CSLLeftMargin{ 2020}%
\CSLRightInline{Let's make the academic job market more humane.
\emph{Anthropology News}.
DOI:\href{https://doi.org/10.1111/AN.1341}{10.1111/AN.1341}.}

\end{CSLReferences}

\newpage

\subsubsection{Colophon}\label{colophon}

This report was generated on 2025-07-08 13:23:12.980024 using the
following computational environment and dependencies:

\begin{verbatim}
- Session info ---------------------------------------------------------------
 setting  value
 version  R version 4.5.1 (2025-06-13)
 os       macOS Ventura 13.3.1
 system   aarch64, darwin20
 ui       X11
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       Asia/Shanghai
 date     2025-07-08
 pandoc   3.4 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.549 @ /usr/local/bin/quarto

- Packages -------------------------------------------------------------------
 ! package           * version date (UTC) lib source
 P askpass             1.2.1   2024-10-04 [?] RSPM (R 4.5.0)
 P beeswarm            0.4.0   2021-06-01 [?] RSPM (R 4.5.0)
 P bit                 4.6.0   2025-03-06 [?] RSPM (R 4.5.0)
 P bit64               4.6.0-1 2025-01-16 [?] RSPM (R 4.5.0)
 P cachem              1.1.0   2024-05-16 [?] RSPM (R 4.5.0)
 P cellranger          1.1.0   2016-07-27 [?] RSPM (R 4.5.0)
 P cli                 3.6.5   2025-04-23 [?] RSPM
 P cowplot           * 1.1.3   2024-01-22 [?] RSPM (R 4.5.0)
 P crayon              1.5.3   2024-06-20 [?] RSPM (R 4.5.0)
 P curl                6.2.2   2025-03-24 [?] RSPM (R 4.5.0)
 P data.table          1.17.0  2025-02-22 [?] RSPM (R 4.5.0)
 P devtools            2.4.5   2022-10-11 [?] RSPM
 P digest              0.6.37  2024-08-19 [?] RSPM (R 4.5.0)
 P dplyr             * 1.1.4   2023-11-17 [?] RSPM (R 4.5.0)
 P ellipsis            0.3.2   2021-04-29 [?] RSPM
 P evaluate            1.0.3   2025-01-10 [?] RSPM (R 4.5.0)
 P farver              2.1.2   2024-05-13 [?] RSPM (R 4.5.0)
 P fastmap             1.2.0   2024-05-15 [?] RSPM (R 4.5.0)
 P fiftystater       * 1.0.1   2025-05-07 [?] Github (wmurphyrd/fiftystater@28e7fa5)
 P flextable         * 0.9.7   2024-10-27 [?] RSPM (R 4.5.0)
 P fontBitstreamVera   0.1.1   2017-02-01 [?] RSPM (R 4.5.0)
 P fontLiberation      0.1.0   2016-10-15 [?] RSPM (R 4.5.0)
 P fontquiver          0.2.1   2017-02-01 [?] RSPM (R 4.5.0)
 P forcats           * 1.0.0   2023-01-29 [?] RSPM (R 4.5.0)
 P fs                  1.6.6   2025-04-12 [?] RSPM
 P gdtools             0.4.2   2025-03-27 [?] RSPM (R 4.5.0)
 P generics            0.1.3   2022-07-05 [?] RSPM (R 4.5.0)
 P ggbeeswarm        * 0.7.2   2023-04-29 [?] RSPM (R 4.5.0)
 P ggplot2           * 3.5.2   2025-04-09 [?] RSPM (R 4.5.0)
 P ggrepel           * 0.9.6   2024-09-07 [?] RSPM (R 4.5.0)
 P glue                1.8.0   2024-09-30 [?] RSPM (R 4.5.0)
 P gridExtra           2.3     2017-09-09 [?] RSPM (R 4.5.0)
 P gtable              0.3.6   2024-10-25 [?] RSPM (R 4.5.0)
 P here              * 1.0.1   2020-12-13 [?] RSPM (R 4.5.0)
 P hms                 1.1.3   2023-03-21 [?] RSPM (R 4.5.0)
 P htmltools           0.5.8.1 2024-04-04 [?] RSPM (R 4.5.0)
 P htmlwidgets         1.6.4   2023-12-06 [?] RSPM
 P httpuv              1.6.16  2025-04-16 [?] RSPM
 P httr                1.4.7   2023-08-15 [?] RSPM (R 4.5.0)
 P janitor             2.2.1   2024-12-22 [?] RSPM
 P jsonlite            2.0.0   2025-03-27 [?] RSPM
 P knitr               1.50    2025-03-16 [?] RSPM
 P labeling            0.4.3   2023-08-29 [?] RSPM (R 4.5.0)
 P later               1.4.2   2025-04-08 [?] RSPM
 P lattice             0.22-7  2025-04-02 [?] CRAN (R 4.5.1)
 P lifecycle           1.0.4   2023-11-07 [?] RSPM (R 4.5.0)
 P lubridate         * 1.9.4   2024-12-08 [?] RSPM (R 4.5.0)
 P magrittr            2.0.3   2022-03-30 [?] RSPM (R 4.5.0)
 P mapproj             1.2.11  2023-01-12 [?] RSPM
 P maps                3.4.2.1 2024-11-10 [?] RSPM
 P Matrix              1.7-3   2025-03-11 [?] RSPM
 P memoise             2.0.1   2021-11-26 [?] RSPM (R 4.5.0)
 P mgcv                1.9-3   2025-04-04 [?] CRAN (R 4.5.1)
 P mime                0.13    2025-03-17 [?] RSPM
 P miniUI              0.1.2   2025-04-17 [?] RSPM
 P nlme                3.1-168 2025-03-31 [?] RSPM
 P officer             0.6.8   2025-03-23 [?] RSPM (R 4.5.0)
 P openssl             2.3.2   2025-02-03 [?] RSPM (R 4.5.0)
 P pillar              1.10.2  2025-04-05 [?] RSPM (R 4.5.0)
 P pkgbuild            1.4.7   2025-03-24 [?] RSPM (R 4.5.0)
 P pkgconfig           2.0.3   2019-09-22 [?] RSPM (R 4.5.0)
 P pkgload             1.4.0   2024-06-28 [?] RSPM
   png                 0.1-8   2022-11-29 [2] CRAN (R 4.5.0)
 P profvis             0.4.0   2024-09-20 [?] RSPM
 P promises            1.3.2   2024-11-28 [?] RSPM
 P purrr             * 1.0.4   2025-02-05 [?] RSPM (R 4.5.0)
 P R6                  2.6.1   2025-02-15 [?] RSPM (R 4.5.0)
 P ragg                1.4.0   2025-04-10 [?] RSPM (R 4.5.0)
 P RColorBrewer        1.1-3   2022-04-03 [?] RSPM (R 4.5.0)
 P Rcpp                1.0.14  2025-01-12 [?] RSPM (R 4.5.0)
 P readr             * 2.1.5   2024-01-10 [?] RSPM (R 4.5.0)
 P readxl            * 1.4.5   2025-03-07 [?] RSPM (R 4.5.0)
 P remotes             2.5.0   2024-03-17 [?] RSPM
 P rgeolocate        * 1.4.2   2025-05-08 [?] Github (ironholds/rgeolocate@d2f550e)
 P rlang               1.1.6   2025-04-11 [?] RSPM
 P rlist             * 0.4.6.2 2021-09-03 [?] RSPM (R 4.5.0)
 P rmarkdown           2.29    2024-11-04 [?] RSPM (R 4.5.0)
 P rprojroot           2.0.4   2023-11-05 [?] RSPM (R 4.5.0)
 P rstudioapi          0.17.1  2024-10-22 [?] RSPM (R 4.5.0)
 P rvest             * 1.0.4   2024-02-12 [?] RSPM (R 4.5.0)
 P scales              1.4.0   2025-04-24 [?] RSPM
 P selectr             0.4-2   2019-11-20 [?] RSPM (R 4.5.0)
 P sessioninfo         1.2.3   2025-02-05 [?] RSPM
 P shiny               1.10.0  2024-12-14 [?] RSPM
 P snakecase           0.11.1  2023-08-27 [?] RSPM (R 4.5.0)
 P stringi           * 1.8.7   2025-03-27 [?] RSPM
 P stringr           * 1.5.1   2023-11-14 [?] RSPM (R 4.5.0)
 P svglite             2.1.3   2023-12-08 [?] RSPM (R 4.5.0)
 P systemfonts         1.2.3   2025-04-30 [?] RSPM
 P textshaping         1.0.1   2025-05-01 [?] RSPM
 P tibble            * 3.2.1   2023-03-20 [?] RSPM (R 4.5.0)
 P tidyr             * 1.3.1   2024-01-24 [?] RSPM (R 4.5.0)
 P tidyselect          1.2.1   2024-03-11 [?] RSPM (R 4.5.0)
 P tidyverse         * 2.0.0   2023-02-22 [?] RSPM (R 4.5.0)
 P timechange          0.3.0   2024-01-18 [?] RSPM (R 4.5.0)
 P tzdb                0.5.0   2025-03-15 [?] RSPM (R 4.5.0)
 P urlchecker          1.0.1   2021-11-30 [?] RSPM
 P usethis             3.1.0   2024-11-26 [?] RSPM
 P uuid                1.2-1   2024-07-29 [?] RSPM (R 4.5.0)
 P vctrs               0.6.5   2023-12-01 [?] RSPM (R 4.5.0)
 P vipor               0.4.7   2023-12-18 [?] RSPM (R 4.5.0)
 P viridis           * 0.6.5   2024-01-29 [?] RSPM (R 4.5.0)
 P viridisLite       * 0.4.2   2023-05-02 [?] RSPM (R 4.5.0)
 P vroom               1.6.5   2023-12-05 [?] RSPM (R 4.5.0)
 P withr               3.0.2   2024-10-28 [?] RSPM (R 4.5.0)
 P xfun                0.52    2025-04-02 [?] RSPM
 P xml2                1.3.8   2025-03-14 [?] RSPM (R 4.5.0)
 P xtable              1.8-4   2019-04-21 [?] RSPM
 P yaml                2.3.10  2024-07-26 [?] RSPM (R 4.5.0)
 P zip                 2.3.2   2025-02-01 [?] RSPM

 [1] /Users/bmarwick/Library/Caches/org.R-project.R/R/renv/library/archyjobads-ba2e7730/macos/R-4.5/aarch64-apple-darwin20
 [2] /Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library

 * -- Packages attached to the search path.
 P -- Loaded and on-disk path mismatch.

------------------------------------------------------------------------------
\end{verbatim}

The current Git commit details are:

\begin{verbatim}
Local:    main /Users/bmarwick/Desktop/archyjobads
Remote:   main @ origin (https://github.com/benmarwick/archyjobads)
Head:     [296ef58] 2025-07-08: revisions for 'accept with revision' decision, add renv
\end{verbatim}

\end{document}

### 2025-07-09 analysis/paper/article.qmd
out-width in the quarto chunk is the key to controlling png figure sizes in the docx

@@
 ```{r}
+#| eval: false
+
+# download the wiki pages so we can have them locally, in case they get deleted or heavily modified
+walk(
+urls_for_each_year[1:9],
+~.x %>%
+  read_html() %>% 
+  xml2::write_xml(., paste0("analysis/data/raw_data/", basename(.x), ".html")))
+
+```
+
+```{r}
 #| label: fig-panel-wiki-self-report-numbers
@@
 
-# get the dat
+# get the data
 # 2021-22 and 2022-2023
@@
 G: Breakdown of self-reported Wiki user categories for each year."
-#| fig-width: 3
-#| fig-height: 2
+#| out-width: "3in"
+#| out-height: "2in"
 
 library(cowplot)
+p_wiki <- 
 plot_grid(
@@
   ncol = 1)
+
+ggsave(plot = p_wiki,
+       filename = here("analysis",
+            "figures", 
+            "fig-panel-wiki-characteristics.png"),
+       bg ="white",
+       h = 5, # experiment with h and w to get the right size and proportion 
+       w = 7,
+       units = "in",
+       dpi = 900) # make the image nice and crisp
+
+# show figure from PNG file
+knitr::include_graphics(here("analysis",
+            "figures", 
+            "fig-panel-wiki-characteristics.png"))
 ```
@@
 
-```{r, fig.width = 1, fig.height = 1}
+```{r}
 #| label: fig-show-basic-plots
@@
 The red line indicates a 1:1 ratio of tenure-track to non-tenure-track positions, bars taller than that line indicate more tenure-track than non-tenure-track positions in that year"
-#| fig-width: 1
-#| fig-height: 1
+#| out-width: "2in"
+#| out-height: "2in"
 

### 2025-07-09 analysis/paper/article.qmd
figures at a good size in the docx now

@@
   fig-dpi: 1000
+  cache: true
 filters:
@@
 #| fig-cap: "Number of users self-reporting in each category of 'current users' per year"
-#| fig-width: 3
 
@@
          )
+
+# set the fig width that we want in the output docx (found this number by trial and error)
+out_width <- "6in"
 ```
@@
 G: Breakdown of self-reported Wiki user categories for each year."
-#| out-width: "3in"
-#| out-height: "2in"
+#| out-width: !expr out_width
 
@@
 The red line indicates a 1:1 ratio of tenure-track to non-tenure-track positions, bars taller than that line indicate more tenure-track than non-tenure-track positions in that year"
-#| out-width: "2in"
-#| out-height: "2in"
+#| out-width: !expr out_width
 
@@
 B: Inset shows map of the United States showing the count of tenure-track job ads posted by all institutions in each state during 2013 -- 2023"
-#| fig-width: 3
+#| out-width: !expr out_width
 
@@
 Individual data points are shown, overlain by a locally weighted regression line for each location to indicate temporal trends."
-#| fig-width: 3
+#| out-width: !expr out_width
 
@@
 #| fig-cap: "Heatmap of topic co-occurrence in job ads."
-#| fig-width: 3
+#| out-width: !expr out_width
 
@@
 #| fig-cap: "Changing requirements in job ads over time. 'Count' refers to the number of each item required."
-#| fig-width: 3
+#| out-width: !expr out_width
 

### 2025-07-09 analysis/paper/article.qmd
resubmitted after minor edits following 'accept with revisions' decisions from editor

@@
 
+# "The Thing" Inspired Palette (8 Colors)
+the_thing_palette <- c(
+  "#6488AC", # 1. Icy Blue
+  "#1A2A3A", # 2. Deep Shadow Blue/Black
+  "#E8702A", # 3. Flamethrower Orange
+  "#9D2323", # 4. Blood Red/Creature Tissue
+  "#8B7355", # 5. Outpost Brown
+  "#E1E1E1", # 6. Snowy Off-White/Grey
+  "#5F7C6A", # 7. Parka Green/Drab
+  "#D9A443"  # 8. Warning Yellow/Orange (Muted)
+)
+
 # how many jobs of each rank per year?
@@
   ylab("Number of\ntenure-track job ads") +
-  scale_fill_brewer(palette = "Dark2") +
+  scale_fill_manual(values = the_thing_palette) +
   guides(fill = guide_legend(nrow=2,
@@
       fill = job_title_simple) +
- # geom_line(linewidth = 2) +
   geom_col() +
@@
 # visualisations ----------------------------------
-small_base_size <- 4
+small_base_size <- 8
 
@@
 
-# "The Thing" Inspired Palette (8 Colors)
-the_thing_palette <- c(
-  "#6488AC", # 1. Icy Blue
-  "#1A2A3A", # 2. Deep Shadow Blue/Black
-  "#E8702A", # 3. Flamethrower Orange
-  "#9D2323", # 4. Blood Red/Creature Tissue
-  "#8B7355", # 5. Outpost Brown
-  "#E1E1E1", # 6. Snowy Off-White/Grey
-  "#5F7C6A", # 7. Parka Green/Drab
-  "#D9A443"  # 8. Warning Yellow/Orange (Muted)
-)
-
 how_many_people_use_the_wiki <- 
@@
        bg ="white",
-       h = 5, # experiment with h and w to get the right size and proportion 
+       h = 7, # experiment with h and w to get the right size and proportion 
        w = 7,

### 2025-07-09 analysis/paper/article.qmd
read from local versions of the wiki pages

@@
 walk(
-urls_for_each_year[1:9],
+urls_for_each_year,
 ~.x %>%
@@
 ```{r}
-#| label: fig-panel-wiki-self-report-numbers
-#| fig-cap: "Number of users self-reporting in each category of 'current users' per year"
-
-# get the data
+# get the data from our local versions of the pages
 # 2021-22 and 2022-2023
@@
 map(
-urls_for_each_year[1:9],
+paste0(here("analysis/data/raw_data/"), 
+       basename(urls_for_each_year)[1:9],
+       ".html"),
 ~.x %>%
@@
 current_user_tbl_2021_22 <-
-    urls_for_each_year[10]  %>%
+    paste0(here("analysis/data/raw_data/"), 
+       basename(urls_for_each_year)[10],
+       ".html")  %>%
       read_html() %>%
@@
 current_user_tbl_2022_2023 <-
-    urls_for_each_year[11] %>%
+   paste0(here("analysis/data/raw_data/"), 
+       basename(urls_for_each_year)[11],
+       ".html") %>%
       read_html() %>%
@@
                         ))
+```
+
+```{r}
+#| label: fig-panel-wiki-self-report-numbers
+#| fig-cap: "Number of users self-reporting in each category of 'current users' per year"
 

### 2025-07-09 README.Rmd
update readme

@@
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/archyjobads/master?urlpath=rstudio) [![DOI:10.17605/OSF.IO/J58K](https://img.shields.io/badge/DOI-10.17605/OSF.IO/J58KH-brightgreen.svg)](http://doi.org/10.17605/OSF.IO/J58KH)
+[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/archyjobads/main?urlpath=rstudio) [![DOI:10.17605/OSF.IO/J58K](https://img.shields.io/badge/DOI-10.17605/OSF.IO/J58KH-brightgreen.svg)](http://doi.org/10.17605/OSF.IO/J58KH)
 
@@
 - open the `.Rproj` file in RStudio
-- run `renv::install()` to ensure you have the packages this analysis depends on 
+- run `renv::restore()` to ensure you have the packages this analysis depends on 
 - finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console

### 2025-07-09 README.Rmd
update readme

@@
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/archyjobads/main?urlpath=rstudio) [![DOI:10.17605/OSF.IO/J58K](https://img.shields.io/badge/DOI-10.17605/OSF.IO/J58KH-brightgreen.svg)](http://doi.org/10.17605/OSF.IO/J58KH)
+[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/archyjobads/main?urlpath=rstudio) [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.15847809.svg)](https://doi.org/10.5281/zenodo.15847809)
 
@@
 > Ben Marwick, Anne Marie Poole, Ailin Zhang, Setareh Shafizadeh, and Jess Beck (2025). Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology. _American Antiquity_
+
+A pre-print of this paper is freely available on SocArXiv: https://osf.io/preprints/socarxiv/whkjs
 
@@
     illustrations
-  - [:file\_folder:
-    supplementary-materials](/analysis/supplementary-materials):
-    Supplementary materials including notes and other documents
-    prepared and collected during the analysis.
 

### 2025-07-09 README.Rmd
update readme

@@
 - run `renv::restore()` to ensure you have the packages this analysis depends on     
-- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console
+- finally, open `analysis/paper/article.qmd` and render to produce the `article.docx`, or run `rmarkdown::render("analysis/paper/article.qmd")` in the R console
 

### 2025-07-09 README.Rmd
add dockerfile and instructions

@@
 
-  - [:file\_folder: paper](/analysis/paper): PDF file of the poster we presented at the SAA. Quarto source document
-    for manuscript. Includes code to reproduce the figures and tables
+  - [:file\_folder: paper](/analysis/paper): PDF file of the poster we presented at the SAA. Quarto source document for manuscript. Includes code to reproduce the figures and tables
     generated by the analysis. 
@@
 
-## How to run in your browser or download and run locally
+## How to run the code in your browser
+
+You can run the R code in this project without downloading or installing anything.
+Click [here](https://mybinder.org/v2/gh/benmarwick/archyjobads/main?urlpath=rstudio) to open a Binder instance with RStudio running in your web broswer.
+Then follow the instructions below as if you were running RStudio locally on your computer (open the `.Rproj` file in RStudio, etc.).
+
+## How to download and run locally
 
@@
 - finally, open `analysis/paper/article.qmd` and render to produce the `article.docx`, or run `rmarkdown::render("analysis/paper/article.qmd")` in the R console
+
+## How to download and run locally in a Docker container
+
+This research compendium includes a [Dockerfile](Dockerfile) that
+specifies the computational environment used to generate the results
+presented in the paper. You can recreate this environment, which
+includes R, RStudio and all necessary R packages, on your computer by
+following these steps:
+
+1.  Install [Docker](https://www.docker.com/get-started/) on your
+    computer
+2.  Download the compendium, e.g. from GitHub by clicking on the green
+    `< > Code` button above, then click `Download ZIP`, then unzip on
+    your computer
+3.  Set your terminal working directory to the compendium, and run
+    `docker build -t archyjobads .` (be sure to include the period) to build 
+    the container on your computer, it will take a few minutes and require 
+    a fast internet connection.
+4.  Run in your terminal
+    `docker run --rm -it -e ROOT=TRUE -e PASSWORD=rstudio -dp 8787:8787 archyjobads`
+    to start the container on your computer
+5.  Go to <http://localhost:8787/> with your browser and enter `rstudio`
+    in both fields to start RStudio in your browser, running in the
+    Docker container with all the required depedencies installed
+6.  In RStudio, run
+    `rstudioapi::openProject("/home/rstudio/archyjobads")`
+    to open the compendium’s project. This project uses
+    [`renv`](https://rstudio.github.io/renv/) to document and manage
+    dependencies, this will be activated when the project is first
+    opened in RStudio.
+7.  Run
+    `rmarkdown::render("analysis/paper/article.qmd", output_dir='/tmp')`
+    to render the Quarto document that is the manuscript submitted for
+    publication
+8.  After the code has executed, view the rendered output document by
+    running `browseURL("/tmp/article.docx")`
+9.  When finished, to delete the Docker container from your computer,
+    run this in the terminal on your desktop:
+    `docker ps -aq | xargs docker stop | xargs docker rm`
 
